---
title: "Model temperature (number of excess hot days) effect on non-partner sexual violence, all female, Mozambique"
format: html
editor: visual
---

# Environment Setup

```{r setup}
# Packages
library(tidyverse)
library(sf)
library(spdep)
library(brms)
options(brms.backend = "cmdstanr", mc.cores = 4)
library(cmdstanr)
library(posterior)
library(ggdist)
library(bayesplot)
library(tidybayes)
# theme_set(bayesplot::theme_default(base_family = "sans", base_size=14))
library(patchwork)
library(paletteer)
library(tidytext)
library(reliabilitydiag)

# Folder Structure
root <- rprojroot::has_file("agile2025.Rproj")$make_fix_file()
temp_dir <- root("temp")
dir.create(temp_dir, showWarnings = FALSE)
raw_data_dir <- root("data/raw")
model_data_dir <- root("data/processed/model_data")
fig_dir <- root("figures")
dir.create(fig_dir, showWarnings = FALSE)
mod_dir <- root("models")
src_dir <- root("src")

# Climate + violence data file
clim_fp_filename <- sort(list.files(model_data_dir, pattern = "climate_fp_\\d", full.names = TRUE), decreasing = TRUE)[[1]]
clim_fp_viol_filename <- sort(list.files(model_data_dir, pattern = "climate_fp_violence", full.names = TRUE), decreasing = TRUE)[[1]]
cat(paste0("Using data files:\n", clim_fp_filename, "\n", clim_fp_viol_filename))
```

# Notebook settings

```{r}
# Settings
nb_name <-  "Model effect of excess number of hot days on non-partner sexual violence, all female, Mozambique"
seed <- sum(sapply(str_split(nb_name, pattern = "", simplify = TRUE), FUN = utf8ToInt))
set.seed(seed)

# Viz settings
col_palette <- c("#dc7726", "#eebb93", "#ecc046", "#f6e0a3", "#0292b7", "#81c9db", "#35A29F", "#97DECE")
label_map <- c(
  live_with_partner = "Live with partner (%)",
  school_level = "Education attainment (%)",
  rural = "Rural area (%)",
  over_18 = "Over 18 (%)",
  value_price_index_hist_anomaly_12m = "Months of food stress",
  value_price_index_reg_anomaly_12m = "Months of food stress",
  value_max_price_index_12m = "Max price of maize",
  value_price_index_stress_12m = "Food price stressed",
  dev_value_price_index_hist_anomaly_12m = "Excess Months of food stress",
  dev_value_price_index_reg_anomaly_12m = "Excess Months of food stress",
  dev_value_max_price_index_12m = "Excess Max price of maize",
  clim = "Excess hot days (std.)",
  sv = "NP-SV victimisation (%)"
)

# Var settings
countries <- c("Mozambique")
clim_var <- "dev_value_temp_above_1sd_12m"
clim_label <- "excess_hot_days"
viol_var <- "sv_np"
viol_label <- "np_viol"
pop_label <- "all_female"
covars <- c(
  "sample_weight",
  "country",
  "adm_name",
  "cluster",
  "rural",
  "over_18",
  "school_level",
  "live_with_partner",
  "value_price_index_hist_anomaly_12m",
  "value_price_index_reg_anomaly_12m",
  "value_max_price_index_12m",
  "value_price_index_stress_12m",
  "dev_value_price_index_hist_anomaly_12m",
  "dev_value_price_index_reg_anomaly_12m",
  "dev_value_max_price_index_12m",
  viol_var
)
vars <- c(covars, clim_var)

bin_covs <- c(
  "rural",
  "over_18",
  "live_with_partner" ,
  "value_price_index_stress_12m"
)
mono_covs <- c(
  "value_price_index_hist_anomaly_12m",
  "value_price_index_reg_anomaly_12m",
  "dev_value_price_index_hist_anomaly_12m",
  "dev_value_price_index_reg_anomaly_12m",
  "school_level"
)
cont_covs <- c(
  "clim",
  "value_max_price_index_12m",
  "dev_value_max_price_index_12m"
)

# Computation settings
adapt_delta = 0.95
iter = 2000
```

# Functions

```{r}
source(file.path(src_dir, "viz.R"))
source(file.path(src_dir, "model_check_utils.R"))

standardize <- function(x) {
  (x - mean(x)) / sd(x)
}
```

# Import Data

## Geometries

### Mozambique

```{r}
sf_moz <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Mozambique 2022 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% st_make_valid()
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Create combined shapes

```{r}
sf_all <- (
  bind_rows(
    sf_moz,
  )
  %>% st_transform(3857)
  %>% mutate(
    centroid = st_centroid(geometry),
    lon = st_coordinates(centroid)[,1],
    lat = st_coordinates(centroid)[,2]
  )
)
```

## Climate & Food Prices

```{r}
df_clim <- (
  read_csv(
    clim_fp_filename,
    show_col_type = FALSE
  )
  %>% filter(
    country %in% countries
  )
)
```

## Survey Data

```{r}
df_clim_viol <- (
  read_csv(
    clim_fp_viol_filename,
    show_col_type = FALSE
  )
  %>% filter(
    !is.na(sample_weight),
    country %in% countries
  )
  %>% group_by(country)
  %>% mutate(
    country_sample_weight = sum(sample_weight, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    sample_weight = (
      1E-3 
      * sample_weight 
      * (
        sum(sample_weight, na.rm = TRUE) 
        / (length(unique(country)) * country_sample_weight)
      )
    ),
    over_18 = as.numeric(age > 18)
  )
)
```

```{r}
df_clim_viol_adm <- (
  df_clim_viol
  %>% group_by(country, adm_name)
  %>% summarise(
    clim = mean(!!sym(clim_var), na.rm = TRUE),
    rural = 100 * sum(rural * sample_weight, na.rm = TRUE) / sum(sample_weight, na.rm = TRUE),
    live_with_partner = 100 *sum(live_with_partner * sample_weight, na.rm = TRUE) / sum(sample_weight, na.rm = TRUE),
    over_18 = 100 * sum(over_18 * sample_weight, na.rm = TRUE) / sum(sample_weight, na.rm = TRUE),
    school_level = sum(school_level * sample_weight, na.rm = TRUE) / sum(sample_weight, na.rm = TRUE),
    value_max_price_index_12m = mean(value_max_price_index_12m, na.rm = TRUE),
    value_price_index_hist_anomaly_12m = mean(value_price_index_hist_anomaly_12m, na.rm = TRUE),
    value_price_index_reg_anomaly_12m = mean(value_price_index_reg_anomaly_12m, na.rm = TRUE),
    value_price_index_stress_12m = mean(value_price_index_stress_12m, na.rm = TRUE),
    dev_value_max_price_index_12m = mean(dev_value_max_price_index_12m, na.rm = TRUE),
    dev_value_price_index_hist_anomaly_12m = mean(dev_value_price_index_hist_anomaly_12m, na.rm = TRUE),
    dev_value_price_index_reg_anomaly_12m = mean(dev_value_price_index_reg_anomaly_12m, na.rm = TRUE),
    dev_value_price_index_stress_12m = mean(dev_value_price_index_stress_12m, na.rm = TRUE),
    sv := 100 * sum(!!sym(viol_var) * sample_weight, na.rm = TRUE) / sum(sample_weight, na.rm = TRUE),
    .groups = "drop"
  )
  %>% merge(
    sf_all,
    by = c("country", "adm_name"),
    all.x = TRUE
  )
  %>% st_as_sf()
)
```

# Viz

## Covariate Maps

```{r}
#| fig-width: 10
#| fig-height: 10
p1 <- (
  ggplot(data = df_clim_viol_adm)
  + geom_sf(
    mapping = aes(fill = over_18),
    color = "transparent"
  )
  + scale_fill_distiller(
    palette = "Greys",
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    legend.background = element_blank(),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14)
  )
  + labs(
    fill = "Over 18 (%)"
  )
)

p2 <- (
  ggplot(data = df_clim_viol_adm)
  + geom_sf(
    mapping = aes(fill = rural),
    color = "transparent"
  )
  + scale_fill_paletteer_c(
    palette = "ggthemes::Orange-Blue Diverging",
    direction = -1
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    legend.background = element_blank(),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14)
  )
  + labs(
    fill = "Rural level (%)"
  )
)

p3 <- (
  ggplot(data = df_clim_viol_adm)
  + geom_sf(
    mapping = aes(fill = live_with_partner),
    color = "transparent"
  )
  + scale_fill_distiller(
    palette = "Blues",
    direction = -1
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    legend.background = element_blank(),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14)
  )
  + labs(
    fill = "Live with partner (%)"
  )
)

p4 <- (
  ggplot(data = df_clim_viol_adm)
  + geom_sf(
    mapping = aes(fill = school_level),
    color = "transparent"
  )
  + scale_fill_distiller(
    palette = "Greens",
    direction = -1
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    legend.background = element_blank(),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14)
  )
  + labs(
    fill = "School Attainment"
  )
)

p5 <- (
  ggplot(data = df_clim_viol_adm)
  + geom_sf(
    mapping = aes(fill = dev_value_price_index_reg_anomaly_12m),
    color = "transparent"
  )
  + scale_fill_viridis_c(
    option = "viridis",
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    legend.background = element_blank(),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14)
  )
  + labs(
    fill = "Food Stress Months"
  )
)

# layout <- "
#   ABE
#   CDE
# "
layout <- "
  ABC
  DEF
"

(
  (p1 + p2 + p3 + p4 + p5 + plot_spacer())
  + plot_layout(
    # widths = c(1, 1, 1.5),
    design = layout
  )
)
  
ggsave(
  path = fig_dir,
  filename = paste0("variable_comparison_", clim_label, "_", pop_label, "_", viol_label, "_food_test.png")
)
```

## Climate

```{r}
#| fig-width: 8
#| fig-height: 6
(
  ggplot()
  + geom_histogram(
    data = df_clim,
    mapping = aes(
      y = after_stat(ncount),
      x = !!sym(clim_var),
      fill = "clim"
    ),
    alpha = 0.7,
    bins = 20
  )
  + geom_histogram(
    data = df_clim_viol,
    mapping = aes(
      y = after_stat(ncount),
      x = !!sym(clim_var),
      fill = "clim_viol"
    ),
    alpha = 0.6,
    bins = 20
  )
  + geom_vline(
    xintercept = 0,
    color = "black",
    linetype = "dashed"
  )
  + facet_wrap(
    ~ adm_name,
    ncol = 4
  )
  + scale_fill_manual(
    values = c(
      "clim" = col_palette[2],
      "clim_viol" = col_palette[5]
    ),
    labels = c(
      "clim" = "Climate variable distribution",
      "clim_viol" = "Climate at survey"
    )
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    plot.background = element_blank(),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13),
    strip.background = element_blank(),
    strip.text = element_text(size = 13),
    legend.background = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 13),
    legend.position = "inside",
    legend.position.inside = c(0.5, 0.1)
  )
  + labs(
    y = "Counts (norm.)",
    x = "Excess hot days"
  )
)
```


```{r}
#| fig-width: 3.5
#| fig-height: 3

rng_clim <- range(df_clim_viol_adm$clim, na.rm = TRUE)

(
  df_clim_viol_adm
  %>% ggplot()
  + geom_sf(
    mapping = aes(fill = clim),
    color = "black",
    linewidth = 0.1
  )
  + paletteer::scale_fill_paletteer_c(
    palette = "ggthemes::Orange-Blue Diverging",
    direction = -1,
    guide = guide_colorbar(
      label.position = "left",
      title.position = "right",
      title.vjust = 1,
      title.hjust = 0.5,
      barheight = unit(0.9, "npc")
    ),
    limits = rng_clim,
    rescaler = ~ scales::rescale_mid(., mid = 0),
    # oob = scales::squish
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.margin = margin(l = 0, r = 25),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    legend.background = element_blank(),
    legend.position = "inside",
    legend.position.inside = c(1.30, 0.5),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14, angle = 90),
    legend.ticks = element_line(color = "black")
  )
  + labs(
    fill = "Excess hot days (std.)"
  )
)

ggsave(
  path = fig_dir,
  filename = paste0("map_heat_", clim_label, "_", pop_label, "_", viol_label, "_moz_food_test.png")
)
```

## Violence

```{r}
#| fig-width: 3.5
#| fig-height: 3

rng_sv <- range(df_clim_viol_adm$sv, na.rm = TRUE)

(
  df_clim_viol_adm
  %>% ggplot()
  + geom_sf(
    mapping = aes(fill = sv),
    color = "black",
    linewidth = 0.1
  )
  + paletteer::scale_fill_paletteer_c(
    palette = "grDevices::Rocket",
    direction = -1,
    guide = guide_colorbar(
      label.position = "left",
      title.position = "right",
      title.vjust = 1,
      title.hjust = 0.5,
      barheight = unit(0.9, "npc")
    ),
    limits = rng_sv,
    # labels = scales::label_percent(accuracy = 1)
    # rescaler = ~ scales::rescale_mid(., mid = 0),
    # oob = scales::squish
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.margin = margin(r = 25),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    legend.background = element_blank(),
    legend.position = "inside",
    legend.position.inside = c(1.30, 0.5),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14, angle = 90),
    legend.ticks = element_line(color = "black")
  )
  + labs(
    fill = label_map[["sv"]]
  )
)

ggsave(
  path = fig_dir,
  filename = paste0("map_sv_", clim_label, "_", pop_label, "_", viol_label, "_moz_food_test.png")
)
```

## Food Insecurity


```{r}
#| fig-width: 3.5
#| fig-height: 3

rng_sv <- range(df_clim_viol_adm$value_price_index_reg_anomaly_12m, na.rm = TRUE)

(
  df_clim_viol_adm
  %>% ggplot()
  + geom_sf(
    mapping = aes(fill = value_price_index_reg_anomaly_12m),
    color = "black",
    linewidth = 0.1
  )
  + paletteer::scale_fill_paletteer_c(
    palette = "grDevices::Purple-Yellow",
    guide = guide_colorbar(
      label.position = "left",
      title.position = "right",
      title.vjust = 1,
      title.hjust = 0.5,
      barheight = unit(0.9, "npc")
    ),
    limits = rng_sv,
    # labels = scales::label_percent(accuracy = 1)
    # rescaler = ~ scales::rescale_mid(., mid = 0),
    # oob = scales::squish
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.margin = margin(r = 25),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    legend.background = element_blank(),
    legend.position = "inside",
    legend.position.inside = c(1.30, 0.5),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14, angle = 90),
    legend.ticks = element_line(color = "black")
  )
  + labs(
    fill = label_map[["value_price_index_reg_anomaly_12m"]]
  )
)

ggsave(
  path = fig_dir,
  filename = paste0("map_fp_", clim_label, "_", pop_label, "_", viol_label, "_moz_food_test.png")
)
```

# Prepare Data

## Restrict locations

```{r}
sf_vacs <- (
  sf_all
  %>% filter(
    country %in% (df_clim_viol$country %>% unique()),
    adm_name %in% (df_clim_viol$adm_name %>% unique())
  )
)
```

## Extract model subsets

```{r}
(
  df_clim_viol
  %>% filter(
    female == 1,
    # age > 14
  )
  %>% select(all_of(vars))
  %>% na.omit()
  %>% merge(
    sf_vacs %>% as.data.frame() %>% select(country, adm_name),
    by = c("country", "adm_name"),
    all.x = TRUE
  )
  %>% ggplot()
  + geom_histogram(
    mapping = aes(x = value_price_index_reg_anomaly_12m)
  )
  
)
```


```{r}
df_model <- (
  df_clim_viol
  %>% filter(
    female == 1,
    # age > 14
  )
  %>% select(all_of(vars))
  %>% na.omit()
  %>% merge(
    sf_vacs %>% as.data.frame() %>% select(country, adm_name),
    by = c("country", "adm_name"),
    all.x = TRUE
  )
  %>% mutate(
    country = factor(country),
    adm_name = factor(adm_name),
    cluster = factor(cluster),
    school_level = factor(
      school_level,
      levels = sort(unique(school_level)),
      ordered = TRUE
    ),
    value_price_index_hist_anomaly_12m = factor(
      value_price_index_hist_anomaly_12m,
      levels = sort(unique(value_price_index_hist_anomaly_12m)),
      ordered = TRUE
    ),
    value_price_index_reg_anomaly_12m = factor(
      value_price_index_reg_anomaly_12m,
      levels = sort(unique(value_price_index_reg_anomaly_12m)),
      ordered = TRUE
    ),
    clim := scale(!!sym(clim_var))
  )
)
```

## Generate adjacency matrix

```{r}
nb_list <- poly2nb(sf_vacs)
W <- nb2mat(nb_list, style = "B", zero.policy = TRUE)
rownames(W) <- sf_vacs$adm_name
colnames(W) <- sf_vacs$adm_name
levs <- levels(df_model$adm_name)
W_ord <- W[levs, levs]
```

```{r}
edges <- (
  geostan::edges(W, unique_pairs_only = TRUE, shape = sf_vacs$geometry)
)
graph <- st_geometry(edges)
```

```{r}
plot(sf_vacs$geometry, lwd = .15)
plot(graph, add = TRUE, type = "p", col = "red")
plot(graph, add = TRUE, type = "l", col = "black")
```

## Spatial correlation in raw data

```{r}
#| fig-width: 3
#| fig-height: 3
plot_moran(df_model, sf_vacs, outcome_var = viol_var)
ggsave(
  path = fig_dir,
  filename = paste0("moran_plot_obs_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_food_test.png")
)
```

# Modeling

## Random Intercepts by admin level and cluster

```{r}
pars_ranint_adm_clu <- c(
  "b_Intercept",
  "b_rural",
  "b_live_with_partner",
  "b_over_18",
  "bsp_moschool_level",
  "sd_adm_name__Intercept",
  "sd_adm_name:cluster__Intercept",
  "b_clim"
)
```

```{r}
stanvars <- stanvar(5, "num_pred")

prior_list_ranint_adm_clu <- c(
  prior(student_t(3, -2, 2), class = "Intercept"),
  prior(normal(0, 0.8 / sqrt(num_pred)), class = "b"),
  prior(exponential(1/0.2), class = "sd"),
  prior(dirichlet(1), class = "simo", coef = "moschool_level1")
)

form_ranint_adm_clu <- bf(
  as.formula(
    paste(
      viol_var,
      "| weights(sample_weight)",
      "~",
      "(
        1 
        + rural
        + live_with_partner
        + over_18
        + mo(school_level)
        + clim
        + (1 | adm_name/cluster)
      )"
    )
  )
)
fit_ranint_adm_clu <- brm(
  formula = form_ranint_adm_clu,
  family = bernoulli("logit"),
  prior = prior_list_ranint_adm_clu,
  data = df_model,
  stanvars = stanvars,
  seed = seed,
  refresh = 1000,
  iter = iter,
  # control = list(adapt_delta = adapt_delta),
  # sample_prior = "only"
  sample_prior = TRUE
)
```

### Model Checks

```{r}
sp <- rstan::get_sampler_params(fit_ranint_adm_clu$fit, inc_warmup = FALSE)
ebfmi_per_chain <- sapply(sp, function(chain_mat) {
  ener <- chain_mat[,"energy__"]
  mean(diff(ener)^2) / var(ener)
})
print(ebfmi_per_chain)
```

```{r}
#| fig-width: 12
#| fig-height: 12
bayesplot::mcmc_pairs(
  as.array(fit_ranint_adm_clu), 
  pars = c(
    "b_Intercept", 
    "sd_adm_name__Intercept", 
    "sd_adm_name:cluster__Intercept", 
    "b_rural",
    "b_live_with_partner",
    "b_over_18",
    "bsp_moschool_level", 
    "b_clim"
  ),
  np = nuts_params(fit_ranint_adm_clu),
  np_style = pairs_style_np(
    div_color = "green",
    div_shape = 20,
    div_size = 3
  )
)
```

```{r}
plot(fit_ranint_adm_clu, variable = "prior_b", regex = TRUE, nvariables = 3)
```

```{r}
plot(fit_ranint_adm_clu, variable = "b_rural", regex = TRUE, nvariables = 3)
```

```{r}
bayestestR::effective_sample(fit_ranint_adm_clu)
```

### Model Output

#### Summary

##### Raw draws

```{r}
bayestestR::describe_posterior(
  fit_ranint_adm_clu,
  effects = "all",
  component = "all",
  test = c("p_direction", "p_significance"),
  centrality = "all",
  ci_method = "HDI"
)
```

##### Transformed draws

```{r}
trans_post_ranint_adm_clu <- (
  fit_ranint_adm_clu
  %>% as.data.frame()
  %>% select(
    -c(
      contains("adm_name"),
      contains("prior"),
      lprior, 
      lp__,
      Intercept
    )
  )
  %>% mutate(
    across(
      -c(contains("simo")),
      ~ exp(.)
    )
  )
)
```

```{r}
bayestestR::describe_posterior(
  trans_post_ranint_adm_clu,
  effects = "all",
  component = "all",
  test = c(),
  centrality = "all",
  ci_method = "HDI"
)
```

#### Climate Marginal Effects

##### Population level

```{r}
cont_marg_ranint_adm_clu <- generate_cont_marginal_df(fit_ranint_adm_clu, df_model, covs = c("clim"))
```

```{r}
#| fig-width: 4
#| fig-height: 3
(
  plot_continuous_marginal(fit_ranint_adm_clu, df_model, covs = c("clim"), grid_points = 20, ndraws = 4000)
  + labs(
    y = "NP violence (%)"
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ranint_adm_clu_food_test.png")
)
```

#### Covariate Posterior Distribution

##### Log odds

```{r}
df_cov_post_adm_clu <- (
  fit_ranint_adm_clu
  %>% as_tibble()
  %>% select(
    b_clim,
    b_rural,
    b_live_with_partner,
    b_over_18,
    bsp_moschool_level
  )
  %>% pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  )
  %>% group_by(variable)
  %>% summarise(
    median_val = median(value),
    lower1_val = bayestestR::hdi(value, ci = 0.5)$CI_low,
    upper1_val = bayestestR::hdi(value, ci = 0.5)$CI_high,
    lower2_val = bayestestR::hdi(value, ci = 0.95)$CI_low,
    upper2_val = bayestestR::hdi(value, ci = 0.95)$CI_high,
  )
  %>% ungroup()
  %>% mutate(
    type = "no_fp"
  )
)
```

##### OR

```{r}
df_cov_or_adm_clu <- (
  fit_ranint_adm_clu
  %>% as_tibble()
  %>% select(
    b_clim,
    b_rural,
    b_live_with_partner,
    b_over_18,
    bsp_moschool_level
  )
  %>% pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  )
  %>% mutate(
    value = exp(value)
  )
  %>% group_by(variable)
  %>% summarise(
    median_val = median(value),
    lower1_val = bayestestR::hdi(value, ci = 0.5)$CI_low,
    upper1_val = bayestestR::hdi(value, ci = 0.5)$CI_high,
    lower2_val = bayestestR::hdi(value, ci = 0.95)$CI_low,
    upper2_val = bayestestR::hdi(value, ci = 0.95)$CI_high,
  )
  %>% ungroup()
  %>% mutate(
    type = "no_fp"
  )
)
```


### Calibration Checks

```{r}
#| fig-width: 4
#| fig-height: 3
plot_calibration(fit_ranint_adm_clu, df_model, outcome_var = viol_var)
ggsave(
  path = fig_dir,
  filename = paste0("calibration_plot_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ranint_adm_clu_food_test.png")
)
```

### Residuals Spatial Correlation Checks

```{r}
#| fig-width: 3.2
#| fig-height: 3
plot_moran(df_model, sf_vacs, fit = fit_ranint_adm_clu, residuals = TRUE)
ggsave(
  path = fig_dir,
  filename = paste0("moran_plot_res_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ranint_adm_clu_food_test.png")
)
```

### Posterior Predictive Checks

```{r}
#| fig-width: 12
#| fig-height: 2.5
(
  plot_ppc_rootogram(fit_ranint_adm_clu, df_model, outcome_var = viol_var)
  + theme(
    legend.position = "inside",
    legend.position.inside = c(0.92, 0.8)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("ppc_by_adm_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ranint_adm_clu_food_test.png")
)
```

## With food prices

```{r}
pars_fp <- c(
  "b_Intercept",
  "b_rural",
  "b_live_with_partner",
  "b_over_18",
  "bsp_moschool_level",
  # "bsp_moprice_index_reg_anomaly_12m1",
  "sd_adm_name__Intercept",
  "sd_adm_name:cluster__Intercept",
  "b_clim"
)
```

```{r}
stanvars <- stanvar(6, "num_pred")

prior_list_fp <- c(
  prior(student_t(3, -3, 1), class = "Intercept"),
  prior(normal(0, 0.8 / sqrt(num_pred)), class = "b"),
  prior(exponential(1/0.2), class = "sd"),
  prior(dirichlet(1), class = "simo", coef = "moschool_level1")
  # prior(dirichlet(1), class = "simo", coef = "moprice_index_reg_anomaly_12m1")
)

form_fp <- bf(
  as.formula(
    paste(
      viol_var,
      "| weights(sample_weight)",
      "~",
      "(
        1 
        + rural
        + live_with_partner
        + over_18
        + mo(school_level)
        + clim
        + (1 | adm_name/cluster)
        + mo(value_price_index_reg_anomaly_12m)
      )"
    )
  )
)
fit_fp <- brm(
  formula = form_fp,
  family = bernoulli("logit"),
  prior = prior_list_fp,
  data = df_model,
  stanvars = stanvars,
  seed = seed,
  refresh = 1000,
  # iter = iter,
  # control = list(adapt_delta = adapt_delta),
  # sample_prior = "only"
  sample_prior = TRUE
)
```

### Model Checks

```{r}
sp <- rstan::get_sampler_params(fit_fp$fit, inc_warmup = FALSE)
ebfmi_per_chain <- sapply(sp, function(chain_mat) {
  ener <- chain_mat[,"energy__"]
  mean(diff(ener)^2) / var(ener)
})
print(ebfmi_per_chain)
```

```{r}
#| fig-width: 12
#| fig-height: 12
bayesplot::mcmc_pairs(
  as.array(fit_fp), 
  pars = c(
    "b_Intercept", 
    "sd_adm_name__Intercept", 
    "sd_adm_name:cluster__Intercept", 
    "b_rural",
    "b_live_with_partner",
    "b_over_18",
    "bsp_moschool_level", 
    "b_dev_value_price_index_reg_anomaly_12m",
    # "bsp_moprice_index_reg_anomaly_12m",
    "b_clim"
  ),
  np = nuts_params(fit_fp),
  np_style = pairs_style_np(
    div_color = "green",
    div_shape = 20,
    div_size = 3
  )
)
```

```{r}
plot(fit_fp, variable = "prior_b", regex = TRUE, nvariables = 3)
```

```{r}
plot(fit_fp, variable = "b_rural", regex = TRUE, nvariables = 3)
```

```{r}
bayestestR::effective_sample(fit_fp)
```

### Model Output

#### Summary

##### Raw draws

```{r}
bayestestR::describe_posterior(
  fit_fp,
  effects = "all",
  component = "all",
  test = c("p_direction", "p_significance"),
  centrality = "all",
  ci_method = "HDI"
)
```

##### Transformed draws

```{r}
trans_post_fp <- (
  fit_fp
  %>% as.data.frame()
  %>% select(
    -c(
      contains("adm_name"),
      contains("prior"),
      lprior, 
      lp__,
      Intercept
    )
  )
  %>% mutate(
    across(
      -c(contains("simo")),
      ~ exp(.)
    )
  )
)
```

```{r}
bayestestR::describe_posterior(
  trans_post_fp,
  effects = "all",
  component = "all",
  test = c(),
  centrality = "all",
  ci_method = "HDI"
)
```

#### Climate Marginal Effects

##### Population level

```{r}
#| fig-width: 5
#| fig-height: 3
(
  plot_continuous_marginal(fit_fp, df_model, covs = c("clim"))
  + scale_y_continuous(
    # limits = c(0, 0.43),
    labels = scales::percent_format(accuracy = 1)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_fp_food_test.png")
)
```

#### Covariate Posterior Distribution

##### Log odds

```{r}
df_cov_post_fp <- (
  fit_fp
  %>% as_tibble()
  %>% select(
    b_clim,
    b_rural,
    b_live_with_partner,
    b_over_18,
    bsp_moschool_level,
    bsp_movalue_price_index_reg_anomaly_12m
  )
  %>% pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  )
  %>% group_by(variable)
  %>% summarise(
    median_val = median(value),
    lower1_val = bayestestR::hdi(value, ci = 0.5)$CI_low,
    upper1_val = bayestestR::hdi(value, ci = 0.5)$CI_high,
    lower2_val = bayestestR::hdi(value, ci = 0.95)$CI_low,
    upper2_val = bayestestR::hdi(value, ci = 0.95)$CI_high,
  )
  %>% ungroup()
  %>% mutate(
    type = "fp"
  )
)
```

```{r}
df_cov_post_comb <- bind_rows(df_cov_post_adm_clu, df_cov_post_fp)
```

```{r}
#| fig-width: 5
#| fig-height: 3

dodge <- ggstance::position_dodgev(height = 0.8)

lims <- c(
  min(df_cov_post_comb$lower2_val, -0.05), 
  max(df_cov_post_comb$upper2_val)
)

p <- (
  df_cov_post_comb
  %>% ggplot(
    mapping = aes(
      y = fct_reorder(variable, desc(median_val)),
      color = type
    )
  )
  + geom_segment(
    mapping = aes(
      x = lower2_val,
      xend = upper2_val
    ),
    position = dodge,
    linewidth = 1
  )
  + geom_segment(
    mapping = aes(
      x = lower1_val,
      xend = upper1_val
    ),
    position = dodge,
    linewidth = 2
  )
  + geom_point(
    mapping = aes(
      x = median_val,
    ),
    size = 3,
    position = dodge,
  )
  + geom_vline(
    xintercept = 0,
    linetype = "dashed",
    color = "black"
  )
  + scale_x_continuous(
    limits = lims
  )
  + scale_y_discrete(
    labels = c(
      "b_clim" = "Heat exposure",
      "b_rural" = "Rural",
      "b_over_18" = "Over 18",
      "b_live_with_partner" = "Live with partner",
      "bsp_moschool_level" = "Schooling level",
      "bsp_movalue_price_index_reg_anomaly_12m" = "Price anomaly"
    )
  )
  + scale_color_manual(
    values = c(
      "no_fp" = col_palette[5],
      "fp" = col_palette[1]
    )
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    plot.background = element_blank(),
    plot.title = element_text(size = 14, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),,
    axis.title.y = element_blank(),
    axis.title.x = element_text(size = 14),
    axis.text = element_text(size = 13),
  )
  + labs(
    y = "Country",
    x = "Effect Size"
  )
  + guides(
    color = "none"
  )
)

p

ggsave(
  path = fig_dir,
  filename = paste0("cov_post_comb_", clim_label, "_", pop_label, "_", viol_label, "_moz_food_test.png")
)
```

##### OR 

```{r}
df_cov_or_fp <- (
  fit_fp
  %>% as_tibble()
  %>% select(
    b_clim,
    b_rural,
    b_live_with_partner,
    b_over_18,
    bsp_moschool_level,
    bsp_movalue_price_index_reg_anomaly_12m
  )
  %>% pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  )
  %>% mutate(
    value = exp(value)
  )
  %>% group_by(variable)
  %>% summarise(
    median_val = median(value),
    lower1_val = bayestestR::hdi(value, ci = 0.5)$CI_low,
    upper1_val = bayestestR::hdi(value, ci = 0.5)$CI_high,
    lower2_val = bayestestR::hdi(value, ci = 0.95)$CI_low,
    upper2_val = bayestestR::hdi(value, ci = 0.95)$CI_high,
  )
  %>% ungroup()
  %>% mutate(
    type = "fp"
  )
)
```

```{r}
df_cov_or_comb <- bind_rows(df_cov_or_adm_clu, df_cov_or_fp)
```


```{r}
#| fig-width: 5
#| fig-height: 3

dodge <- ggstance::position_dodgev(height = 0.8)

lims <- c(
  min(df_cov_or_comb$lower2_val, 0.95), 
  max(df_cov_or_comb$upper2_val)
)

p <- (
  df_cov_or_comb
  %>% ggplot(
    mapping = aes(
      y = fct_reorder(variable, desc(median_val)),
      color = type
    )
  )
  + geom_segment(
    mapping = aes(
      x = lower2_val,
      xend = upper2_val
    ),
    position = dodge,
    linewidth = 1
  )
  + geom_segment(
    mapping = aes(
      x = lower1_val,
      xend = upper1_val
    ),
    position = dodge,
    linewidth = 2
  )
  + geom_point(
    mapping = aes(
      x = median_val,
    ),
    size = 3,
    position = dodge,
  )
  + geom_vline(
    xintercept = 1,
    linetype = "dashed",
    color = "black"
  )
  + scale_x_continuous(
    limits = lims
  )
  + scale_y_discrete(
    labels = c(
      "b_clim" = "Heat exposure",
      "b_rural" = "Rural",
      "b_over_18" = "Over 18",
      "b_live_with_partner" = "Live with partner",
      "bsp_moschool_level" = "Schooling level",
      "bsp_movalue_price_index_reg_anomaly_12m" = "Price anomaly"
    )
  )
  + scale_color_manual(
    values = c(
      "no_fp" = col_palette[5],
      "fp" = col_palette[1]
    )
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    plot.background = element_blank(),
    plot.title = element_text(size = 14, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),,
    axis.title.y = element_blank(),
    axis.title.x = element_text(size = 14),
    axis.text = element_text(size = 13),
  )
  + labs(
    y = "Country",
    x = "Odds Ratio"
  )
  + guides(
    color = "none"
  )
)

p

ggsave(
  path = fig_dir,
  filename = paste0("cov_or_comb_", clim_label, "_", pop_label, "_", viol_label, "_moz_food_test.png")
)
```

### Calibration Checks

```{r}
#| fig-width: 4
#| fig-height: 3
plot_calibration(fit_fp, df_model, outcome_var = viol_var)
ggsave(
  path = fig_dir,
  filename = paste0("calibration_plot_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_fp_food_test.png")
)
```

### Residuals Spatial Correlation Checks

```{r}
#| fig-width: 3.2
#| fig-height: 3
plot_moran(df_model, sf_vacs, fit = fit_fp, residuals = TRUE)
ggsave(
  path = fig_dir,
  filename = paste0("moran_plot_res_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_fp_food_test.png")
)
```

### Posterior Predictive Checks

```{r}
#| fig-width: 12
#| fig-height: 2.5
(
  plot_ppc_rootogram(fit_fp, df_model, outcome_var = viol_var)
  + theme(
    legend.position = "inside",
    legend.position.inside = c(0.92, 0.8)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("ppc_by_adm_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_fp_food_test.png")
)
```

