---
title: "UTCI Analysis"
format: html
editor: visual
---

# Environment Setup

```{r setup}
# Packages
library(tidyverse)
library(tidyterra)
library(sf)
library(paletteer)
library(patchwork)

# Path Structure
root <- rprojroot::has_file("agile2025.Rproj")$make_fix_file()
raw_data_dir <- root("data/raw")
proc_data_dir <- root("data/processed/UTCI")
dir.create(proc_data_dir, showWarnings = FALSE)
fig_dir <- root("figures")

# Country Mapping
country_map <- sort(c(
  "KEN" = "Kenya",
  "LSO" = "Lesotho",
  "MOZ" = "Mozambique",
  "NAM" = "Namibia",
  "UGA" = "Uganda",
  "ZMB" = "Zambia",
  "ZWE" = "Zimbabwe"
))
```

# Functions

```{r}
build_utci_daily <- function(data_dir, country_map, adm_shapes) {
  print(paste0("Starting UTCI data processing"))
  rast_root_dir <- file.path(data_dir, "UTCI/daily/")
  country_dir <- list.dirs(rast_root_dir)
  thresholds <- round(quantile(seq_along(rast_files), probs = c(0, 0.25, 0.5, 0.75, 1)))
  dfs_utci <- list()
  for (c in seq_along(country_iso)) {
    print(paste0("Processing UTCI data for: ", country_iso[[c]]))
    shape <- adm_shapes %>% filter(country == country_map[[country_iso[[c]]]])
    rast_files <- list.files(country_dir[[c]], pattern = "^ECMWF_utci_daily_stats", full.names = TRUE, recursive = TRUE)
    thresholds <- round(quantile(seq_along(rast_files), probs = c(0, 0.25, 0.5, 0.75, 1)))
    dfs_utci_country <- list()
    for (i in seq_along(rast_files)) {
      if (i %in% thresholds) {
        thresh_idx <- which(thresholds == i)
        print(paste0(names(thresh_idx), " complete"))
      }
      rast_utci <- terra::rast(rast_files[[i]])["utci_daily_max"]
      date <- str_split_i(terra::time(rast_utci)[[1]], " ", 1)
      dfs_utci_country[[i]] <- (
        rast_utci
        %>% exactextractr::exact_extract(
          shape,
          fun = "weighted_mean",
          weights = "area",
          max_cells_in_memory = 176067000,
          append_cols = c("country", "adm_code", "adm_name"),
          progress = FALSE
        )
        %>% mutate(
          date = as.Date(date, format = "%Y-%m-%d"),
          utci_max = weighted_mean - 273.15
        )
      )
    }
    dfs_utci[[c]] <- (
      bind_rows(dfs_utci_country)
      %>% select(-weighted_mean)
      %>% mutate(
        country = country_map[[country_iso[[c]]]]
      )
      %>% relocate(country)
    )
  }
  df_utci <- bind_rows(dfs_utci)
}

build_weighted_utci_daily_country <- function(data_dir, iso3, rast_pop, adm_pop, adm_shapes) {
  print(paste0("Starting processing UTCI data for: ", iso3))
  dfs_utci_daily <- list()
  rast_root_dir <- file.path(data_dir, paste0("UTCI/daily/", iso3))
  rast_files <- list.files(rast_root_dir, pattern = "^ECMWF_utci_daily_stats", full.names = TRUE, recursive = TRUE)
  thresholds <- round(quantile(seq_along(rast_files), probs = c(0, 0.25, 0.5, 0.75, 1)))
  for (i in seq_along(rast_files)) {
    if (i %in% thresholds) {
      thresh_idx <- which(thresholds == i)
      print(paste0(names(thresh_idx), " complete"))
    }
    rast_utci <- terra::rast(rast_files[[i]])["utci_daily_max"]
    date <- terra::time(rast_utci)[[1]]
    rast_utci_fine <- terra::resample(rast_utci, rast_pop)
    rast_utci_weighted <- rast_utci_fine * rast_pop
    dfs_utci_daily[[i]] <- (
      rast_utci_weighted
      %>% exactextractr::exact_extract(
        adm_shapes,
        fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
        max_cells_in_memory = 176067000,
        append_cols = c("country", "adm_code"),
        progress = FALSE
      )
      %>% merge(
        adm_pop,
        by = "adm_code",
        all.x = TRUE
      )
      %>% mutate(
        date = as.Date(date, format = "%Y-%m-%d"),
        utci_max = (result / pop) - 273.15
      )
    )
  }
  df_utci <- (
    bind_rows(dfs_utci_daily)
    %>% select(-result)
    %>% mutate(
      country = country_map[[iso3]]
    )
    %>% relocate(country)
  )
  df_utci %>% write_csv(file.path(proc_data_dir, paste0("utci_max_daily_", iso3, ".csv")))
  df_utci
}
```

# Data Import

## Geographies

### Mozambique

```{r}
sf_moz <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Mozambique 2022 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Zimbabwe

```{r}
sf_zwe <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Zimbabwe 2015 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Lesotho

```{r}
sf_lso <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Lesotho 2023 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Namibia

```{r}
sf_nam <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Namibia 2013 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Zambia

```{r}
sf_zmb <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Zambia 2018 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Uganda

```{r}
# sf_uga <- (
#   st_read(
#     file.path(raw_data_dir, "geographies", "uga_admbnda_ubos_20200824.gdb", fsep = "/"),
#     layer = "uga_admbnda_adm2_ubos_20200824"
#   )
#   %>% rename_with(tolower)
#   %>% mutate(
#     id = row_number()
#   )
#   %>% select(
#     id,
#     adm_name = admin1name_en,
#     adm_name = admin2name_en,
#     adm2_pcode = admin2pcode,
#     # adm3_name = admin3name_en,
#     # adm4_name = admin4name_en,
#     # adm4_pcode = admin4pcode,
#     geometry = shape
#   )
# )
sf_uga <- (
  st_read(
    file.path(
      raw_data_dir,
      "geographies",
      "uganda_herbert",
      "uganda_districts.shp"
    ),
  )
  %>% rename_with(tolower)
  %>% mutate(
    country = "Uganda",
    district = case_match(
      district,
      "SSEMBABULE" ~ "Sembabule",
      .default = str_to_title(district)
    )
  )
  %>% select(
    country,
    adm_code = objectid,
    adm_name = district
  )
  %>% st_transform(4326)
)
```

### Kenya

```{r}
sf_ken <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Kenya 2022 DHS - sdr_subnational_boundaries_2025-11-13", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Create combined shapes

```{r}
sf_all <- (
  bind_rows(
    sf_moz,
    sf_zwe,
    sf_lso,
    sf_nam,
    sf_zmb,
    sf_uga,
    sf_ken
  )
)
```

## Population

### Mozambique

```{r}
rast_pop_moz_2022 <- (
  terra::rast(
    file.path(raw_data_dir, "population", "moz_pop_2022_CN_1km_R2025A_UA_v1.tif")
  )
  %>% rename(
    pop = global_pop_2022_CN_1km_R2025A_v1
  )
)
```

```{r}
adm_pop_moz_2022 <- (
  rast_pop_moz_2022
  %>% exactextractr::exact_extract(
    sf_moz,
    fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
    max_cells_in_memory = 176067000,
    append_cols = c("adm_code", "adm_name")
  )
  %>% rename(
    pop = result
  )
)
```

### Zimbabwe

```{r}
rast_pop_zwe_2022 <- (
  terra::rast(  
    file.path(raw_data_dir, "population", "zwe_pop_2022_CN_1km_R2025A_UA_v1.tif")
  )
  %>% rename(
    pop = global_pop_2022_CN_1km_R2025A_v1
  )
)
```

```{r}
adm_pop_zwe_2022 <- (
  rast_pop_zwe_2022
  %>% exactextractr::exact_extract(
    sf_zwe,
    fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
    max_cells_in_memory = 176067000,
    append_cols = c("adm_code", "adm_name")
  )
  %>% rename(
    pop = result
  )
)
```

### Lesotho

```{r}
rast_pop_lso_2022 <- (
  terra::rast(
    file.path(raw_data_dir, "population", "lso_pop_2022_CN_1km_R2025A_UA_v1.tif")
  )
  %>% rename(
    pop = global_pop_2022_CN_1km_R2025A_v1
  )
)
```

```{r}
adm_pop_lso_2022 <- (
  rast_pop_lso_2022
  %>% exactextractr::exact_extract(
    sf_lso,
    fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
    max_cells_in_memory = 176067000,
    append_cols = c("adm_code", "adm_name")
  )
  %>% rename(
    pop = result
  )
)
```

### Namibia

```{r}
rast_pop_nam_2022 <- (
  terra::rast(
    file.path(raw_data_dir, "population", "nam_pop_2022_CN_1km_R2025A_UA_v1.tif")
  )
  %>% rename(
    pop = global_pop_2022_CN_1km_R2025A_v1
  )
)
```

```{r}
adm_pop_nam_2022 <- (
  rast_pop_nam_2022
  %>% exactextractr::exact_extract(
    sf_nam,
    fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
    max_cells_in_memory = 176067000,
    append_cols = c("adm_code", "adm_name")
  )
  %>% rename(
    pop = result
  )
)
```

### Zambia

```{r}
rast_pop_zmb_2022 <- (
  terra::rast(
    file.path(raw_data_dir, "population", "zmb_pop_2022_CN_1km_R2025A_UA_v1.tif")
  )
  %>% rename(
    pop = global_pop_2022_CN_1km_R2025A_v1
  )
)
```

```{r}
adm_pop_zmb_2022 <- (
  rast_pop_zmb_2022
  %>% exactextractr::exact_extract(
    sf_zmb,
    fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
    max_cells_in_memory = 176067000,
    append_cols = c("adm_code", "adm_name")
  )
  %>% rename(
    pop = result
  )
)
```

### Uganda

```{r}
rast_pop_uga_2022 <- (
  terra::rast(
    file.path(raw_data_dir, "population", "uga_pop_2022_CN_1km_R2025A_UA_v1.tif")
  )
  %>% rename(
    pop = global_pop_2022_CN_1km_R2025A_v1
  )
)
```

```{r}
adm_pop_uga_2022 <- (
  rast_pop_uga_2022
  %>% exactextractr::exact_extract(
    sf_uga,
    fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
    max_cells_in_memory = 176067000,
    append_cols = c("adm_code", "adm_name")
  )
  %>% rename(
    pop = result
  )
)
```

### Kenya

```{r}
rast_pop_ken_2022 <- (
  terra::rast(
    file.path(raw_data_dir, "population", "ken_pop_2022_CN_1km_R2025A_UA_v1.tif")
  )
  %>% rename(
    pop = global_pop_2022_CN_1km_R2025A_v1
  )
)
```

```{r}
adm_pop_ken_2022 <- (
  rast_pop_ken_2022
  %>% exactextractr::exact_extract(
    sf_ken,
    fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
    max_cells_in_memory = 176067000,
    append_cols = c("adm_code", "adm_name")
  )
  %>% rename(
    pop = result
  )
)
```

# Generate UTCI dataset

```{r}
df_utci <- build_utci_daily(raw_data_dir, country_map, sf_all)
```

# Generate Population Weighted UTCI datasets

Data obtained using the Climate Data Store API

## Method Outline

```{r}
# For plotting
xlims <- c(32.1, 33.05)
ylims <- c(-26.9, -25.5)
plot_extent <- terra::ext(xlims[1] - 0.5, xlims[2] + 0.5, ylims[1] - 0.5, ylims[2] + 0.5)
```

### UTCI in country

```{r}
rast_root_dir <- file.path(raw_data_dir, "UTCI/daily/MOZ")
rast_files <- list.files(rast_root_dir, pattern = "^ECMWF_utci_daily_stats", full.names = TRUE, recursive = TRUE)
rast_utci_moz <- terra::rast(rast_files[[1]])["utci_daily_max"]
date <- terra::time(rast_utci_moz)[[1]]
```

```{r}
#| fig-width: 5
#| fig-height: 4
(
  ggplot()
  + geom_spatraster(
    data = (rast_utci_moz - 273.15),
    mapping = aes(fill = utci_daily_max)
  )
  + scale_fill_paletteer_c(
    "scico::vik",
    na.value = "transparent",
  )
  + geom_sf(
    data = sf_moz,
    fill = NA,
    color = "black",
    # linewidth = 1.1
  )
  + theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.key.height = unit(50, "pt"),
    legend.title = element_text(size = 18, angle = 90, hjust = 0.5),
    legend.text = element_text(size = 16),
    legend.background = element_blank(),
    legend.title.position = "right",
  )
  + labs(
    fill = "UTCI"
  )
)
ggsave(
  path = fig_dir,
  filename = "utci_grid_moz.png",
  width = 5,
  height = 4
)
```

### Downsample one time step

```{r}
rast_utci_fine_moz <- terra::resample(rast_utci_moz, rast_pop_moz_2022)
```

```{r}
#| fig-width: 5
#| fig-height: 4
(
  ggplot()
  + geom_spatraster(
    data = rast_utci_fine_moz %>% terra::crop(plot_extent),
    mapping = aes(fill = utci_daily_max)
  )
  + scale_fill_paletteer_c(
    "scico::vik",
    na.value = "transparent",
  )
  + geom_sf(
    data = sf_moz,
    fill = NA,
    color = "black",
  )
  + theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.key.height = unit(50, "pt"),
    legend.title = element_text(size = 18, angle = 90, hjust = 0.5),
    legend.text = element_text(size = 16),
    legend.background = element_blank(),
    legend.title.position = "right",
  )
  + coord_sf(
    xlim = xlims,
    ylim = ylims
  )
  + labs(
    fill = "UTCI"
  )
)
ggsave(
  path = fig_dir,
  filename = "utci_fine_moz_zoom.png",
  width = 5,
  height = 4
)
```

### Population scaled UTCI

```{r}
rast_utci_weighted_moz <- rast_utci_fine_moz * rast_pop_moz_2022
```

```{r}
#| fig-width: 8
#| fig-height: 4

p1 <- (
  ggplot()
  + geom_spatraster(
    data = rast_pop_moz_2022 %>% terra::crop(plot_extent),
    mapping = aes(fill = pop)
  )
  + scale_fill_paletteer_c(
    "scico::devon",
    na.value = "transparent",
    direction = -1
  )
  + geom_sf(
    data = sf_moz,
    fill = NA,
    color = "black",
  )
  + theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.key.height = unit(50, "pt"),
    legend.title = element_text(size = 18, angle = 90, hjust = 0.5),
    legend.text = element_text(size = 16),
    legend.background = element_blank(),
    legend.title.position = "right",
  )
  + coord_sf(
    xlim = xlims,
    ylim = ylims
  )
  + labs(
    fill = "Population"
  )
)

p2 <- (
  ggplot()
  + geom_spatraster(
    data = rast_utci_weighted_moz %>% terra::crop(plot_extent),
    mapping = aes(fill = utci_daily_max)
  )
  + scale_fill_paletteer_c(
    "scico::vik",
    na.value = "transparent",
  )
  + geom_sf(
    data = sf_moz,
    fill = NA,
    color = "black",
  )
  + theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.key.height = unit(50, "pt"),
    legend.title = element_text(size = 18, angle = 90, hjust = 0.5),
    legend.text = element_text(size = 16),
    legend.background = element_blank(),
    legend.title.position = "right",
  )
  + coord_sf(
    xlim = xlims,
    ylim = ylims
  )
  + labs(
    fill = "UTCI * Population"
  )
)

p1 + p2

ggsave(
  path = fig_dir,
  filename = "utci_pop.png",
  width = 6.5,
  height = 4
)
```

### Weighted UTCI

```{r}
df_utci_daily_moz <- (
  rast_utci_weighted_moz
  %>% exactextractr::exact_extract(
    sf_moz,
    fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
    max_cells_in_memory = 176067000,
    append_cols = c("adm_code"),
    progress = FALSE
  )
  %>% merge(
    adm_pop_moz_2022,
    by = "adm_code",
    all.x = TRUE
  )
  %>% mutate(
    date = as.Date(date, format = "%Y-%m-%d"),
    utci_max = (result / pop) - 273.15
  )
  %>% select(-result)
)
```

```{r}
sf_utci_daily_moz <- (
  sf_moz
  %>% select(adm_code)
  %>% merge(
    df_utci_daily_moz,
    by = "adm_code",
    all.x = TRUE
  )
)
```

```{r}
#| fig-width: 4
#| fig-height: 5
(
  ggplot()
  + geom_sf(
    data = sf_utci_daily_moz %>% filter(adm_name != "Maputo City"),
    mapping = aes(fill = utci_max),
    color = "black",
  )
  + theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5, margin = margin(b = 0)),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.key.height = unit(50, "pt"),
    legend.title = element_text(size = 16, angle = 90, hjust = 0.5),
    legend.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.title.position = "right",
  )
  + scale_fill_paletteer_c(
    "scico::vik",
    na.value = "transparent",
  )
  + labs(
    title = "Weighted UTCI Index",
    fill = "Max daily UTCI"
  )
)
ggsave(
  path = fig_dir,
  filename = "utci_weighted_moz.png",
  width = 4,
  height = 5
)
```

## Mozambique

```{r}
df_utci_moz <- build_weighted_utci_daily_country(raw_data_dir, "MOZ", rast_pop_moz_2022, adm_pop_moz_2022, sf_moz)
```

## Zimbabwe

```{r}
df_weighted_utci_zwe <- build_weighted_utci_daily_country(raw_data_dir, "ZWE", rast_pop_zwe_2022, adm_pop_zwe_2022, sf_zwe)
```

## Lesotho

```{r}
df_weighted_utci_lso <- build_weighted_utci_daily_country(raw_data_dir, "LSO", rast_pop_lso_2022, adm_pop_lso_2022, sf_lso)
```

### Namibia

```{r}
df_weighted_utci_nam <- build_weighted_utci_daily_country(raw_data_dir, "NAM", rast_pop_nam_2022, adm_pop_nam_2022, sf_nam)
```

## Zambia

```{r}
df_weighted_utci_zmb <- build_weighted_utci_daily_country(raw_data_dir, "ZMB", rast_pop_zmb_2022, adm_pop_zmb_2022, sf_zmb)
```

### Uganda

```{r}
df_weighted_utci_uga <- build_weighted_utci_daily_country(raw_data_dir, "UGA", rast_pop_uga_2022, adm_pop_uga_2022, sf_uga)
```

### Kenya

```{r}
df_weighted_utci_ken <- build_weighted_utci_daily_country(raw_data_dir, "KEN", rast_pop_ken_2022, adm_pop_ken_2022, sf_ken)
```

# Save

```{r}
df_utci %>% write_csv(file.path(proc_data_dir, "utci_max_daily_all_countries.csv"))
```

```{r}
df_weighted_utci <- bind_rows(
  df_weighted_utci_moz,
  df_weighted_utci_zwe,
  df_weighted_utci_lso,
  df_weighted_utci_nam,
  df_weighted_utci_zmb,
  df_weighted_utci_uga,
  df_weighted_utci_ken
)
 
df_weighted_utci %>% write_csv(file.path(proc_data_dir, "weighted_utci_max_daily_all_countries.csv"))
```
