---
title: "Model main UTCI effect on non-partner sexual violence - children"
format: html
editor: visual
---

# Environment Setup

```{r setup}
# Packages
library(tidyverse)
library(sf)
library(spdep)
library(brms)
options(brms.backend = "cmdstanr", mc.cores = 4)
library(cmdstanr)
library(posterior)
library(ggdist)
library(bayesplot)
library(tidybayes)
# theme_set(bayesplot::theme_default(base_family = "sans", base_size=14))
library(patchwork)
library(tidytext)
library(reliabilitydiag)

# Folder Structure
root <- rprojroot::has_file("agile2025.Rproj")$make_fix_file()
temp_dir <- root("temp")
dir.create(temp_dir, showWarnings = FALSE)
raw_data_dir <- root("data/raw")
model_data_dir <- root("data/processed/model_data")
fig_dir <- root("figures")
dir.create(fig_dir, showWarnings = FALSE)
mod_dir <- root("models")
src_dir <- root("src")
```

# Notebook settings

```{r}
# Settings
nb_name <-  "Model main UTCI effect on non-partner sexual violence - children"
seed <- sum(sapply(str_split(nb_name, pattern = "", simplify = TRUE), FUN = utf8ToInt))
set.seed(seed)

# Viz settings
col_palette <- c("#dc7726", "#eebb93", "#ecc046", "#f6e0a3", "#0292b7", "#81c9db", "#35A29F", "#97DECE")
label_map <- c(
  tolerate_women_violence = "Tolerate violence",
  live_with_partner = "Live with partner",
  live_with_father = "Live with father",
  live_with_parent = "Live with parent",
  in_school = "In school",
  school_level = "Schooling level",
  work = "Has worked",
  rural = "Rural area",
  age = "Age",
  rel_age_ord = "Excess Age (years)",
  log_pop_density = "Population density",
  clim = "Num. Excess Hot Days (past year)",
  clim_stress = "Climate Stress"
)

# Var settings
clim_var <- "dev_value_pop_weighted_utci_max_above_32_12m"
clim_label <- "utci_above_32"
viol_var <- "sv_np"
viol_label <- "np_viol"
pop_label <- "child"
covars <- c(
  "sample_weight",
  "country",
  "adm_name",
  "rural",
  "age",
  "in_school",
  "live_with_parent",
  "tolerate_women_violence",
  viol_var
)
vars <- c(covars, clim_var)

bin_covs <- c(
  "rural",
  "in_school",
  "live_with_parent", 
  "tolerate_women_violence"
)
mono_covs <- c(
  "rel_age_ord"
)
cont_covs <- c(
  "clim"
)

# Computation settings
adapt_delta = 0.95
```

# Functions

```{r}
source(file.path(src_dir, "viz.R"))
source(file.path(src_dir, "model_check_utils.R"))
```

# Import Data

## Geometries

### Mozambique

```{r}
sf_moz <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Mozambique 2022 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% st_make_valid()
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Zimbabwe

```{r}
sf_zwe <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Zimbabwe 2015 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% st_make_valid()
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Lesotho

```{r}
sf_lso <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Lesotho 2023 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% st_make_valid()
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Namibia

```{r}
sf_nam <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Namibia 2013 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% st_make_valid()
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Zambia

```{r}
sf_zmb <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Zambia 2018 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% st_make_valid()
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Create combined shapes

```{r}
sf_all <- (
  bind_rows(
    sf_moz,
    sf_zwe,
    sf_lso,
    sf_nam,
    sf_zmb,
  )
  %>% st_transform(3857)
  %>% mutate(
    centroid = st_centroid(geometry),
    lon = st_coordinates(centroid)[,1],
    lat = st_coordinates(centroid)[,2]
  )
)
```

## Climate & Violence

```{r}
df_clim_viol <- (
  read_csv(
    file.path(
      model_data_dir,
      "climate_food_prices_violence.csv"
    ),
    show_col_type = FALSE
  )
  %>% filter(
    !is.na(sample_weight),
    country != "Kenya"
  )
  %>% group_by(country)
  %>% mutate(
    country_sample_weight = sum(sample_weight, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    sample_weight = (
      1E-3 
      * sample_weight 
      * (
        sum(sample_weight, na.rm = TRUE) 
        / (length(unique(country)) * country_sample_weight)
      )
    ),
    live_with_parent = as.numeric((live_with_father == 1) | (live_with_mother == 1)),
  )
  %>% select(-c(live_with_father, live_with_mother))
  %>% merge(
    sf_all %>% as.data.frame() %>% select(country, adm_name, lon, lat),
    by = c("country", "adm_name"),
    all.x = TRUE
  )
)
```

# Prepare Data

## Restrict locations

```{r}
sf_vacs <- (
  sf_all
  %>% filter(
    country %in% (df_clim_viol$country %>% unique()),
    adm_name %in% (df_clim_viol$adm_name %>% unique())
  )
)
```

## Extract model subsets

```{r}
df_model <- (
  df_clim_viol
  %>% filter(
    female == 1,
    age < 19
  )
  %>% select(all_of(vars))
  %>% na.omit()
  %>% mutate(
    adm_name = factor(adm_name),
    rel_age = as.integer(age - min(age)),
    rel_age_ord = factor(
      rel_age,
      levels = sort(unique(rel_age)),
      ordered = TRUE
    ),
    clim := scale(!!sym(clim_var))
  )
)
```

## Generate adjacency matrix

```{r}
nb_list <- poly2nb(sf_vacs)
W <- nb2mat(nb_list, style = "B", zero.policy = TRUE)
rownames(W) <- sf_vacs$adm_name
colnames(W) <- sf_vacs$adm_name
levs <- levels(df_model$adm_name)
W_ord <- W[levs, levs]
```

```{r}
edges <- (
  geostan::edges(W, unique_pairs_only = TRUE, shape = sf_vacs$geometry)
)
graph <- st_geometry(edges)
```

```{r}
plot(sf_vacs$geometry, lwd = .15)
plot(graph, add = TRUE, type = "p", col = "red")
plot(graph, add = TRUE, type = "l", col = "black")
```

# Modeling

## No spatial correlation

```{r}
stanvars <- stanvar(6, "num_pred")

prior_list_ns <- c(
  prior(student_t(3, -1, 0.4), class = "Intercept"),
  prior(normal(0, 0.8 / sqrt(num_pred)), class = "b"),
  prior(exponential(1/0.3), class = "sd"),
  prior(dirichlet(1), class = "simo", coef = "morel_age_ord1")
)

form_ns <- bf(
  as.formula(
    paste(
      viol_var,
      "| weights(sample_weight)",
      "~",
      "(
        1 
        + live_with_parent
        + tolerate_women_violence
        + rural
        + in_school
        + mo(rel_age_ord)
        + clim
        + (1 + clim | country)
      )"
    )
  )
)
fit_ns <- brm(
  formula = form_ns,
  family = bernoulli("logit"),
  prior = prior_list_ns,
  data = df_model,
  data2 = list(W = W_ord),
  stanvars = stanvars,
  seed = seed,
  refresh = 1000,
  control = list(adapt_delta = adapt_delta),
  # sample_prior = "only"
  sample_prior = TRUE
)
```

### Model Checks

```{r}
sp <- rstan::get_sampler_params(fit_ns$fit, inc_warmup = FALSE)
ebfmi_per_chain <- sapply(sp, function(chain_mat) {
  ener <- chain_mat[,"energy__"]
  mean(diff(ener)^2) / var(ener)
})
print(ebfmi_per_chain)
```

```{r}
#| fig-width: 12
#| fig-height: 12
bayesplot::mcmc_pairs(
  as.array(fit_ns), 
  pars = c(
    "b_Intercept", 
    "sd_country__Intercept", 
    "b_clim",
    "bsp_morel_age_ord" 
  ),
  np = nuts_params(fit_ns),
  np_style = pairs_style_np(
    div_color = "green",
    div_shape = 20,
    div_size = 3
  )
)
```

```{r}
plot(fit_ns, variable = "prior_b", regex = TRUE, nvariables = 3)
```

```{r}
plot(fit_ns, variable = "^b_rural", regex = TRUE, nvariables = 3)
```

### Model Output

#### Summary

```{r}
fit_ns
```

#### Baseline

```{r}
#| fig-width: 5
#| fig-height: 3
plot_baseline_levels(fit_ns, lims = c(0, 0.13))
```

#### Marginal Effects

##### Categorical variables

```{r}
#| fig-width: 5
#| fig-height: 3
plot_bin_marginal(fit_ns, df_model, covs = bin_covs)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_bin_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ns.png")
)
```

##### Ordered factors

```{r}
#| fig-width: 4.5
#| fig-height: 3
plot_mono_marginal(fit_ns, df_model, covs = mono_covs)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_mono_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ns.png")
)
```

##### Climate

```{r}
#| fig-width: 5
#| fig-height: 3
plot_continuous_marginal(fit_ns, df_model, covs = cont_covs)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ns.png")
)
```

#### Marginal Effects by Country

##### Ordered factors

```{r}
#| fig-width: 5.5
#| fig-height: 2.5
(
  plot_mono_marginal_by_group(fit_ns, df_model, covs = mono_covs, dodge = 0.6)
  # + theme(
  #   legend.position = "top"
  # )
)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_mono_main_effect_by_country_", clim_label, "_", pop_label, "_", viol_label, "_ns.png")
)
```

##### Climate

```{r}
#| fig-width: 12
#| fig-height: 3
plot_single_continuous_marginal_by_group(fit_ns, df_model, cov = "clim")
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_by_country_", clim_label, "_", pop_label, "_", viol_label, "_ns.png")
)
```

### Calibration Checks

```{r}
#| fig-width: 4
#| fig-height: 4
plot_calibration(fit_ns, df_model, outcome_var = viol_var)
ggsave(
  path = fig_dir,
  filename = paste0("calibration_plot_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ns.png")
)
```

### Spatial Correlation Checks

##### Raw values

```{r}
#| fig-width: 5
#| fig-height: 2.5
plot_moran(df_model, sf_vacs, outcome_var = viol_var)
ggsave(
  path = fig_dir,
  filename = paste0("moran_plot_obs_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ns.png")
)
```

##### Residuals

```{r}
#| fig-width: 5
#| fig-height: 2.5
plot_moran(df_model, sf_vacs, fit = fit_ns, residuals = TRUE)
ggsave(
  path = fig_dir,
  filename = paste0("moran_plot_res_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ns.png")
)
```

### Posterior Predictive Checks

```{r}
#| fig-width: 11
#| fig-height: 2.5
(
  plot_ppc_rootogram(fit_ns, df_model, outcome_var = viol_var)
  + theme(
    legend.position = "inside",
    legend.position.inside = c(0.92, 0.8)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("ppc_by_adm_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ns.png")
)
```

### Inspect Climate Variable Posterior

```{r}
df_clim_pop_ns <- (
  as_draws_df(fit_ns)
  %>% select(clim_pop = b_clim)
)
df_clim_ranef_ns <- (
  as_tibble(ranef(fit_ns, summary = FALSE)$country[, , 2])
  %>% bind_cols(df_clim_pop_ns)
  %>% mutate(
    across(
      .cols = -clim_pop, 
      .fns = ~ (.) + clim_pop,
      .names = "{.col}"
    )
  )
  %>% select(-clim_pop)
)

df_clim_ranef_plot_ns <- (
  df_clim_ranef_ns
  %>% pivot_longer(
    cols = everything(),
    names_to = "country",
    values_to = "value"
  )
)

df_clim_ranef_plot_ns_stats <- (
  df_clim_ranef_plot_ns
  %>% group_by(country)
  %>% summarise(
    median_val = median(value),
    mean_val = mean(value),
    lower1_val = bayestestR::hdi(value, ci = 0.5)$CI_low,
    upper1_val = bayestestR::hdi(value, ci = 0.5)$CI_high,
    lower2_val = bayestestR::hdi(value, ci = 0.95)$CI_low,
    upper2_val = bayestestR::hdi(value, ci = 0.95)$CI_high,
    .groups = "drop"
  )
)

```

```{r}
#| fig-width: 4.1
#| fig-height: 3
(
  ggplot(
    data = df_clim_pop_ns,
    mapping = aes(
      x = clim_pop
    )
  )
  + stat_dotsinterval(
    quantiles = 100,
    slab_color = col_palette[3],
    slab_fill = col_palette[4],
    .width = c(0.5, 0.95),
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    plot.background = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
  )
  + labs(
    x = "Climate Posterior Value",
    y = "Density"
  )
)

ggsave(
  path = fig_dir,
  filename = paste0("clim_post_covariate_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ns.png")
)
```

```{r}
#| fig-width: 4.1
#| fig-height: 3
(
  ggplot(
    data = df_clim_ranef_plot_ns,
    mapping = aes(
      x = value,
      y = fct_reorder(country, value, .fun = ~ -median(., na.rm = T))
    )
  )
  + ggridges::geom_density_ridges(
    aes(
      height = after_stat(scaled),
    ),
    stat = "density",
    orientation = "x",
    scale = 0.6,
    # rel_min_height = 0.01,
    alpha = 0.7,
    fill = "#FDD0A2FF",
    color = NA
  )
  + geom_segment(
    data = df_clim_ranef_plot_ns_stats,
    aes(y = country, yend = country, x = lower2_val, xend = upper2_val),
    linewidth = 1,
    alpha = 0.4,
    color = "#A63603FF",
    inherit.aes = FALSE
  )
  + geom_segment(
    data = df_clim_ranef_plot_ns_stats,
    aes(y = country, yend = country, x = lower1_val, xend = upper1_val),
    linewidth = 1.5,
    alpha = 0.7,
    color = "#A63603FF",
    inherit.aes = FALSE
  )
  + geom_point(
    data = df_clim_ranef_plot_ns_stats,
    aes(y = country, x = median_val, color = country),
    size = 3,
    color = "#A63603FF",
    inherit.aes = FALSE
  )
  + geom_vline(xintercept = 0, linetype = "dashed", color = "gray50")
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 0.5),
    plot.background = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
  )
  + scale_x_continuous(
    limits = c(-1.2, 1.8)
  )
  + scale_y_discrete(
      expand = expansion(mult = c(0.1, 0.25))
    )
  + labs(
    x = "Climate Posterior Term",
    y = "Density"
  )
)

ggsave(
  path = fig_dir,
  filename = paste0("clim_post_covariate_main_effect_by_country_", clim_label, "_", pop_label, "_", viol_label, "_ns.png")
)
```

## Full Model

```{r}
stanvars <- stanvar(6, "num_pred")

prior_list <- c(
  prior(student_t(3, -1, 0.4), class = "Intercept"),
  prior(normal(0, (0.8 / sqrt(num_pred))), class = "b"),
  prior(exponential(1/0.3), class = "sd"),
  prior(dirichlet(1), class = "simo", coef = "morel_age_ord1"),
  prior(normal(0, 0.5), class = "car"),
  prior(exponential(1/0.3), class = "sdcar")
)

form <- bf(
  as.formula(
    paste(
      viol_var,
      "| weights(sample_weight)",
      "~",
      "(
        1 
        + live_with_parent
        + tolerate_women_violence
        + rural
        + in_school
        + mo(rel_age_ord)
        + clim
        + (1 + clim | country)
        + car(W, gr = adm_name, type = 'escar')
      )"
    )
  )
)
fit <- brm(
  formula = form,
  family = bernoulli("logit"),
  prior = prior_list,
  data = df_model,
  data2 = list(W = W_ord),
  stanvars = stanvars,
  seed = seed,
  refresh = 1000,
  control = list(adapt_delta = adapt_delta),
  # sample_prior = "only"
  sample_prior = TRUE
)
```

### Model Checks

```{r}
sp <- rstan::get_sampler_params(fit$fit, inc_warmup = FALSE)
ebfmi_per_chain <- sapply(sp, function(chain_mat) {
  ener <- chain_mat[,"energy__"]
  mean(diff(ener)^2) / var(ener)
})
print(ebfmi_per_chain)
```

```{r}
mcmc_trace(as_draws_df(fit), pars = c("car", "sdcar", "rcar[1]"))
```

```{r}
#| fig-width: 12
#| fig-height: 12
bayesplot::mcmc_pairs(
  as.array(fit), 
  pars = c(
    "b_Intercept", 
    "sd_country__Intercept", 
    "b_clim",
    "bsp_morel_age_ord", 
    "car", 
    "sdcar", 
    "rcar[1]"
  ),
  np = nuts_params(fit),
  np_style = pairs_style_np(
    div_color = "green",
    div_shape = 20,
    div_size = 3
  )
)
```

```{r}
mcmc_parcoord(
  as.array(fit), 
  pars = c(
    "b_Intercept", 
    "sd_country__Intercept", 
    "b_clim",
    "bsp_morel_age_ord", 
    "car", 
    "sdcar", 
    "rcar[1]"
  ),
  np = nuts_params(fit),
  np_style = parcoord_style_np(
    div_color = "green",
    div_alpha = 0.6,
    div_size = 1.1
  )
)
```

```{r}
plot(
  fit, 
  variable = "prior_b",
  regex = TRUE, 
  nvariables = 3
)
```

```{r}
plot(fit, variable = "^b_rural", regex = TRUE, nvariables = 3)
```

### Model Output

#### Summary

```{r}
fit
```

#### Baseline

```{r}
#| fig-width: 5
#| fig-height: 3
plot_baseline_levels(fit, lims = c(0, 0.3))
```

#### Marginal Effects

##### Categorical variables

```{r}
#| fig-width: 5
#| fig-height: 3
plot_bin_marginal(fit, df_model, covs = bin_covs)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_bin_main_effect_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

##### Ordered factors

```{r}
#| fig-width: 8
#| fig-height: 3
plot_mono_marginal(fit, df_model, covs = mono_covs)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_mono_main_effect_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

##### Climate

```{r}
#| fig-width: 5
#| fig-height: 3
plot_continuous_marginal(fit, df_model, covs = cont_covs)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

#### Marginal Effects by Country

##### Ordered factors

```{r}
#| fig-width: 5.5
#| fig-height: 2.5
(
  plot_mono_marginal_by_group(fit, df_model, covs = mono_covs, dodge = 0.6)
  # + theme(
  #   legend.position = "top"
  # )
)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_mono_main_effect_by_country_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

##### Climate

```{r}
#| fig-width: 12
#| fig-height: 3
plot_single_continuous_marginal_by_group(fit, df_model, cov = "clim")
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_by_country_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

### Calibration Checks

```{r}
#| fig-width: 4
#| fig-height: 4
plot_calibration(fit, df_model, outcome_var = viol_var)
ggsave(
  path = fig_dir,
  filename = paste0("calibration_plot_main_effect_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

### Spatial Correlation Checks

##### Raw values

```{r}
#| fig-width: 5
#| fig-height: 2.5
plot_moran(df_model, sf_vacs, outcome_var = viol_var)
ggsave(
  path = fig_dir,
  filename = paste0("moran_plot_obs_main_effect_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

##### Residuals

```{r}
#| fig-width: 5
#| fig-height: 2.5
plot_moran(df_model, sf_vacs, fit = fit, residuals = TRUE)
ggsave(
  path = fig_dir,
  filename = paste0("moran_plot_res_main_effect_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

### Posterior Predictive Checks

```{r}
#| fig-width: 11
#| fig-height: 2.5
(
  plot_ppc_rootogram(fit, df_model, outcome_var = viol_var)
  + theme(
    legend.position = "inside",
    legend.position.inside = c(0.93, 0.82)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("ppc_by_adm_main_effect_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

### Inspect Climate Variable Posterior

```{r}
df_clim_pop <- (
  as_draws_df(fit)
  %>% select(clim_pop = b_clim)
)

df_clim_ranef <- (
  as_tibble(ranef(fit, summary = FALSE)$country[, , 2])
  %>% bind_cols(df_clim_pop)
  %>% mutate(
    across(
      .cols = -clim_pop, 
      .fns = ~ (.) + clim_pop,
      .names = "{.col}"
    )
  )
  %>% select(-clim_pop)
)


df_clim_ranef_plot <- (
  df_clim_ranef
  %>% pivot_longer(
    cols = everything(),
    names_to = "country",
    values_to = "value"
  )
)

df_clim_ranef_plot_stats <- (
  df_clim_ranef_plot
  %>% group_by(country)
  %>% summarise(
    median_val = median(value),
    mean_val = mean(value),
    lower1_val = bayestestR::hdi(value, ci = 0.5)$CI_low,
    upper1_val = bayestestR::hdi(value, ci = 0.5)$CI_high,
    lower2_val = bayestestR::hdi(value, ci = 0.95)$CI_low,
    upper2_val = bayestestR::hdi(value, ci = 0.95)$CI_high,
    .groups = "drop"
  )
)
```

```{r}
#| fig-width: 4.1
#| fig-height: 3
(
  ggplot(
    data = df_clim_pop,
    mapping = aes(
      x = clim_pop
    )
  )
  + stat_dotsinterval(
    quantiles = 100,
    slab_color = col_palette[3],
    slab_fill = col_palette[4],
    .width = c(0.5, 0.95),
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    plot.background = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
  )
  + labs(
    x = "Climate Posterior Value",
    y = "Density"
  )
)

ggsave(
  path = fig_dir,
  filename = paste0("clim_post_covariate_main_effect_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

```{r}
#| fig-width: 4.1
#| fig-height: 3
(
  ggplot(
    data = df_clim_ranef_plot,
    mapping = aes(
      x = value,
      y = country
    )
  )
  + ggridges::geom_density_ridges(
    aes(
      height = after_stat(scaled),
    ),
    stat = "density",
    orientation = "x",
    scale = 0.6,
    # rel_min_height = 0.01,
    alpha = 0.7,
    fill = "#FDD0A2FF",
    color = NA
  )
  + geom_segment(
    data = df_clim_ranef_plot_stats,
    aes(y = country, yend = country, x = lower2_val, xend = upper2_val),
    linewidth = 1,
    alpha = 0.4,
    color = "#A63603FF",
    inherit.aes = FALSE
  )
  + geom_segment(
    data = df_clim_ranef_plot_stats,
    aes(y = country, yend = country, x = lower1_val, xend = upper1_val),
    linewidth = 1.5,
    alpha = 0.7,
    color = "#A63603FF",
    inherit.aes = FALSE
  )
  + geom_point(
    data = df_clim_ranef_plot_stats,
    aes(y = country, x = median_val, color = country),
    size = 3,
    color = "#A63603FF",
    inherit.aes = FALSE
  )
  + geom_vline(xintercept = 0, linetype = "dashed", color = "gray50")
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 0.5),
    plot.background = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
  )
  + scale_x_continuous(
    limits = c(-0.9, 1.1)
  )
  + scale_y_discrete(
      expand = expansion(mult = c(0.1, 0.25))
    )
  + labs(
    x = "Climate Posterior Term",
    y = "Density"
  )
)

ggsave(
  path = fig_dir,
  filename = paste0("clim_post_covariate_main_effect_by_country_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

### Check Spatial Component

```{r}
df_car_vars <- (
  as_draws_df(
    fit,
    variable = c("car", "sdcar")
  )
  %>% select(-c(.chain, .iteration, .draw))
  %>% pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  )
)

df_car_vars_stats <- (
  df_car_vars
  %>% group_by(variable)
  %>% summarise(
    median = median(value, na.rm = T),
    low_ci1 = bayestestR::hdi(value, ci = 0.5)$CI_low,
    high_ci1 = bayestestR::hdi(value, ci = 0.5)$CI_high,
    low_ci2 = bayestestR::hdi(value, ci = 0.95)$CI_low,
    high_ci2 = bayestestR::hdi(value, ci = 0.95)$CI_high
  )
)
```

```{r}
#| fig-width: 4
#| fig-height: 2.5
(
  ggplot(
    data = df_car_vars,
    mapping = aes(
      y = fct_reorder(variable, value, .fun = ~ -median(.)),
      x = value,
    )
  )
  + stat_dotsinterval(
    quantiles = 100,
    slab_color = col_palette[3],
    slab_fill = col_palette[4],
    .width = c(0.5, 0.95),
  )
  + geom_vline(xintercept = 0, linetype = "dashed", color = "gray50")
  + facet_wrap(
    ~variable,
    ncol = 1,
    scales = "free_y",
    strip.position = "left"
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    plot.background = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    axis.title.y = element_blank(),
    axis.title.x = element_text(size = 16),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 15),
    axis.ticks.y = element_blank(),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text = element_text(size = 15)
  )
  + labs(
    x = "Posterior Value",
  )
  + scale_y_discrete(
    labels = label_map,
    expand = expansion(mult = c(0.2, 0.75))
  )
  # + scale_x_continuous(
  #   limits = c(-0.1, 1.8)
  # )
)

ggsave(
  path = fig_dir,
  filename = paste0("car_parameters_main_effect_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

```{r}
spatial_draws <- (
  as_draws_df(
    fit,
    variable = "^rcar",
    regex = TRUE
  )
  %>% select(-c(.chain, .iteration, .draw))
  %>% pivot_longer(
    cols = everything(),
    names_to = "param",
    values_to = "value"
  )
  %>% mutate(
    adm_code = as.numeric(gsub(".*\\[(.*?)\\].*", "\\1", param)),
    adm_name = sf_vacs$adm_name[adm_code]
  )
  %>% group_by(adm_name)
  %>% summarise(
    mean = mean(value, na.rm = TRUE),
    median = median(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    q10 = quantile(value, 0.1),
    q90 = quantile(value, 0.9),
    .groups = "drop"
  )
)
```

```{r}
sf_plot <- (
  sf_vacs
  %>% left_join(
    spatial_draws,
    by = "adm_name"
  )
)
```

```{r}
#| fig-width: 5
#| fig-height: 4.5
(
  ggplot(data = sf_plot)
  + geom_sf(
    mapping = aes(fill = mean),
    color = "black",
    linewidth = 0.1
  )
  + scale_fill_viridis_c()
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_blank(),
    plot.background = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    legend.background = element_blank(),
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 16),
    legend.title.position = "top",
    legend.position = "inside",
    legend.position.inside = c(0.2, 0.9),
    legend.direction = "horizontal"
  )
  + labs(
    fill = "Mean CAR term"
  )
)

ggsave(
  path = fig_dir,
  filename = paste0("spatial_term_map_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

# Compare Models

```{r}
fit_ns <- add_criterion(
  fit_ns, 
  "loo",
  # moment_match = TRUE
)
fit <- add_criterion(
  fit, 
  "loo", 
  # moment_match = TRUE
)
loo::loo_compare(
  fit_ns,
  fit
)
```

```{r}
bind_rows(
  get_calibration_scores(fit_ns, df_model, outcome_var = viol_var, model_name = "No SAR"),
  get_calibration_scores(fit, df_model, outcome_var = viol_var, model_name = "Full Model")
)
```

# Save in Temp files

```{r}
save(
  list = ls(envir = .GlobalEnv),
  file = file.path(temp_dir, paste0("workspace_backup_main_effect_", clim_label, "_", pop_label, "_", viol_label, ".RData")),
  compress = TRUE
)
```
