---
title: "Test Weighted UTCI"
format: html
editor: visual
---

# Environment Setup

```{r setup}
# Packages
library(tidyverse)
library(sf)
library(patchwork)

# Folder Structure
root <- rprojroot::has_file("agile2025.Rproj")$make_fix_file()
raw_data_dir <- root("data/raw")
proc_data_dir <- root("data/processed")
fig_dir <- root("figures")
dir.create(fig_dir, showWarnings = FALSE)
mod_dir <- root("models")

# Settings
nb_name <-  "Build climate-violence dataset - new"
seed <- sum(sapply(str_split(nb_name, pattern = "", simplify = TRUE), FUN = utf8ToInt))
set.seed(seed)
```

# Viz Settings

```{r}
feature_map <- c(
  "log_pop_density" = "Pop. Density (log)",
  "in_school" = "Enrolled in school",
  "work" = "Worked",
  "rural" = "Rural Area",
  "value_utci_days_above_26_daily_max_sum_12m" = "UTCI days max above 26",
  "value_utci_days_above_26_daily_min_sum_12m" = "UTCI days min above 26",
  "value_utci_days_above_32_daily_min_sum_12m" = "UTCI days min above 32",
  "value_utci_days_above_32_daily_max_sum_12m" = "UTCI days max above 32",
  "value_utci_days_above_38_daily_max_sum_12m" = "UTCI days max above 38",
  "value_utci_days_below_0_daily_min_sum_12m" = "UTCI days min below 0",
  "value_utci_days_below_9_daily_max_sum_12m" = "UTCI days max below 9",
  "value_utci_days_below_9_daily_min_sum_12m" = "UTCI days min below 9",
  "value_utci_monthly_max_max_12m" = "Max UTCI high",
  "value_utci_monthly_max_mean_12m" = "Mean UTCI high",
  "value_utci_monthly_min_max_12m" = "Max UTCI low",
  "value_spei_mean_12m" = "Mean SPEI",
  "value_spei_max_12m" = "Max SPEI",
  "value_spei_below_15q_sum_12m" = "SPEI months below 15th quantile",
  "value_spei_above_85q_sum_12m" = "SPEI months above 85th quantile",
  "dev_value_utci_days_above_26_daily_max_sum_12m" = "UTCI days max above 26\n(dev from mean)",
  "dev_value_utci_days_above_26_daily_min_sum_12m" = "UTCI days min above 26\n(dev from mean)",
  "dev_value_utci_days_above_32_daily_max_sum_12m" = "UTCI days max above 32\n(dev from mean)",
  "dev_value_utci_days_above_38_daily_max_sum_12m" = "UTCI days max above 38\n(dev from mean)",
  "dev_value_utci_days_below_0_daily_min_sum_12m" = "UTCI days min below 0\n(dev from mean)",
  "dev_value_utci_days_below_9_daily_max_sum_12m" = "UTCI days max below 9\n(dev from mean)",
  "dev_value_utci_days_below_9_daily_min_sum_12m" = "UTCI days min below 9\n(dev from mean)",
  "dev_value_utci_monthly_max_max_12m" = "Max UTCI high\n(dev from mean)",
  "dev_value_utci_monthly_max_mean_12m" = "Mean UTCI high\n(dev from mean)",
  "dev_value_utci_monthly_min_max_12m" = "Max UTCI low\n(dev from mean)",
  "dev_value_spei_mean_12m" = "Mean SPEI\n(dev from mean)",
  "dev_value_spei_max_12m" = "Max SPEI\n(dev from mean)",
  "dev_value_spei_below_15q_sum_12m" = "SPEI days below 15th quantile\n(dev from mean)",
  "dev_value_spei_above_85q_sum_12m" = "SPEI days above 85th quantile\n(dev from mean)"
)
```

# Data Import

## Geographies

### Mozambique

```{r}
sf_moz <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Mozambique 2022 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    adm1_name = dhsregen,
    adm1_code = regcode
  )
  %>% arrange(adm1_code)
)
```

### Zimbabwe

```{r}
sf_zwe <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Zimbabwe 2015 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    adm1_name = dhsregen,
    adm1_code = regcode
  )
  %>% arrange(adm1_code)
)
```

### Lesotho

```{r}
sf_lso <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Lesotho 2023 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    adm1_name = dhsregen,
    adm1_code = regcode
  )
  %>% arrange(adm1_code)
)
```

### Namibia

```{r}
sf_nam <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Namibia 2013 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    adm1_name = dhsregen,
    adm1_code = regcode
  )
  %>% arrange(adm1_code)
)
```

### Zambia

```{r}
sf_zmb <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Zambia 2018 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    adm1_name = dhsregen,
    adm1_code = regcode
  )
  %>% arrange(adm1_code)
)
```

### Kenya

```{r}
sf_ken <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Kenya 2022 DHS - sdr_subnational_boundaries_2025-11-13", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    adm1_name = dhsregen,
    adm1_code = regcode
  )
  %>% arrange(adm1_code)
)
```


## UTCI

There are 10 UTCI thermal stress categories that correspond to specific human physiological responses to the thermal environment. The categories relate to UTCI values as follows: above +46: extreme heat stress; +38 to +46: very strong heat stress; +32 to +38: strong heat stress; +26 to +32: moderate heat stress; +9 to +26: no thermal stress; +9 to 0: slight cold stress; 0 to -13: moderate cold stress; -13 to -27: strong cold stress; -27 to -40: very strong cold stress; below -40: extreme cold stress.

### Mozambique

```{r}
df_utci_moz <- (
  read_csv(
    file.path(proc_data_dir, "utci_MOZ.csv"),
    show_col_types = FALSE
  )
  %>% mutate(
    date = ceiling_date(date, "month") - days(1)
  )
)
```

### Zimbabwe

```{r}
df_utci_zwe <- (
  read_csv(
    file.path(proc_data_dir, "utci_ZWE.csv"),
    show_col_types = FALSE
  )
  %>% mutate(
    date = ceiling_date(date, "month") - days(1)
  )
)
```

### Lesotho

```{r}
df_utci_lso <- (
  read_csv(
    file.path(proc_data_dir, "utci_LSO.csv"),
    show_col_types = FALSE
  )
  %>% mutate(
    date = ceiling_date(date, "month") - days(1)
  )
)
```

### Namibia

```{r}
df_utci_nam <- (
  read_csv(
    file.path(proc_data_dir, "utci_NAM.csv"),
    show_col_types = FALSE
  )
  %>% mutate(
    date = ceiling_date(date, "month") - days(1)
  )
)
```

### Zambia

```{r}
df_utci_zmb <- (
  read_csv(
    file.path(proc_data_dir, "utci_ZMB.csv"),
    show_col_types = FALSE
  )
  %>% mutate(
    date = ceiling_date(date, "month") - days(1)
  )
)
```


### Kenya

```{r}
df_utci_ken <- (
  read_csv(
    file.path(proc_data_dir, "utci_KEN.csv"),
    show_col_types = FALSE
  )
  %>% mutate(
    date = ceiling_date(date, "month") - days(1)
  )
  %>% rename(adm1_code = id)
)
```


## Crop cycles

```{r}
df_crops <- (
  read_csv(
    file.path(
      raw_data_dir, 
      "crop",
      "harvestStat",
      "hvstat_africa_data_v1.0.csv"
    ),
    show_col_types = FALSE
  )
  %>% filter(
    country %in% c("Mozambique", "Zimbabwe", "Lesotho", "Namibia", "Zambia", "Kenya"),
    qc_flag == 0,
    product %in% c("Maize", "Sorghum", "Wheat", "Cassava", "Millet", "Rice", "Sweet Potatoes"),
    !is.na(production),
    harvest_year >= 2000,
    !(season_name %in% c("Annual", "Short"))
  )
  %>% group_by(country, admin_1, harvest_year)
  %>% slice_max(
    order_by = production,
    n = 1,
    with_ties = FALSE,
  )
  %>% ungroup()
  %>% mutate(
    planting_date = make_date(year = planting_year, month = planting_month, day = 1),
    harvest_date = make_date(year = harvest_year, month = harvest_month, day = 1),
  )
  %>% rowwise()
  %>% mutate(
    date = list(
      seq(
        from = planting_date,
        to = harvest_date,
        by = "month"
      )
    )
  )
  %>% unnest(date)
  %>% ungroup()
  %>% mutate(
    date = ceiling_date(date, unit = "month") - days()
  )
  %>% select(
    country,
    adm1_name = admin_1,
    date
  )
  %>% mutate(
    crop_season = 1
  )
)
```


```{r}
(
  df_crops
  %>% group_by(country)
  %>% summarise(
    max_date = max(date),
    min_date = min(date)
  )
  # %>% filter(
  #   country == "Mozambique",
  #   date == "2019-07-31"
  # )
)
```

## SPEI

### Mozambique

```{r}
df_spei_moz <- (
  read_csv(
    file.path(proc_data_dir, "spei_MOZ.csv"),
    show_col_types = FALSE
  )
  %>% rename(
    adm1_code = id
  )
)
```

### Zimbabwe

```{r}
df_spei_zwe <- (
  read_csv(
    file.path(proc_data_dir, "spei_ZWE.csv"),
    show_col_types = FALSE
  )
  %>% rename(
    adm1_code = id
  )
)
```

### Lesotho

```{r}
df_spei_lso <- (
  read_csv(
    file.path(proc_data_dir, "spei_LSO.csv"),
    show_col_types = FALSE
  )
  %>% rename(
    adm1_code = id
  )
)
```

### Namibia

```{r}
df_spei_nam <- (
  read_csv(
    file.path(proc_data_dir, "spei_NAM.csv"),
    show_col_types = FALSE
  )
  %>% rename(
    adm1_code = id
  )
)
```

### Zambia

```{r}
df_spei_zmb <- (
  read_csv(
    file.path(proc_data_dir, "spei_ZMB.csv"),
    show_col_types = FALSE
  )
  %>% rename(
    adm1_code = id
  )
)
```

### Kenya

```{r}
df_spei_ken <- (
  read_csv(
    file.path(proc_data_dir, "spei_KEN.csv"),
    show_col_types = FALSE
  )
  %>% rename(
    adm1_code = id
  )
)
```

# Process Data

## Combine UTCI, SPEI and Geography

### Mozambique

```{r}
df_clim_moz_adm1 <- (
  df_utci_moz
  %>% merge(
    sf_moz %>% select(adm1_code, adm1_name),
    by = "adm1_code",
    all.x = TRUE
  )
  %>% merge(
    df_spei_moz %>% select(-pop),
    by = c("adm1_code", "date"),
    all.x = TRUE
  )
  %>% mutate(
    country = "Mozambique",
    adm1_area = unclass(st_area(geometry)) / 1E6, #km^2
    pop_density = pop / adm1_area,
    log_pop_density = log10(pop_density)
  )
)
```

### Zimbabwe

```{r}
df_clim_zwe_adm1 <- (
  df_utci_zwe
  %>% merge(
    sf_zwe %>% select(adm1_code, adm1_name),
    by = "adm1_code",
    all.x = TRUE
  )
  %>% merge(
    df_spei_zwe %>% select(-pop),
    by = c("adm1_code", "date"),
    all.x = TRUE
  )
  %>% mutate(
    country = "Zimbabwe",
    adm1_area = unclass(st_area(geometry)) / 1E6,
    pop_density = pop / adm1_area,
    log_pop_density = log10(pop_density)
  )
)
```

### Lesotho

```{r}
df_clim_lso_adm1 <- (
  df_utci_lso
  %>% merge(
    sf_lso %>% select(adm1_code, adm1_name),
    by = "adm1_code",
    all.x = TRUE
  )
  %>% merge(
    df_spei_lso %>% select(-pop),
    by = c("adm1_code", "date"),
    all.x = TRUE
  )
  %>% mutate(
    country = "Lesotho",
    adm1_area = unclass(st_area(st_make_valid(geometry))) / 1E6,
    pop_density = pop / adm1_area,
    log_pop_density = log10(pop_density)
  )
)
```

### Namibia

```{r}
df_clim_nam_adm1 <- (
  df_utci_nam
  %>% merge(
    sf_nam %>% select(adm1_code, adm1_name),
    by = "adm1_code",
    all.x = TRUE
  )
  %>% merge(
    df_spei_nam %>% select(-pop),
    by = c("adm1_code", "date"),
    all.x = TRUE
  )
  %>% mutate(
    country = "Namibia",
    adm1_area = unclass(st_area(geometry)) / 1E6,
    pop_density = pop / adm1_area,
    log_pop_density = log10(pop_density)
  )
)
```

### Zambia

```{r}
df_clim_zmb_adm1 <- (
  df_utci_zmb
  %>% merge(
    sf_zmb %>% select(adm1_code, adm1_name),
    by = "adm1_code",
    all.x = TRUE
  )
  %>% merge(
    df_spei_zmb %>% select(-pop),
    by = c("adm1_code", "date"),
    all.x = TRUE
  )
  %>% mutate(
    country = "Zambia",
    adm1_area = unclass(st_area(st_make_valid(geometry))) / 1E6,
    pop_density = pop / adm1_area,
    log_pop_density = log10(pop_density)
  )
)
```

### Kenya

```{r}
df_clim_ken_adm1 <- (
  df_utci_ken
  %>% merge(
    sf_ken %>% select(adm1_code, adm1_name),
    by = "adm1_code",
    all.x = TRUE
  )
  %>% merge(
    df_spei_ken %>% select(-pop),
    by = c("adm1_code", "date"),
    all.x = TRUE
  )
  %>% mutate(
    country = "Kenya",
    adm1_area = unclass(st_area(st_make_valid(geometry))) / 1E6,
    pop_density = pop / adm1_area,
    log_pop_density = log10(pop_density)
  )
)
```


### Combine

```{r}
df_clim_adm1 <- (
  bind_rows(df_clim_moz_adm1, df_clim_zwe_adm1, df_clim_lso_adm1, df_clim_nam_adm1, df_clim_zmb_adm1, df_clim_ken_adm1)
  %>% merge(
    df_crops,
    by = c("country", "adm1_name", "date"),
    all.x = TRUE
  )
  %>% mutate(
    crop_season = if_else(is.na(crop_season), 0, 1),
    adm1_name = case_when(
      adm1_name == "Zambézia" ~ "Zambezia",
      adm1_name == "Maputo Provincia" ~ "Maputo",
      adm1_name == "Cidade de Maputo" ~ "Maputo City",
      adm1_name == "Harare Chitungwiza" ~ "Harare",
      adm1_name == "//Karas" ~ "!Karas",
      .default = adm1_name
    ),
  )
  %>% group_by(country, adm1_name, adm1_code)
  %>% mutate(
    dist_from_last_crop_season = {
      ones_pos <- which(crop_season == 1L)
      vapply(
        seq_along(crop_season),
        function(i) {
          if (crop_season[i] == 1L) return (0L)
          prev <- ones_pos[ones_pos < i]
          if (length(prev) == 0L) return (NA_integer_)
          i - max(prev)
        },
        integer(1)
      )
    }
  )
  %>% ungroup()
  %>% select(
    country,
    adm1_code,
    adm1_name,
    log_pop_density,
    date,
    utci_monthly_max,
    utci_days_above_32_daily_min,
    utci_days_above_32_daily_max,
    spei,
    crop_season,
    dist_from_last_crop_season
  )
  %>% arrange(country, adm1_name, date)
)
```

## Calculate Stats in previous year

```{r}
df_clim_adm1_roll <- (
  df_clim_adm1
  %>% mutate(
    spei_crop = spei * crop_season
  )
  %>% group_by(country, adm1_code, adm1_name, log_pop_density)
  %>% mutate(
    across(
      -c(date, crop_season, dist_from_last_crop_season, utci_monthly_max, spei, spei_crop),
      ~ zoo::rollsum((.), 12, fill = NA, align = "right"),
      .names = "{.col}_sum_12m"
    ),
    utci_monthly_max_mean_12m = zoo::rollmean(utci_monthly_max, 12, fill = NA, align = "right"),
    utci_monthly_max_max_12m = zoo::rollmax(utci_monthly_max, 12, fill = NA, align = "right"),
    spei_crop_sum_temp = zoo::rollsum(spei_crop, 12, fill = NA, align = "right"),
    spei_mean_12m = zoo::rollmean(spei, 12, fill = NA, align = "right"),
    spei_max_12m = zoo::rollmax(spei, 12, fill = NA, align = "right"),
    spei_crop_max_12m = zoo::rollmax(spei_crop, 12, fill = NA, align = "right"),
    spei_min_12m = zoo::rollapply(spei, 12, min, fill = NA, align = "right"),
    spei_crop_min_12m = zoo::rollapply(spei_crop, 12, min, fill = NA, align = "right"),
    spei_below_15q = if_else(
      spei < quantile(spei, 0.15),
      1,
      0
    ),
    spei_crop_below_15q = if_else(
      (spei_crop < quantile(spei_crop[crop_season == 1], 0.15)) & (crop_season == 1),
      1,
      0
    ),
    spei_above_85q = if_else(
      spei > quantile(spei, 0.85),
      1,
      0
    ),
    spei_crop_above_85q = if_else(
      (spei_crop > quantile(spei_crop[crop_season == 1], 0.85)) & (crop_season == 1),
      1,
      0
    ),
    spei_below_15q_sum_12m = zoo::rollsum(spei_below_15q, 12, fill = NA, align = "right"),
    spei_crop_below_15q_sum_12m = zoo::rollsum(spei_crop_below_15q, 12, fill = NA, align = "right"),
    spei_above_85q_sum_12m = zoo::rollsum(spei_above_85q, 12, fill = NA, align = "right"),
    spei_crop_above_85q_sum_12m = zoo::rollsum(spei_crop_above_85q, 12, fill = NA, align = "right"),
    n_crop_months = zoo::rollsum(crop_season, 12, fill = NA, align = "right")
  )
  %>% ungroup()
  %>% mutate(
    spei_crop_mean_12m = if_else(
      n_crop_months == 0, 
      0,
      spei_crop_sum_temp / n_crop_months
    )
  )
  %>% filter(!is.na(spei_mean_12m))
  %>% select(country, adm1_code, adm1_name, log_pop_density, date, dist_from_last_crop_season, contains("12m"))
)
```


```{r}
col_palette <- c("#dc7726", "#eebb93", "#ecc046","#0292b7", "#81c9db", "#35A29F")
(
  ggplot(
    data = df_clim_adm1_roll %>% filter(date > "2015-01-01"), 
    aes(
      x = utci_days_above_32_daily_max_sum_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    position = "stack"
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_text(size = 14),
      legend.text = element_text(size = 13),
    )
  + scale_fill_manual(
    values = col_palette
  )
)
```


## Calculate deviations from average

```{r}
df_clim_adm1_roll_dev <- (
  df_clim_adm1_roll
  %>% pivot_longer(
    cols = -c(country, adm1_code, adm1_name, log_pop_density, date, dist_from_last_crop_season),
    names_to = "variable",
    values_to = "value"
  )
  %>% group_by(country, adm1_code, adm1_name, log_pop_density, variable)
  %>% mutate(
    mean_value = mean(value, na.rm = TRUE),
  )
  %>% ungroup()
  %>% mutate(
    dev_value = (value - mean_value) #/ mean_value
  )
  %>% pivot_wider(
    id_cols = c(country, adm1_code, adm1_name, log_pop_density, date, dist_from_last_crop_season),
    names_from = "variable",
    values_from = c(value, dev_value)
  )
)
```


```{r}
col_palette <- c("#dc7726", "#eebb93", "#ecc046","#0292b7", "#81c9db", "#35A29F")
(
  ggplot(
    data = df_clim_adm1_roll_dev %>% filter(date > "2015-01-01", country != "Kenya"), 
    aes(
      x = dev_value_utci_days_above_32_daily_max_sum_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    position = "stack"
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_text(size = 14),
      legend.text = element_text(size = 13),
    )
  + scale_fill_manual(
    values = col_palette
  )
)
```

```{r}
colSums(is.na(df_clim_adm1_roll_dev))
```

```{r}
(
  df_clim_adm1_roll_dev
  %>% filter(is.na(dev_value_spei_above_85q_sum_12m))
)
```

# Visualizations

#### UTCI

```{r}
#| fig-width: 9
#| fig-height: 5

(
  df_clim_adm1
  %>% filter(
    date > "2010-01-01",
    date < "2025-01-01",
    country == "Mozambique",
    !(adm1_name %in% c("Tete", "Zambezia"))
  )
  %>% select(
    date,
    adm1_name,
    utci_days_above_32_daily_max,
  )
  %>% ggplot(
    mapping = aes(x = date, y = utci_days_above_32_daily_max)
  )
  + geom_line(
    color = "#0292b7",
    linewidth = 1.3
  )
  + facet_wrap(
    ~ adm1_name,
    ncol = 3
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    panel.spacing.x = unit(25, "pt"),
    strip.background = element_blank(),
    strip.text = element_text(size = 14),
    plot.background = element_blank(),
    plot.title = element_text(size = 16),
    plot.margin = margin(t = 0, r = 20, b = 0, l = 10, unit = "pt"),
    axis.ticks = element_line(linewidth = 1),
    axis.ticks.length = unit(4, "pt"),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 15),
  )
  + labs(
    x = "Date",
    y = "# days UTCI max > 32 (month)"
  )
)
ggsave(
  path = fig_dir,
  filename = "utci_max_above_32_month_moz.png",
  width = 9,
  height = 5
)
```

```{r}
#| fig-width: 9
#| fig-height: 5

(
  df_clim_adm1_roll
  %>% filter(
    date > "2010-01-01",
    date < "2025-01-01",
    country == "Mozambique",
    !(adm1_name %in% c("Tete", "Zambezia"))
  )
  %>% select(
    date,
    adm1_name,
    utci_days_above_32_daily_max_sum_12m,
  )
  %>% ggplot(
    mapping = aes(x = date, y = utci_days_above_32_daily_max_sum_12m)
  )
  + geom_line(
    color = "#dc7726",
    linewidth = 1.3
  )
  + facet_wrap(
    ~ adm1_name,
    ncol = 3
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    panel.spacing.x = unit(25, "pt"),
    strip.background = element_blank(),
    strip.text = element_text(size = 14),
    plot.background = element_blank(),
    plot.title = element_text(size = 16),
    plot.margin = margin(t = 0, r = 20, b = 0, l = 10, unit = "pt"),
    axis.ticks = element_line(linewidth = 1),
    axis.ticks.length = unit(4, "pt"),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 15),
  )
  + labs(
    x = "Date",
    y = "# days UTCI max > 32 (year)"
  )
)
ggsave(
  path = fig_dir,
  filename = "utci_max_above_32_year_moz.png",
  width = 9,
  height = 5
)
```

```{r}
#| fig-width: 9
#| fig-height: 5

(
  df_clim_adm1_roll_dev
  %>% filter(
    date > "2010-01-01",
    date < "2025-01-01",
    country == "Mozambique",
    !(adm1_name %in% c("Tete", "Zambezia"))
  )
  %>% select(
    date,
    adm1_name,
    dev_value_utci_days_above_32_daily_max_sum_12m,
  )
  %>% ggplot(
    mapping = aes(x = date, y = dev_value_utci_days_above_32_daily_max_sum_12m)
  )
  + geom_line(
    color = "#ecc046",
    linewidth = 1.3
  )
  + facet_wrap(
    ~ adm1_name,
    ncol = 3
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    panel.spacing.x = unit(25, "pt"),
    strip.background = element_blank(),
    strip.text = element_text(size = 14),
    plot.background = element_blank(),
    plot.title = element_text(size = 16),
    plot.margin = margin(t = 0, r = 20, b = 0, l = 10, unit = "pt"),
    axis.ticks = element_line(linewidth = 1),
    axis.ticks.length = unit(4, "pt"),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 15),
  )
  + labs(
    x = "Date",
    y = "# days UTCI max > 32 (dev. from mean year)"
  )
)
ggsave(
  path = fig_dir,
  filename = "dev_utci_max_above_32_year_moz.png",
  width = 9,
  height = 5
)
```

#### SPEI

```{r}
#| fig-width: 9
#| fig-height: 5

(
  df_clim_adm1_roll
  %>% filter(
    date > "2010-01-01",
    date < "2025-01-01",
    country == "Mozambique",
    !(adm1_name %in% c("Tete", "Zambezia"))
  )
  %>% ggplot(
    mapping = aes(x = date, y = spei_crop_max_12m)
  )
  + geom_line(
    color = "#0292b7",
    linewidth = 1.3
  )
  + facet_wrap(
    ~ adm1_name,
    ncol = 3
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    panel.spacing.x = unit(25, "pt"),
    strip.background = element_blank(),
    strip.text = element_text(size = 14),
    plot.background = element_blank(),
    plot.title = element_text(size = 16),
    plot.margin = margin(t = 0, r = 20, b = 0, l = 10, unit = "pt"),
    axis.ticks = element_line(linewidth = 1),
    axis.ticks.length = unit(4, "pt"),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 15),
  )
  + labs(
    x = "Date",
    y = "SPEI"
  )
)
ggsave(
  path = fig_dir,
  filename = "spei_moz.png",
  width = 9,
  height = 5
)
```

```{r}
#| fig-width: 9
#| fig-height: 5

(
  df_clim_adm1_roll
  %>% filter(
    date > "2010-01-01",
    date < "2025-01-01",
    country == "Mozambique",
    !(adm1_name %in% c("Tete", "Zambezia"))
  )
  %>% ggplot(
    mapping = aes(x = date, y = spei_crop_above_85q_sum_12m)
  )
  + geom_line(
    color = "#dc7726",
    linewidth = 1.3
  )
  + facet_wrap(
    ~ adm1_name,
    ncol = 3
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    panel.spacing.x = unit(25, "pt"),
    strip.background = element_blank(),
    strip.text = element_text(size = 14),
    plot.background = element_blank(),
    plot.title = element_text(size = 16),
    plot.margin = margin(t = 0, r = 20, b = 0, l = 10, unit = "pt"),
    axis.ticks = element_line(linewidth = 1),
    axis.ticks.length = unit(4, "pt"),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 15),
  )
  + labs(
    x = "Date",
    y = "SPEI (year)"
  )
)
ggsave(
  path = fig_dir,
  filename = "spei_above_85q_sum_12m_moz.png",
  width = 9,
  height = 5
)
```

```{r}
#| fig-width: 9
#| fig-height: 5

(
  df_clim_adm1_roll_dev
  %>% filter(
    date > "2010-01-01",
    date < "2025-01-01",
    country == "Mozambique",
    !(adm1_name %in% c("Tete", "Zambezia"))
  )
  %>% ggplot(
    mapping = aes(x = date, y = dev_value_spei_crop_above_85q_sum_12m)
  )
  + geom_line(
    color = "#ecc046",
    linewidth = 1.3
  )
  + facet_wrap(
    ~ adm1_name,
    ncol = 3
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    panel.spacing.x = unit(25, "pt"),
    strip.background = element_blank(),
    strip.text = element_text(size = 14),
    plot.background = element_blank(),
    plot.title = element_text(size = 16),
    plot.margin = margin(t = 0, r = 20, b = 0, l = 10, unit = "pt"),
    axis.ticks = element_line(linewidth = 1),
    axis.ticks.length = unit(4, "pt"),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 15),
  )
  + labs(
    x = "Date",
    y = "# months SPEI above 85th quantile (dev. from mean year)"
  )
)
ggsave(
  path = fig_dir,
  filename = "dev_value_spei_above_85q_sum_12m_moz.png",
  width = 9,
  height = 5
)
```

# Save data

```{r}
df_clim_adm1_roll_dev %>% write_csv(file.path(proc_data_dir, "processed_utci_spei.csv"))
```
