---
title: "Uganda Disaster Analysis"
format: html
editor: visual
---

# Environment Setup

```{r setup}
library(tidyverse)
library(tidyterra)
library(readxl)
library(sf)
library(patchwork)
library(paletteer)
library(ggtext)
library(scales)

# Set random seed
nb_name <- "Agile 2025 - Data Exploration"
seed <- sum(sapply(str_split(nb_name, pattern = "", simplify = TRUE), FUN = utf8ToInt))

# Folder structure
root <- rprojroot::has_file("agile2025.Rproj")$make_fix_file()
raw_data_dir <- root("data/raw")
proc_data_dir <- root("data/processed")
dir.create(proc_data_dir, showWarnings = FALSE)
fig_dir <- root("figures")
dir.create(fig_dir, showWarnings = FALSE)

# Viz settings
col_palette <- c("#dc7726", "#eebb93", "#ecc046", "#f6e0a3", "#0292b7", "#81c9db", "#97DECE", "#35A29F")
theme_set(
  theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = 15, face = "bold"),
    plot.background = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
  )
)
```

# Functions

```{r}
clean_column_name <- function(cols) {
  cols <- tolower(cols)
  cols <- gsub("[.:,'()-/?]", "", cols)
  cols <- str_squish(cols)
  for (i in 1:length(cols)) {
    cols[i] <- paste0(unlist(strsplit(cols[i], " ")), collapse = "_")
  }
  cols
}

load_dhis_viol_data <- function(filename) {
  filepath <- file.path(raw_data_dir, "Uganda", "DHIS2", filename)
  sheet_names <- excel_sheets(filepath)
  dfs <- list()
  for (i in 1:length(sheet_names)) {
    dhis_raw <- read_xlsx(
      file.path(raw_data_dir, "Uganda", "DHIS2", filename),
      sheet = sheet_names[i],
      col_names = FALSE,
      .name_repair = "unique_quiet"
    )
    headers <- as_tibble(t(zoo::na.locf(t(dhis_raw[1:4, 1:(dim(dhis_raw)[2] -1)]))))
    combined_names <- apply(headers, 2, function(x) paste(x, collapse = "_"))
  
    df_dhis <- (
      read_xlsx(
        file.path(raw_data_dir, "Uganda", "DHIS2", filename),
        sheet = sheet_names[i],
        skip = 5,
        col_names = FALSE,
        .name_repair = "unique_quiet"
      )
      %>% select(-last_col())
      %>% head(-1)
    )
    colnames(df_dhis) <- combined_names
  
    dfs[[i]] <- (
      df_dhis
      %>% rename(
        district = 1
      )
      %>% pivot_longer(
        cols = -district,
        names_to = c("variable", "month", "sex", "age_bracket"),
        names_sep = "_",
        values_to = "cases"
      )
      %>% mutate(
        year = sheet_names[i],
        sex = if_else(
          variable == "Abortions due to Gender-Based Violence",
          "Female",
          sex
        )
      )
    )
  }
  (
    bind_rows(dfs)
    %>% mutate(
      district = str_squish(str_remove_all(district, "District")),
      date = lubridate::ceiling_date(
        as.Date(paste(year, month, "01", sep="-"), "%Y-%B-%d"),
        "month"
      ) - lubridate::days(1),
      cases = replace_na(cases, 0)
    )
    # %>% group_by(district, date, variable, sex, age_bracket)
    # %>% summarise(
    #   cases = sum(cases, na.rm = TRUE),
    #   .groups = "drop"
    # )
    %>% arrange(district, date, variable, sex, age_bracket)
  )
}

load_dhis_maln_data <- function(filename) {
  filepath <- file.path(raw_data_dir, "Uganda", "DHIS2", filename)
  sheet_names <- excel_sheets(filepath)
  dfs <- list()
  for (i in 1:length(sheet_names)) {
    dhis_raw <- read_xlsx(
      file.path(raw_data_dir, "Uganda", "DHIS2", filename),
      sheet = sheet_names[i],
      col_names = FALSE,
      .name_repair = "unique_quiet"
    )
    headers <- as_tibble(t(zoo::na.locf(t(dhis_raw[1:2, 1:(dim(dhis_raw)[2] -1)]))))
    combined_names <- apply(headers, 2, function(x) paste(x, collapse = "_"))
  
    df_dhis <- (
      read_xlsx(
        file.path(raw_data_dir, "Uganda", "DHIS2", filename),
        sheet = sheet_names[i],
        skip = 2,
        col_names = FALSE,
        .name_repair = "unique_quiet"
      )
      %>% select(-last_col())
    )
    colnames(df_dhis) <- combined_names
  
    dfs[[i]] <- (
      df_dhis
      %>% rename(
        district = "Data element_District/Month"
      )
      %>% pivot_longer(
        cols = -district,
        names_to = c("variable", "month"),
        names_sep = "_",
        values_to = "cases"
      )
      %>% mutate(
        year = sheet_names[i],
        # sex = if_else(
        #   variable == "Abortions due to Gender-Based Violence",
        #   "Female",
        #   sex
        # )
      )
    )
  }
  (
    bind_rows(dfs)
    %>% mutate(
      district = str_squish(str_remove_all(district, "District")),
      date = lubridate::ceiling_date(
        as.Date(paste(year, month, "01", sep="-"), "%Y-%B-%d"),
        "month"
      ) - lubridate::days(1),
      cases = replace_na(cases, 0)
    )
    %>% arrange(district, date, variable)
  )
}
```

# Parameters

```{r params}

regions_with_operations <- c(
  "Mayuge",
  "Budaka",
  "Nakapiripiti",
  "Tororo",
  "Bulambuli",
  "Soroti",
  "Lira",
  "Napak",
  "Kamuli",
  "Kayunga",
  "Luuka",
  "Sironko"
)

elgon_region <- c(
  "Bududa", 
  "Bukedea",
  "Bukwo",
  "Bulambuli",
  "Butaleja",
  "Kapchorwa",
  "Kween",
  "Manafwa", 
  "Mbale", 
  "Namisindwa",
  "Sironko", 
  "Tororo"
)

karamoja_region <- c(
  "Abim",
  "Amudat",
  "Kaabong", 
  "Karenga",
  "Kotido", 
  "Moroto",
  "Nabilatuk",
  "Nakapiripirit",
  "Napak"
)

rwenzori_region <- c(
  "Bundibugyo",
  "Bunyangabu",
  "Kasese"
)

teso_region <- c(
  "Amuria",
  "Kapelebyong",
  "Katakwi",
  "Kumi",
  "Bukedea",
  "Butebo",
  "Budaka"
)

west_nile_region <- c(
  "Adjumani",
  "Madi Okollo",
  "Zombo"
)
  

commodities_to_plot <- c("Maize (white)", "Maize flour", "Cassava flour", "Sorghum")
```

# Load data

## Geometries

Data from: https://data.humdata.org/dataset/cod-ab-uga

```{r load_data geom}
# sf_uga <- (
#   st_read(
#     file.path(raw_data_dir, "geographies", "uga_admbnda_ubos_20200824.gdb", fsep = "/"),
#     layer = "uga_admbnda_adm2_ubos_20200824"
#   )
#   %>% rename_with(tolower)
#   %>% st_make_valid()
#   %>% st_simplify(
#     dTolerance = 100
#   )
#   %>% mutate(
#     id = row_number(),
#     admin2name_en = case_when(
#       admin2name_en == "Madi Okollo" ~ "Madi-Okollo",
#       admin2name_en == "Ssembabule" ~ "Sembabule",
#       .default = admin2name_en
#     )
#   )
#   # %>% select(
#   #   id,
#   #   adm1_name = admin1name_en,
#   #   adm2_name = admin2name_en,
#   #   adm2_pcode = admin2pcode,
#   #   # adm3_name = admin3name_en,
#   #   # adm4_name = admin4name_en,
#   #   # adm4_pcode = admin4pcode,
#   #   geometry = shape
#   # )
# )
```

```{r}
(
  sf_uga
  %>% st_crop(
    xmin = 30.88,
    ymin = 2.98,
    xmax = 31.13,
    ymax = 3.27
  )
)$adm2_name %>% unique()
    
```

From data sent by Herbert

```{r load_data geom}
# Sys.setenv(SHAPE_RESTORE_SHX='YES')
sf_uga <- (
  st_read(
    file.path(
      raw_data_dir,
      "geographies",
      "uganda_herbert",
      "uganda_districts.shp"
    ),
  )
  %>% rename_with(tolower)
  # %>% mutate(
  #   district = str_to_title(district)
  # )
  %>% mutate(
    district = case_match(
      district,
      "SSEMBABULE" ~ "Sembabule",
      .default = str_to_title(district)
    )
  )
  %>% select(district)
)
```


## Population

https://data.humdata.org/dataset/cod-ps-global

```{r}
df_pop_cols <- (
  read_csv(
    file.path(raw_data_dir, "population", "cod_population_admin2.csv"),
    show_col_types = FALSE,
    n_max = 0
  )
  %>% names()
)

df_pop <- (
  read_csv(
    file.path(raw_data_dir, "population", "cod_population_admin2.csv"),
    show_col_types = FALSE,
    col_names = df_pop_cols,
    skip = 2
  )
  %>% rename_with(tolower)
  %>% filter(iso3 == "UGA")
  %>% select(
    adm2_code = adm2_pcode,
    adm2_name,
    population_group,
    population
  )
  %>% pivot_wider(
    id_cols = c(adm2_code, adm2_name),
    names_from = population_group,
    values_from = population,
    names_prefix = "pop_"
  )
  %>% rename_with(tolower)
)
```

Data from: https://hub.worldpop.org/geodata/listing?id=136

```{r}
# rast_pop_2022 <- (
#   terra::rast(
#     file.path(raw_data_dir, "population", "uga_pop_2022_CN_1km_R2025A_UA_v1.tif")
#   )
#   %>% rename(
#     pop = global_pop_2022_CN_1km_R2025A_v1
#   )
# )
```
 
```{r}
# df_pop_adm_2022 <- (
#   rast_pop_2022
#   %>% terra::extract(
#     sf_uga,
#     na.rm = TRUE,
#     cells = TRUE,
#     exact = TRUE
#   )
#   %>% group_by(cell)
#   %>% mutate(
#     fraction = fraction / sum(fraction, na.rm = TRUE)
#   )
#   %>% group_by(id = ID)
#   %>% summarise(
#     pop = sum(pop * fraction, na.rm = TRUE)
#   )
# )
```


## Crop cycles

```{r}
df_crops <- (
  read_csv(
    file.path(
      raw_data_dir, 
      "crop",
      "harvestStat",
      "hvstat_africa_data_v1.0.csv"
    ),
    show_col_types = FALSE
  )
  %>% filter(
    country %in% c("Uganda"),
    qc_flag == 0,
    product %in% c("Maize", "Sorghum", "Wheat", "Cassava", "Millet", "Rice", "Sweet Potatoes"),
    !is.na(production),
    harvest_year >= 2000
  )
  %>% group_by(country, admin_1, harvest_year)
  %>% slice_max(
    order_by = production,
    n = 1,
    with_ties = FALSE,
  )
  %>% ungroup()
  %>% mutate(
    planting_date = make_date(year = planting_year, month = planting_month, day = 1),
    harvest_date = make_date(year = harvest_year, month = harvest_month, day = 1),
  )
  %>% rowwise()
  %>% mutate(
    date = list(
      seq(
        from = planting_date,
        to = harvest_date,
        by = "month"
      )
    )
  )
  %>% unnest(date)
  %>% ungroup()
  %>% mutate(
    date = ceiling_date(date, unit = "month") - days()
  )
  %>% select(
    country,
    adm1_name = admin_1,
    date
  )
  %>% mutate(
    crop_season = 1
  )
)
```

## Food Prices

Data from: https://data.humdata.org/dataset/global-wfp-food-prices

```{r load data food prices}
food_prices_files <- list.files(file.path(raw_data_dir, "food_prices", "wfp_food_prices", fsep = "/"))
list_food_prices <- list()
for (i in seq(length(food_prices_files))) {
  list_food_prices[[i]] <- (
    read_csv(
      file.path(raw_data_dir, "food_prices", "wfp_food_prices", food_prices_files[i], fsep = "/"),
      show_col_types = FALSE
    )
    %>% filter(
    # category == "cereals and tubers"
      pricetype == "Retail",
      !is.na(latitude),
      category %in% c("cereals and tubers", "pulses and nuts", "vegetables and fruits")
    )
    %>% mutate(
      admin2 = str_replace(admin2, " Municipality", "")
    )
  )
}
```

```{r}
sf_food_prices_wfp_uga <- (
  bind_rows(list_food_prices)
  %>% filter(
    countryiso3 == "UGA"
  )
  %>% select(
    date,
    admin1,
    admin2,
    market,
    latitude,
    longitude,
    category,
    commodity,
    unit,
    price
  )
  %>% mutate(
    date = as.Date(date, format = "%Y-%m-%d"),
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude),
    price = as.numeric(price)
  )
  %>% st_as_sf(
    coords = c("longitude", "latitude"),
    crs = 4326
  )
)
```

## Crop and Rangeland Warnings

data from: https://data.jrc.ec.europa.eu/collection/id-0112

```{r}
df_crop_warn <- (
  read_delim(
    file.path(raw_data_dir, "crop", "warnings_l2_ts", "warnings_l2_ts.csv"),
    show_col_types = FALSE,
    delim = ";",
    lazy = TRUE
  )
  %>% filter(
    asap0_name == "Uganda",
    date >= "2020-01-01"
  )
  %>% group_by(asap2_name, month = floor_date(date, "month"))
  %>% summarise(
    w_crop = mean(w_crop, na.rm = TRUE),
    w_range = mean(w_range, na.rm = TRUE),
    .groups = "drop"
  )
  %>% mutate(
    date = ceiling_date(month, "month") - days(1)
  )
  %>% select(
    adm2_name = asap2_name,
    date,
    w_crop,
    w_range
  )
)
```


## Food Insecurity

https://data.humdata.org/dataset/uganda-acute-food-insecurity-country-data

```{r}
fs_cols <- (
  read_csv(
    file.path(raw_data_dir, "food_security", "ipc_uga_area_long.csv"),
    show_col_types = FALSE,
    n_max = 0
  )
  %>% names()
)

df_fs <- (
  read_csv(
    file.path(raw_data_dir, "food_security", "ipc_uga_area_long.csv"),
    show_col_types = FALSE,
    col_names = fs_cols,
    skip = 2
  )
  %>% rename_with(tolower)
  %>% filter(
    `validity period` == "current",
    !(phase %in% c("all", "3+")),
    is.na(`level 1`)
  )
  %>% select(
    adm2_name = area,
    date = to,
    phase,
    percentage
  )
  %>% pivot_wider(
    id_cols = c(adm2_name, date),
    names_from = phase,
    values_from = percentage,
    names_prefix = "fs_level_"
  )
  %>% mutate(
    fs_level_above_3 = rowSums(select(., c(fs_level_3, fs_level_4, fs_level_5)))
  )
)
```


## Malnutrition

```{r}
df_maln <- load_dhis_maln_data("Nutrition_Data.xlsx")
```



## NDVI 

https://data.humdata.org/dataset/uga-ndvi-subnational

```{r}
df_ndvi <- (
  read_csv(
    file.path(raw_data_dir, "NDVI", "uga-ndvi-subnat-full.csv"),
    show_col_types = FALSE
  )
  %>% filter(
    date >= "2020-01-01"
  )
  %>% mutate(
    adm2_code = paste0("UG", adm_id)
  )
  %>% select(
    adm2_code,
    date,
    vim,      # 10-day NDVI
    vim_avg,  # NDVI long-term average
    viq       # 10 day NDVI anomaly
  )
)
```


## UTCI

```{r}
df_utci <- (
  read_csv(
    file.path(proc_data_dir, "utci_UGA.csv"),
    show_col_types = FALSE
  )
  %>% mutate(
    date = ceiling_date(date, "month") - days(1)
  )
  %>% rename(district = id)
)
```

## SPEI

```{r}
df_spei <- (
  read_csv(
    file.path(proc_data_dir, "spei_UGA.csv"),
    show_col_types = FALSE
  )
  %>% mutate(
    date = ceiling_date(date, "month") - days(1)
  )
  %>% rename(district = id)
)
```



## Disasters - EMDAT

Raw data from: https://public.emdat.be/. Cleaned by: - removing disasters with unclear or very widespread location - removing duplicates in location names (i.e. location appearing both as districts and regions) - limited the spatial information to the district level

```{r}
df_dis_emdat <- (
  read_csv(
    file.path(proc_data_dir, "uganda_relevant_disasters_2010_2025_cleaned.csv"),
    show_col_types = FALSE
  )
  %>% filter(
    !(is.na(start_date) | is.na(end_date) | is.na(location))
  )
  %>% mutate(
    start_date = as.Date.character(start_date, format = "%m/%d/%y"),
    end_date = as.Date.character(end_date, format = "%m/%d/%y")
  )
  %>% arrange(desc(total_affected))
)
```

```{r}
df_dis_emdat_by_day <- (
  df_dis_emdat
  %>% filter(
    !is.na(total_affected),
    total_affected > 0
  )
  %>% select(
    disno,
    disaster_type,
    start_date,
    end_date,
    location,
    total_affected
  )
  %>% mutate(
    location = str_split(location, pattern = ", ")
  )
  %>% unnest_longer(
    col = location
  )
  %>% group_by(disno)
  %>% mutate(
    n_loc_affected = n()
  )
  %>% mutate(
    duration = as.numeric(end_date - start_date) + 1,
    total_affected = floor(total_affected / (n_loc_affected * duration)),
  )
  %>% group_by(disno, disaster_type, location, total_affected)
  %>% reframe(
    date = as.Date(start_date:end_date),
  )
  %>% ungroup()
  %>% select(date, location, total_affected)
  %>% group_by(location)
  %>% complete(
    date = seq(min(df_dis_emdat$start_date), max(df_dis_emdat$end_date), by = "1 day"),
    fill = list(total_affected = 0)
  )
  %>% ungroup()
  %>% arrange(location, date)
)

df_dis_emdat_month <- (
  df_dis_emdat_by_day
  %>% group_by(date = cut(date, "month"), location)
  %>% summarise(
    total_affected = sum(total_affected, na.rm = TRUE),
    .groups = "drop"
  )
  # %>% mutate(
  #   total_affected = zoo::rollmean(total_affected, k = 30, fill = NA, align = "right"),
  # )
  %>% mutate(
    date = as.Date(date)
  )
  %>% ungroup()
  %>% na.omit()
)
```

## Disasters - IOM

Data from: https://data.humdata.org/dataset/uganda-displacement-data-multi-hazard-response-drr-platform-iom-dtm

```{r}
dis_to_remove <- c(
  "Transport Related Accidents",
  "Lightning",
  "Land Conflicts",
  "Accident",
  "Collapsed Structure/Buildings/ Earth.Quarry",
  "Plane Crash",
  "Human Epidemics",
  "Disease",
  "Cholera",
  "Other Retrogressive Cultural Practices",
  "Terrorism",
  "Boating Accident",
  "Disease Outbreak (Suspected)",
  "Public Riots"
)

df_dis_iom <- (
  read_excel(
    file.path(proc_data_dir, "uganda_multihazard_cleaned.xlsx"),
    sheet = "list_of_events"
  )
  %>% rename_with(clean_column_name)
  %>% mutate(
    across(where(is.numeric), as.integer),
    across(
      c(
        "what_is_the_level_of_damage_on_crops",
        "health", 
        "water_supply", 
        "sanitation", 
        "shelter", 
        "food_assistance", 
        "nutrition", 
        "education",
        "nfi",
        "protection",
        "child_protection",
        "sgbv",
        "livelihood"
      ),
      tolower
    ),
    across(
      c("region", "subregion", "district"),
      str_to_title
    ),
      
    reported_date = as.Date(reported_date),
    date = if_else(
      is.na(date),
      reported_date,
      as.Date(date)
    )
  )
  %>% filter(
    !(type_of_emergency_disaster %in% dis_to_remove),
    date >= "2022-01-01"
  )
  %>% select(
    date,
    disaster = type_of_emergency_disaster,
    region,
    subregion,
    district,
    affected = affected_people_total,
    homeless,
    hospitalised,
    evacuated,
    households_affected,
    households_displaced,
    crops_affected = have_crops_been_affected,
    crops_damage = what_is_the_level_of_damage_on_crops,
    business_affected = has_usual_businesslivelihood_of_people_in_this_area_been_affected,
    prepositioned_stock = are_there_any_prepositioned_stock_at_the_branch_that_can_facilitate_immediate_response,
    other_actors = are_other_actors_already_taking_response_actions,
    worsening_situation = is_the_situation_likely_to_get_better_or_worsen,
    main_need = what_is_the_most_pressing_need,
    health,
    water_supply,
    sanitation,
    shelter,
    food_assistance,
    nutrition,
    education,
    nfi,
    protection,
    child_protection,
    sgbv,
    livelihood
  )
)
```

```{r}
df_dis_iom
```



## Violence

```{r}
dhis_raw <- read_xlsx(
  file.path(raw_data_dir, "Uganda", "DHIS2", "DHIS2_GBV_Data.xlsx"),
  sheet = "2020",
  col_names = FALSE,
  .name_repair = "unique_quiet"
)
headers <- as_tibble(t(zoo::na.locf(t(dhis_raw[1:4, 1:(dim(dhis_raw)[2] -1)]))))
combined_names <- apply(headers, 2, function(x) paste(x, collapse = "_"))
```

```{r}
df_dhis <- (
  read_xlsx(
    file.path(raw_data_dir, "Uganda", "DHIS2", "DHIS2_GBV_Data.xlsx"),
    sheet = "2020",
    skip = 5,
    col_names = TRUE,
    .name_repair = "unique_quiet"
  )
  %>% select(-last_col())
  %>% head(-1)
)
  
```

```{r}
df_sgbv <- load_dhis_viol_data("DHIS2_GBV_Data.xlsx")
```

# Data processing

```{r}
df_clim_uga_adm1 <- (
  df_utci_uga
  %>% merge(
    sf_uga %>% select(adm1_code, adm1_name),
    by = "adm1_code",
    all.x = TRUE
  )
  %>% merge(
    df_spei_zmb %>% select(-pop),
    by = c("adm1_code", "date"),
    all.x = TRUE
  )
  %>% mutate(
    country = "Zambia",
    adm1_area = unclass(st_area(st_make_valid(geometry))) / 1E6,
    pop_density = pop / adm1_area,
    log_pop_density = log10(pop_density)
  )
)
```


## Combine climate, geographies and population

```{r}
df_clim_adm2 <- (
  df_uga
  %>% merge(
    df_utci,
    by = "id",
    all.x = TRUE
  )
  %>% merge(
    df_spei_zwe %>% select(-pop),
    by = c("adm1_code", "date"),
    all.x = TRUE
  )
  %>% mutate(
    country = "Zimbabwe",
    adm1_area = unclass(st_area(geometry)) / 1E6,
    pop_density = pop / adm1_area,
    log_pop_density = log10(pop_density)
  )
)
```


```{r}
df_clim_adm2 <- (
  sf_uga
  %>% merge(
    df_utci,
    by = "id",
    all.x = TRUE
  )
  %>% as.data.frame()
  %>% select(-c(id, adm2_pcode, geometry))
  %>% arrange(adm2_name, date)
  %>% group_by(
    adm1_name, adm2_name, pop
  )
  %>% mutate(
    across(
      -c(date, utci_monthly_max, utci_monthly_min), 
      ~ zoo::rollsum((.), 12, fill = NA, align = "right"),
      .names = "{.col}_sum_12m"
    ),
    across(
      c(utci_monthly_max, utci_monthly_min),
      ~ zoo::rollmax((.), 12, fill = NA, align = "right"),
      .names = "{.col}_max_12m"
    ),
  )
  %>% ungroup()
  %>% mutate(
    month = month(date)
  )
  %>% group_by(adm2_name, month)
  %>% mutate(
    across(
      (contains("utci") & !(contains("12m"))),
      ~ (.) - mean(., na.rm = TRUE),
      .names = "dev_from_adm2_month_mean_{.col}"
    )
  )
  %>% ungroup()
  %>% group_by(adm2_name)
  %>% mutate(
    across(
      contains("utci") & (contains("12m")), 
      ~ (.) - mean(., na.rm = TRUE),
      .names = "dev_from_adm2_mean_{.col}"
    )
  )
  %>% ungroup()
)
```

## Combine UTCI and violence

```{r}
df_sgbv_adm2 <- (
  df_sgbv
  %>% filter(
    variable == "Injuries due to Gender based violence"
  )
  %>% group_by(adm2_name, date)
  %>% summarise(
    cases = sum(cases, na.rm = TRUE),
    .groups = "drop"
  )
)
```

```{r}
df_utci_sgbv_adm2 <- (
  df_utci_adm2
  %>% merge(
    df_sgbv_adm2,
    by = c("adm2_name", "date"),
    all.x = TRUE
  )
  %>% filter(date >= "2020-01-31")
  %>% mutate(
    cases_per_1000 = 1000 * cases / pop,
  )
  %>% group_by(adm2_name)
  %>% mutate(
    dev_from_adm_mean_cases = cases - mean(cases, na.rm = TRUE),
    dev_from_adm_mean_cases_per_1000 = cases_per_1000 - mean(cases_per_1000, na.rm = TRUE),
  )
  %>% ungroup()
)
```

```{r}
df_utci_sgbv_reg <- (
  df_utci_adm2
  %>% merge(
    df_sgbv_adm2,
    by = c("adm2_name", "date"),
    all.x = TRUE
  )
  %>% filter(date >= "2020-01-31")
  %>% mutate(
    region = case_when(
      adm2_name %in% karamoja_region ~"Karamoja",
      adm2_name %in% elgon_region ~ "Elgon",
      adm2_name %in% rwenzori_region ~ "Rwenzori",
      .default = "Rest of the country"
    )
  )
  %>% group_by(region, date)
  %>% summarise(
    pop = sum(pop),
    cases = sum(cases, na.rm = TRUE),
    cases_per_1000 = 1000 * cases / pop,
    across(contains("dev_from_adm2_month"), ~ mean(., na.rm = TRUE)),
    .groups = "drop"
  )
  %>% group_by(region)
  %>% mutate(
    dev_from_reg_mean_cases = cases - mean(cases, na.rm = TRUE),
    dev_from_reg_mean_cases_per_1000 = cases_per_1000 - mean(cases_per_1000, na.rm = TRUE),
  )
)
```

# Viz

## Violence

```{r}
#| fig-width: 6
#| fig-height: 4.5

min_year <- min(year(df_utci_sgbv_reg$date))
max_year <- max(year(df_utci_sgbv_reg$date))
years <- seq(min_year, max_year, by = 1)
num_years <- length(years)
(
  df_utci_sgbv_reg
  %>% pivot_longer(
    cols = c(dev_from_reg_mean_cases_per_1000),
    names_to = "variable",
    values_to = "value"
  )
  %>% filter(
    region != "Rwenzori"
  )
  %>% mutate(
    region = fct_relevel(
      region,
      "Karamoja",
      "Elgon",
      "Rest of the country"
    )
  )
  # %>% ggplot(
  #   aes(x = date, y = value, color = region, linetype = region)
  # )
  # + geom_line(linewidth = 1.2)
  # + scale_color_manual(
  #   values = c(
  #     "Karamoja" = col_palette[1],
  #     "Elgon" = col_palette[5],
  #     "Rest of the country" = "black"
  #   )
  # )
  # + scale_linetype_manual(
  #   values = c("Karamoja" = "solid", "Elgon" = "solid", "Rest of the country" = "dashed"),
  # )
  # + scale_x_date(
  #   breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
  #   labels = c(rep(c(""), num_years + 1), rep(years)),
  #   limits = c(as.Date(paste0(years[1], "-01-01")), as.Date("2025-01-01"))
  # )
  # + theme(
  #   panel.background = element_blank(),
  #   panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
  #   panel.grid = element_blank(),
  #   plot.background = element_blank(),
  #   plot.title = element_text(size = 15, hjust = 0.5),
  #   plot.subtitle = element_text(size = 13, hjust = 0.5),
  #   axis.title = element_text(size = 16),
  #   axis.text.y = element_text(size = 15),
  #   axis.text.x = element_text(size = 15),
  #   legend.background = element_blank(),
  #   legend.title = element_text(size = 14),
  #   legend.text = element_text(size = 13),
  #   legend.position = "top",
  #   legend.direction = "vertical",
  #   legend.justification = "left",
  #   # legend.position = "inside",
  #   # legend.position.inside = c(0.8, 0.72),
  #   axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  # )
  # + labs(
  #   x = "Date",
  #   y = "Cases per 1000 people\n(dev. from regional mean)",
  #   color = "Region"
  # )
  # + guides(
  #   linetype = "none"
  # )
)

ggsave(file.path(fig_dir, "viol_cases_over_time_by_region.png"))
```

```{r}
df_utci_sgbv_adm2_month_summary <- (
  df_utci_sgbv_adm2
  %>% group_by(month)
  %>% summarise(
    utci_month = mean(utci_days_above_32_daily_max, na.rm = TRUE),
    mean_cases = mean(cases, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    dev_cases = mean_cases - mean(mean_cases, na.rm = TRUE),
    dev_utci_month = utci_month - mean(utci_month)
  )
)
```

```{r}
(
  df_utci_sgbv_adm2_month_summary
  %>% pivot_longer(
    cols = c("dev_cases", "dev_utci_month"),
    names_to = "variable",
    values_to = "value"
  )
  %>% mutate(
    month = as.factor(month)
  )
  %>% ggplot(aes(x = month, y = value, fill = variable))
  + geom_col()
  + facet_wrap(
    ~ variable,
    ncol = 1,
    scales = "free_y",
    labeller = as_labeller(c(
      "dev_cases" = "Dev. cases from mean",
      "dev_utci_month" = "Dev. UTCI from mean"
    ))
  )
  + scale_fill_manual(
    values = c(
      "dev_cases" = col_palette[4],
      "dev_utci_month" = col_palette[6]
    )
  )
  + scale_x_discrete(
    breaks = seq(1, 12),
    labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title.x = element_text(size = 16),
    axis.title.y = element_blank(),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.8, 0.72),
  )
  + guides(
    fill = "none"
  )
  + labs(
    x = "Month",
  )
)

ggsave(
  file.path(fig_dir, "viol_utci_month_comp.png")
)
```

```{r}
cor(df_utci_sgbv_adm2_month_summary$dev_cases, df_utci_sgbv_adm2_month_summary$dev_utci_month)
```

## Disaster locations

```{r}
#| fig-width: 5
#| fig-height: 5
(
  sf_uga
  %>% mutate(
    region = case_when(
      adm2_name %in% karamoja_region ~"Karamoja",
      adm2_name %in% elgon_region ~ "Elgon",
      adm2_name %in% rwenzori_region ~ "Rwenzori"
    )
  )
  %>% mutate(
    region = fct_relevel(region, "Karamoja", "Elgon", "Rwenzori")
  )
  %>% ggplot()
  + geom_sf(
    aes(fill = region)
  )
  + scale_fill_manual(
    values = c("Karamoja" = "#dc7726", "Elgon" = "#ecc046", "Rwenzori" = "#35A29F"),
    na.value = "whitesmoke",
    na.translate = FALSE
  )
  + theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.title = element_blank(),
    legend.text = element_text(size = 16),
    legend.background = element_blank(),
    legend.position = "right",
    legend.direction = "vertical",
    legend.key.spacing.y = unit(3, "pt"),
    legend.margin = margin(0, 0, 0, 0, "pt")
  )
)
ggsave(
  path = fig_dir,
  filename = "affected_locations.png",
  width = 6,
  height = 5,
  unit = "in"
)
```

## Market locations

```{r}
#| fig-width: 5
#| fig-height: 5

sf_food_prices_wfp_uga_maize_latest <- (
  sf_food_prices_wfp_uga
  %>% filter(
    commodity == "Maize (white)",
    # date == max(date, na.rm = TRUE)  # filter for latest date
  )
  %>% group_by(market)
  %>% slice(which.max(date))  # select latest date for each market
)
(
  ggplot()
  + geom_sf(data = sf_uga, color = "black")
  + geom_sf(
    data = sf_food_prices_wfp_uga_maize_latest,
    # mapping = aes(size = usdprice),
    color = "dodgerblue",
    size = 5,
  )
  + scale_fill_manual(
    values = c("TRUE" = "salmon", "FALSE" = "whitesmoke")
  )
  + theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.key.height = unit(50, "pt"),
    legend.title = element_text(size = 16, angle = 90, hjust = 0.5),
    legend.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.title.position = "right",
  )
  + labs(
    title = "WFP Market Price Data - Uganda",
    subtitle = "Market location"
  )
)
ggsave(
  filename = file.path(fig_dir, "uga_map_market_location.png"),
  height = 5,
  width = 5,
  unit = "in"
)
```

# Disaster Analysis

## General

### EM-DAT data

```{r}
#| fig-width: 6
#| fig-height: 3.5

min_year <- 2020
max_year <- max(year(df_dis_emdat_month$date))
years <- seq(min_year, max_year, by = 1)
num_years <- length(years)

(
  df_dis_emdat_month
  %>% filter(date >= "2020-01-01")
  %>% mutate(
    region = case_when(
      location %in% karamoja_region ~"Karamoja",
      location %in% elgon_region ~ "Elgon",
      location %in% rwenzori_region ~ "Rwenzori",
      .default = "Rest of the country"
    )
  )
  %>% mutate(
    region = fct_relevel(region, "Karamoja", "Elgon", "Rwenzori", "Rest of the country"),
  )
  %>% group_by(date, region)
  %>% summarise(
    total_affected = sum(total_affected, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ggplot()
  + geom_line(
    aes(x = date, y = total_affected, color = region, linetype = region),
    linewidth = 1.2
  )
  + scale_color_manual(
    values = c("Karamoja" = "#dc7726", "Elgon" = "#ecc046", "Rwenzori" = "#35A29F", "Rest of the country" = "gray"),
    na.value = "whitesmoke",
    na.translate = FALSE
  )
  + scale_linetype_manual(
    values = c("Karamoja" = "solid", "Elgon" = "solid", "Rwenzori" = "solid", "Rest of the country" = "dashed"),
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    legend.position = "inside",
    legend.position.inside = c(0.8, 0.72),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
    breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
    labels = c(rep(c(""), num_years + 1), rep(years)),
    limits = c(as.Date(paste0(years[1], "-01-01")), as.Date("2025-01-01"))
  )
  + labs(
    x = "Date",
    y = "Count",
    color = "Region",
    title = "Monthly number of people affected by disasters"
  )
  + guides(
    linetype = "none"
  )
)

ggsave(
  path = fig_dir,
  filename = "disaster_numbers.png"
)
```

```{r}
(
  df_dis_emdat
  %>% filter(
    start_date >= "2022-01-01",
    end_date <= "2023-07-01"
  )
  %>% arrange(desc(total_deaths))
)
```

### IOM Data

```{r}
df_dis_iom$disaster %>% unique()
```

#### People affected per disaster and riks cluster

```{r}
df_dis_iom_by_disaster_and_priority <- (
  df_dis_iom
  %>% mutate(
    region = case_when(
      district %in% karamoja_region ~"Karamoja",
      district %in% elgon_region ~ "Elgon",
      district %in% rwenzori_region ~ "Rwenzori",
      district %in% teso_region ~ "Teso",
      district %in% west_nile_region ~ "West Nile",
      .default = "Rest of the country"
    ),
    disaster = case_when(
      disaster %in% c("Floods", "Floods and mudslides", "Landslides and Mudslides") ~ "Floods & Mudslides",
      disaster %in% c("Drought", "Drought and Famine", "drought/famine", "Famine / Food Insecurity") ~ "Droughts",
      disaster %in% c("Heavy Storms/ Hailstorms", "EVD (Heavy Storms/ Hailstorms)", "Heavy winds and rains") ~ "Storms",
      disaster %in% c("Fires", "EVD (Fires)", "Fire", "fire") ~ "Fires"
    ),
    priority = case_when(
      (
        str_detect(child_protection, "medium") 
        | str_detect(child_protection, "high")
        & !(
          str_detect(sgbv, "medium")
          | str_detect(sgbv, "high")
          )
      ) ~ "Child Protection",
      (
        str_detect(sgbv, "medium") 
        | str_detect(sgbv, "high")
        & !(
          str_detect(child_protection, "medium")
          | str_detect(child_protection, "high")
          )
      )~ "SGBV",
      (
        str_detect(sgbv, "medium") 
        | str_detect(sgbv, "high")
        & (
          str_detect(child_protection, "medium")
          | str_detect(child_protection, "high")
          )
      )~ "Child Protection & SGBV",
    ),
    disaster = fct_relevel(disaster, c("Floods & Mudslides", "Storms", "Fires", "Droughts")),
    priority = fct_relevel(priority, c("Child Protection", "SGBV", "Child Protection & SGBV"))
  )
  %>% filter(
    !is.na(priority),
    !is.na(disaster)
  )
)
```

```{r}
#| fig-width: 7
#| fig-height: 3.5

(
  df_dis_iom_by_disaster_and_priority
  %>% group_by(disaster, priority)
  %>% summarise(
    affected = sum(affected, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ggplot()
  + geom_col(
    aes(x = disaster, y = affected, fill = priority),
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    axis.title.y = element_text(size = 16),
    axis.title.x = element_blank(),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    legend.position = "inside",
    legend.position.inside = c(0.72, 0.72),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_fill_manual(
    values = c("Child Protection" = "#0292b7", "SGBV" = "#ecc046", "Child Protection & SGBV" = "#dc7726")
  )
  + labs(
    # y = "none",
    y = "Counts",
    fill = "Response Priority",
    title = "People affected per disaster and risk cluster",
    subtitle = "Medium or high response priority, Jan 2023 - May 2025"
  )
)

ggsave(
  path = fig_dir, 
  filename = "affected_by_disaster_and_risk.png"
)
```

#### People affected per disaster and crop failure

```{r}
df_dis_iom_by_disaster_and_crop <- (
  df_dis_iom
  %>% mutate(
    disaster = case_when(
      disaster %in% c("Floods", "Floods and mudslides", "Landslides and Mudslides") ~ "Floods & Mudslides",
      disaster %in% c("Drought", "Drought and Famine", "drought/famine", "Famine / Food Insecurity") ~ "Droughts",
      disaster %in% c("Heavy Storms/ Hailstorms", "EVD (Heavy Storms/ Hailstorms)", "Heavy winds and rains") ~ "Storms",
      disaster %in% c("Fires", "EVD (Fires)", "Fire", "fire") ~ "Fires"
    ),
    crops_damage = case_when(
      crops_damage %in% c("low", "40 per cent during dry season", "45 percent")  ~ "Low",
      crops_damage %in% c("medium", "mdium", "50 percent its was dry season", "50 percent it was dry season", "60 percent") ~ "Medium",
      crops_damage == "high" ~ "High",
    ),
    disaster = fct_relevel(disaster, c("Floods & Mudslides", "Storms", "Fires", "Droughts")),
    crops_damage = fct_relevel(crops_damage, c("High", "Medium", "Low"))
  )
  %>% filter(
    !is.na(disaster),
    !is.na(crops_damage)
  )
)
```

```{r}
#| fig-width: 7
#| fig-height: 3.5

(
  df_dis_iom_by_disaster_and_crop
  %>% group_by(disaster, crops_damage)
  %>% summarise(
    affected = sum(affected, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ggplot()
  + geom_col(
    aes(x = disaster, y = affected, fill = crops_damage),
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    axis.title.y = element_text(size = 16),
    axis.title.x = element_blank(),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    legend.position = "inside",
    legend.position.inside = c(0.72, 0.72),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_fill_manual(
    values = c("Low" = "#0292b7", "Medium" = "#ecc046", "High" = "#dc7726")
  )
  + labs(
    # y = "none",
    y = "Counts",
    fill = "Crops Damage",
    title = "People affected per disaster and crops damage level",
    subtitle = "Jan 2023 - May 2025"
  )
)

ggsave(
  path = fig_dir, 
  filename = "affected_by_disaster_and_crop.png"
)
```

#### People affected by region

```{r}
df_dis_iom_high_priority <- (
  df_dis_iom
  %>% filter(
    str_detect(child_protection, "medium")
    | str_detect(child_protection, "high")
    | str_detect(sgbv, "medium")
    | str_detect(sgbv, "high")
  )
  %>% mutate(
    region = case_when(
      district %in% karamoja_region ~"Karamoja",
      district %in% elgon_region ~ "Elgon",
      district %in% rwenzori_region ~ "Rwenzori",
      district %in% teso_region ~ "Teso",
      district %in% west_nile_region ~ "West Nile",
      .default = "Rest of the country"
    )
  )
  %>% mutate(
    region = fct_relevel(region, "Karamoja", "Elgon", "Teso", "Rwenzori", "Rest of the country"),
  )
)
```

```{r}
#| fig-width: 6
#| fig-height: 3.5

min_year <- min(year(df_dis_iom_high_priority$date))
max_year <- max(year(df_dis_iom_high_priority$date))
years <- seq(min_year, max_year, by = 1)
num_years <- length(years)


(
  df_dis_iom_high_priority
  %>% select(date, region, affected)
  %>% group_by(region)
  %>% complete(
    date = seq(min(df_dis_iom_high_priority$date), max(df_dis_iom_high_priority$date), by = "1 day"),
    fill = list(affected = 0)
  )
  %>% ungroup()
  %>% group_by(date = cut(date, "month"), region)
  %>% summarise(
    affected = sum(affected, na.rm = TRUE),
    .groups = "drop"
  )
  %>% mutate(
    date = as.Date(date)
  )
  %>% ggplot()
  + geom_line(
    aes(x = date, y = affected, color = region, linetype = region),
    linewidth = 1.2
  )
  + scale_color_manual(
    values = c("Karamoja" = "#dc7726", "Elgon" = "#ecc046", "Teso" = "#0292b7", "Rwenzori" = "#35A29F", "Rest of the country" = "gray"),
    na.value = "whitesmoke",
    na.translate = FALSE
  )
  + scale_linetype_manual(
    values = c("Karamoja" = "solid", "Elgon" = "solid", "Teso" = "solid", "Rwenzori" = "solid", "Rest of the country" = "dashed"),
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    legend.position = "inside",
    legend.position.inside = c(0.8, 0.72),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
    breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
    labels = c(rep(c(""), num_years + 1), rep(years)),
    limits = c(as.Date(paste0(years[1], "-01-01")), as.Date("2025-01-01"))
  )
  + labs(
    x = "Date",
    y = "Count",
    color = "Region",
    title = "Monthly number of people affected by disasters"
  )
  + guides(
    linetype = "none"
  )
)


  
```

```{r}
(
  df_dis_iom_high_priority
  %>% filter(
    date >= "2023-09-01",
    date <= "2024-01-01"
  )
  %>% arrange(desc(affected))
)
```

## Karamoja, 2022 Drought

```{r}
karamoja_2022_distr <- c(
  "Napak", 
  "Kaabong", 
  "Kotido", 
  "Moroto"
)

other_affected_areas <- c(
  "Mbale",
  "Kapchorwa",
  "Sironko",
  "Bulambuli",
  "Bukedea",
  "Bududa",
  "Butaleja"
)
  
```

### Prices

```{r}
sf_food_prices_wfp_karamoja_2022 <- (
  sf_food_prices_wfp_uga
  %>% as.data.frame()
  %>% filter(
    commodity %in% commodities_to_plot,
    !(admin2 %in% other_affected_areas),
    date >= "2021-01-01",
    date <= "2024-12-31"
  )
  %>% mutate(
    karamoja_region = if_else(
      admin2 %in% karamoja_2022_distr,
      "Affected Areas",
      "Rest of the country"
    )
  )
  %>% mutate(
    karamoja_region = factor(karamoja_region, c("Rest of the country", "Affected Areas"))
  )
  %>% group_by(date, karamoja_region, commodity)
  %>% summarise(
    price_q07 = quantile(price, probs = 0.07, na.rm = TRUE),
    price_q50 = quantile(price, probs = 0.5, na.rm = TRUE),
    price_q93 = quantile(price, probs = 0.93, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ungroup()
)

```

```{r}
#| fig-width: 8
#| fig-height: 4

min_year <- min(year(sf_food_prices_wfp_karamoja_2022$date))
max_year <- max(year(sf_food_prices_wfp_karamoja_2022$date))
years <- seq(min_year, max_year, by = 1)
num_years <- length(years)

(
  sf_food_prices_wfp_karamoja_2022
  %>% filter(commodity == "Maize (white)")
  %>% ggplot()
  + geom_rect(
    data = sf_food_prices_wfp_karamoja_2022[1,],
    xmin = as.Date("2022-07-01"),
    xmax = as.Date("2022-12-01"),
    ymin = -Inf, 
    ymax = Inf,
    fill = "gray",
    alpha = 0.4
  )
  + geom_ribbon(
    aes(x = date, ymin = price_q07, ymax = price_q93, fill = karamoja_region),
    alpha = 0.5
  )
  + geom_point(
    aes(x = date, y = price_q50, color = karamoja_region, alpha = karamoja_region),
    size = 2
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.83),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
      breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
      labels = c(rep(c(""), num_years + 1), rep(years)),
      limits = c(as.Date(paste0(years[1], "-01-01")), as.Date("2025-01-01"))
  )
  + scale_color_manual(
    values = c("Affected Areas" = "#dc7726", "Rest of the country" = "#97DECE"),
  )
  + scale_fill_manual(
    values = c("Affected Areas" = "#dc7726", "Rest of the country" = "#97DECE"),
  )
  + scale_alpha_manual(
    values = c("Affected Areas" = 1, "Rest of the country" = 0.8)
  )
  + labs(
    fill = "Location",
    x = "Date",
    y = "Price (UGX/Kg)",
    subtitle = "Median and 87% interval coverage",
    title = "Maize price comparison"
  )
  + guides(
    color = "none",
    alpha = "none",
    fill = guide_legend(override.aes = list(alpha = 1))
  )
)
ggsave(
  filename = file.path(fig_dir, "uganda_maize_price_karamoja_2022.png")
)
```

```{r}
#| fig-width: 8
#| fig-height: 8

(
  sf_food_prices_wfp_karamoja_2022
  %>% ggplot()
  + geom_rect(
    data = sf_food_prices_wfp_karamoja_2022 %>% distinct(commodity),
    xmin = as.Date("2022-07-01"),
    xmax = as.Date("2022-12-01"),
    ymin = -Inf, 
    ymax = Inf,
    fill = "gray",
    alpha = 0.4
  )
  + geom_ribbon(
    aes(x = date, ymin = price_q07, ymax = price_q93, fill = karamoja_region),
    alpha = 0.5
  )
  + geom_point(
    aes(x = date, y = price_q50, color = karamoja_region, alpha = karamoja_region),
    size = 2
  )
  + facet_wrap(
    ~commodity,
    ncol = 1,
    # strip.position = "right",
    scales = "free_y"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = 15, face = "bold"),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.83),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
      breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
      labels = c(rep(c(""), num_years + 1), rep(years)),
      limits = c(as.Date(paste0(years[1], "-01-01")), as.Date("2025-01-01"))
  )
  + scale_color_manual(
    values = c("Affected Areas" = "#dc7726", "Rest of the country" = "#97DECE"),
  )
  + scale_fill_manual(
    values = c("Affected Areas" = "#dc7726", "Rest of the country" = "#97DECE"),
  )
  + scale_alpha_manual(
    values = c("Affected Areas" = 1, "Rest of the country" = 0.8)
  )
  + labs(
    fill = "Location",
    x = "Date",
    y = "Price (UGX/Kg)",
    subtitle = "Median and 87% interval coverage",
    title = "Maize price comparison"
  )
  + guides(
    color = "none",
    alpha = "none",
    fill = guide_legend(override.aes = list(alpha = 1))
  )
)
ggsave(
  filename = file.path(fig_dir, "uganda_all_prices_karamoja_2022.png")
)
```

### Price differential

```{r}
sf_food_price_diff_wfp_karamoja_2022 <- (
  sf_food_prices_wfp_karamoja_2022
  %>% select(
    date,
    karamoja_region,
    commodity,
    price_q50
  )
  %>% pivot_wider(
    id_cols = c(date, commodity),
    names_from = karamoja_region,
    values_from = price_q50
  )
  %>% mutate(
    price_diff = `Affected Areas` - `Rest of the country`
  )
  %>% filter(!is.na(price_diff))
  %>% select(date, commodity, price_diff)
)
```

### Violence VS Drought

```{r}
#| fig-width: 8
#| fig-height: 4

(
  sf_food_price_diff_wfp_karamoja_2022
  %>% filter(commodity == "Maize (white)")
  %>% ggplot()
  + geom_hline(
    yintercept = 0,
    linetype = "dashed",
    color = "black"
  )
  + geom_rect(
    data = sf_food_price_diff_wfp_karamoja_2022[1,],
    xmin = as.Date("2022-07-01"),
    xmax = as.Date("2022-12-01"),
    ymin = -Inf, 
    ymax = Inf,
    fill = "gray",
    alpha = 0.4
  )
  # + geom_point(
  #   aes(x = date, y = price_diff),
  #   size = 2,
  #   color = "#0292b7"
  # )
  + geom_line(
    data = df_utci_sgbv_reg  %>% filter(region == "Karamoja"),
    aes(x = date, y = dev_from_reg_mean_cases_per_1000),
    color = "#dc7726",
    linewidth = 1.3
  )
  # + scale_y_continuous(
  #   "Price diff. (UGX/Kg)",
  #   sec.axis = sec_axis(~ ./10000, name = "Cases per 1000 people (dev.)")
  # )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.83),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
      breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
      labels = c(rep(c(""), num_years + 1), rep(years)),
      limits = c(as.Date(paste0(years[1], "-01-01")), as.Date("2025-01-01"))
  )
  + labs(
    fill = "Location",
    x = "Date",
    y = "Price diff. (UGX/Kg)",
    title = "Maize price comparison"
  )
  + guides(
    color = "none",
    alpha = "none",
  )
)
ggsave(
  filename = file.path(fig_dir, "uganda_maize_price_diff_karamoja_2022.png")
)
```

```{r}
#| fig-width: 5.5
#| fig-height: 10

(
  sf_food_price_diff_wfp_karamoja_2022
  %>% ggplot()
  + geom_hline(
    yintercept = 0,
    linetype = "dashed",
    color = "black"
  )
  + geom_rect(
    data = sf_food_price_diff_wfp_karamoja_2022 %>% distinct(commodity),
    xmin = as.Date("2022-07-01"),
    xmax = as.Date("2022-12-01"),
    ymin = -Inf, 
    ymax = Inf,
    fill = "gray",
    alpha = 0.4
  )
  + geom_point(
    aes(x = date, y = price_diff),
    size = 2,
    color = "#0292b7"
  )
  + facet_wrap(
    ~commodity,
    ncol = 1,
    # strip.position = "right",
    scales = "free_y"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = 15, face = "bold"),
    plot.background = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.83),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
      breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
      labels = c(rep(c(""), num_years + 1), rep(years)),
      limits = c(as.Date(paste0(years[1], "-01-01")), as.Date("2025-01-01"))
  )
  + labs(
    fill = "Location",
    x = "Date",
    y = "\u0394Price (UGX/Kg)",
    title = "Price differential",
    subtitle = "Karamoja vs rest of the country (mean)"
  )
  + guides(
    color = "none",
    alpha = "none",
  )
)
ggsave(
  filename = file.path(fig_dir, "uganda_all_price_diff_karamoja_2022.png")
)
```

## Elgon

```{r}
affected_areas <- c(
  "Mbale",
  "Kapchorwa",
  "Sironko",
  "Bulambuli",
  "Bukedea",
  "Bududa",
  "Butaleja"
)
```

```{r}
df_dis_emdat_elgon <- (
  df_dis_emdat_month
  %>% mutate(
    affected = if_else(
      location %in% affected_areas,
      TRUE,
      FALSE
    ),
    # date = as.Date(month)
  )
  %>% group_by(date, affected)
  %>% summarise(
    total_affected = sum(total_affected),
    .groups = "drop"
  )
)
```

### Prices

```{r}
sf_food_prices_wfp_elgon <- (
  sf_food_prices_wfp_uga
  %>% as.data.frame()
  %>% filter(
    commodity %in% commodities_to_plot,
    date >= "2018-01-01"
  )
  %>% mutate(
    affected_areas = if_else(
      admin2 %in% affected_areas,
      "Affected Areas",
      "Rest of the country"
    )
  )
  %>% mutate(
    affected_areas = factor(affected_areas, c("Rest of the country", "Affected Areas"))
  )
  %>% group_by(date, affected_areas, commodity)
  %>% summarise(
    price_q07 = quantile(price, probs = 0.07, na.rm = TRUE),
    price_q50 = quantile(price, probs = 0.5, na.rm = TRUE),
    price_q93 = quantile(price, probs = 0.93, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ungroup()
)
min_year <- min(year(sf_food_prices_wfp_elgon$date))
max_year <- max(year(sf_food_prices_wfp_elgon$date))
years <- seq(min_year, max_year, by = 1)
num_years <- length(years)
```

```{r}
#| fig-width: 8
#| fig-height: 4

coeff <- 30


(
  sf_food_prices_wfp_elgon
  %>% filter(commodity == "Maize (white)")
  %>% ggplot()
  + geom_col(
    data = df_dis_emdat_elgon %>% filter(affected),
    mapping = aes(x = date, y = total_affected / coeff),
    width = 20
  )
  + geom_ribbon(
    aes(x = date, ymin = price_q07, ymax = price_q93, fill = affected_areas),
    alpha = 0.5
  )
  + geom_point(
    aes(x = date, y = price_q50, color = affected_areas, alpha = affected_areas),
    size = 2
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    legend.position = "inside",
    legend.position.inside = c(0.25, 0.83),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
      breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
      labels = c(rep(c(""), num_years + 1), rep(years)),
      limits = c(as.Date(paste0(years[1], "-01-01")), as.Date("2025-01-01"))
  )
  + scale_y_continuous(
    "Price (UGX/Kg)",
    sec.axis = sec_axis(transform = ~.*coeff, name = "# Affected")
  )
  + scale_color_manual(
    values = c("Affected Areas" = "#dc7726", "Rest of the country" = "#97DECE"),
  )
  + scale_fill_manual(
    values = c("Affected Areas" = "#dc7726", "Rest of the country" = "#97DECE"),
  )
  + scale_alpha_manual(
    values = c("Affected Areas" = 1, "Rest of the country" = 0.8)
  )
  + labs(
    fill = "Location",
    x = "Date",
    y = "Price (UGX/Kg)",
    subtitle = "Median and 87% interval coverage",
    title = "Maize price comparison"
  )
  + guides(
    color = "none",
    alpha = "none",
    fill = guide_legend(override.aes = list(alpha = 1))
  )
)
ggsave(
  filename = file.path(fig_dir, "uganda_maize_price_elgon_2022.png")
)
```

## Elgon, 2022 Flood

```{r}
other_affected_areas <- c(
  "Napak", 
  "Kaabong", 
  "Kotido", 
  "Moroto"
)

elgon_2022_distr <- c(
  "Mbale",
  "Kapchorwa",
  "Sironko",
  "Bulambuli",
  "Bukedea",
  "Bududa",
  "Butaleja"
)
  
```

### Prices

```{r}
sf_food_prices_wfp_elgon_2022 <- (
  sf_food_prices_wfp_uga
  %>% as.data.frame()
  %>% filter(
    commodity %in% commodities_to_plot,
    !(admin2 %in% other_affected_areas),
    date >= "2021-01-01",
    date <= "2024-12-31"
  )
  %>% mutate(
    elgon_region = if_else(
      admin2 %in% elgon_2022_distr,
      "Affected Areas",
      "Rest of the country"
    )
  )
  %>% mutate(
    elgon_region = factor(elgon_region, c("Rest of the country", "Affected Areas"))
  )
  %>% group_by(date, elgon_region, commodity)
  %>% summarise(
    price_q07 = quantile(price, probs = 0.07, na.rm = TRUE),
    price_q50 = quantile(price, probs = 0.5, na.rm = TRUE),
    price_q93 = quantile(price, probs = 0.93, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ungroup()
)
min_year <- min(year(sf_food_prices_wfp_elgon_2022$date))
max_year <- max(year(sf_food_prices_wfp_elgon_2022$date))
years <- seq(min_year, max_year, by = 1)
num_years <- length(years)
```

```{r}
#| fig-width: 8
#| fig-height: 4

(
  sf_food_prices_wfp_elgon_2022
  %>% filter(commodity == "Maize (white)")
  %>% ggplot()
  + geom_rect(
    data = sf_food_prices_wfp_elgon_2022[1,],
    xmin = as.Date("2022-07-30"),
    xmax = as.Date("2022-08-05"),
    ymin = -Inf, 
    ymax = Inf,
    fill = "gray",
    alpha = 0.4
  )
  + geom_ribbon(
    aes(x = date, ymin = price_q07, ymax = price_q93, fill = elgon_region),
    alpha = 0.5
  )
  + geom_point(
    aes(x = date, y = price_q50, color = elgon_region, alpha = elgon_region),
    size = 2
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.83),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
      breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
      labels = c(rep(c(""), num_years + 1), rep(years)),
      limits = c(as.Date(paste0(years[1], "-01-01")), as.Date("2025-01-01"))
  )
  + scale_color_manual(
    values = c("Affected Areas" = "#dc7726", "Rest of the country" = "#97DECE"),
  )
  + scale_fill_manual(
    values = c("Affected Areas" = "#dc7726", "Rest of the country" = "#97DECE"),
  )
  + scale_alpha_manual(
    values = c("Affected Areas" = 1, "Rest of the country" = 0.8)
  )
  + labs(
    fill = "Location",
    x = "Date",
    y = "Price (UGX/Kg)",
    subtitle = "Median and 87% interval coverage",
    title = "Maize price comparison"
  )
  + guides(
    color = "none",
    alpha = "none",
    fill = guide_legend(override.aes = list(alpha = 1))
  )
)
ggsave(
  filename = file.path(fig_dir, "uganda_maize_price_elgon_2022.png")
)
```

```{r}
#| fig-width: 8
#| fig-height: 8

(
  sf_food_prices_wfp_elgon_2022
  %>% ggplot()
  + geom_rect(
    data = sf_food_prices_wfp_elgon_2022 %>% distinct(commodity),
    xmin = as.Date("2022-07-30"),
    xmax = as.Date("2022-08-05"),
    ymin = -Inf, 
    ymax = Inf,
    fill = "gray",
    alpha = 0.4
  )
  + geom_ribbon(
    aes(x = date, ymin = price_q07, ymax = price_q93, fill = elgon_region),
    alpha = 0.5
  )
  + geom_point(
    aes(x = date, y = price_q50, color = elgon_region, alpha = elgon_region),
    size = 2
  )
  + facet_wrap(
    ~commodity,
    ncol = 1,
    # strip.position = "right",
    scales = "free_y"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = 15, face = "bold"),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.83),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
      breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
      labels = c(rep(c(""), num_years + 1), rep(years)),
      limits = c(as.Date(paste0(years[1], "-01-01")), as.Date("2025-01-01"))
  )
  + scale_color_manual(
    values = c("Affected Areas" = "#dc7726", "Rest of the country" = "#97DECE"),
  )
  + scale_fill_manual(
    values = c("Affected Areas" = "#dc7726", "Rest of the country" = "#97DECE"),
  )
  + scale_alpha_manual(
    values = c("Affected Areas" = 1, "Rest of the country" = 0.8)
  )
  + labs(
    fill = "Location",
    x = "Date",
    y = "Price (UGX/Kg)",
    subtitle = "Median and 87% interval coverage",
    title = "Maize price comparison"
  )
  + guides(
    color = "none",
    alpha = "none",
    fill = guide_legend(override.aes = list(alpha = 1))
  )
)
ggsave(
  filename = file.path(fig_dir, "uganda_all_prices_elgon_2022.png")
)
```

### Price differential

```{r}
sf_food_price_diff_wfp_elgon_2022 <- (
  sf_food_prices_wfp_elgon_2022
  %>% select(
    date,
    elgon_region,
    commodity,
    price_q50
  )
  %>% pivot_wider(
    id_cols = c(date, commodity),
    names_from = elgon_region,
    values_from = price_q50
  )
  %>% mutate(
    price_diff = `Affected Areas` - `Rest of the country`
  )
  %>% filter(!is.na(price_diff))
  %>% select(date, commodity, price_diff)
)
```

```{r}
#| fig-width: 8
#| fig-height: 4

(
  sf_food_price_diff_wfp_elgon_2022
  %>% filter(commodity == "Maize (white)")
  %>% ggplot()
  + geom_hline(
    yintercept = 0,
    linetype = "dashed",
    color = "black"
  )
  + geom_rect(
    data = sf_food_price_diff_wfp_elgon_2022[1,],
    xmin = as.Date("2022-07-30"),
    xmax = as.Date("2022-08-05"),
    ymin = -Inf, 
    ymax = Inf,
    fill = "gray",
    alpha = 0.4
  )
  + geom_point(
    aes(x = date, y = price_diff),
    size = 2,
    color = "cornflowerblue"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.83),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
      breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
      labels = c(rep(c(""), num_years + 1), rep(years)),
      limits = c(as.Date(paste0(years[1], "-01-01")), as.Date("2025-01-01"))
  )
  + labs(
    fill = "Location",
    x = "Date",
    y = "Price diff. (UGX/Kg)",
    title = "Maize price comparison"
  )
  + guides(
    color = "none",
    alpha = "none",
  )
)
ggsave(
  filename = file.path(fig_dir, "uganda_maize_price_diff_elgon_2022.png")
)
```

```{r}
#| fig-width: 8
#| fig-height: 8

(
  sf_food_price_diff_wfp_elgon_2022
  %>% ggplot()
  + geom_hline(
    yintercept = 0,
    linetype = "dashed",
    color = "black"
  )
  + geom_rect(
    data = sf_food_price_diff_wfp_elgon_2022 %>% distinct(commodity),
    xmin = as.Date("2022-07-30"),
    xmax = as.Date("2022-08-05"),
    ymin = -Inf, 
    ymax = Inf,
    fill = "gray",
    alpha = 0.4
  )
  + geom_point(
    aes(x = date, y = price_diff),
    size = 2,
    color = "cornflowerblue"
  )
  + facet_wrap(
    ~commodity,
    ncol = 1,
    # strip.position = "right",
    scales = "free_y"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = 15, face = "bold"),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.83),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
      breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
      labels = c(rep(c(""), num_years + 1), rep(years)),
      limits = c(as.Date(paste0(years[1], "-01-01")), as.Date("2025-01-01"))
  )
  + labs(
    fill = "Location",
    x = "Date",
    y = "Price diff. (UGX/Kg)",
    title = "Price diff. comparison"
  )
  + guides(
    color = "none",
    alpha = "none",
  )
)
ggsave(
  filename = file.path(fig_dir, "uganda_all_price_diff_elgon_2022.png")
)
```

### Violence VS Flood

```{r}
df_dis_iom_mbale <- (
  df_dis_iom 
  # %>% filter(district == "Mbale")
  %>% filter(district %in% elgon_region)
  %>% merge(
    df_utci_adm2 %>% select(district = adm2_name, pop),
    by = "district",
    all.x = TRUE
  )
  %>% group_by(date)
  %>% summarise(
    affected = sum(affected, na.rm = TRUE),
    pop = sum(pop, na.rm = TRUE)
  )
  %>% mutate(
    affected_per_1000 = 1000 * affected / pop
  )
)

df_utci_sgbv_mbale <- (
  df_utci_sgbv_adm2
  %>% filter(
    # region == "Elgon", 
    # adm2_name != "Butaleja"
    adm2_name == "Mbale"
  )
  %>% mutate(
    cases_per_1000 = 1000 * cases / pop
  )
  %>% select(date, adm2_name, cases_per_1000)
)
  
```

```{r}
(
  ggplot()
  + geom_point(
    data = df_dis_iom_mbale,
    aes(x = date, y = affected_per_1000),
    size = 3,
    color = "#dc7726"
  )
  + geom_line(
    data = df_utci_sgbv_reg %>% filter(region == "Elgon"),
    aes(x = date, y = dev_from_reg_mean_cases_per_1000 * 1000),
    color = "#0292b7"
  )
  + theme(
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  # + scale_x_date(
  #     breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
  #     labels = c(rep(c(""), num_years + 1), rep(years)),
  #     limits = c(as.Date(paste0(years[1], "-01-01")), as.Date("2025-01-01"))
  # )
  + scale_y_continuous(
    "Cases per 1000",
    sec.axis = sec_axis(~ ./1000, "# Affected People")
  )
)
```

```{r}
#| fig-width: 8
#| fig-height: 4

data <- df_utci_sgbv_reg %>% filter(region == "Elgon")

(
  ggplot()
  + geom_hline(
    yintercept = 0,
    linetype = "dashed",
    color = "black"
  )
  + geom_rect(
    data = data[1,],
    xmin = as.Date("2022-07-30"),
    xmax = as.Date("2022-08-05"),
    ymin = -Inf, 
    ymax = Inf,
    fill = "darkgray",
    alpha = 1
  )
  + geom_line(
    data = df_utci_sgbv_reg,
    aes(x = date, y = dev_from_reg_mean_cases_per_1000),
    color = "#dc7726",
    linewidth = 1.3
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.83),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
      breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
      labels = c(rep(c(""), num_years + 1), rep(years)),
      limits = c(as.Date(paste0(years[1], "-01-01")), as.Date("2025-01-01"))
  )
  + labs(
    x = "Date",
    y = "STIs from SGBV (dev. cases per 1000)",
    # title = "Maize price comparison"
  )
  + guides(
    color = "none",
    alpha = "none",
  )
)
# ggsave(
#   filename = file.path(fig_dir, "uganda_maize_price_diff_karamoja_2022.png")
# )
```

## Elgon, June 2019 Flood

```{r}
# elgon_2019_distr <- c(
#   "Bududa",
#   # "Bushenyi",
#   # "Rukiga",
#   "Sironko",
#   "Mbale",
#   "Butaleja"
# )

affected_distr <- c(
  "Bududa", 
  "Bukwo",
  "Bulambuli",
  "Kapchorwa",
  "Kween",
  "Manafwa", 
  "Mbale", 
  "Namisindwa",
  "Sironko", 
  "Tororo"
)

other_affected_distr <- c(
  "Bushenyi",
  "Rukiga"
)
  
```

```{r}
(
  ggplot(sf_uga %>% mutate(affected = if_else(adm2_name %in% affected_distr, TRUE, FALSE)))
  + geom_sf(
    aes(fill = affected)
  )
  + geom_sf(
    data = sf_food_prices_wfp_uga_maize_latest,
    color = "dodgerblue",
    size = 5,
  )
  + scale_fill_manual(
    values = c("TRUE" = "salmon", "FALSE" = "whitesmoke")
  )
  + theme_classic()
)
```

### Prices

```{r}
sf_food_prices_wfp_elgon_2019 <- (
  sf_food_prices_wfp_uga
  %>% as.data.frame()
  %>% filter(
    commodity %in% commodities_to_plot,
    date >= "2018-01-01",
    date <= "2021-12-31"
  )
  %>% mutate(
    affected = if_else(
      admin2 %in% affected_distr,
      "Affected Areas",
      "Rest of the country"
    )
  )
  %>% mutate(
    affected = factor(affected, c("Rest of the country", "Affected Areas"))
  )
  %>% group_by(date, affected, commodity)
  %>% summarise(
    price_q07 = quantile(price, probs = 0.07, na.rm = TRUE),
    price_q50 = quantile(price, probs = 0.5, na.rm = TRUE),
    price_q93 = quantile(price, probs = 0.93, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ungroup()
)
min_year <- min(year(sf_food_prices_wfp_elgon_2019$date))
max_year <- max(year(sf_food_prices_wfp_elgon_2019$date))
years <- seq(min_year, max_year, by = 1)
num_years <- length(years)
```

```{r}
#| fig-width: 8
#| fig-height: 4

(
  sf_food_prices_wfp_elgon_2019
  %>% filter(commodity == "Maize (white)")
  %>% ggplot()
  + geom_rect(
    data = sf_food_prices_wfp_elgon_2019[1,],
    xmin = as.Date("2019-06-04"),
    xmax = as.Date("2019-06-08"),
    ymin = -Inf, 
    ymax = Inf,
    fill = "gray",
    alpha = 0.4
  )
  + geom_ribbon(
    aes(x = date, ymin = price_q07, ymax = price_q93, fill = affected),
    alpha = 0.5
  )
  + geom_point(
    aes(x = date, y = price_q50, color = affected, alpha = affected),
    size = 2
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
      breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
      labels = c(rep(c(""), num_years + 1), rep(years)),
      limits = c(as.Date(paste0(years[1], "-01-01")), as.Date(paste0(years[num_years], "-01-01")))
  )
  + scale_color_manual(
    values = c("Affected Areas" = "#dc7726", "Rest of the country" = "#97DECE"),
  )
  + scale_fill_manual(
    values = c("Affected Areas" = "#dc7726", "Rest of the country" = "#97DECE"),
  )
  + scale_alpha_manual(
    values = c("Affected Areas" = 1, "Rest of the country" = 0.8)
  )
  + labs(
    fill = "Location",
    x = "Date",
    y = "Price (UGX/Kg)",
    subtitle = "Median and 87% interval coverage",
    title = "Maize price comparison"
  )
  + guides(
    color = "none",
    alpha = "none",
    fill = guide_legend(override.aes = list(alpha = 1))
  )
)
ggsave(
  filename = file.path(fig_dir, "uganda_maize_price_elgon_2019.png")
)
```

```{r}
#| fig-width: 8
#| fig-height: 8

(
  sf_food_prices_wfp_elgon_2019
  %>% ggplot()
  + geom_rect(
    data = sf_food_prices_wfp_elgon_2019 %>% distinct(commodity),
    xmin = as.Date("2019-06-04"),
    xmax = as.Date("2019-06-08"),
    ymin = -Inf, 
    ymax = Inf,
    fill = "gray",
    alpha = 0.4
  )
  + geom_ribbon(
    aes(x = date, ymin = price_q07, ymax = price_q93, fill = affected),
    alpha = 0.5
  )
  + geom_point(
    aes(x = date, y = price_q50, color = affected, alpha = affected),
    size = 2
  )
  + facet_wrap(
    ~commodity,
    ncol = 1,
    # strip.position = "right",
    scales = "free_y"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = 15, face = "bold"),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.83),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
      breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
      labels = c(rep(c(""), num_years + 1), rep(years)),
      limits = c(as.Date(paste0(years[1], "-01-01")), as.Date(paste0(years[num_years], "-01-01")))
  )
  + scale_color_manual(
    values = c("Affected Areas" = "#dc7726", "Rest of the country" = "#97DECE"),
  )
  + scale_fill_manual(
    values = c("Affected Areas" = "#dc7726", "Rest of the country" = "#97DECE"),
  )
  + scale_alpha_manual(
    values = c("Affected Areas" = 1, "Rest of the country" = 0.8)
  )
  + labs(
    fill = "Location",
    x = "Date",
    y = "Price (UGX/Kg)",
    subtitle = "Median and 87% interval coverage",
    title = "Maize price comparison"
  )
  + guides(
    color = "none",
    alpha = "none",
    fill = guide_legend(override.aes = list(alpha = 1))
  )
)
ggsave(
  filename = file.path(fig_dir, "uganda_all_prices_elgon_2019.png")
)
```

### Price differential

```{r}
sf_food_price_diff_wfp_elgon_2019 <- (
  sf_food_prices_wfp_elgon_2019
  %>% select(
    date,
    affected,
    commodity,
    price_q50
  )
  %>% pivot_wider(
    id_cols = c(date, commodity),
    names_from = affected,
    values_from = price_q50
  )
  %>% mutate(
    price_diff = `Affected Areas` - `Rest of the country`
  )
  %>% filter(!is.na(price_diff))
  %>% select(date, commodity, price_diff)
)
```

```{r}
#| fig-width: 8
#| fig-height: 4

(
  sf_food_price_diff_wfp_elgon_2019
  %>% filter(commodity == "Maize (white)")
  %>% ggplot()
  + geom_hline(
    yintercept = 0,
    linetype = "dashed",
    color = "black"
  )
  + geom_rect(
    data = sf_food_price_diff_wfp_elgon_2019[1,],
    xmin = as.Date("2019-06-04"),
    xmax = as.Date("2019-06-08"),
    ymin = -Inf, 
    ymax = Inf,
    fill = "gray",
    alpha = 0.4
  )
  + geom_point(
    aes(x = date, y = price_diff),
    size = 2,
    color = "cornflowerblue"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.83),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
      breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
      labels = c(rep(c(""), num_years + 1), rep(years)),
      limits = c(as.Date(paste0(years[1], "-01-01")), as.Date(paste0(years[num_years], "-01-01")))
  )
  + labs(
    fill = "Location",
    x = "Date",
    y = "Price diff. (UGX/Kg)",
    title = "Maize price comparison"
  )
  + guides(
    color = "none",
    alpha = "none",
  )
)
ggsave(
  filename = file.path(fig_dir, "uganda_maize_price_diff_elgon_2019.png")
)
```

```{r}
#| fig-width: 8
#| fig-height: 8

(
  sf_food_price_diff_wfp_elgon_2019
  %>% ggplot()
  + geom_hline(
    yintercept = 0,
    linetype = "dashed",
    color = "black"
  )
  + geom_rect(
    data = sf_food_price_diff_wfp_elgon_2019 %>% distinct(commodity),
    xmin = as.Date("2019-06-04"),
    xmax = as.Date("2019-06-08"),
    ymin = -Inf, 
    ymax = Inf,
    fill = "gray",
    alpha = 0.4
  )
  + geom_point(
    aes(x = date, y = price_diff),
    size = 2,
    color = "cornflowerblue"
  )
  + facet_wrap(
    ~commodity,
    ncol = 1,
    # strip.position = "right",
    scales = "free_y"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = 15, face = "bold"),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.83),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
      breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
      labels = c(rep(c(""), num_years + 1), rep(years)),
      limits = c(as.Date(paste0(years[1], "-01-01")), as.Date(paste0(years[num_years], "-01-01")))
  )
  + labs(
    fill = "Location",
    x = "Date",
    y = "Price diff. (UGX/Kg)",
    title = "Price diff. comparison"
  )
  + guides(
    color = "none",
    alpha = "none",
  )
)
ggsave(
  filename = file.path(fig_dir, "uganda_all_price_diff_elgon_2019.png")
)
```

## Elgon, April 2024 Flood

```{r}
affected_distr <- c(
  "Mbale",
  "Kapchorwa",
  "Bulambuli",
  "Bukedea", 
  "Butaleja",
  "Sironko",
  "Bududa",
  "Namisindwa"
)

# other_affected_distr <- c(
#   "Bushenyi",
#   "Rukiga"
# )
  
```

```{r}
(
  ggplot(sf_uga %>% mutate(affected = if_else(adm2_name %in% affected_distr, TRUE, FALSE)))
  + geom_sf(
    aes(fill = affected)
  )
  + geom_sf(
    data = sf_food_prices_wfp_uga_maize_latest,
    color = "dodgerblue",
    size = 5,
  )
  + scale_fill_manual(
    values = c("TRUE" = "salmon", "FALSE" = "whitesmoke")
  )
  + theme_classic()
)
```

### Prices

```{r}
sf_food_prices_wfp_elgon_2024 <- (
  sf_food_prices_wfp_uga
  %>% as.data.frame()
  %>% filter(
    commodity %in% commodities_to_plot,
    date >= "2023-01-01"
  )
  %>% mutate(
    affected = if_else(
      admin2 %in% affected_distr,
      "Affected Areas",
      "Rest of the country"
    )
  )
  %>% mutate(
    affected = factor(affected, c("Rest of the country", "Affected Areas"))
  )
  %>% group_by(date, affected, commodity)
  %>% summarise(
    price_q07 = quantile(price, probs = 0.07, na.rm = TRUE),
    price_q50 = quantile(price, probs = 0.5, na.rm = TRUE),
    price_q93 = quantile(price, probs = 0.93, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ungroup()
)
min_year <- min(year(sf_food_prices_wfp_elgon_2024$date))
max_year <- max(year(sf_food_prices_wfp_elgon_2024$date))
years <- seq(min_year, max_year, by = 1)
num_years <- length(years)
```

```{r}
#| fig-width: 8
#| fig-height: 4

(
  sf_food_prices_wfp_elgon_2024
  %>% filter(commodity == "Maize (white)")
  %>% ggplot()
  + geom_rect(
    data = sf_food_prices_wfp_elgon_2024[1,],
    xmin = as.Date("2024-04-01"),
    xmax = as.Date("2024-05-07"),
    ymin = -Inf, 
    ymax = Inf,
    fill = "gray",
    alpha = 0.4
  )
  + geom_ribbon(
    aes(x = date, ymin = price_q07, ymax = price_q93, fill = affected),
    alpha = 0.5
  )
  + geom_point(
    aes(x = date, y = price_q50, color = affected, alpha = affected),
    size = 2
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
      breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
      labels = c(rep(c(""), num_years + 1), rep(years)),
      limits = c(as.Date(paste0(years[1], "-01-01")), as.Date(paste0(years[num_years] + 1, "-01-01")))
  )
  + scale_color_manual(
    values = c("Affected Areas" = "#dc7726", "Rest of the country" = "#97DECE"),
  )
  + scale_fill_manual(
    values = c("Affected Areas" = "#dc7726", "Rest of the country" = "#97DECE"),
  )
  + scale_alpha_manual(
    values = c("Affected Areas" = 1, "Rest of the country" = 0.8)
  )
  + labs(
    fill = "Location",
    x = "Date",
    y = "Price (UGX/Kg)",
    subtitle = "Median and 87% interval coverage",
    title = "Maize price comparison"
  )
  + guides(
    color = "none",
    alpha = "none",
    fill = guide_legend(override.aes = list(alpha = 1))
  )
)
ggsave(
  filename = file.path(fig_dir, "uganda_maize_price_elgon_2024.png")
)
```

```{r}
#| fig-width: 8
#| fig-height: 8

(
  sf_food_prices_wfp_elgon_2024
  %>% ggplot()
  + geom_rect(
    data = sf_food_prices_wfp_elgon_2024 %>% distinct(commodity),
    xmin = as.Date("2024-04-01"),
    xmax = as.Date("2024-05-07"),
    ymin = -Inf, 
    ymax = Inf,
    fill = "gray",
    alpha = 0.4
  )
  + geom_ribbon(
    aes(x = date, ymin = price_q07, ymax = price_q93, fill = affected),
    alpha = 0.5
  )
  + geom_point(
    aes(x = date, y = price_q50, color = affected, alpha = affected),
    size = 2
  )
  + facet_wrap(
    ~commodity,
    ncol = 1,
    # strip.position = "right",
    scales = "free_y"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = 15, face = "bold"),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.83),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
      breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
      labels = c(rep(c(""), num_years + 1), rep(years)),
      limits = c(as.Date(paste0(years[1], "-01-01")), as.Date(paste0(years[num_years] + 1, "-01-01")))
  )
  + scale_color_manual(
    values = c("Affected Areas" = "#dc7726", "Rest of the country" = "#97DECE"),
  )
  + scale_fill_manual(
    values = c("Affected Areas" = "#dc7726", "Rest of the country" = "#97DECE"),
  )
  + scale_alpha_manual(
    values = c("Affected Areas" = 1, "Rest of the country" = 0.8)
  )
  + labs(
    fill = "Location",
    x = "Date",
    y = "Price (UGX/Kg)",
    subtitle = "Median and 87% interval coverage",
    title = "Maize price comparison"
  )
  + guides(
    color = "none",
    alpha = "none",
    fill = guide_legend(override.aes = list(alpha = 1))
  )
)
ggsave(
  filename = file.path(fig_dir, "uganda_all_prices_elgon_2024.png")
)
```

### Price differential

```{r}
sf_food_price_diff_wfp_elgon_2024 <- (
  sf_food_prices_wfp_elgon_2024
  %>% select(
    date,
    affected,
    commodity,
    price_q50
  )
  %>% pivot_wider(
    id_cols = c(date, commodity),
    names_from = affected,
    values_from = price_q50
  )
  %>% mutate(
    price_diff = `Affected Areas` - `Rest of the country`
  )
  %>% filter(!is.na(price_diff))
  %>% select(date, commodity, price_diff)
)
```

```{r}
#| fig-width: 8
#| fig-height: 4

(
  sf_food_price_diff_wfp_elgon_2024
  %>% filter(commodity == "Maize (white)")
  %>% ggplot()
  + geom_hline(
    yintercept = 0,
    linetype = "dashed",
    color = "black"
  )
  + geom_rect(
    data = sf_food_price_diff_wfp_elgon_2024[1,],
    xmin = as.Date("2024-04-01"),
    xmax = as.Date("2024-04-07"),
    ymin = -Inf, 
    ymax = Inf,
    fill = "gray",
    alpha = 0.4
  )
  + geom_point(
    aes(x = date, y = price_diff),
    size = 2,
    color = "cornflowerblue"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.83),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
      breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
      labels = c(rep(c(""), num_years + 1), rep(years)),
      limits = c(as.Date(paste0(years[1], "-01-01")), as.Date(paste0(years[num_years] + 1, "-01-01")))
  )
  + labs(
    fill = "Location",
    x = "Date",
    y = "Price diff. (UGX/Kg)",
    title = "Maize price comparison"
  )
  + guides(
    color = "none",
    alpha = "none",
  )
)
ggsave(
  filename = file.path(fig_dir, "uganda_maize_price_diff_elgon_2024.png")
)
```

```{r}
#| fig-width: 8
#| fig-height: 8

(
  sf_food_price_diff_wfp_elgon_2024
  %>% ggplot()
  + geom_hline(
    yintercept = 0,
    linetype = "dashed",
    color = "black"
  )
  + geom_rect(
    data = sf_food_price_diff_wfp_elgon_2024 %>% distinct(commodity),
    xmin = as.Date("2024-04-01"),
    xmax = as.Date("2024-04-07"),
    ymin = -Inf, 
    ymax = Inf,
    fill = "gray",
    alpha = 0.4
  )
  + geom_point(
    aes(x = date, y = price_diff),
    size = 2,
    color = "cornflowerblue"
  )
  + facet_wrap(
    ~commodity,
    ncol = 1,
    # strip.position = "right",
    scales = "free_y"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = 15, face = "bold"),
    plot.background = element_blank(),
    plot.title = element_text(size = 15, hjust = 0.5),
    plot.subtitle = element_text(size = 13, hjust = 0.5),
    axis.title = element_text(size = 16),
    axis.text.y = element_text(size = 15),
    axis.text.x = element_text(size = 15),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.83),
    axis.ticks.x = element_line(color = c(rep("black", num_years + 1), rep(NA, num_years + 1)))
  )
  + scale_x_date(
      breaks = c(as.Date(paste0(c(years, years[length(years)] + 1), "-01-01")), as.Date(paste0(years, "-07-01"))),
      labels = c(rep(c(""), num_years + 1), rep(years)),
      limits = c(as.Date(paste0(years[1], "-01-01")), as.Date(paste0(years[num_years] + 1, "-01-01")))
  )
  + labs(
    fill = "Location",
    x = "Date",
    y = "Price diff. (UGX/Kg)",
    title = "Price diff. comparison"
  )
  + guides(
    color = "none",
    alpha = "none",
  )
)
ggsave(
  filename = file.path(fig_dir, "uganda_all_price_diff_elgon_2024.png")
)
```

```{r}
df_dis
```
