---
title: "Model main temperature effect on all sexual violence, all female"
format: html
editor: visual
---

# Environment Setup

```{r setup}
# Packages
library(tidyverse)
library(sf)
library(spdep)
library(brms)
options(brms.backend = "cmdstanr", mc.cores = 4)
library(cmdstanr)
library(posterior)
library(ggdist)
library(bayesplot)
library(tidybayes)
# theme_set(bayesplot::theme_default(base_family = "sans", base_size=14))
library(patchwork)
library(paletteer)
library(tidytext)
library(reliabilitydiag)

# Folder Structure
root <- rprojroot::has_file("agile2025.Rproj")$make_fix_file()
temp_dir <- root("temp")
dir.create(temp_dir, showWarnings = FALSE)
raw_data_dir <- root("data/raw")
model_data_dir <- root("data/processed/model_data")
fig_dir <- root("figures")
dir.create(fig_dir, showWarnings = FALSE)
mod_dir <- root("models")
src_dir <- root("src")
```

# Notebook settings

```{r}
# Settings
nb_name <-  "Model main temperature effect on all sexual violence, all female"
seed <- sum(sapply(str_split(nb_name, pattern = "", simplify = TRUE), FUN = utf8ToInt))
set.seed(seed)

# Viz settings
col_palette <- c("#dc7726", "#eebb93", "#ecc046", "#f6e0a3", "#0292b7", "#81c9db", "#35A29F", "#97DECE")
label_map <- c(
  live_with_male_relative = "Live with male relative",
  school_level = "Schooling level",
  rural = "Rural area",
  age = "Age",
  over_18 = "Over 18",
  clim = "Num. Excess Hot Days (past year)",
  clim_between = "Climate between",
  clim_within = "Climate within"
)

# Var settings
countries <- c(
  "Lesotho",
  "Mozambique", 
  "Namibia",
  "Zambia",
  "Zimbabwe"
)
clim_var <- "dev_value_pop_weighted_temp_2m_max_above_1sd_12m"
clim_label <- "temp_max_above_1sd"
viol_var <- "sv"
viol_label <- "all_viol"
pop_label <- "all_female"
covars <- c(
  "sample_weight",
  "country",
  "adm_name",
  "rural",
  "age",
  "school_level",
  "live_with_male_relative",
  viol_var
)
vars <- c(covars, clim_var)

bin_covs <- c(
  "rural",
  "live_with_male_relative" 
)
mono_covs <- c(
  "school_level"
)
cont_covs <- c(
  "age",
  "clim"
)

vars_to_comp <- c(
  # "b_a",
  # "a_coa"
  # "a_coa[1,1]",
  # "a_coa[1,2]",
  # "a_coa[1,3]",
  # "a_coa[1,4]",
  # "a_coa[1,5]",
  # "a_coa[2,1]",
  # "a_coa[2,2]",
  # "a_coa[2,3]",
  # "a_coa[2,4]",
  # "a_coa[2,5]",
  "a_coa[1]",
  "a_coa[2]",
  "a_coa[3]",
  "a_coa[4]",
  "a_coa[5]",
  "b_sc",
  "b_r",
  "b_l",
  "b_cl_mu",
  "b_cl[1,1]",
  "b_cl[1,2]",
  "b_cl[1,3]",
  "b_cl[1,4]",
  "b_cl[1,5]",
  "b_cl[2,1]",
  "b_cl[2,2]",
  "b_cl[2,3]",
  "b_cl[2,4]",
  "b_cl[2,5]"
)
  

# Computation settings
adapt_delta = 0.95
```

# Functions

```{r}
source(file.path(src_dir, "viz.R"))
source(file.path(src_dir, "model_check_utils.R"))

standardize <- function(x) {
  (x - mean(x)) / sd(x)
}
```

# Import Data

## Geometries

### Mozambique

```{r}
sf_moz <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Mozambique 2022 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% st_make_valid()
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Zimbabwe

```{r}
sf_zwe <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Zimbabwe 2015 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% st_make_valid()
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Lesotho

```{r}
sf_lso <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Lesotho 2023 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% st_make_valid()
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Namibia

```{r}
sf_nam <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Namibia 2013 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% st_make_valid()
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Zambia

```{r}
sf_zmb <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Zambia 2018 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% st_make_valid()
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Create combined shapes

```{r}
sf_all <- (
  bind_rows(
    sf_moz,
    sf_zwe,
    sf_lso,
    sf_nam,
    sf_zmb,
  )
  %>% st_transform(3857)
  %>% mutate(
    centroid = st_centroid(geometry),
    lon = st_coordinates(centroid)[,1],
    lat = st_coordinates(centroid)[,2]
  )
)
```

## Climate

```{r}
df_clim <- (
  read_csv(
    file.path(
      model_data_dir,
      "climate_food_prices.csv"
    ),
    show_col_type = FALSE
  )
  %>% filter(
    country %in% countries
  )
)
```

## Climate & Violence

```{r}
df_clim_viol <- (
  read_csv(
    file.path(
      model_data_dir,
      "climate_food_prices_violence.csv"
    ),
    show_col_type = FALSE
  )
  %>% filter(
    !is.na(sample_weight),
    country %in% countries
  )
  %>% group_by(country)
  %>% mutate(
    country_sample_weight = sum(sample_weight, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    sample_weight = (
      1E-3 
      * sample_weight 
      * (
        sum(sample_weight, na.rm = TRUE) 
        / (length(unique(country)) * country_sample_weight)
      )
    ),
    live_with_male_relative = as.numeric(
      (live_with_father == 1) 
      | (live_with_mother == 1) 
      | (live_with_partner == 1)
    ),
  )
  %>% select(
    -c(
      live_with_father, 
      live_with_mother,
      live_with_partner
    )
  )
)
```

# Prepare Data

## Restrict locations

```{r}
sf_vacs <- (
  sf_all
  %>% filter(
    country %in% (df_clim_viol$country %>% unique()),
    adm_name %in% (df_clim_viol$adm_name %>% unique())
  )
  %>% mutate(
    centroid = st_centroid(geometry),
    lon = standardize(st_coordinates(centroid)[,1]),
    lat = standardize(st_coordinates(centroid)[,2])
  )
)
```

## Extract model subsets

```{r}
df_model <- (
  df_clim_viol
  %>% filter(
    female == 1,
  )
  %>% select(all_of(vars))
  # %>% na.omit()
  %>% filter(!is.na(sample_weight))
  %>% mutate(
    rel_age = as.integer(age - min(age))
  )
  %>% group_by(country, adm_name)
  %>% summarise(
    rural = sum(rural * sample_weight, na.rm = TRUE) / sum(sample_weight, na.rm = TRUE),
    age = sum(rel_age * sample_weight, na.rm = TRUE) / sum(sample_weight, na.rm = TRUE),
    school_level = sum(school_level * sample_weight, na.rm = TRUE) / sum(sample_weight, na.rm = TRUE),
    live_with_male_relative = sum(live_with_male_relative * sample_weight, na.rm = TRUE) / sum(sample_weight, na.rm = TRUE),
    sv = round(sum(sv * sample_weight, na.rm = TRUE)),
    N_p = round(sum(sample_weight, na.rm = TRUE)),
    dev_value_pop_weighted_temp_2m_max_above_1sd_12m = sum(dev_value_pop_weighted_temp_2m_max_above_1sd_12m * sample_weight, na.rm = TRUE) / sum(sample_weight, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ungroup()
  %>% merge(
    sf_vacs %>% as.data.frame() %>% select(country, adm_name, lon, lat),
    by = c("country", "adm_name"),
    all.x = TRUE
  )
  %>% mutate(
    country = factor(country),
    adm_name = factor(adm_name),
    clim := scale(!!sym(clim_var)),
    x = standardize(lon),
    y = standardize(lat)
  )
)
```

## Final data for stan models

```{r}
num_hsgp_bases <- 16
indices <- matrix(data = NA, num_hsgp_bases^2, 2)
mm = 0
for (m1 in 1:num_hsgp_bases) {
  for (m2 in 1:num_hsgp_bases) {
    mm = mm + 1
    indices[mm,] = c(m1, m2)
  }
}

data_list <- list(
  N = nrow(df_model),
  N_c = length(unique(df_model$country)),
  N_adm = length(unique(df_model$adm_name)),
  N_p = df_model$N_p,
  co = as.numeric(df_model$country),
  adm = as.numeric(df_model$adm_name),
  R = df_model$rural,
  A = df_model$age,
  sc = df_model$school_level, 
  rel = df_model$live_with_male_relative, 
  cl = as.numeric(df_model$clim),
  sv = df_model$sv,
  M = num_hsgp_bases,
  M_nD = num_hsgp_bases^2,
  L = c(5/2*max(df_model$x), 5/2*max(df_model$y)),
  indices = indices,
  X = cbind(df_model$x, df_model$y)
)
```

# Viz

```{r}
#| fig-width: 6
#| fig-height: 7
(
  ggplot()
  + geom_histogram(
    data = df_clim,
    mapping = aes(
      y = after_stat(ncount),
      x = !!sym(clim_var),
      fill = "clim"
    ),
    alpha = 0.7,
    binwidth = 5
  )
  + geom_histogram(
    data = df_clim_viol,
    mapping = aes(
      y = after_stat(ncount),
      x = !!sym(clim_var),
      fill = "clim_viol"
    ),
    alpha = 0.6,
    binwidth = 5
  )
  + geom_vline(
    xintercept = 0,
    color = "black",
    linetype = "dashed"
  )
  + facet_wrap(
    ~ country,
    ncol = 1
  )
  + scale_fill_manual(
    values = c(
      "clim" = col_palette[2],
      "clim_viol" = col_palette[5]
    ),
    labels = c(
      "clim" = "Climate variable distribution",
      "clim_viol" = "Climate at survey"
    )
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    plot.background = element_blank(),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13),
    strip.background = element_blank(),
    strip.text = element_text(size = 13),
    legend.background = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 13),
  )
  + labs(
    y = "Counts (norm.)",
    x = "Excess hot days"
  )
)
```

```{r}
#| fig-width: 8
#| fig-height: 6
(
  ggplot()
  + geom_histogram(
    data = df_clim %>% filter(country == "Namibia"),
    mapping = aes(
      y = after_stat(ncount),
      x = !!sym(clim_var),
      fill = "clim"
    ),
    alpha = 0.7,
    binwidth = 5
  )
  + geom_histogram(
    data = df_clim_viol %>% filter(country == "Namibia"),
    mapping = aes(
      y = after_stat(ncount),
      x = !!sym(clim_var),
      fill = "clim_viol"
    ),
    alpha = 0.6,
    binwidth = 5
  )
  + geom_vline(
    xintercept = 0,
    color = "black",
    linetype = "dashed"
  )
  + facet_wrap(
    ~ adm_name,
    ncol = 4
  )
  + scale_fill_manual(
    values = c(
      "clim" = col_palette[2],
      "clim_viol" = col_palette[5]
    ),
    labels = c(
      "clim" = "Climate variable distribution",
      "clim_viol" = "Climate at survey"
    )
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    plot.background = element_blank(),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13),
    strip.background = element_blank(),
    strip.text = element_text(size = 13),
    legend.background = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 13),
    legend.position = "inside",
    legend.position.inside = c(0.5, 0.1)
  )
  + labs(
    y = "Counts (norm.)",
    x = "Excess hot days"
  )
)
```

# Modeling

## Fixed effects: country \* climate \* age

```{r}
pars_fixed_country_age <- c(
  "a_co",
  "b_r",
  "b_l",
  "b_a",
  "b_sc",
  "b_cl_mu",
  "b_cl"
)
```

```{r}
m_fixed_country_age <- rstan::stan(
  file = file.path(mod_dir, "fixed_effects_country_age_agg.stan"),
  data = data_list,
  model_name = "m_fixed_country_age",
  seed = seed,
  # control = list(adapt_delta = adapt_delta),
  refresh = 1000
)
```

### Model Checks

```{r}
sp <- rstan::get_sampler_params(m_fixed_country_age, inc_warmup = FALSE)
ebfmi_per_chain <- sapply(sp, function(chain_mat) {
  ener <- chain_mat[,"energy__"]
  mean(diff(ener)^2) / var(ener)
})
print(ebfmi_per_chain)
```

```{r}
#| fig-width: 12
#| fig-height: 12
bayesplot::mcmc_pairs(
  as.array(m_fixed_country_age), 
  pars = c(
    "a_co[2]",
    "b_cl[2]"
  ),
  np = nuts_params(m_fixed_country_age),
  np_style = pairs_style_np(
    div_color = "green",
    div_shape = 20,
    div_size = 3
  )
)
```

### Summary

```{r}
print(
  m_fixed_country_age, 
  pars = pars_fixed_country_age,
  probs = c(0.025, 0.95)
)
```

### Variable analysis

```{r}
(
  m_fixed_country_age
  %>% as_draws_df()
  %>% select(all_of(vars_to_comp))
  %>% pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  )
  %>% mutate(
    variable = factor(variable, levels = vars_to_comp)
  )
  %>% group_by(variable)
  %>% summarise(
    mean = mean(value, na.rm = TRUE),
    pl0 = mean(value < 0, na.rm = TRUE),
    pg0 = mean(value > 0, na.rm = TRUE)
  )
)
```

## Random intercept and slope by country

```{r}
pars_ranef_country_age <- c(
  "a",
  # "a_coa_mu",
  # "a_coa_sd",
  "a_coa",
  "b_r",
  "b_l",
  "b_a",
  "b_sc",
  "b_cl_mu",
  "b_cl_sd",
  "b_cl"
)
```

```{r}
m_ranef_country_age <- rstan::stan(
  file = file.path(mod_dir, "random_effects_country_age.stan"),
  data = data_list,
  model_name = "m_ranef_country_age",
  seed = seed,
  # control = list(adapt_delta = adapt_delta),
  refresh = 1000
)
```

### Model Checks

```{r}
sp <- rstan::get_sampler_params(m_ranef_country_age, inc_warmup = FALSE)
ebfmi_per_chain <- sapply(sp, function(chain_mat) {
  ener <- chain_mat[,"energy__"]
  mean(diff(ener)^2) / var(ener)
})
print(ebfmi_per_chain)
```

```{r}
#| fig-width: 12
#| fig-height: 12
bayesplot::mcmc_pairs(
  as.array(m_ranef_country_age), 
  pars = c(
    # "b_cosv[2]",
    "a_coa[1]",
    # "a_coa[1,2]",
    "b_cl[1,2]"
  ),
  np = nuts_params(m_ranef_country_age),
  np_style = pairs_style_np(
    div_color = "green",
    div_shape = 20,
    div_size = 3
  )
)
```

### Summary

```{r}
print(
  m_ranef_country_age, 
  pars = pars_ranef_country_age,
  probs = c(0.025, 0.95)
)
```

### Variable analysis

```{r}
(
  m_ranef_country_age
  %>% as_draws_df()
  %>% select(all_of(vars_to_comp))
  %>% pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "value"
  )
  %>% mutate(
    variable = factor(variable, levels = vars_to_comp)
  )
  %>% group_by(variable)
  %>% summarise(
    mean = mean(value, na.rm = TRUE),
    pl0 = mean(value < 0, na.rm = TRUE),
    pg0 = mean(value > 0, na.rm = TRUE)
  )
)
```

## Random intercept and slope by country, correlated intercepts and slopes

```{r}
pars_ranef_country_age_corr <- c(
  "a",
  "b_coa",
  "b_coa_sd",
  "b_r",
  "b_l",
  "b_a",
  "b_sc",
  "b_cl_mu",
  "b_cl_sd",
  "b_cl",
  "rho[1]",
  "rho[2]"
)
```

```{r}
m_ranef_country_age_corr <- rstan::stan(
  file = file.path(mod_dir, "random_effects_country_age_corr.stan"),
  data = data_list,
  model_name = "m_ranef_country_age_corr",
  seed = seed,
  # control = list(adapt_delta = adapt_delta),
  refresh = 1000
)
```

### Model Checks

```{r}
sp <- rstan::get_sampler_params(m_ranef_country_age_corr, inc_warmup = FALSE)
ebfmi_per_chain <- sapply(sp, function(chain_mat) {
  ener <- chain_mat[,"energy__"]
  mean(diff(ener)^2) / var(ener)
})
print(ebfmi_per_chain)
```

```{r}
#| fig-width: 12
#| fig-height: 12
bayesplot::mcmc_pairs(
  as.array(m_ranef_country_age_corr), 
  pars = c(
    "b_coa[1,2]",
    "b_cl[1,2]"
  ),
  np = nuts_params(m_ranef_country_age_corr),
  np_style = pairs_style_np(
    div_color = "green",
    div_shape = 20,
    div_size = 3
  )
)
```

### Summary

```{r}
print(
  m_ranef_country_age_corr, 
  pars = pars_ranef_country_age_corr,
  probs = c(0.025, 0.95)
)
```

## Random effects + HSGP

```{r}
pars_ranef_gp <- c(
  "b_co",
  "b_r",
  "b_l",
  "b_a",
  "b_sc",
  "b_cl_mu",
  "b_cl_sd",
  "b_cl",
  "delta",
  "sigma"
)
```

```{r}
m_ranef_gp <- rstan::stan(
  file = file.path(mod_dir, "random_effects_country_age_gp.stan"),
  data = data_list,
  model_name = "m_ranef_gp",
  seed = seed,
  # control = list(adapt_delta = adapt_delta),
  refresh = 1000
)
```

### Model Checks

```{r}
sp <- rstan::get_sampler_params(m_ranef_gp, inc_warmup = FALSE)
ebfmi_per_chain <- sapply(sp, function(chain_mat) {
  ener <- chain_mat[,"energy__"]
  mean(diff(ener)^2) / var(ener)
})
print(ebfmi_per_chain)
```

```{r}
#| fig-width: 12
#| fig-height: 12
bayesplot::mcmc_pairs(
  as.array(m_ranef_gp), 
  pars = c(
    "b_co[2]",
    "b_cl[1,2]"
  ),
  np = nuts_params(m_ranef_gp),
  np_style = pairs_style_np(
    div_color = "green",
    div_shape = 20,
    div_size = 3
  )
)
```

### Summary

```{r}
print(
  m_ranef_gp, 
  pars = pars_ranef_gp,
  probs = c(0.025, 0.95)
)
```

## Total climate effect

```{r}
pars_ranef_tot_eff <- c(
  "a",
  "b_co_mu",
  "b_co_sd",
  "b_co",
  "b_r",
  "b_cl_mu",
  "b_cl_sd",
  "b_cl",
  "rho"
)
```

```{r}
m_ranef_tot_eff <- rstan::stan(
  file = file.path(mod_dir, "random_effects_total_effect.stan"),
  data = data_list,
  model_name = "m_ranef_tot_eff",
  seed = seed,
  # control = list(adapt_delta = adapt_delta),
  refresh = 1000
)
```

### Model Checks

```{r}
sp <- rstan::get_sampler_params(m_ranef_tot_eff, inc_warmup = FALSE)
ebfmi_per_chain <- sapply(sp, function(chain_mat) {
  ener <- chain_mat[,"energy__"]
  mean(diff(ener)^2) / var(ener)
})
print(ebfmi_per_chain)
```

```{r}
#| fig-width: 12
#| fig-height: 12
bayesplot::mcmc_pairs(
  as.array(m_ranef_tot_eff), 
  pars = c(
    "b_co[1]",
    "b_cl[1]"
  ),
  np = nuts_params(m_ranef_tot_eff),
  np_style = pairs_style_np(
    div_color = "green",
    div_shape = 20,
    div_size = 3
  )
)
```

### Summary

```{r}
print(
  m_ranef_tot_eff, 
  pars = pars_ranef_tot_eff,
  probs = c(0.025, 0.95)
)
```

## Total climate effect, with adm random effects

```{r}
pars_ranef_tot_eff_adm <- c(
  "a",
  "b_co_mu",
  "b_co_sd",
  "b_co",
  "b_adm_mu",
  "b_adm_sd",
  "b_r",
  "b_cl_mu",
  "b_cl_sd",
  "b_cl",
  "rho"
)
```

```{r}
m_ranef_tot_eff_adm <- rstan::stan(
  file = file.path(mod_dir, "random_effects_total_effect_adm.stan"),
  data = data_list,
  model_name = "m_ranef_tot_eff_adm",
  seed = seed,
  # control = list(adapt_delta = adapt_delta),
  refresh = 1000
)
```

### Model Checks

```{r}
sp <- rstan::get_sampler_params(m_ranef_tot_eff_adm, inc_warmup = FALSE)
ebfmi_per_chain <- sapply(sp, function(chain_mat) {
  ener <- chain_mat[,"energy__"]
  mean(diff(ener)^2) / var(ener)
})
print(ebfmi_per_chain)
```

```{r}
#| fig-width: 12
#| fig-height: 12
bayesplot::mcmc_pairs(
  as.array(m_ranef_tot_eff_adm), 
  pars = c(
    "b_co[1]",
    "b_cl[1]"
  ),
  np = nuts_params(m_ranef_tot_eff_adm),
  np_style = pairs_style_np(
    div_color = "green",
    div_shape = 20,
    div_size = 3
  )
)
```

### Summary

```{r}
print(
  m_ranef_tot_eff_adm, 
  pars = pars_ranef_tot_eff_adm,
  probs = c(0.025, 0.95)
)
```
