---
title: "Model main temperature effect on all sexual violence, all female"
format: html
editor: visual
---

# Environment Setup

```{r setup}
# Packages
library(tidyverse)
library(sf)
library(spdep)
library(brms)
options(brms.backend = "cmdstanr", mc.cores = 4)
library(cmdstanr)
library(posterior)
library(ggdist)
library(bayesplot)
library(tidybayes)
# theme_set(bayesplot::theme_default(base_family = "sans", base_size=14))
library(patchwork)
library(paletteer)
library(tidytext)
library(reliabilitydiag)

# Folder Structure
root <- rprojroot::has_file("agile2025.Rproj")$make_fix_file()
temp_dir <- root("temp")
dir.create(temp_dir, showWarnings = FALSE)
raw_data_dir <- root("data/raw")
model_data_dir <- root("data/processed/model_data")
fig_dir <- root("figures")
dir.create(fig_dir, showWarnings = FALSE)
mod_dir <- root("models")
src_dir <- root("src")
```

# Notebook settings

```{r}
# Settings
nb_name <-  "Model main temperature effect on all sexual violence, all female"
seed <- sum(sapply(str_split(nb_name, pattern = "", simplify = TRUE), FUN = utf8ToInt))
set.seed(seed)

# Viz settings
col_palette <- c("#dc7726", "#eebb93", "#ecc046", "#f6e0a3", "#0292b7", "#81c9db", "#35A29F", "#97DECE")
label_map <- c(
  live_with_male_relative = "Live with male relative",
  school_level = "Schooling level",
  rural = "Rural area",
  age = "Age",
  over_18 = "Over 18",
  clim = "Num. Excess Hot Days (past year)",
  clim_between = "Climate between",
  clim_within = "Climate within"
)

# Var settings
countries <- c("Lesotho", "Mozambique", "Namibia", "Zambia", "Zimbabwe")
clim_var <- "dev_value_pop_weighted_temp_2m_max_above_1sd_12m"
clim_label <- "temp_max_above_1sd"
viol_var <- "sv"
viol_label <- "all_viol"
pop_label <- "all_female"
covars <- c(
  "sample_weight",
  "country",
  "adm_name",
  "rural",
  "over_18",
  "school_level",
  "live_with_male_relative",
  viol_var
)
vars <- c(covars, clim_var)

bin_covs <- c(
  "rural",
  "over_18",
  "live_with_male_relative" 
)
mono_covs <- c(
  "school_level"
)
cont_covs <- c(
  "clim"
)

# Computation settings
adapt_delta = 0.95
```

# Functions

```{r}
source(file.path(src_dir, "viz.R"))
source(file.path(src_dir, "model_check_utils.R"))

standardize <- function(x) {
  (x - mean(x)) / sd(x)
}
```

# Import Data

## Geometries

### Mozambique

```{r}
sf_moz <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Mozambique 2022 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% st_make_valid()
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Zimbabwe

```{r}
sf_zwe <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Zimbabwe 2015 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% st_make_valid()
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Lesotho

```{r}
sf_lso <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Lesotho 2023 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% st_make_valid()
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Namibia

```{r}
sf_nam <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Namibia 2013 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% st_make_valid()
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Zambia

```{r}
sf_zmb <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Zambia 2018 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% st_make_valid()
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Create combined shapes

```{r}
sf_all <- (
  bind_rows(
    sf_moz,
    sf_zwe,
    sf_lso,
    sf_nam,
    sf_zmb,
  )
  %>% st_transform(3857)
  %>% mutate(
    centroid = st_centroid(geometry),
    lon = st_coordinates(centroid)[,1],
    lat = st_coordinates(centroid)[,2]
  )
)
```

## Climate

```{r}
df_clim <- (
  read_csv(
    file.path(
      model_data_dir,
      "climate_food_prices.csv"
    ),
    show_col_type = FALSE
  )
  %>% filter(
    country %in% countries
  )
)
```

## Climate & Violence

```{r}
df_clim_viol <- (
  read_csv(
    file.path(
      model_data_dir,
      "climate_food_prices_violence.csv"
    ),
    show_col_type = FALSE
  )
  %>% filter(
    !is.na(sample_weight),
    country %in% countries
  )
  %>% group_by(country)
  %>% mutate(
    country_sample_weight = sum(sample_weight, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    sample_weight = (
      1E-3 
      * sample_weight 
      * (
        sum(sample_weight, na.rm = TRUE) 
        / (length(unique(country)) * country_sample_weight)
      )
    ),
    live_with_male_relative = as.numeric(
      (live_with_father == 1) 
      | (live_with_mother == 1) 
      | (live_with_partner == 1)
    ),
    over_18 = as.numeric(age > 18)
  )
  %>% select(
    -c(
      live_with_father, 
      live_with_mother,
      live_with_partner
    )
  )
)
```

# Viz

```{r}
#| fig-width: 6
#| fig-height: 7
(
  ggplot()
  + geom_histogram(
    data = df_clim,
    mapping = aes(
      y = after_stat(ncount),
      x = !!sym(clim_var),
      fill = "clim"
    ),
    alpha = 0.7,
    binwidth = 5
  )
  + geom_histogram(
    data = df_clim_viol,
    mapping = aes(
      y = after_stat(ncount),
      x = !!sym(clim_var),
      fill = "clim_viol"
    ),
    alpha = 0.6,
    binwidth = 5
  )
  + geom_vline(
    xintercept = 0,
    color = "black",
    linetype = "dashed"
  )
  + facet_wrap(
    ~ country,
    ncol = 1
  )
  + scale_fill_manual(
    values = c(
      "clim" = col_palette[2],
      "clim_viol" = col_palette[5]
    ),
    labels = c(
      "clim" = "Climate variable distribution",
      "clim_viol" = "Climate at survey"
    )
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    plot.background = element_blank(),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13),
    strip.background = element_blank(),
    strip.text = element_text(size = 13),
    legend.background = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 13),
  )
  + labs(
    y = "Counts (norm.)",
    x = "Excess hot days"
  )
)
```

```{r}
#| fig-width: 8
#| fig-height: 6
(
  ggplot()
  + geom_histogram(
    data = df_clim %>% filter(country == "Namibia"),
    mapping = aes(
      y = after_stat(ncount),
      x = !!sym(clim_var),
      fill = "clim"
    ),
    alpha = 0.7,
    binwidth = 5
  )
  + geom_histogram(
    data = df_clim_viol %>% filter(country == "Namibia"),
    mapping = aes(
      y = after_stat(ncount),
      x = !!sym(clim_var),
      fill = "clim_viol"
    ),
    alpha = 0.6,
    binwidth = 5
  )
  + geom_vline(
    xintercept = 0,
    color = "black",
    linetype = "dashed"
  )
  + facet_wrap(
    ~ adm_name,
    ncol = 4
  )
  + scale_fill_manual(
    values = c(
      "clim" = col_palette[2],
      "clim_viol" = col_palette[5]
    ),
    labels = c(
      "clim" = "Climate variable distribution",
      "clim_viol" = "Climate at survey"
    )
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    plot.background = element_blank(),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 13),
    strip.background = element_blank(),
    strip.text = element_text(size = 13),
    legend.background = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 13),
    legend.position = "inside",
    legend.position.inside = c(0.5, 0.1)
  )
  + labs(
    y = "Counts (norm.)",
    x = "Excess hot days"
  )
)
```

```{r}
df_clim_viol_summ <- (
  df_clim_viol
  %>% group_by(country, adm_name)
  %>% summarise(
    clim = mean(!!sym(clim_var), na.rm = TRUE),
    rural = sum(rural * sample_weight, na.rm = TRUE) / sum(sample_weight, na.rm = TRUE),
    live_with_male_relative = sum(live_with_male_relative * sample_weight, na.rm = TRUE) / sum(sample_weight, na.rm = TRUE),
    over_18 = sum(over_18 * sample_weight, na.rm = TRUE) / sum(sample_weight, na.rm = TRUE),
    school_level = sum(school_level * sample_weight, na.rm = TRUE) / sum(sample_weight, na.rm = TRUE),
    sv = sum(sv * sample_weight, na.rm = TRUE) / sum(sample_weight, na.rm = TRUE),
    .groups = "drop"
  )
  %>% merge(
    sf_all,
    by = c("country", "adm_name"),
    all.x = TRUE
  )
  %>% st_as_sf()
)
```

```{r}
#| fig-width: 14
#| fig-height: 6
p1 <- (
  ggplot(data = df_clim_viol_summ)
  + geom_sf(
    mapping = aes(fill = clim),
    color = "transparent"
  )
  + scale_fill_viridis_c(
    option = "magma"
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    legend.background = element_blank(),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14)
  )
  + labs(
    fill = "Excess heat days"
  )
)

p2 <- (
  ggplot(data = df_clim_viol_summ)
  + geom_sf(
    mapping = aes(fill = over_18),
    color = "transparent"
  )
  + scale_fill_distiller(
    palette = "Greys",
    labels = scales::label_percent(accuracy = 1)
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    legend.background = element_blank(),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14)
  )
  + labs(
    fill = "Over 18"
  )
)

p3 <- (
  ggplot(data = df_clim_viol_summ)
  + geom_sf(
    mapping = aes(fill = rural),
    color = "transparent"
  )
  + scale_fill_paletteer_c(
    palette = "ggthemes::Orange-Blue Diverging",
    labels = scales::label_percent(accuracy = 1),
    direction = -1
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    legend.background = element_blank(),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14)
  )
  + labs(
    fill = "Rural level"
  )
)

p4 <- (
  ggplot(data = df_clim_viol_summ)
  + geom_sf(
    mapping = aes(fill = live_with_male_relative),
    color = "transparent"
  )
  + scale_fill_distiller(
    palette = "Blues",
    direction = -1,
    labels = scales::label_percent(accuracy = 1)
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    legend.background = element_blank(),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14)
  )
  + labs(
    fill = "Live with male relative"
  )
)

p5 <- (
  ggplot(data = df_clim_viol_summ)
  + geom_sf(
    mapping = aes(fill = school_level),
    color = "transparent"
  )
  + scale_fill_distiller(
    palette = "Greens",
    direction = -1
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    legend.background = element_blank(),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14)
  )
  + labs(
    fill = "School Attainment"
  )
)

p6 <- (
  ggplot(data = df_clim_viol_summ)
  + geom_sf(
    mapping = aes(fill = sv),
    color = "transparent"
  )
  + scale_fill_viridis_c(
    option = "viridis",
    labels = scales::label_percent(accuracy = 1)
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    legend.background = element_blank(),
    legend.text = element_text(size = 13),
    legend.title = element_text(size = 14)
  )
  + labs(
    fill = "Mean SV"
  )
)

# layout <- "
#   ABE
#   CDE
# "
layout <- "
  ABC
  DEF
"

(
  p3 + p2 + p5 + p4 + p1 + p6
  + plot_layout(
    # widths = c(1, 1, 1.5),
    design = layout
  )
)
  
ggsave(
  path = fig_dir,
  filename = paste0("variable_comparison_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

# Prepare Data

## Restrict locations

```{r}
sf_vacs <- (
  sf_all
  %>% filter(
    country %in% (df_clim_viol$country %>% unique()),
    adm_name %in% (df_clim_viol$adm_name %>% unique())
  )
  %>% mutate(
    centroid = st_centroid(geometry),
    lon = standardize(st_coordinates(centroid)[,1]),
    lat = standardize(st_coordinates(centroid)[,2])
  )
)
```

## Extract model subsets

```{r}
df_model <- (
  df_clim_viol
  %>% filter(
    female == 1,
  )
  %>% select(all_of(vars))
  %>% na.omit()
  %>% merge(
    sf_vacs %>% as.data.frame() %>% select(country, adm_name, lon, lat),
    by = c("country", "adm_name"),
    all.x = TRUE
  )
  %>% mutate(
    country = factor(country),
    adm_name = factor(adm_name),
    school_level = factor(
      school_level, 
      levels = sort(unique(school_level)),
      ordered = TRUE
    ),
    clim := scale(!!sym(clim_var))
  )
  %>% group_by(country)
  %>% mutate(
    clim_between = mean(!!sym(clim_var), na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    clim_within = scale(!!sym(clim_var) - clim_between),
    clim_between = scale(clim_between)
  )
)
```

```{r}
data_list <- list(
  N = nrow(df_model),
  N_c = length(unique(df_model$country)),
  N_sc = length(unique(df_model$school_level)),
  W = df_model$sample_weight,
  co = as.numeric(df_model$country),
  R = df_model$rural,
  A = df_model$over_18,
  sc = as.numeric(df_model$school_level), 
  L = df_model$live_with_male_relative, 
  cl = as.numeric(df_model$clim),
  sv = df_model$sv
)
```

## Generate adjacency matrix

```{r}
nb_list <- poly2nb(sf_vacs)
W <- nb2mat(nb_list, style = "B", zero.policy = TRUE)
rownames(W) <- sf_vacs$adm_name
colnames(W) <- sf_vacs$adm_name
levs <- levels(df_model$adm_name)
W_ord <- W[levs, levs]
```

```{r}
edges <- (
  geostan::edges(W, unique_pairs_only = TRUE, shape = sf_vacs$geometry)
)
graph <- st_geometry(edges)
```

```{r}
plot(sf_vacs$geometry, lwd = .15)
plot(graph, add = TRUE, type = "p", col = "red")
plot(graph, add = TRUE, type = "l", col = "black")
```

## Spatial correlation in raw data

```{r}
#| fig-width: 3
#| fig-height: 3
plot_moran(df_model, sf_vacs, outcome_var = viol_var)
ggsave(
  path = fig_dir,
  filename = paste0("moran_plot_obs_main_effect_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

# Modeling

## Fixed intercepts by country

```{r}
stanvars <- stanvar(6, "num_pred")

prior_list_fix_country <- c(
  prior(student_t(3, -1, 0.4), class = "Intercept"),
  prior(normal(0, 0.8 / sqrt(num_pred)), class = "b"),
  prior(dirichlet(1), class = "simo", coef = "moschool_level1")
)

form_fix_country <- bf(
  as.formula(
    paste(
      viol_var,
      "| weights(sample_weight)",
      "~",
      "(
        1 
        + country
        + rural
        + live_with_male_relative
        + over_18
        + mo(school_level)
        + clim
      )"
    )
  )
)
fit_fix_country <- brm(
  formula = form_fix_country,
  family = bernoulli("logit"),
  prior = prior_list_fix_country,
  data = df_model,
  data2 = list(W = W_ord),
  stanvars = stanvars,
  seed = seed,
  refresh = 1000,
  control = list(adapt_delta = adapt_delta),
)
```

### Model Checks

```{r}
sp <- rstan::get_sampler_params(fit_fix_country$fit, inc_warmup = FALSE)
ebfmi_per_chain <- sapply(sp, function(chain_mat) {
  ener <- chain_mat[,"energy__"]
  mean(diff(ener)^2) / var(ener)
})
print(ebfmi_per_chain)
```

```{r}
#| fig-width: 12
#| fig-height: 12
bayesplot::mcmc_pairs(
  as.array(fit_fix_country), 
  pars = c(
    "b_Intercept", 
    "b_rural",
    "b_live_with_male_relative",
    "b_over_18",
    "bsp_moschool_level", 
    "b_clim"
  ),
  np = nuts_params(fit_fix_country),
  np_style = pairs_style_np(
    div_color = "green",
    div_shape = 20,
    div_size = 3
  )
)
```

```{r}
plot(fit_fix_country, variable = "^b_rural", regex = TRUE, nvariables = 3)
```

### Model Output

#### Summary

```{r}
fit_fix_country
```

#### Baseline

```{r}
df_base_sv_by_country_fix_country <- (
  fit_fix_country
  %>% as_draws_df()
  %>% select(
    b_Intercept,
    contains("b_country")
  )
  %>% mutate(
    b_countryLesotho = 0
  )
  %>% pivot_longer(
    cols = -b_Intercept,
    names_to = c("variable", "country"),
    names_pattern = "(b_country)(.*)",
    values_to = "b_country"
  )
  %>% mutate(
    logitp_baseline_sv = b_Intercept + b_country,
    p_baseline_sv = 1 / (1 + exp(-logitp_baseline_sv))
  )
  %>% select(country, p_baseline_sv)
)

df_base_sv_by_country_fix_country_stats <- (
  df_base_sv_by_country_fix_country
  %>% group_by(country)
  %>% summarise(
    median_p_baseline_sv = median(p_baseline_sv, na.rm = TRUE),
    mean_p_baseline_sv = mean(p_baseline_sv, na.rm = TRUE),
    lower1_p = bayestestR::hdi(p_baseline_sv, ci = 0.5)$CI_low,
    upper1_p = bayestestR::hdi(p_baseline_sv, ci = 0.5)$CI_high,
    lower2_p = bayestestR::hdi(p_baseline_sv, ci = 0.95)$CI_low,
    upper2_p = bayestestR::hdi(p_baseline_sv, ci = 0.95)$CI_high,
  )
  %>% ungroup()
)
```

```{r}
#| fig-width: 4.5
#| fig-heigh: 3
(
  ggplot(
    data = df_base_sv_by_country_fix_country,
    mapping = aes(
      y = fct_reorder(country, p_baseline_sv, .fun = ~ -median(.)),
      x = p_baseline_sv
    )
  )
  + ggridges::geom_density_ridges(
      aes(height = after_stat(scaled)),
      stat = "density",
      orientation = "x",
      scale = 0.6,
      rel_min_height = 0.01,
      alpha = 0.7,
      color = NA,
      fill = "#FDD0A2FF"
  )
  + geom_segment(
      data = df_base_sv_by_country_fix_country_stats,
      aes(y = country, yend = country, x = lower2_p, xend = upper2_p),
      linewidth = 1,
      alpha = 0.4,
      color = "#A63603FF",
      inherit.aes = FALSE
  )
  + geom_segment(
      data = df_base_sv_by_country_fix_country_stats,
      aes(y = country, yend = country, x = lower1_p, xend = upper1_p),
      linewidth = 1.5,
      alpha = 0.7,
      color = "#A63603FF",
      inherit.aes = FALSE
  )
  + geom_point(
      data = df_base_sv_by_country_fix_country_stats,
      aes(y = country, x = median_p_baseline_sv),
      color = "#A63603FF",
      size = 3,
      inherit.aes = FALSE
  )
  + geom_vline(xintercept = 0, linetype = "dashed", color = "gray50")
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title.y = element_blank(),
      axis.title.x = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_text(size = 14),
      legend.text = element_text(size = 13),
  )
  + labs(
      x = "Baseline Sexual Violence (%)"
  )
  + scale_y_discrete(
      expand = expansion(mult = c(0.1, 0.25))
  )
  + scale_x_continuous(
      # limits = lims,
      labels = scales::percent_format(accuracy = 1)
  )
) 
```

#### Climate Marginal Effects

##### Population level

```{r}
#| fig-width: 5
#| fig-height: 3
(
  plot_continuous_marginal(fit_fix_country, df_model, covs = cont_covs)
  + scale_y_continuous(
    limits = c(0, 0.6),
    labels = scales::percent_format(accuracy = 1)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_fix_country.png")
)
```

##### By Country

```{r}
#| fig-width: 12
#| fig-height: 2.5
(
  plot_single_continuous_marginal_by_group(fit_fix_country, df_model, cov = "clim")
  + scale_y_continuous(
    limits = c(0, 0.35),
    labels = scales::percent_format(accuracy = 1)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_by_country_", clim_label, "_", pop_label, "_", viol_label, "_fix_country.png")
)
```

### Calibration Checks

```{r}
#| fig-width: 4
#| fig-height: 3
plot_calibration(fit_fix_country, df_model, outcome_var = viol_var)
ggsave(
  path = fig_dir,
  filename = paste0("calibration_plot_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_fix_country.png")
)
```

### Residuals Spatial Correlation Checks

```{r}
#| fig-width: 3.2
#| fig-height: 3
plot_moran(df_model, sf_vacs, fit = fit_fix_country, residuals = TRUE)
ggsave(
  path = fig_dir,
  filename = paste0("moran_plot_res_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_fix_country.png")
)
```

### Posterior Predictive Checks

```{r}
#| fig-width: 12
#| fig-height: 2.5
(
  plot_ppc_rootogram(fit_fix_country, df_model, outcome_var = viol_var)
  + theme(
    legend.position = "inside",
    legend.position.inside = c(0.92, 0.8)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("ppc_by_adm_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_fix_country.png")
)
```

## Random Intercept by country

```{r}
stanvars <- stanvar(5, "num_pred")

prior_list_ranint_country <- c(
  prior(student_t(3, -1, 0.4), class = "Intercept"),
  prior(normal(0, 0.8 / sqrt(num_pred)), class = "b"),
  prior(exponential(1/0.2), class = "sd"),
  prior(dirichlet(1), class = "simo", coef = "moschool_level1")
)

form_ranint_country <- bf(
  as.formula(
    paste(
      viol_var,
      "| weights(sample_weight)",
      "~",
      "(
        1 
        + rural
        + live_with_male_relative
        + over_18
        + mo(school_level)
        + clim
        + (1 | country)
      )"
    )
  )
)
fit_ranint_country <- brm(
  formula = form_ranint_country,
  family = bernoulli("logit"),
  prior = prior_list_ranint_country,
  data = df_model,
  data2 = list(W = W_ord),
  stanvars = stanvars,
  seed = seed,
  refresh = 1000,
  # control = list(adapt_delta = adapt_delta),
  # sample_prior = "only"
  sample_prior = TRUE
)
```

### Model Checks

```{r}
sp <- rstan::get_sampler_params(fit_ranint_country$fit, inc_warmup = FALSE)
ebfmi_per_chain <- sapply(sp, function(chain_mat) {
  ener <- chain_mat[,"energy__"]
  mean(diff(ener)^2) / var(ener)
})
print(ebfmi_per_chain)
```

```{r}
#| fig-width: 12
#| fig-height: 12
bayesplot::mcmc_pairs(
  as.array(fit_ranint_country), 
  pars = c(
    "b_Intercept", 
    "sd_country__Intercept", 
    "b_rural",
    "b_live_with_male_relative",
    "b_over_18",
    "bsp_moschool_level", 
    "b_clim"
  ),
  np = nuts_params(fit_ranint_country),
  np_style = pairs_style_np(
    div_color = "green",
    div_shape = 20,
    div_size = 3
  )
)
```

```{r}
plot(fit_ranint_country, variable = "prior_b", regex = TRUE, nvariables = 3)
```

```{r}
plot(fit_ranint_country, variable = "b_rural", regex = TRUE, nvariables = 3)
```

### Model Output

#### Summary

```{r}
fit_ranint_country
```

#### Baseline

```{r}
#| fig-width: 5
#| fig-height: 3
plot_baseline_levels(fit_ranint_country, lims = c(0, 0.3))
```

#### Climate Marginal Effects

##### Population level

```{r}
#| fig-width: 5
#| fig-height: 3
(
  plot_continuous_marginal(fit_ranint_country, df_model, covs = c("clim"))
  + scale_y_continuous(
    limits = c(0, 0.43),
    labels = scales::percent_format(accuracy = 1)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ranint_country.png")
)
```

##### By Country

```{r}
#| fig-width: 12
#| fig-height: 3
(
  plot_single_continuous_marginal_by_group(fit_ranint_country, df_model, cov = "clim")
  + scale_y_continuous(
    limits = c(0, 0.35),
    labels = scales::percent_format(accuracy = 1)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_by_country_", clim_label, "_", pop_label, "_", viol_label, "_ranint_country.png")
)
```

### Calibration Checks

```{r}
#| fig-width: 4
#| fig-height: 3
plot_calibration(fit_ranint_country, df_model, outcome_var = viol_var)
ggsave(
  path = fig_dir,
  filename = paste0("calibration_plot_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ranint_country.png")
)
```

### Residuals Spatial Correlation Checks

```{r}
#| fig-width: 3.2
#| fig-height: 3
plot_moran(df_model, sf_vacs, fit = fit_ranint_country, residuals = TRUE)
ggsave(
  path = fig_dir,
  filename = paste0("moran_plot_res_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ranint_country.png")
)
```

### Posterior Predictive Checks

```{r}
#| fig-width: 12
#| fig-height: 2.5
(
  plot_ppc_rootogram(fit_ranint_country, df_model, outcome_var = viol_var)
  + theme(
    legend.position = "inside",
    legend.position.inside = c(0.92, 0.8)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("ppc_by_adm_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ranint_country.png")
)
```

## Random Intercept by admin level

```{r}
stanvars <- stanvar(5, "num_pred")

prior_list_ranint_adm <- c(
  prior(student_t(3, -1, 0.4), class = "Intercept"),
  prior(normal(0, 0.8 / sqrt(num_pred)), class = "b"),
  prior(exponential(1/0.3), class = "sd"),
  prior(dirichlet(1), class = "simo", coef = "moschool_level1")
)

form_ranint_adm <- bf(
  as.formula(
    paste(
      viol_var,
      "| weights(sample_weight)",
      "~",
      "(
        1 
        + country
        + (1 | adm_name)
        + rural
        + live_with_male_relative
        + over_18
        + mo(school_level)
        + clim
      )"
    )
  )
)
fit_ranint_adm <- brm(
  formula = form_ranint_adm,
  family = bernoulli("logit"),
  prior = prior_list_ranint_adm,
  data = df_model,
  data2 = list(W = W_ord),
  stanvars = stanvars,
  seed = seed,
  refresh = 1000,
  # control = list(adapt_delta = adapt_delta),
  # sample_prior = "only"
  sample_prior = TRUE
)
```

### Model Checks

```{r}
sp <- rstan::get_sampler_params(fit_ranint_adm$fit, inc_warmup = FALSE)
ebfmi_per_chain <- sapply(sp, function(chain_mat) {
  ener <- chain_mat[,"energy__"]
  mean(diff(ener)^2) / var(ener)
})
print(ebfmi_per_chain)
```

```{r}
#| fig-width: 12
#| fig-height: 12
bayesplot::mcmc_pairs(
  as.array(fit_ranint_adm), 
  pars = c(
    "b_Intercept", 
    "sd_adm_name__Intercept", 
    "b_rural",
    "b_live_with_male_relative",
    "b_over_18",
    "bsp_moschool_level", 
    "b_clim"
  ),
  np = nuts_params(fit_ranint_adm),
  np_style = pairs_style_np(
    div_color = "green",
    div_shape = 20,
    div_size = 3
  )
)
```

```{r}
plot(fit_ranint_adm, variable = "prior_b", regex = TRUE, nvariables = 3)
```

```{r}
plot(fit_ranint_adm, variable = "b_rural", regex = TRUE, nvariables = 3)
```

### Model Output

#### Summary

```{r}
fit_ranint_adm
```

#### Climate Marginal Effects

##### Population level

```{r}
#| fig-width: 5
#| fig-height: 3
(
  plot_continuous_marginal(fit_ranint_adm, df_model, covs = c("clim"))
  + scale_y_continuous(
    limits = c(0, 0.43),
    labels = scales::percent_format(accuracy = 1)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ranint_adm.png")
)
```

##### By Country

```{r}
#| fig-width: 12
#| fig-height: 3
(
  plot_single_continuous_marginal_by_group(fit_ranint_adm, df_model, cov = "clim")
  + scale_y_continuous(
    limits = c(0, 0.35),
    labels = scales::percent_format(accuracy = 1)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_by_country_", clim_label, "_", pop_label, "_", viol_label, "_ranint_adm.png")
)
```

### Calibration Checks

```{r}
#| fig-width: 4
#| fig-height: 3
plot_calibration(fit_ranint_adm, df_model, outcome_var = viol_var)
ggsave(
  path = fig_dir,
  filename = paste0("calibration_plot_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ranint_adm.png")
)
```

### Residuals Spatial Correlation Checks

```{r}
#| fig-width: 3.2
#| fig-height: 3
plot_moran(df_model, sf_vacs, fit = fit_ranint_adm, residuals = TRUE)
ggsave(
  path = fig_dir,
  filename = paste0("moran_plot_res_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ranint_adm.png")
)
```

### Posterior Predictive Checks

```{r}
#| fig-width: 12
#| fig-height: 2.5
(
  plot_ppc_rootogram(fit_ranint_adm, df_model, outcome_var = viol_var)
  + theme(
    legend.position = "inside",
    legend.position.inside = c(0.92, 0.8)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("ppc_by_adm_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ranint_adm.png")
)
```

## Fixed interaction: country \* climate

```{r}
stanvars <- stanvar(5, "num_pred")

prior_list_fix_int_cc <- c(
  prior(student_t(3, -1, 0.4), class = "Intercept"),
  prior(normal(0, 0.8 / sqrt(num_pred)), class = "b"),
  prior(dirichlet(1), class = "simo", coef = "moschool_level1")
)

form_fix_int_cc <- bf(
  as.formula(
    paste(
      viol_var,
      "| weights(sample_weight)",
      "~",
      "(
        1 
        + rural
        + live_with_male_relative
        + over_18
        + mo(school_level)
        + country * clim
      )"
    )
  )
)
fit_fix_int_cc <- brm(
  formula = form_fix_int_cc,
  family = bernoulli("logit"),
  prior = prior_list_fix_int_cc,
  data = df_model,
  data2 = list(W = W_ord),
  stanvars = stanvars,
  seed = seed,
  refresh = 1000,
  # control = list(adapt_delta = adapt_delta),
  # sample_prior = "only"
  sample_prior = TRUE
)
```

### Model Checks

```{r}
sp <- rstan::get_sampler_params(fit_fix_int_cc$fit, inc_warmup = FALSE)
ebfmi_per_chain <- sapply(sp, function(chain_mat) {
  ener <- chain_mat[,"energy__"]
  mean(diff(ener)^2) / var(ener)
})
print(ebfmi_per_chain)
```

```{r}
#| fig-width: 12
#| fig-height: 12
bayesplot::mcmc_pairs(
  as.array(fit_fix_int_cc), 
  pars = c(
    "b_Intercept", 
    "b_rural",
    "b_live_with_male_relative",
    "b_over_18",
    "bsp_moschool_level", 
    "b_clim"
  ),
  np = nuts_params(fit_fix_int_cc),
  np_style = pairs_style_np(
    div_color = "green",
    div_shape = 20,
    div_size = 3
  )
)
```

```{r}
plot(fit_fix_int_cc, variable = "prior_b", regex = TRUE, nvariables = 3)
```

```{r}
plot(fit_fix_int_cc, variable = "b_rural", regex = TRUE, nvariables = 3)
```

### Model Output

#### Summary

```{r}
fit_fix_int_cc
```

#### Climate Marginal Effects

##### Population level

```{r}
#| fig-width: 5
#| fig-height: 3
(
  plot_continuous_marginal(fit_fix_int_cc, df_model, covs = c("clim"))
  + scale_y_continuous(
    limits = c(0, 0.43),
    labels = scales::percent_format(accuracy = 1)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_fix_int_cc.png")
)
```

##### By Country

```{r}
#| fig-width: 12
#| fig-height: 3
(
  plot_single_continuous_marginal_by_group(fit_fix_int_cc, df_model, cov = "clim")
  + scale_y_continuous(
    limits = c(0, 0.35),
    labels = scales::percent_format(accuracy = 1)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_by_country_", clim_label, "_", pop_label, "_", viol_label, "_fix_int_cc.png")
)
```

### Calibration Checks

```{r}
#| fig-width: 4
#| fig-height: 3
plot_calibration(fit_fix_int_cc, df_model, outcome_var = viol_var)
ggsave(
  path = fig_dir,
  filename = paste0("calibration_plot_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_fix_int_cc.png")
)
```

### Residuals Spatial Correlation Checks

```{r}
#| fig-width: 3.2
#| fig-height: 3
plot_moran(df_model, sf_vacs, fit = fit_fix_int_cc, residuals = TRUE)
ggsave(
  path = fig_dir,
  filename = paste0("moran_plot_res_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_fix_int_cc.png")
)
```

### Posterior Predictive Checks

```{r}
#| fig-width: 12
#| fig-height: 2.5
(
  plot_ppc_rootogram(fit_fix_int_cc, df_model, outcome_var = viol_var)
  + theme(
    legend.position = "inside",
    legend.position.inside = c(0.92, 0.8)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("ppc_by_adm_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_fix_int_cc.png")
)
```

## Fixed interaction: country \* climate \* age

```{r}
stanvars <- stanvar(5, "num_pred")

prior_list_fix_int_cca <- c(
  prior(student_t(3, -1, 0.4), class = "Intercept"),
  prior(normal(0, 0.8 / sqrt(num_pred)), class = "b"),
  prior(dirichlet(1), class = "simo", coef = "moschool_level1")
)

form_fix_int_cca <- bf(
  as.formula(
    paste(
      viol_var,
      "| weights(sample_weight)",
      "~",
      "(
        1 
        + country
        + rural
        + live_with_male_relative
        + mo(school_level)
        + over_18
        + clim
        + clim:over_18
      )"
    )
  )
)
fit_fix_int_cca <- brm(
  formula = form_fix_int_cca,
  family = bernoulli("logit"),
  prior = prior_list_fix_int_cca,
  data = df_model,
  data2 = list(W = W_ord),
  stanvars = stanvars,
  seed = seed,
  refresh = 1000,
  # control = list(adapt_delta = adapt_delta),
  # sample_prior = "only"
  sample_prior = TRUE
)
```

### Model Checks

```{r}
sp <- rstan::get_sampler_params(fit_fix_int_cca$fit, inc_warmup = FALSE)
ebfmi_per_chain <- sapply(sp, function(chain_mat) {
  ener <- chain_mat[,"energy__"]
  mean(diff(ener)^2) / var(ener)
})
print(ebfmi_per_chain)
```

```{r}
#| fig-width: 12
#| fig-height: 12
bayesplot::mcmc_pairs(
  as.array(fit_fix_int_cca), 
  pars = c(
    "b_Intercept", 
    "b_rural",
    "b_live_with_male_relative",
    "b_over_18",
    "bsp_moschool_level", 
    "b_clim"
  ),
  np = nuts_params(fit_fix_int_cca),
  np_style = pairs_style_np(
    div_color = "green",
    div_shape = 20,
    div_size = 3
  )
)
```

```{r}
plot(fit_fix_int_cca, variable = "prior_b", regex = TRUE, nvariables = 3)
```

```{r}
plot(fit_fix_int_cca, variable = "b_rural", regex = TRUE, nvariables = 3)
```

### Model Output

#### Summary

```{r}
fit_fix_int_cca
```

#### Climate Marginal Effects

##### Population level

```{r}
#| fig-width: 5
#| fig-height: 3
(
  plot_continuous_marginal(fit_fix_int_cca, df_model, covs = c("clim"))
  + scale_y_continuous(
    limits = c(0, 0.43),
    labels = scales::percent_format(accuracy = 1)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_fix_int_cca.png")
)
```

##### By Country

```{r}
#| fig-width: 12
#| fig-height: 3
(
  plot_single_continuous_marginal_by_group(fit_fix_int_cca, df_model, cov = "clim")
  + scale_y_continuous(
    limits = c(0, 0.35),
    labels = scales::percent_format(accuracy = 1)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_by_country_", clim_label, "_", pop_label, "_", viol_label, "_fix_int_cca.png")
)
```

### Calibration Checks

```{r}
#| fig-width: 4
#| fig-height: 3
plot_calibration(fit_fix_int_cca, df_model, outcome_var = viol_var)
ggsave(
  path = fig_dir,
  filename = paste0("calibration_plot_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_fix_int_cca.png")
)
```

### Residuals Spatial Correlation Checks

```{r}
#| fig-width: 3.2
#| fig-height: 3
plot_moran(df_model, sf_vacs, fit = fit_fix_int_cca, residuals = TRUE)
ggsave(
  path = fig_dir,
  filename = paste0("moran_plot_res_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_fix_int_cca.png")
)
```

### Posterior Predictive Checks

```{r}
#| fig-width: 12
#| fig-height: 2.5
(
  plot_ppc_rootogram(fit_fix_int_cca, df_model, outcome_var = viol_var)
  + theme(
    legend.position = "inside",
    legend.position.inside = c(0.92, 0.8)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("ppc_by_adm_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_fix_int_cca.png")
)
```

## Random intercept and slope by country

```{r}
pars_to_plot <- c(
  "b_cosv",
  "b_rsv",
  "b_lsv",
  "b_asv",
  "b_scsv",
  "b_clsv_mu",
  "b_clsv_sd",
  "b_clsv"
)
```

```{r}
m_ranef_country_age <- rstan::stan(
  file = file.path(mod_dir, "random_effects_country_age.stan"),
  data = data_list,
  model_name = "m_ranef_country_age",
  seed = seed,
  # control = list(adapt_delta = adapt_delta),
  refresh = 1000
)
```

```{r}
print(m_ranef_country_age, pars = pars_to_plot)
```

```{r}
stanvars <- stanvar(6, "num_pred")

prior_list_ran_country <- c(
  prior(student_t(3, -1, 0.4), class = "Intercept"),
  prior(normal(0, 0.8 / sqrt(num_pred)), class = "b"),
  prior(exponential(1/0.2), class = "sd"),
  prior(dirichlet(1), class = "simo", coef = "moschool_level1")
)

form_ran_country <- bf(
  as.formula(
    paste(
      viol_var,
      "| weights(sample_weight)",
      "~",
      "(
        1 
        + rural
        + live_with_male_relative
        + over_18
        + mo(school_level)
        + clim
        + (1 + clim | country)
      )"
    )
  )
)
fit_ran_country <- brm(
  formula = form_ran_country,
  family = bernoulli("logit"),
  prior = prior_list_ran_country,
  data = df_model,
  data2 = list(W = W_ord),
  stanvars = stanvars,
  seed = seed,
  refresh = 1000,
  control = list(adapt_delta = adapt_delta),
  # sample_prior = "only"
  sample_prior = TRUE
)
```

### Model Checks

```{r}
sp <- rstan::get_sampler_params(fit_ran_country$fit, inc_warmup = FALSE)
ebfmi_per_chain <- sapply(sp, function(chain_mat) {
  ener <- chain_mat[,"energy__"]
  mean(diff(ener)^2) / var(ener)
})
print(ebfmi_per_chain)
```

```{r}
#| fig-width: 12
#| fig-height: 12
bayesplot::mcmc_pairs(
  as.array(fit_ran_country), 
  pars = c(
    "b_Intercept", 
    "sd_country__Intercept", 
    "b_rural",
    "b_live_with_male_relative",
    "b_over_18",
    "bsp_moschool_level", 
    "b_clim"
  ),
  np = nuts_params(fit_ran_country),
  np_style = pairs_style_np(
    div_color = "green",
    div_shape = 20,
    div_size = 3
  )
)
```

```{r}
plot(fit_ran_country, variable = "prior_b", regex = TRUE, nvariables = 3)
```

```{r}
plot(fit_ran_country, variable = "^b_rural", regex = TRUE, nvariables = 3)
```

### Model Output

#### Summary

```{r}
fit_ran_country
```

#### Baseline

```{r}
#| fig-width: 5
#| fig-height: 3
plot_baseline_levels(fit_ran_country, lims = c(0, 0.3))
```

#### Climate Marginal Effects

##### Population level

```{r}
#| fig-width: 5
#| fig-height: 3
(
  plot_continuous_marginal(fit_ran_country, df_model, covs = c("clim"))
  + scale_y_continuous(
    limits = c(0, 0.6),
    labels = scales::percent_format(accuracy = 1)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ran_country.png")
)
```

##### By Country

```{r}
#| fig-width: 12
#| fig-height: 2.5
(
  plot_single_continuous_marginal_by_group(fit_ran_country, df_model, cov = "clim")
  + scale_y_continuous(
    limits = c(0, 1),
    labels = scales::percent_format(accuracy = 1)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_by_country_", clim_label, "_", pop_label, "_", viol_label, "_ran_country.png")
)
```

### Calibration Checks

```{r}
#| fig-width: 4
#| fig-height: 3
plot_calibration(fit_ran_country, df_model, outcome_var = viol_var)
ggsave(
  path = fig_dir,
  filename = paste0("calibration_plot_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ran_country.png")
)
```

### Residuals Spatial Correlation Checks

```{r}
#| fig-width: 3.2
#| fig-height: 3
plot_moran(df_model, sf_vacs, fit = fit_ran_country, residuals = TRUE)
ggsave(
  path = fig_dir,
  filename = paste0("moran_plot_res_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ran_country.png")
)
```

### Posterior Predictive Checks

```{r}
#| fig-width: 12
#| fig-height: 2.5
(
  plot_ppc_rootogram(fit_ran_country, df_model, outcome_var = viol_var)
  + theme(
    legend.position = "inside",
    legend.position.inside = c(0.92, 0.8)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("ppc_by_adm_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_ran_country.png")
)
```

## Mundlak

```{r}
stanvars <- stanvar(7, "num_pred")

prior_list_mund <- c(
  prior(student_t(3, -1, 0.4), class = "Intercept"),
  prior(normal(0, 0.8 / sqrt(num_pred)), class = "b"),
  prior(dirichlet(1), class = "simo", coef = "moschool_level1")
)

form_mund <- bf(
  as.formula(
    paste(
      viol_var,
      "| weights(sample_weight)",
      "~",
      "(
        1 
        + country
        + rural
        + live_with_male_relative
        + over_18
        + mo(school_level)
        + clim_between
        + clim_within
      )"
    )
  )
)
fit_mund <- brm(
  formula = form_mund,
  family = bernoulli("logit"),
  prior = prior_list_mund,
  data = df_model,
  data2 = list(W = W_ord),
  stanvars = stanvars,
  seed = seed,
  refresh = 1000,
  # control = list(adapt_delta = adapt_delta),
  # sample_prior = "only"
  sample_prior = TRUE
)
```

### Model Checks

```{r}
sp <- rstan::get_sampler_params(fit_mund$fit, inc_warmup = FALSE)
ebfmi_per_chain <- sapply(sp, function(chain_mat) {
  ener <- chain_mat[,"energy__"]
  mean(diff(ener)^2) / var(ener)
})
print(ebfmi_per_chain)
```

```{r}
#| fig-width: 12
#| fig-height: 12
bayesplot::mcmc_pairs(
  as.array(fit_mund), 
  pars = c(
    "b_Intercept", 
    "b_rural",
    "b_live_with_male_relative",
    "b_over_18",
    "bsp_moschool_level", 
    "b_clim_between",
    "b_clim_within"
  ),
  np = nuts_params(fit_mund),
  np_style = pairs_style_np(
    div_color = "green",
    div_shape = 20,
    div_size = 3
  )
)
```

```{r}
plot(fit_mund, variable = "prior_b", regex = TRUE, nvariables = 3)
```

```{r}
plot(fit_mund, variable = "b_rural", regex = TRUE, nvariables = 3)
```

### Model Output

#### Summary

```{r}
fit_mund
```

#### Baseline

```{r}
df_base_sv_by_country_mund <- (
  fit_mund
  %>% as_draws_df()
  %>% select(
    b_Intercept,
    contains("b_country")
  )
  %>% mutate(
    b_countryLesotho = 0
  )
  %>% pivot_longer(
    cols = -b_Intercept,
    names_to = c("variable", "country"),
    names_pattern = "(b_country)(.*)",
    values_to = "b_country"
  )
  %>% mutate(
    logitp_baseline_sv = b_Intercept + b_country,
    p_baseline_sv = 1 / (1 + exp(-logitp_baseline_sv))
  )
  %>% select(country, p_baseline_sv)
)

df_base_sv_by_country_mund_stats <- (
  df_base_sv_by_country_mund
  %>% group_by(country)
  %>% summarise(
    median_p_baseline_sv = median(p_baseline_sv, na.rm = TRUE),
    mean_p_baseline_sv = mean(p_baseline_sv, na.rm = TRUE),
    lower1_p = bayestestR::hdi(p_baseline_sv, ci = 0.5)$CI_low,
    upper1_p = bayestestR::hdi(p_baseline_sv, ci = 0.5)$CI_high,
    lower2_p = bayestestR::hdi(p_baseline_sv, ci = 0.95)$CI_low,
    upper2_p = bayestestR::hdi(p_baseline_sv, ci = 0.95)$CI_high,
  )
  %>% ungroup()
)
```

```{r}
#| fig-width: 4.5
#| fig-heigh: 3
(
  ggplot(
    data = df_base_sv_by_country_mund,
    mapping = aes(
      y = fct_reorder(country, p_baseline_sv, .fun = ~ -median(.)),
      x = p_baseline_sv
    )
  )
  + ggridges::geom_density_ridges(
      aes(height = after_stat(scaled)),
      stat = "density",
      orientation = "x",
      scale = 0.6,
      rel_min_height = 0.01,
      alpha = 0.7,
      color = NA,
      fill = "#FDD0A2FF"
  )
  + geom_segment(
      data = df_base_sv_by_country_mund_stats,
      aes(y = country, yend = country, x = lower2_p, xend = upper2_p),
      linewidth = 1,
      alpha = 0.4,
      color = "#A63603FF",
      inherit.aes = FALSE
  )
  + geom_segment(
      data = df_base_sv_by_country_mund_stats,
      aes(y = country, yend = country, x = lower1_p, xend = upper1_p),
      linewidth = 1.5,
      alpha = 0.7,
      color = "#A63603FF",
      inherit.aes = FALSE
  )
  + geom_point(
      data = df_base_sv_by_country_mund_stats,
      aes(y = country, x = median_p_baseline_sv),
      color = "#A63603FF",
      size = 3,
      inherit.aes = FALSE
  )
  + geom_vline(xintercept = 0, linetype = "dashed", color = "gray50")
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title.y = element_blank(),
      axis.title.x = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_text(size = 14),
      legend.text = element_text(size = 13),
  )
  + labs(
      x = "Baseline Sexual Violence (%)"
  )
  + scale_y_discrete(
      expand = expansion(mult = c(0.1, 0.25))
  )
  + scale_x_continuous(
      # limits = lims,
      labels = scales::percent_format(accuracy = 1)
  )
) 
```

#### Climate Marginal Effects

##### Population level

```{r}
#| fig-width: 5
#| fig-height: 3
(
  plot_continuous_marginal(fit_mund, df_model, covs = c("clim_between"))
  + scale_y_continuous(
    # limits = c(0, 0.43),
    labels = scales::percent_format(accuracy = 1)
  )
)
# ggsave(
#   path = fig_dir,
#   filename = paste0("marginal_prob_clim_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_mund.png")
# )
```

```{r}
#| fig-width: 5
#| fig-height: 3
(
  plot_continuous_marginal(fit_mund, df_model, covs = c("clim_within"))
  + scale_y_continuous(
    # limits = c(0, 0.43),
    labels = scales::percent_format(accuracy = 1)
  )
)
# ggsave(
#   path = fig_dir,
#   filename = paste0("marginal_prob_clim_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_mund.png")
# )
```

##### By Country

```{r}
#| fig-width: 12
#| fig-height: 3
(
  plot_single_continuous_marginal_by_group(fit_mund, df_model, cov = "clim_between")
  + scale_y_continuous(
    limits = c(0, 0.35),
    labels = scales::percent_format(accuracy = 1)
  )
)
# ggsave(
#   path = fig_dir,
#   filename = paste0("marginal_prob_clim_main_effect_by_country_", clim_label, "_", pop_label, "_", viol_label, "_mund.png")
# )
```

```{r}
#| fig-width: 12
#| fig-height: 3
(
  plot_single_continuous_marginal_by_group(fit_mund, df_model, cov = "clim_within")
  + scale_y_continuous(
    limits = c(0, 0.35),
    labels = scales::percent_format(accuracy = 1)
  )
)
# ggsave(
#   path = fig_dir,
#   filename = paste0("marginal_prob_clim_main_effect_by_country_", clim_label, "_", pop_label, "_", viol_label, "_mund.png")
# )
```

### Calibration Checks

```{r}
#| fig-width: 4
#| fig-height: 3
plot_calibration(fit_mund, df_model, outcome_var = viol_var)
ggsave(
  path = fig_dir,
  filename = paste0("calibration_plot_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_mund.png")
)
```

### Residuals Spatial Correlation Checks

```{r}
#| fig-width: 3.2
#| fig-height: 3
plot_moran(df_model, sf_vacs, fit = fit_mund, residuals = TRUE)
ggsave(
  path = fig_dir,
  filename = paste0("moran_plot_res_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_mund.png")
)
```

### Posterior Predictive Checks

```{r}
#| fig-width: 12
#| fig-height: 2.5
(
  plot_ppc_rootogram(fit_mund, df_model, outcome_var = viol_var)
  + theme(
    legend.position = "inside",
    legend.position.inside = c(0.92, 0.8)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("ppc_by_adm_main_effect_", clim_label, "_", pop_label, "_", viol_label, "_mund.png")
)
```

## Full Model

```{r}
stanvars <- stanvar(6, "num_pred")

prior_list <- c(
  prior(student_t(3, -1, 0.4), class = "Intercept"),
  prior(normal(0, (0.8 / sqrt(num_pred))), class = "b"),
  prior(exponential(1/0.3), class = "sd"),
  prior(dirichlet(1), class = "simo", coef = "moschool_level1"),
  # prior(beta(1, 1), class = "rhocar"),
  prior(beta(1, 1), class = "car"),
  prior(exponential(1), class = "sdcar")
)

form <- bf(
  as.formula(
    paste(
      viol_var,
      "| weights(sample_weight)",
      "~",
      "(
        1
        + country
        + rural
        + live_with_male_relative
        + over_18
        + mo(school_level)
        + clim
        + (-1 + clim | country)
        + car(W, gr = adm_name, type = 'escar')
      )"
    )
  )
)
fit <- brm(
  formula = form,
  family = bernoulli("logit"),
  prior = prior_list,
  data = df_model,
  data2 = list(W = W_ord),
  stanvars = stanvars,
  seed = seed,
  refresh = 1000,
  control = list(
    adapt_delta = adapt_delta
  ),
  # sample_prior = "only"
  sample_prior = TRUE
)
```

### Model Checks

```{r}
sp <- rstan::get_sampler_params(fit$fit, inc_warmup = FALSE)
ebfmi_per_chain <- sapply(sp, function(chain_mat) {
  ener <- chain_mat[,"energy__"]
  mean(diff(ener)^2) / var(ener)
})
print(ebfmi_per_chain)
```

```{r}
mcmc_trace(
  as_draws_df(fit), 
  pars = c(
    "car",
    # "rhocar",
    "sdcar",
    "rcar[1]"
  )
)
```

```{r}
#| fig-width: 12
#| fig-height: 12
bayesplot::mcmc_pairs(
  as.array(fit),
  pars = c(
    "b_Intercept", 
    "sd_country__Intercept", 
    "b_rural",
    "b_live_with_male_relative",
    "b_over_18",
    "bsp_moschool_level", 
    "b_clim",
    "car",
    # "rhocar",
    "sdcar",
    "rcar[1]"
    # "sdgp_gplonlat",
    # "lscale_gplonlat",
    # "zgp_gplonlat[1]"
  ),
  np = nuts_params(fit),
  np_style = pairs_style_np(
    div_color = "green",
    div_shape = 20,
    div_size = 3
  )
)
```

```{r}
mcmc_parcoord(
  as.array(fit),
  pars = c(
    "b_Intercept", 
    "sd_country__Intercept", 
    "b_rural",
    "b_live_with_male_relative",
    "b_over_18",
    "bsp_moschool_level", 
    "b_clim",
    "car",
    # "rhocar",
    "sdcar",
    "rcar[1]"
  ),
  np = nuts_params(fit),
  np_style = parcoord_style_np(
    div_color = "green",
    div_alpha = 0.6,
    div_size = 1.1
  )
)
```

```{r}
plot(
  fit,
  variable = "prior_car",
  # variable = "prior_rhocar",
  # variable = "prior_lscale__1_gplonlat",
  # variable = "prior_sdgp",
  regex = TRUE,
  nvariables = 3
)
```

```{r}
plot(
  fit, 
  variable = "^car",
  # variable = "^rhocar",
  # variable = "^lscale",
  # variable = "sdgp_gplonlat",
  regex = TRUE, 
  nvariables = 3
)
```

### Model Output

#### Summary

```{r}
fit
```

#### Baseline

```{r}
#| fig-width: 5
#| fig-height: 3
plot_baseline_levels(fit, lims = c(0, 0.35))
```

#### Climate Marginal Effects

##### Population level

```{r}
#| fig-width: 5
#| fig-height: 3
(
  plot_continuous_marginal(fit, df_model, covs = c("clim"))
  + scale_y_continuous(
    limits = c(0, 0.35),
    labels = scales::percent_format(accuracy = 1)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

##### By Country

```{r}
#| fig-width: 12
#| fig-height: 2.5
(
  plot_single_continuous_marginal_by_group(fit, df_model, cov = "clim")
  + scale_y_continuous(
    limits = c(0, 1),
    labels = scales::percent_format(accuracy = 1)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("marginal_prob_clim_main_effect_by_country_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

### Calibration Checks

```{r}
#| fig-width: 4
#| fig-height: 3
(
  plot_calibration(fit, df_model, outcome_var = viol_var)
  + coord_cartesian(clip = "off")
)
ggsave(
  path = fig_dir,
  filename = paste0("calibration_plot_main_effect_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

### Residuals Spatial Correlation Checks

```{r}
#| fig-width: 3.2
#| fig-height: 3
(
  plot_moran(df_model, sf_vacs, fit = fit, residuals = TRUE)
  # + theme(
  #   margins = margin(r = 10, l = 10)
  # )
  + theme(
    scale_x_continuous(
      breaks = scales::pretty_breaks()
    )
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("moran_plot_res_main_effect_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

### Posterior Predictive Checks

```{r}
#| fig-width: 12
#| fig-height: 2.5
(
  plot_ppc_rootogram(fit, df_model, outcome_var = viol_var)
  + theme(
    legend.position = "inside",
    legend.position.inside = c(0.93, 0.82)
  )
)
ggsave(
  path = fig_dir,
  filename = paste0("ppc_by_adm_main_effect_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```

# Compare Models

```{r}
fit_fix_country <- add_criterion(
  fit_fix_country, 
  "loo",
  # moment_match = TRUE
)
fit_ranint_country <- add_criterion(
  fit_ranint_country, 
  "loo",
  # moment_match = TRUE
)
fit_ranint_adm <- add_criterion(
  fit_ranint_adm, 
  "loo",
  # moment_match = TRUE
)
fit_fix_int_cc <- add_criterion(
  fit_fix_int_cc, 
  "loo",
  # moment_match = TRUE
)
fit_fix_int_cca <- add_criterion(
  fit_fix_int_cca, 
  "loo",
  # moment_match = TRUE
)
fit_ran_country <- add_criterion(
  fit_ran_country, 
  "loo",
  # moment_match = TRUE
)
fit_mund <- add_criterion(
  fit_mund, 
  "loo",
  # moment_match = TRUE
)
fit <- add_criterion(
  fit, 
  "loo", 
  moment_match = TRUE
)
loo::loo_compare(
  fit_fix_country,
  fit_ranint_country,
  fit_ranint_adm,
  fit_fix_int_cc,
  fit_fix_int_cca,
  fit_ran_country,
  fit_mund,
  fit
)
```

```{r}
bind_rows(
  get_calibration_scores(fit_fix_country, df_model, outcome_var = viol_var, model_name = "Fixed country intercept"),
  get_calibration_scores(fit_ranint_country, df_model, outcome_var = viol_var, model_name = "Random intercept by country"),
  get_calibration_scores(fit_ranint_adm, df_model, outcome_var = viol_var, model_name = "Random intercept by admin"),
  get_calibration_scores(fit_fix_int_cc, df_model, outcome_var = viol_var, model_name = "Fixed interaction, country * clim"),
  get_calibration_scores(fit_fix_int_cca, df_model, outcome_var = viol_var, model_name = "Fixed interaction, country * clim * age"),
  get_calibration_scores(fit_ran_country, df_model, outcome_var = viol_var, model_name = "Random by country"),
  get_calibration_scores(fit_mund, df_model, outcome_var = viol_var, model_name = "Mundlak"),
  get_calibration_scores(fit, df_model, outcome_var = viol_var, model_name = "Full Model"),
)
```

```{r}
clim_draws_fix_country <- (
  as_draws_df(fit_fix_country)
  %>% select(clim = b_clim)
  %>% mutate(
    model = "Fixed country intercept"
  )
)

clim_draws_ranint_country <- (
  as_draws_df(fit_ranint_country)
  %>% select(clim = b_clim)
  %>% mutate(
    model = "Random intercept by country"
  )
)

clim_draws_ranint_adm <- (
  as_draws_df(fit_ranint_adm)
  %>% select(clim = b_clim)
  %>% mutate(
    model = "Random intercept by admin"
  )
)

clim_draws_fix_int_cc <- (
  as_draws_df(fit_fix_int_cc)
  %>% select(clim = b_clim)
  %>% mutate(
    model = "Fixed interaction, country * clim"
  )
)

clim_draws_fix_int_cca <- (
  as_draws_df(fit_fix_int_cca)
  %>% select(clim = b_clim)
  %>% mutate(
    model = "Fixed interaction, country * clim * age"
  )
)

clim_draws_ran_country <- (
  as_draws_df(fit_ran_country)
  %>% select(clim = b_clim)
  %>% mutate(
    model = "Random by country"
  )
)

clim_draws_mund <- (
  as_draws_df(fit_mund)
  %>% select(clim = b_clim)
  %>% mutate(
    model = "Mundlak"
  )
)

clim_draws_sp <- (
  as_draws_df(fit)
  %>% select(clim = b_clim)
  %>% mutate(
    model = "With Spatial Corr."
  )
)

clim_draws <- (
  bind_rows(
    clim_draws_fix_country,
    clim_draws_ranint_country,
    clim_draws_ranint_admin,
    clim_draws_fix_int_cc,
    clim_draws_ran_country, 
    clim_draws_mund,
    clim_draws_sp
  )
  %>% mutate(
    model = factor(
      model, 
      levels = c(
        "Fixed country intercept",
        "Random intercept by country",
        "Random intercept by admin",
        "Fixed interaction",
        "Random by country", 
        "Mundlak",
        "With Spatial Corr."
      )
    )
  )
)
```

```{r}
#| fig-width: 3
#| fig-height: 7
(
  ggplot(
    data = clim_draws,
    mapping = aes(
      x = clim,
    )
  )
  + stat_dotsinterval(
    quantiles = 100,
    slab_color = col_palette[3],
    slab_fill = col_palette[4],
    .width = c(0.5, 0.95),
  )
  + geom_vline(xintercept = 0, linetype = "dashed", color = "gray50")
  + facet_wrap(
    ~model,
    ncol = 1,
    scales = "free_y",
    # strip.position = "left"
  )
  + theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    plot.background = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    axis.title.y = element_blank(),
    axis.title.x = element_text(size = 16),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 15),
    axis.ticks.y = element_blank(),
    legend.background = element_blank(),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 13),
    strip.background = element_blank(),
    # strip.placement = "outside",
    strip.text = element_text(size = 15)
  )
  + labs(
    x = "Posterior Value",
    title = "Model Comparison"
  )
  + scale_y_discrete(
    expand = expansion(mult = c(0.2, 0.75))
  )
  # + scale_x_continuous(
  #   limits = c(-0.1, 1.8)
  # )
)

ggsave(
  path = fig_dir,
  filename = paste0("clim_post_comp_", clim_label, "_", pop_label, "_", viol_label, ".png")
)
```
