---
title: "Heat Analysis"
format: html
editor: visual
---

# Environment Setup

```{r setup}
# Packages
library(tidyverse)
library(sf)
library(brms)
options(brms.backend = "cmdstanr", mc.cores = 4)
library(cmdstanr)
library(posterior)
library(bayesplot)
library(tidybayes)
# theme_set(bayesplot::theme_default(base_family = "sans", base_size=14))
library(patchwork)

# Folder Structure
root <- rprojroot::has_file("agile2025.Rproj")$make_fix_file()
raw_data_dir <- root("data/raw")
proc_data_dir <- root("data/processed")
fig_dir <- root("figures")
dir.create(fig_dir, showWarnings = FALSE)
mod_dir <- root("models")
```

# Notebook settings

```{r}
# Settings
nb_name <-  "Heat analysis and modelling"
seed <- sum(sapply(str_split(nb_name, pattern = "", simplify = TRUE), FUN = utf8ToInt))
set.seed(seed)

# Viz settings
col_palette <- c("#dc7726", "#eebb93", "#ecc046", "#f6e0a3", "#0292b7", "#81c9db", "#35A29F", "#97DECE")
label_map <- c(
  tolerate_women_violence = "Tolerate violence",
  live_with_partner = "Live with partner",
  live_with_father = "Live with father",
  in_school = "In school",
  work = "Has worked",
  rural = "Rural area",
  log_pop_density = "Population density",
  clim = "Climate factor"
)

# Var settings
clim_var <- "value_spei_crop_mean_12m"
covars_fem <- c(
  "country",
  "cluster",
  "adm1_name",
  "rural",
  "log_pop_density",
  "work",
  "live_with_partner",
  "tolerate_women_violence",
  "sv",
  "sv_np"
)
vars_fem <- c(covars_fem, "clim")
covars_ch <- c(
  "country",
  "cluster",
  "adm1_name",
  "rural",
  "log_pop_density",
  "in_school",
  "live_with_partner",
  "live_with_father",
  "tolerate_women_violence",
  "sv",
  "sv_np"
)
vars_ch <- c(covars_ch, "clim")

bin_cov_fem <- c(
  # "rural", 
  "work", 
  # "in_school",
  "live_with_partner", 
  "tolerate_women_violence"
)
bin_cov_ch <- c(
  # "rural",
  "in_school", 
  "live_with_partner",
  "live_with_father", 
  "tolerate_women_violence"
)
cont_covs <- c("log_pop_density", "clim")
```

# Functions

```{r}

# calculate_marginal_values <- function(fit, data, covariate, value) {
#   newdata <- data
#   newdata[[covariate]] <- value
#   
#   epreds <- posterior_epred(fit, newdata = newdata, re_formula = NA)
#   marginal_means <- rowMeans(epreds)
#   
#   tibble(
#     covariate = covariate,
#     covariate_value = as.factor(value),
#     .draw = seq_along(marginal_means),
#     expected_viol = marginal_means
#   )
# }


create_ranef_df <- function(ranef_arr, group_name) {
  df <- as.data.frame.table(ranef_arr, responseName = "value", stringsAsFactors = FALSE)
  colnames(df)[1:3] <- c("iter", "level", "term")
  df$group <- group_name
  df$value <- as.numeric(df$value)
  df
}


calculate_marginal_difference <- function(fit, data, covariate) {
  newdata0 <- data
  newdata1 <- data
  newdata0[[covariate]] <- 0
  newdata1[[covariate]] <- 1
  
  epreds0 <- posterior_epred(fit, newdata = newdata0, re_formula = NA)
  epreds1 <- posterior_epred(fit, newdata = newdata1, re_formula = NA)
  
  diff_means <- rowMeans(epreds1 - epreds0)
  
  tibble(
    covariate = covariate,
    expected_diff = diff_means
  )
}

create_df_marginal_differences <- function(fit, data, covs) {
  (
    lapply(covs, function(cov) {
      calculate_marginal_difference(fit, data, cov)
    })
    %>% bind_rows()
    %>% mutate(
      covariate = label_map[covariate],
    )
  )
}

create_df_marginal_differences_stats <- function(marg_diff) {
  (
    marg_diff
    %>% group_by(covariate)
    %>% summarise(
      median_exp_diff = median(expected_diff),
      mean_exp_diff = mean(expected_diff),
      lower95_exp_diff = quantile(expected_diff, 0.025),
      upper95_exp_diff = quantile(expected_diff, 0.975),
      .groups = "drop"
    )
    %>% arrange(median_exp_diff)
  )
}

# calculate_marginal_grid <- function(fit, data, cov, grid_vals, rural_value) {
#   draws_list <- vector("list", length(grid_vals))
#   for (i in seq_along(grid_vals)) {
#     newdata <- data
#     newdata[[cov]] <- grid_vals[i]
#     newdata[["rural"]] <- rural_value
#     
#     ep <- posterior_epred(fit, newdata = newdata, re_formula = NA)
#     
#     draws_list[[i]] <- rowMeans(ep)
#   }
#   
#   list(
#     covariate_value = grid_vals,
#     expected_viol = draws_list
#   )
# }

calculate_marginal_grid <- function(fit, data, cov, grid_vals) {
  draws_list <- vector("list", length(grid_vals))
  for (i in seq_along(grid_vals)) {
    newdata <- data
    newdata[[cov]] <- grid_vals[i]
    ep <- posterior_epred(fit, newdata = newdata, re_formula = NA)
    draws_list[[i]] <- rowMeans(ep)
  }
  
  list(
    covariate_value = grid_vals,
    expected_viol = draws_list
  )
}

summarize_draws <- function(grid, draws_list, cred = 0.95) {
   df_list <- lapply(seq_along(grid), function(i) {
    draws_i <- draws_list[[i]]
    hdi_i <- bayestestR::hdi(draws_i, ci = cred)
    tibble(
      covariate_value = grid[i],
      median = median(draws_i),
      mean = mean(draws_i),
      lower = hdi_i$CI_low,
      upper = hdi_i$CI_high
    )
  })
  bind_rows(df_list)
}

# plot_continuous_marginal <- function(df_summary, covar_label = NULL) {
#   (
#     ggplot(
#       df_summary %>% filter(covariate == covar_label), 
#       aes(
#         x = covariate_value, 
#         y = median, 
#         color = factor(rural), 
#         fill = factor(rural)
#       )
#     ) 
#     + geom_ribbon(aes(ymin = lower, ymax = upper, fill = factor(rural)), alpha = 0.25, color = NA) 
#     + geom_line(size = 1)
#     + scale_color_manual(
#       name = "Rural", 
#       values = c("0" = col_palette[1], "1" = col_palette[5]),
#       labels = c("0" = "Rural", "1" = "Urban")
#     ) 
#     + scale_fill_manual(
#       name = "Rural", 
#       values = c("0" = col_palette[1], "1" = col_palette[5]),
#       labels = c("0" = "Rural", "1" = "Urban")
#     ) 
#     + labs(
#       x = covar_label %||% unique(df_summary$covariate),
#       y = "Marginal expected P(violence)",
#     ) 
#     + scale_y_continuous(labels = scales::percent_format(accuracy = 1))
#     + theme(
#       panel.background = element_blank(),
#       panel.grid = element_blank(),
#       panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
#       plot.background = element_blank(),
#       plot.title = element_text(size = 16, hjust = 0.5),
#       plot.subtitle = element_text(size = 14, hjust = 0.5),,
#       axis.title = element_text(size = 16),
#       axis.text.y = element_text(size = 15),
#       axis.text.x = element_text(size = 15),
#       legend.background = element_blank(),
#       legend.title = element_blank(),
#       legend.text = element_text(size = 15),
#       legend.position = "inside",
#       legend.position.inside = c(0.15, 0.8)
#     )
#   )
# }

generate_marginal_df <- function(fit, data, covs) {
  results <- list()
  for (cov in covs) {
    rng <- range(data[[cov]], na.rm = TRUE)
    grid_vals <- seq(rng[1], rng[2], length.out = 60)  # 60 grid points
    mg <- calculate_marginal_grid(fit, data, cov, grid_vals)
    s <- (
      summarize_draws(
        mg$covariate_value, 
        mg$expected_viol, 
        cred = 0.95
      )
      %>% mutate(
        covariate = cov
      )
    )
    results[[cov]] <- s
  }

  (
    bind_rows(results) 
    %>% mutate(
      covariate = label_map[covariate],
    )
  )
}


plot_bin_marginal <- function(ppc, ppc_stats, lims = NULL) {
    (
      ggplot(
        data = ppc,
        mapping = aes(
          y = fct_reorder(covariate, expected_diff, .fun = ~ -median(.)),
          x = expected_diff,
        )
      )
      + ggridges::geom_density_ridges(
        aes(height = ..scaled..),
        stat = "density",
        orientation = "x",
        scale = 0.6,
        rel_min_height = 0.01,
        alpha = 0.7,
        color = col_palette[1],
        fill = col_palette[2]
      )
      + geom_segment(
        data = ppc_stats,
        aes(y = covariate, yend = covariate, x = lower95_exp_diff, xend = upper95_exp_diff),
        linewidth = 0.8,
        color = "black",
        inherit.aes = FALSE
      )
      + geom_point(
        data = ppc_stats,
        aes(y = covariate, x = median_exp_diff),
        color = "black",
        size = 2,
        inherit.aes = FALSE
      )
      + geom_vline(xintercept = 0, linetype = "dashed", color = "gray50")
      + theme(
        panel.background = element_blank(),
        panel.grid = element_blank(),
        panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
        plot.background = element_blank(),
        plot.title = element_text(size = 16, hjust = 0.5),
        plot.subtitle = element_text(size = 14, hjust = 0.5),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 16),
        axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        legend.background = element_blank(),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 13),
      )
      + labs(
        x = "Average marginal effect\n(p(violence|1) − p(violence|0))",
      )
      + scale_y_discrete(expand = expansion(mult = c(0.1, 0.35)))
      + scale_x_continuous(
        limits = lims,
        labels = scales::percent_format(accuracy = 1)
      )
    )
}


plot_continuous_marginal <- function(df_summary, covar_label = NULL, lims = NULL) {
  (
    ggplot(
      df_summary %>% filter(covariate == covar_label), 
      aes(
        x = covariate_value, 
        y = median, 
      )
    ) 
    + geom_ribbon(
      aes(ymin = lower, ymax = upper), 
      alpha = 0.25, 
      fill = col_palette[5],
      color = NA
    ) 
    + geom_line(
      size = 1,
      color = col_palette[5]
    )
    + labs(
      x = covar_label,
      y = "Marginal expected\np(violence)",
    ) 
    + scale_y_continuous(
      limits = lims,
      labels = scales::percent_format(accuracy = 1)
    )
    + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),,
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 15),
      legend.position = "inside",
      legend.position.inside = c(0.15, 0.8)
    )
  )
}
```

# Import Data

```{r}
df_clim_viol <- (
  read_csv(
    file.path(
      proc_data_dir,
      "climate_violence_new_vacs.csv"
    ),
    show_col_type = FALSE
  )
  %>% mutate(
    clim := scale(!!sym(clim_var)),
    log_pop_density = scale(log_pop_density)
  )
)

if (str_detect(clim_var, "crop")) {
  df_clim_viol <- (
    df_clim_viol
    %>% filter(
      !is.na(dist_from_last_crop_season)
    )
  )
}
  
```

```{r}
(
  df_clim_viol
  %>% filter(is.na(dist_from_last_crop_season))
)
```

# Prepare Data

```{r}
df_model_fem <- (
  df_clim_viol
  %>% filter(
    female == 1,
  )
  %>% select(all_of(vars_fem))
  %>% na.omit()
)

df_model_ch <- (
  df_clim_viol
  %>% filter(
    female == 1,
    age < 17
  )
  %>% select(all_of(vars_ch))
  %>% na.omit()
)
```

# Modeling

## All Female

### Any violence

```{r}
form_fem_all_viol <- bf(
  sv ~ (
    1 
    + work
    + live_with_partner
    + tolerate_women_violence
    # + rural
    + log_pop_density
    + clim
    + (1 + clim | country/cluster)
    # + (1 + clim | country/adm1_name)
  )
)
fit_fem_all_viol <- brm(
  formula = form_fem_all_viol,
  family = bernoulli("logit"),
  prior = c(
    prior(student_t(6, -4, 1), class = "Intercept"),
    prior(normal(0, 0.2), class = "b"),
    prior(normal(0, 0.2), class = "sd")
  ),
  data = df_model_fem,
  seed = seed,
  # sample_prior = "only"
)
```

#### Prior Predictive

```{r}
pp_check(fit_fem_all_viol)
```

#### Model Output

```{r}
fit_fem_all_viol
```

#### Random Effects

```{r}
ranefs_fem_all_viol <- ranef(fit_fem_all_viol, summary = FALSE)
ranef_fem_all_viol_names <- names(ranefs_fem_all_viol)
```

```{r}
df_ranefs_country_fem_all_viol <- create_ranef_df(ranefs_fem_all_viol[["country"]], group_name = "country")
df_ranefs_cluster_fem_all_viol <- create_ranef_df(ranefs_fem_all_viol[["country:cluster"]], group_name = "cluster")
df_ranefs_groups_fem_all_viol <- bind_rows(df_ranefs_country_fem_all_viol, df_ranefs_cluster_fem_all_viol)
```

```{r}
df_ranef_fem_all_viol_summary <- (
  df_ranefs_groups_fem_all_viol
  %>% group_by(group, level, term) 
  %>% summarise(
    median = median(value),
    lower = quantile(value, probs = 0.025),
    upper = quantile(value, probs = 0.975),
    .groups = "drop"
  )
  %>% mutate(
    level_label = paste0(level, " (", group, ")")
  )
  %>% group_by(term)
  %>% arrange(median)
  %>% mutate(level_order = row_number())
  %>% ungroup()
)
```

```{r}
(
  ggplot(
    df_ranef_fem_all_viol_summary, 
    aes(
      x = median, 
      y = fct_reorder(level_label, median)
    )
  )
  + geom_point(position = position_dodge(width = 0.7), size = 2) 
  + geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0, position = position_dodge(width = 0.7))
  + facet_wrap(~ term + group, scales = "free_y", ncol = 2)  
  + labs(
    x = "Effect (posterior median and 95% CrI)",
    y = NULL,
  ) 
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      strip.background = element_blank(),
      strip.text = element_text(size = 16),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.ticks.y = element_blank(),
      axis.title.y = element_blank(),
      axis.title.x = element_text(size = 16),
      axis.text.y = element_blank(),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_text(size = 14),
      legend.text = element_text(size = 13),
    )
)

```

#### Posterior Predictive Checks

##### Marginal Average Effect for Categorical variables

```{r}
# marginal_posteriors <- (
#   lapply(bin_cov, function(cov) {
#     bind_rows(
#       marginal_for_covariate_value(fit_fem_all_viol, df_model_fem, cov, 0),
#       marginal_for_covariate_value(fit_fem_all_viol, df_model_fem, cov, 1)
#     )
#   })
#   %>% bind_rows()
#   %>% mutate(
#     covariate = factor(covariate, levels= bin_cov),
#     covariate_value = factor(covariate_value, levels = c("0", "1"))
#   )
# )
```

```{r}
expected_diffs_fem_all_viol <- create_df_marginal_differences(fit_fem_all_viol, df_model_fem, bin_cov_fem)
expected_diffs_fem_all_viol_stats <- create_df_marginal_differences_stats(expected_diffs_fem_all_viol)
```

```{r}
#| fig-width: 7
#| fig-height: 3

plot_bin_marginal(expected_diffs_fem_all_viol, expected_diffs_fem_all_viol_stats, lims = c(-0.09, 0.09))
ggsave(
  path = fig_dir,
  filename = "marginal_predicted_prob_bin_fem_all_viol.png"
)
```

##### Marginal Effect of Continuous Variables

```{r}
df_marg_cont_fem_all_viol <- generate_marginal_df(fit_fem_all_viol, df_model_fem, cont_covs)
```

```{r}
#| fig-width: 7
#| fig-height: 3
plot_continuous_marginal(df_marg_cont_fem_all_viol, covar_label = "Population density", lims = c(0.01, 0.25))

ggsave(
  path = fig_dir,
  filename = "marginal_predicted_prob_log_density_ur_fem_all_viol.png"
)
```

```{r}
#| fig-width: 7
#| fig-height: 3
plot_continuous_marginal(df_marg_cont_fem_all_viol, covar_label = "Climate factor", lims = c(0.02, 0.5))

ggsave(
  path = fig_dir,
  filename = "marginal_predicted_prob_climate_ur_fem_all_viol.png"
)
```

### Non-partner violence

```{r}
form_fem_np_viol <- bf(
  sv_np ~ (
    1 
    + work
    + live_with_partner
    + tolerate_women_violence
    # + rural
    + log_pop_density
    + clim
    # + (1 + clim | country/rural/cluster)
    + (1 + clim | country/cluster)
    # + (1 + clim | country/adm1_name)
  )
)
fit_fem_np_viol <- brm(
  formula = form_fem_np_viol,
  family = bernoulli("logit"),
  prior = c(
    prior(student_t(6, -4, 1), class = "Intercept"),
    prior(normal(0, 0.2), class = "b"),
    prior(normal(0, 0.2), class = "sd")
  ),
  data = df_model_fem,
  seed = seed,
  # sample_prior = "only"
)
```

#### Prior Predictive

```{r}
pp_check(fit_fem_np_viol)
```

#### Model Output

```{r}
fit_fem_np_viol
```

#### Posterior Predictive Checks

##### Marginal Average Effect for Categorical variables

```{r}
expected_diffs_fem_np_viol <- create_df_marginal_differences(fit_fem_np_viol, df_model_fem, bin_cov_fem)
expected_diffs_fem_np_viol_stats <- create_df_marginal_differences_stats(expected_diffs_fem_np_viol)
```

```{r}
#| fig-width: 7
#| fig-height: 3

plot_bin_marginal(expected_diffs_fem_np_viol, expected_diffs_fem_np_viol_stats, lims = c(-0.09, 0.09))
ggsave(
  path = fig_dir,
  filename = "marginal_predicted_prob_bin_fem_np_viol.png"
)
```

##### Marginal Effect of Continuous Variables

```{r}
df_marg_cont_fem_np_viol <- generate_marginal_df(fit_fem_np_viol, df_model_fem, cont_covs)
```

```{r}
#| fig-width: 7
#| fig-height: 3
plot_continuous_marginal(df_marg_cont_fem_np_viol, covar_label = "Population density", lims = c(0.01, 0.20))

ggsave(
  path = fig_dir,
  filename = "marginal_predicted_prob_log_density_ur_fem_np_viol.png"
)
```

```{r}
#| fig-width: 7
#| fig-height: 3
plot_continuous_marginal(df_marg_cont_fem_np_viol, covar_label = "Climate factor", lims = c(0.01, 0.4))

ggsave(
  path = fig_dir,
  filename = "marginal_predicted_prob_climate_ur_fem_np_viol.png"
)
```

## Children

### Any violence

```{r}
form_ch_all_viol <- bf(
  sv ~ (
    1 
    + in_school
    + live_with_father
    + live_with_partner
    + tolerate_women_violence
    # + rural
    + log_pop_density
    + clim
    + (1 + clim | country/cluster)
    # + (1 + clim | country/adm1_name)
  )
)
fit_ch_all_viol <- brm(
  formula = form_ch_all_viol,
  family = bernoulli("logit"),
  prior = c(
    prior(student_t(6, -4, 1), class = "Intercept"),
    prior(normal(0, 0.2), class = "b"),
    prior(normal(0, 0.2), class = "sd")
  ),
  data = df_model_ch,
  seed = seed,
  # sample_prior = "only"
)
```

#### Prior Predictive

```{r}
pp_check(fit_ch_all_viol)
```

#### Model Output

```{r}
fit_ch_all_viol
```

#### Posterior Predictive Checks

##### Marginal Average Effect for Categorical variables

```{r}
expected_diffs_ch_all_viol <- create_df_marginal_differences(fit_ch_all_viol, df_model_ch, bin_cov_ch)
expected_diffs_ch_all_viol_stats <- create_df_marginal_differences_stats(expected_diffs_ch_all_viol)
```

```{r}
#| fig-width: 7
#| fig-height: 3

plot_bin_marginal(expected_diffs_ch_all_viol, expected_diffs_ch_all_viol_stats, lims = c(-0.09, 0.09))
ggsave(
  path = fig_dir,
  filename = "marginal_predicted_prob_bin_ch_all_viol.png"
)
```

##### Marginal Effect of Continuous Variables

```{r}
df_marg_cont_ch_all_viol <- generate_marginal_df(fit_ch_all_viol, df_model_ch, cont_covs)
```

```{r}
#| fig-width: 7
#| fig-height: 3
plot_continuous_marginal(df_marg_cont_ch_all_viol, covar_label = "Population density")

ggsave(
  path = fig_dir,
  filename = "marginal_predicted_prob_log_density_ur_ch_all_viol.png"
)
```

```{r}
#| fig-width: 7
#| fig-height: 3
plot_continuous_marginal(df_marg_cont_ch_all_viol, covar_label = "Climate factor")

ggsave(
  path = fig_dir,
  filename = "marginal_predicted_prob_climate_ur_ch_all_viol.png"
)
```

### Non-partner violence

```{r}
form_ch_np_viol <- bf(
  sv_np ~ (
    1 
    + in_school
    + live_with_father
    + live_with_partner
    + tolerate_women_violence
    # + rural
    + log_pop_density
    + clim
    + (1 + clim | country/cluster)
    # + (1 + clim | country/adm1_name)
  )
)
fit_ch_np_viol <- brm(
  formula = form_ch_np_viol,
  family = bernoulli("logit"),
  prior = c(
    prior(student_t(6, -4, 1), class = "Intercept"),
    prior(normal(0, 0.2), class = "b"),
    prior(normal(0, 0.2), class = "sd")
  ),
  data = df_model_ch,
  seed = seed,
  # sample_prior = "only"
)
```

#### Prior Predictive

```{r}
pp_check(fit_ch_np_viol)
```

#### Model Output

```{r}
fit_ch_np_viol
```

#### Posterior Predictive Checks

##### Marginal Average Effect for Categorical variables

```{r}
expected_diffs_ch_np_viol <- create_df_marginal_differences(fit_ch_np_viol, df_model_ch, bin_cov_ch)
expected_diffs_ch_np_viol_stats <- create_df_marginal_differences_stats(expected_diffs_ch_np_viol)
```

```{r}
#| fig-width: 7
#| fig-height: 3

plot_bin_marginal(expected_diffs_ch_np_viol, expected_diffs_ch_np_viol_stats, lims = c(-0.09, 0.09))
ggsave(
  path = fig_dir,
  filename = "marginal_predicted_prob_bin_ch_np_viol.png"
)
```

##### Marginal Effect of Continuous Variables

```{r}
df_marg_cont_ch_np_viol <- generate_marginal_df(fit_ch_np_viol, df_model_ch, cont_covs)
```

```{r}
#| fig-width: 7
#| fig-height: 3
plot_continuous_marginal(df_marg_cont_ch_np_viol, covar_label = "Population density")

ggsave(
  path = fig_dir,
  filename = "marginal_predicted_prob_log_density_ur_ch_np_viol.png"
)
```

```{r}
#| fig-width: 7
#| fig-height: 3
plot_continuous_marginal(df_marg_cont_ch_np_viol, covar_label = "Climate factor")

ggsave(
  path = fig_dir,
  filename = "marginal_predicted_prob_climate_ur_ch_np_viol.png"
)
```
