---
title: "Create SPEI dataset at admin 2 level"
format: html
editor: visual
---

# Environment Setup

```{r setup}
# Packages
library(tidyverse)
library(tidyterra)
library(sf)
library(paletteer)
library(patchwork)

# Path Structure
root <- rprojroot::has_file("agile2025.Rproj")$make_fix_file()
raw_data_dir <- root("data/raw")
proc_data_dir <- root("data/processed/climate/spei")
dir.create(proc_data_dir, showWarnings = FALSE)
fig_dir <- root("figures")
```

# Parameters

```{r}
coords <- list(
  "ZWE" = c(25.2, 33.1, -22.5, -15.45),
  "ZMB" = c(21.95, 33.73, -18.1, -8.15)
)

# Country Mapping
country_map <- sort(c(
  "ZMB" = "Zambia",
  "ZWE" = "Zimbabwe"
))
```

# Functions

```{r}
build_spei <- function(rast_spei, adm_shapes, start_date) {
  print(paste0("Starting SPEI data processing"))
  rast_times <- terra::time(rast_spei)
  ids <- which(rast_times >= start_date)
  rast_spei <- rast_spei %>% terra::subset(ids)
  dates <- as.character(terra::time(rast_spei))
  df_spei <- (
      rast_spei
      %>% exactextractr::exact_extract(
        adm_shapes,
        fun = "weighted_mean",
        weights = "area",
        max_cells_in_memory = 176067000,
        append_cols = c("country", "adm_code", "adm_name"),
        progress = FALSE
      )
    )
  colnames(df_spei) <- c("country", "adm_code", "adm_name", dates)
  df_spei <- (
    df_spei
    %>% pivot_longer(
      cols = -c("country", "adm_code", "adm_name"),
      names_to = "date",
      values_to = "spei"
    )
    %>% mutate(
      date = as.Date(date, format = "%Y-%m-%d")
    )
  )
}

build_weighted_spei_country <- function(
    iso3, 
    rast_spei, 
    rast_pop, 
    adm_pop,
    adm_shapes, 
    start_date, 
    age_disag = FALSE
  ) {
  print(paste0("Starting processing SPEI data for: ", iso3))
  if (age_disag) {
    file_name = "age_weighted_spei_"
  } else {
    file_name = "weighted_spei_"
  }
  dfs_spei_times <- list()
  rast_times <- terra::time(rast_spei)
  ids <- which(rast_times >= start_date)
  start_id <- ids[1]
  rast_times <- rast_times[ids]
  thresholds <- round(quantile(seq_along(rast_times), probs = c(0, 0.25, 0.5, 0.75, 1)))
  for (i in seq_along(rast_times)) {
    if (i %in% thresholds) {
      thresh_idx <- which(thresholds == i)
      print(paste0(names(thresh_idx), " complete"))
    }
    rast_spei_time  <- (
      rast_spei[[paste0("spei_", start_id + i - 1)]]
      %>% terra::crop(terra::ext(unlist(coords[iso3])))
    )
    rast_spei_time_fine <- terra::resample(rast_spei_time, rast_pop)
    rast_spei_weighted <- rast_spei_time_fine * rast_pop
    dfs_spei_times[[i]] <- (
      rast_spei_weighted
      %>% exactextractr::exact_extract(
        adm_shapes,
        fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
        max_cells_in_memory = 176067000,
        append_cols = c("adm_code"),
        progress = FALSE
      )
      %>% merge(
        adm_pop,
        by = "adm_code",
        all.x = TRUE
      )
      %>% mutate(
        date = as.Date(rast_times[[i]], format = "%Y-%m-%d"),
        spei = (result / pop)
      )
    )
  }
  df_spei <- (
    bind_rows(dfs_spei_times)
    %>% select(-c(pop, result))
    %>% mutate(
      country = country_map[[iso3]]
    )
    %>% relocate(country)
  )
  df_spei %>% write_csv(file.path(proc_data_dir, paste0(file_name, iso3, ".csv")))
  df_spei
}
```

# Data Import

## Geographies

### Zimbabwe

```{r}
sf_zwe <- (
  st_read(
    file.path(raw_data_dir, "geographies", "zwe_admin_boundaries", "zwe_admin2.geojson")
  )
  %>% rename_with(tolower)
  %>% mutate(
    country = "Zimbabwe",
    adm_code = as.numeric(gsub("ZW1", "", adm2_pcode))
  )
  %>% select(
    country,
    adm_name = adm2_name,
    adm_code
  )
  %>% arrange(adm_name)
)
```

### Zambia

```{r}
sf_zmb <- (
  st_read(
    file.path(raw_data_dir, "geographies", "zmb_admin_boundaries", "zmb_admin2_em.shp")
  )
  %>% rename_with(tolower)
  %>% mutate(
    country = "Zambia",
    adm_code = as.numeric(gsub("ZM1", "", adm2_pcode))
  )
  %>% select(
    country,
    adm_name = adm2_name,
    adm_code
  )
  %>% arrange(adm_name)
)
```

### Create combined shapes

```{r}
sf_all <- (
  bind_rows(
    sf_zwe,
    sf_zmb
  )
)
```

## All Population

### Zimbabwe

```{r}
rast_pop_zwe <- (
  terra::rast(  
    file.path(raw_data_dir, "population", "zwe_pop_2017_CN_1km_R2025A_UA_v1.tif")
  )
  %>% rename(
    pop = global_pop_2017_CN_1km_R2025A_v1
  )
)
```

```{r}
adm_pop_zwe <- (
  rast_pop_zwe
  %>% exactextractr::exact_extract(
    sf_zwe,
    fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
    max_cells_in_memory = 176067000,
    append_cols = c("adm_code", "adm_name")
  )
  %>% rename(
    pop = result
  )
)
```

### Zambia

```{r}
rast_pop_zmb <- (
  terra::rast(
    file.path(raw_data_dir, "population", "zmb_pop_2015_CN_1km_R2025A_UA_v1.tif")
  )
  %>% rename(
    pop = global_pop_2015_CN_1km_R2025A_v1
  )
)
```

```{r}
adm_pop_zmb <- (
  rast_pop_zmb
  %>% exactextractr::exact_extract(
    sf_zmb,
    fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
    max_cells_in_memory = 176067000,
    append_cols = c("adm_code", "adm_name")
  )
  %>% rename(
    pop = result
  )
)
```

## 10-19 Female Population

For each country we consider the population at the year closest to that of the survey.

```{r}
# Age brackets to consider
ages <- c(10, 15)
```

### Zimbabwe

```{r}
rast_pops <- list()
for (i in seq_along(ages)) {
  rast_pops[[i]] <- (
    terra::rast(
      file.path(
        raw_data_dir, 
        "population", 
        "disaggregated",
        "zwe_agesex_structures_2017_CN_1km_R2025A_UA_v1",
        sprintf("zwe_f_%02d_2017_CN_1km_R2025A_UA_v1.tif", ages[[i]])
      )
    )
    %>% rename(
      pop = sprintf("global_f_%02d_2017_CN_1km_R2025A_v1", ages[[i]])
    )
  )
}

rast_popf_10_19_zwe <- (
  sum(terra::rast(rast_pops))
  %>% rename(
    pop = sum
  )
)
```

```{r}
adm_popf_10_19_zwe <- (
  rast_popf_10_19_zwe
  %>% exactextractr::exact_extract(
    sf_zwe,
    fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
    max_cells_in_memory = 176067000,
    append_cols = c("adm_code", "adm_name")
  )
  %>% rename(
    pop = result
  )
)
```

### Zambia

```{r}
rast_pops <- list()
for (i in seq_along(ages)) {
  rast_pops[[i]] <- (
    terra::rast(
      file.path(
        raw_data_dir, 
        "population", 
        "disaggregated",
        "zmb_agesex_structures_2015_CN_1km_R2025A_UA_v1",
        sprintf("zmb_f_%02d_2015_CN_1km_R2025A_UA_v1.tif", ages[[i]])
      )
    )
    %>% rename(
      pop = sprintf("global_f_%02d_2015_CN_1km_R2025A_v1", ages[[i]])
    )
  )
}

rast_popf_10_19_zmb <- (
  sum(terra::rast(rast_pops))
  %>% rename(
    pop = sum
  )
)
```

```{r}
adm_popf_10_19_zmb <- (
  rast_popf_10_19_zmb
  %>% exactextractr::exact_extract(
    sf_zmb,
    fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
    max_cells_in_memory = 176067000,
    append_cols = c("adm_code", "adm_name")
  )
  %>% rename(
    pop = result
  )
)
```

## SPEI

Data obtained using the Climate Data Store API

```{r}
rast_spei <- terra::rast(
  file.path(
    raw_data_dir, 
    "climate/spei",
    "spei01.nc"
  )
)
n_layers <- terra::nlyr(rast_spei)
```

# Method Outline

```{r}
# For plotting
xlims <- c(25.22, 33.1)
ylims <- c(-22.46, -15.59)
plot_extent <- terra::ext(coords[["ZWE"]] + c(-0.5, 0.5, -0.5, 0.5))
date <- terra::time(rast_spei)[[n_layers]]
```

## SPEI in country

```{r}
rast_spei_zwe  <- (
  rast_spei[[paste0("spei_", n_layers)]]
  %>% terra::crop(plot_extent)
)
```

```{r}
#| fig-width: 5
#| fig-height: 4
(
  ggplot()
  + geom_spatraster(
    data = rast_spei_zwe,
    mapping = aes(fill = !!sym(paste0("spei_", n_layers)))
  )
  + scale_fill_paletteer_c(
    "scico::vik",
    na.value = "transparent",
  )
  + geom_sf(
    data = sf_zwe,
    fill = NA,
    color = "black",
    # linewidth = 1.1
  )
  + theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.key.height = unit(50, "pt"),
    legend.title = element_text(size = 18, angle = 90, hjust = 0.5),
    legend.text = element_text(size = 16),
    legend.background = element_blank(),
    legend.title.position = "right",
  )
  + labs(
    fill = "SPEI"
  )
)
ggsave(
  path = fig_dir,
  filename = "spei_zwe_adm2.png",
  width = 5,
  height = 4
)
```

## Downsample one time step

```{r}
rast_spei_fine_zwe <- terra::resample(rast_spei_zwe, rast_pop_zwe)
```

```{r}
#| fig-width: 5
#| fig-height: 4
(
  ggplot()
  + geom_spatraster(
    data = rast_spei_fine_zwe %>% terra::crop(plot_extent),
    mapping = aes(fill = !!sym(paste0("spei_", n_layers)))
  )
  + scale_fill_paletteer_c(
    "scico::vik",
    na.value = "transparent",
  )
  + geom_sf(
    data = sf_zwe,
    fill = NA,
    color = "black",
  )
  + theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.key.height = unit(50, "pt"),
    legend.title = element_text(size = 18, angle = 90, hjust = 0.5),
    legend.text = element_text(size = 16),
    legend.background = element_blank(),
    legend.title.position = "right",
  )
  + labs(
    fill = "SPEI"
  )
)
ggsave(
  path = fig_dir,
  filename = "spei_fine_zwe_adm2.png",
  width = 5,
  height = 4
)
```

## Population scaled UTCI

```{r}
rast_spei_weighted_zwe <- rast_spei_fine_zwe * rast_pop_zwe
```

```{r}
#| fig-width: 8
#| fig-height: 4

p1 <- (
  ggplot()
  + geom_spatraster(
    data = rast_pop_zwe %>% terra::crop(plot_extent),
    mapping = aes(fill = pop)
  )
  + scale_fill_paletteer_c(
    "scico::devon",
    na.value = "transparent",
    direction = -1
  )
  + geom_sf(
    data = sf_zwe,
    fill = NA,
    color = "black",
  )
  + theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.key.height = unit(50, "pt"),
    legend.title = element_text(size = 18, angle = 90, hjust = 0.5),
    legend.text = element_text(size = 16),
    legend.background = element_blank(),
    legend.title.position = "right",
  )
  # + coord_sf(
  #   xlim = xlims,
  #   ylim = ylims
  # )
  + labs(
    fill = "Population"
  )
)

p2 <- (
  ggplot()
  + geom_spatraster(
    data = rast_spei_weighted_zwe %>% terra::crop(plot_extent),
    mapping = aes(fill = !!sym(paste0("spei_", n_layers)))
  )
  + scale_fill_paletteer_c(
    "scico::vik",
    na.value = "transparent",
  )
  + geom_sf(
    data = sf_zwe,
    fill = NA,
    color = "black",
  )
  + theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.key.height = unit(50, "pt"),
    legend.title = element_text(size = 18, angle = 90, hjust = 0.5),
    legend.text = element_text(size = 16),
    legend.background = element_blank(),
    legend.title.position = "right",
  )
  + coord_sf(
    xlim = xlims,
    ylim = ylims
  )
  + labs(
    fill = "SPEI * Population"
  )
)

p1 + p2

ggsave(
  path = fig_dir,
  filename = "spei_pop_zwe_adm2.png",
  width = 6.5,
  height = 4
)
```

## Weighted SPEI

```{r}
df_spei_zwe <- (
  rast_spei_weighted_zwe
  %>% exactextractr::exact_extract(
    sf_zwe,
    fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
    max_cells_in_memory = 176067000,
    append_cols = c("adm_code"),
    progress = FALSE
  )
  %>% merge(
    adm_pop_zwe,
    by = "adm_code",
    all.x = TRUE
  )
  %>% mutate(
    date = as.Date(date, format = "%Y-%m-%d"),
    spei = (result / pop)
  )
  %>% select(-result)
)
```

```{r}
sf_spei_zwe <- (
  sf_zwe
  %>% merge(
    df_spei_zwe,
    by = c("adm_code", "adm_name"),
    all.x = TRUE
  )
)
```

```{r}
#| fig-width: 4
#| fig-height: 5
(
  ggplot()
  + geom_sf(
    data = sf_spei_zwe %>% filter(adm_name != "Maputo City"),
    mapping = aes(fill = spei),
    color = "black",
  )
  + theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5, margin = margin(b = 0)),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.key.height = unit(50, "pt"),
    legend.title = element_text(size = 16, angle = 90, hjust = 0.5),
    legend.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.title.position = "right",
  )
  + scale_fill_paletteer_c(
    "scico::vik",
    na.value = "transparent",
  )
  + labs(
    title = "Weighted SPEI Index",
    fill = "SPEI"
  )
)
ggsave(
  path = fig_dir,
  filename = "spei_weighted_zwe_adm2.png",
  width = 4,
  height = 5
)
```

# Generate SPEI dataset

```{r}
df_spei <- build_spei(rast_spei, sf_all, "2000-01-01")
```

# Generate Population Weighted UTCI datasets

## Zimbabwe

```{r}
df_weighted_spei_zwe <- build_weighted_spei_country("ZWE", rast_spei, rast_pop_zwe, adm_pop_zwe, sf_zwe, "2000-01-01")
```

## Zambia

```{r}
df_weighted_spei_zmb <- build_weighted_spei_country("ZMB", rast_spei, rast_pop_zmb, adm_pop_zmb, sf_zmb, "2000-01-01")
```

# Generate Population Weighted UTCI datasets

## Zimbabwe

```{r}
df_age_weighted_spei_zwe <- build_weighted_spei_country("ZWE", rast_spei, rast_popf_10_19_zwe, adm_popf_10_19_zwe, sf_zwe, "2000-01-01", age_disag = TRUE)
```

## Zambia

```{r}
df_age_weighted_spei_zmb <- build_weighted_spei_country("ZMB", rast_spei, rast_popf_10_19_zmb, adm_popf_10_19_zmb, sf_zmb, "2000-01-01", age_disag = TRUE)
```

# Save

```{r}
df_spei %>% write_csv(file.path(proc_data_dir, "spei_monthly_all_countries_admin2.csv"))
```

```{r}
df_weighted_spei <- bind_rows(
  df_weighted_spei_zwe,
  df_weighted_spei_zmb
)
 
df_weighted_spei %>% write_csv(file.path(proc_data_dir, "weighted_spei_monthly_all_countries_admin2.csv"))
```

```{r}
df_age_weighted_spei <- bind_rows(
  df_age_weighted_spei_zwe,
  df_age_weighted_spei_zmb
)
 
df_age_weighted_spei %>% write_csv(file.path(proc_data_dir, "age_weighted_spei_monthly_all_countries_admin2.csv"))
```
