---
title: "temp Analysis"
format: html
editor: visual
---

# Environment Setup

```{r setup}
# Packages
library(tidyverse)
library(tidyterra)
library(sf)
library(paletteer)
library(patchwork)

# Path Structure
root <- rprojroot::has_file("agile2025.Rproj")$make_fix_file()
raw_data_dir <- root("data/raw")
proc_data_dir <- root("data/processed/temp_2m")
dir.create(proc_data_dir, showWarnings = FALSE)
fig_dir <- root("figures")

# Country Mapping
country_map <- c(
  "ZWE" = "Zimbabwe",
  "ZMB" = "Zambia"
)
```

# Functions

```{r}
build_temp_daily <- function(rast_temp, adm_shapes) {
  df_temp <- (
    rast_temp
    %>% exactextractr::exact_extract(
      adm_shapes,
      fun = "weighted_mean",
      weights = "area",
      max_cells_in_memory = 176067000,
      append_cols = c("country", "adm_code", "adm_name")
    )
  )
  colnames(df_temp) <- c("country", "adm_code", "adm_name", as.character(terra::time(rast_temp)))
  
  df_temp <- (
    df_temp
    %>% pivot_longer(
      cols = -c("country", "adm_code", "adm_name"),
      names_to = "date",
      values_to = "temp_2m_max"
    )
    %>% mutate(
      date = as.Date(date, format = "%Y-%m-%d"),
      temp_2m_max = temp_2m_max - 273.15
    )
  )
  df_temp
}
  
build_weighted_temp_daily_country <- function(iso3, country_map, rast_temp, rast_pop, adm_pop, adm_shapes) {
  print(paste0("Starting processing temperature data for: ", iso3))
  rast_temp_country <- rast_temp %>% terra::crop(terra::vect(adm_shapes %>% st_make_valid() %>% st_union()))
  dfs_temp_daily <- list()
  times <- terra::time(rast_temp_country)
  thresholds <- round(quantile(seq_along(times), probs = c(0, 0.25, 0.5, 0.75, 1)))
  for (i in seq_along(times)) {
    if (i %in% thresholds) {
      thresh_idx <- which(thresholds == i)
      print(paste0(names(thresh_idx), " complete"))
    }
    rast_temp_fine <- terra::resample(rast_temp[[i]], rast_pop)
    rast_temp_weighted <- rast_temp_fine * rast_pop
    dfs_temp_daily[[i]] <- (
      rast_temp_weighted
      %>% exactextractr::exact_extract(
        adm_shapes,
        fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
        max_cells_in_memory = 176067000,
        append_cols = c("country", "adm_code"),
        progress = FALSE
      )
      %>% merge(
        adm_pop,
        by = "adm_code",
        all.x = TRUE
      )
      %>% mutate(
        date = as.Date(times[i], format = "%Y-%m-%d"),
        temp_2m_max = (result / pop) - 273.15
      )
    )
  }
  df_temp <- (
    bind_rows(dfs_temp_daily)
    %>% select(-c(pop, result))
    %>% mutate(
      country = country_map[[iso3]]
    )
    %>% relocate(country)
  )
  df_temp %>% write_csv(file.path(proc_data_dir, paste0("temp_2m_max_daily_", iso3, ".csv")))
  df_temp
}
```

# Data Import

## Geographies

### Zimbabwe

```{r}
sf_zwe <- (
  st_read(
    file.path(raw_data_dir, "geographies", "zwe_admin_boundaries", "zwe_admin2.geojson")
  )
  %>% rename_with(tolower)
  %>% mutate(
    country = "Zimbabwe",
    adm_code = as.numeric(gsub("ZW1", "", adm2_pcode))
  )
  %>% select(
    country,
    adm_name = adm2_name,
    adm_code
  )
  %>% arrange(adm_name)
)
```

### Zambia

```{r}
sf_zmb <- (
  st_read(
    file.path(raw_data_dir, "geographies", "zmb_admin_boundaries", "zmb_admin2_em.shp")
  )
  %>% rename_with(tolower)
  %>% mutate(
    country = "Zambia",
    adm_code = as.numeric(gsub("ZM1", "", adm2_pcode))
  )
  %>% select(
    country,
    adm_name = adm2_name,
    adm_code
  )
  %>% arrange(adm_name)
)
```

### Create combined shapes

```{r}
sf_all <- (
  bind_rows(
    sf_zwe,
    sf_zmb
  )
)
```

## Population

### Zimbabwe

```{r}
rast_pop_zwe <- (
  terra::rast(  
    file.path(raw_data_dir, "population", "zwe_pop_2017_CN_1km_R2025A_UA_v1.tif")
  )
  %>% rename(
    pop = global_pop_2017_CN_1km_R2025A_v1
  )
)
```

```{r}
adm_pop_zwe <- (
  rast_pop_zwe
  %>% exactextractr::exact_extract(
    sf_zwe,
    fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
    max_cells_in_memory = 176067000,
    append_cols = c("adm_code", "adm_name")
  )
  %>% rename(
    pop = result
  )
)
```

### Zambia

```{r}
rast_pop_zmb <- (
  terra::rast(
    file.path(raw_data_dir, "population", "zmb_pop_2015_CN_1km_R2025A_UA_v1.tif")
  )
  %>% rename(
    pop = global_pop_2015_CN_1km_R2025A_v1
  )
)
```

```{r}
adm_pop_zmb <- (
  rast_pop_zmb
  %>% exactextractr::exact_extract(
    sf_zmb,
    fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
    max_cells_in_memory = 176067000,
    append_cols = c("adm_code", "adm_name")
  )
  %>% rename(
    pop = result
  )
)
```

## 10-19 Female Population

For each country we consider the population at the year closest to that of the survey.

```{r}
# Age brackets to consider
ages <- c(10, 15)
```

### Zimbabwe

```{r}
rast_pops <- list()
for (i in seq_along(ages)) {
  rast_pops[[i]] <- (
    terra::rast(
      file.path(
        raw_data_dir, 
        "population", 
        "disaggregated",
        "zwe_agesex_structures_2017_CN_1km_R2025A_UA_v1",
        sprintf("zwe_f_%02d_2017_CN_1km_R2025A_UA_v1.tif", ages[[i]])
      )
    )
    %>% rename(
      pop = sprintf("global_f_%02d_2017_CN_1km_R2025A_v1", ages[[i]])
    )
  )
}

rast_popf_10_19_zwe <- (
  sum(terra::rast(rast_pops))
  %>% rename(
    pop = sum
  )
)
```

```{r}
adm_popf_10_19_zwe <- (
  rast_popf_10_19_zwe
  %>% exactextractr::exact_extract(
    sf_zwe,
    fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
    max_cells_in_memory = 176067000,
    append_cols = c("adm_code", "adm_name")
  )
  %>% rename(
    pop = result
  )
)
```

### Zambia

```{r}
rast_pops <- list()
for (i in seq_along(ages)) {
  rast_pops[[i]] <- (
    terra::rast(
      file.path(
        raw_data_dir, 
        "population", 
        "disaggregated",
        "zmb_agesex_structures_2015_CN_1km_R2025A_UA_v1",
        sprintf("zmb_f_%02d_2015_CN_1km_R2025A_UA_v1.tif", ages[[i]])
      )
    )
    %>% rename(
      pop = sprintf("global_f_%02d_2015_CN_1km_R2025A_v1", ages[[i]])
    )
  )
}

rast_popf_10_19_zmb <- (
  sum(terra::rast(rast_pops))
  %>% rename(
    pop = sum
  )
)
```

```{r}
adm_popf_10_19_zmb <- (
  rast_popf_10_19_zmb
  %>% exactextractr::exact_extract(
    sf_zmb,
    fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
    max_cells_in_memory = 176067000,
    append_cols = c("adm_code", "adm_name")
  )
  %>% rename(
    pop = result
  )
)
```

## Temperature

Data from ERA5 received from Neil

```{r}
rast_temp <- terra::rast(
  file.path(
    raw_data_dir, 
    "temp_2m",
    "era5_t2m_max_2000-2022.nc"
  )
)
```

# Method Outline

```{r}
# For plotting
xlims <- c(25.22, 33.1)
ylims <- c(-22.46, -15.59)
plot_extent <- terra::ext(xlims[1] - 0.5, xlims[2] + 0.5, ylims[1] - 0.5, ylims[2] + 0.5)
```

## Temperature in country

```{r}
rast_temp_zwe <- rast_temp %>% terra::crop(terra::vect(sf_zwe %>% st_union()))
rast_times <- terra::time(rast_temp_zwe)
```

```{r}
#| fig-width: 5
#| fig-height: 4
(
  ggplot()
  + geom_spatraster(
    data = rast_temp_zwe[[1]],
    mapping = aes(fill = t2m_max_1)
  )
  + scale_fill_paletteer_c(
    "scico::vik",
    na.value = "transparent",
  )
  + geom_sf(
    data = sf_zwe,
    fill = NA,
    color = "black",
    # linewidth = 1.1
  )
  + theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.key.height = unit(50, "pt"),
    legend.title = element_text(size = 18, angle = 90, hjust = 0.5),
    legend.text = element_text(size = 16),
    legend.background = element_blank(),
    legend.title.position = "right",
  )
  + labs(
    fill = "Temperature"
  )
)
ggsave(
  path = fig_dir,
  filename = "temp_grid_zwe.png",
  width = 5,
  height = 4
)
```

## Downsample one time step

```{r}
rast_temp_fine_zwe <- terra::resample(rast_temp_zwe[[1]], rast_pop_zwe)
```

```{r}
#| fig-width: 5
#| fig-height: 4
(
  ggplot()
  + geom_spatraster(
    data = rast_temp_fine_zwe %>% terra::crop(plot_extent),
    mapping = aes(fill = t2m_max_1)
  )
  + scale_fill_paletteer_c(
    "scico::vik",
    na.value = "transparent",
  )
  + geom_sf(
    data = sf_zwe,
    fill = NA,
    color = "black",
  )
  + theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.key.height = unit(50, "pt"),
    legend.title = element_text(size = 18, angle = 90, hjust = 0.5),
    legend.text = element_text(size = 16),
    legend.background = element_blank(),
    legend.title.position = "right",
  )
  + coord_sf(
    xlim = xlims,
    ylim = ylims
  )
  + labs(
    fill = "temp"
  )
)
ggsave(
  path = fig_dir,
  filename = "temp_fine_zwe_zoom.png",
  width = 5,
  height = 4
)
```

## Population scaled temperature

```{r}
rast_temp_weighted_zwe <- rast_temp_fine_zwe * rast_pop_zwe
```

```{r}
#| fig-width: 7
#| fig-height: 4

p1 <- (
  ggplot()
  + geom_spatraster(
    data = rast_pop_zwe %>% terra::crop(plot_extent),
    mapping = aes(fill = pop)
  )
  + scale_fill_paletteer_c(
    "scico::devon",
    na.value = "transparent",
    direction = -1
  )
  + geom_sf(
    data = sf_zwe,
    fill = NA,
    color = "black",
  )
  + theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.key.height = unit(50, "pt"),
    legend.title = element_text(size = 18, angle = 90, hjust = 0.5),
    legend.text = element_text(size = 16),
    legend.background = element_blank(),
    legend.title.position = "right",
  )
  + coord_sf(
    xlim = xlims,
    ylim = ylims
  )
  + labs(
    fill = "Population"
  )
)

p2 <- (
  ggplot()
  + geom_spatraster(
    data = rast_temp_weighted_zwe %>% terra::crop(plot_extent),
    mapping = aes(fill = t2m_max_1)
  )
  + scale_fill_paletteer_c(
    "scico::vik",
    na.value = "transparent",
  )
  + geom_sf(
    data = sf_zwe,
    fill = NA,
    color = "black",
  )
  + theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.key.height = unit(50, "pt"),
    legend.title = element_text(size = 18, angle = 90, hjust = 0.5),
    legend.text = element_text(size = 16),
    legend.background = element_blank(),
    legend.title.position = "right",
  )
  + coord_sf(
    xlim = xlims,
    ylim = ylims
  )
  + labs(
    fill = "Temperature * Population"
  )
)

p1 + p2

ggsave(
  path = fig_dir,
  filename = "temp_pop.png",
  width = 6.5,
  height = 4
)
```

## Weighted temp

```{r}
df_temp_daily_zwe <- (
  rast_temp_weighted_zwe
  %>% exactextractr::exact_extract(
    sf_zwe,
    fun = function(values, coverage_fraction) sum(values * coverage_fraction, na.rm = TRUE),
    max_cells_in_memory = 176067000,
    append_cols = c("adm_code"),
    progress = FALSE
  )
  %>% merge(
    adm_pop_zwe,
    by = "adm_code",
    all.x = TRUE
  )
  %>% mutate(
    date = as.Date(rast_times[1], format = "%Y-%m-%d"),
    temp_2m_max = (result / pop) - 273.15
  )
)
```

```{r}
sf_temp_daily_zwe <- (
  sf_zwe
  %>% merge(
    df_temp_daily_zwe,
    by = c("adm_code", "adm_name"),
    all.x = TRUE
  )
)
```

```{r}
#| fig-width: 4
#| fig-height: 5
(
  ggplot()
  + geom_sf(
    data = sf_temp_daily_zwe,
    mapping = aes(fill = temp_2m_max),
    color = "black",
  )
  + theme(
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.background = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5, margin = margin(b = 0)),
    plot.subtitle = element_text(size = 14, hjust = 0.5),
    legend.key.height = unit(50, "pt"),
    legend.title = element_text(size = 16, angle = 90, hjust = 0.5),
    legend.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.title.position = "right",
  )
  + scale_fill_paletteer_c(
    "scico::vik",
    na.value = "transparent",
  )
  + labs(
    title = "Weighted Temperature",
    fill = "Max daily temperature"
  )
)
ggsave(
  path = fig_dir,
  filename = "temp_weighted_zwe.png",
  width = 4,
  height = 5
)
```

# Generate datasets

## Raw temperature dataset

```{r}
df_temp <- build_temp_daily(rast_temp, sf_all)
```

## All Population Weighted Temperature

### Zimbabwe

```{r}
df_weighted_temp_zwe <- build_weighted_temp_daily_country("ZWE", country_map, rast_temp, rast_pop_zwe, adm_pop_zwe, sf_zwe)
```

### Zambia

```{r}
df_weighted_temp_zmb <- build_weighted_temp_daily_country("ZMB", country_map, rast_temp, rast_pop_zmb, adm_pop_zmb, sf_zmb)
```

## Girl Population Weighted Temperature

### Zimbabwe

```{r}
df_girl_weighted_temp_zwe <- build_weighted_temp_daily_country("ZWE", country_map, rast_temp, rast_popf_10_19_zwe, adm_popf_10_19_zwe, sf_zwe)
```

### Zambia

```{r}
df_girl_weighted_temp_zmb <- build_weighted_temp_daily_country("ZMB", country_map, rast_temp, rast_popf_10_19_zmb, adm_popf_10_19_zmb, sf_zmb)
```

# Save

```{r}
df_temp %>% write_csv(file.path(proc_data_dir, "temp_max_daily_all_countries_admin2.csv"))
```

```{r}
df_weighted_temp <- bind_rows(
  df_weighted_temp_zwe,
  df_weighted_temp_zmb
)
 
df_weighted_temp %>% write_csv(file.path(proc_data_dir, "weighted_temp_max_daily_all_countries_admin2.csv"))
```

```{r}
df_girl_weighted_temp <- bind_rows(
  df_girl_weighted_temp_zwe,
  df_girl_weighted_temp_zmb
)
 
df_girl_weighted_temp %>% write_csv(file.path(proc_data_dir, "age_weighted_temp_max_daily_all_countries_admin2.csv"))
```
