---
title: "DHS Zambia Data Processing"
format: html
editor: visual
---

# Environment Setup

```{r setup}
library(tidyverse)
library(sf)
library(readxl)
library(patchwork)
library(paletteer)
library(ggtext)
library(scales)

# Set random seed
nb_name <- "Agile 2025 - Analysis of the Uganda call centre data"
seed <- sum(sapply(str_split(nb_name, pattern = "", simplify = TRUE), FUN = utf8ToInt))

# Folder structure
root <- rprojroot::has_file("agile2025.Rproj")$make_fix_file()
raw_data_dir <- root("data/raw")
proc_data_dir <- root("data/processed/DHS")
dir.create(proc_data_dir, showWarnings = FALSE)
fig_dir <- root("figures")

# Viz settings
col_palette <- c("#dc7726", "#eebb93", "#ecc046", "#f6e0a3", "#0292b7", "#81c9db", "#97DECE", "#35A29F")
```

# Parameters

```{r}
coords <- list(
  "MOZ" = c(30, 41.2, -26.9, -10.3),
  "ZWE" = c(25.2, 33.1, -22.5, -15.45),
  "LSO" = c(26.99, 29.46, -30.68, -28.55),
  "NAM" = c(11.65, 25.27, -29, -16.9),
  "ZMB" = c(21.95, 33.8, -18.1, -8.15),
  "UGA" = c(29.55, 35.05, -1.49, 4.22),
  "KEN" = c(33.90, 41.92, -4.68, 5.05)
)
```


# Import Data

## Geometries

```{r}
sf_zmb <- (
  st_read(
    file.path(
      raw_data_dir,
      "geographies",
      "zmb_admin_boundaries",
      "zmb_admin2.shp"
    )
  )
  %>% select(
    country = adm0_name,
    adm1_name,
    adm2_name
  )
)
```

## Geolocations

```{r}
sf_dhs_geo_zmb <- (
  st_read(
    file.path(raw_data_dir, "DHS", "ZMGE81FL", "ZMGE81FL.shp")
  )
  %>% rename_with(str_to_lower)
  %>% select(
    cluster = dhsclust
  )
)
```

## Survey Data

```{r}
wealth_map <- c(
  "poorest" = 1,
  "poorer" = 2,
  "middle" = 3,
  "richer" = 4,
  "richest" = 5
)

df_dhs_surv_zmb <- (
  read_csv(
    file.path(
      raw_data_dir,
      "DHS",
      "data_from_Diana_2.csv"
    ),
    show_col_type = FALSE
  )
  %>% mutate(
    date = as.Date(
      paste0(
        year_of_interview,
        sprintf("%02d", month_of_interview),
        sprintf("%02d", day_of_interview)
      ),
      format = "%Y%m%d"
    ),
    date = lubridate::ceiling_date(date, unit = "month") - days(1),
    female = if_else(
      sex == "female",
      1,
      0
    ),
    rural = if_else(
      residence_type == "rural",
      1,
      0
    ),
    wealth_quintile = wealth_map[wealth_quintile],
    num_of_residence_years = as.numeric(case_when(
      num_of_residence_years == "Always" ~ "100",
      num_of_residence_years == "Visitor" ~ "0",
      .default = num_of_residence_years
    )),
    years_since_sexual_debut = age_respondent - age_sexual_debut,
    year_of_sexual_debut = year(date %m-% years(years_since_sexual_debut))
  )
  %>% select(
    cluster = clustre_num,
    date,
    household = household_num,
    weight = weight_household,
    rural,
    female,
    age = age_respondent,
    education,
    marital_status,
    wealth_quintile,
    num_of_residence_years,
    age_sexual_debut,
    years_since_sexual_debut,
    year_of_sexual_debut,
    num_sex_partners,
    condom_last_sex,
    ever_hiv_tested
  )
)
```

```{r}
(
  ggplot(data =   df_dhs_surv_zmb)
  + geom_histogram(
    mapping = aes(
      x = years_since_sexual_debut
    )
  )
)
```

## SPEI

Data obtained using the Climate Data Store API

```{r}
rast_spei <- terra::rast(
  file.path(
    raw_data_dir, 
    "SPEI",
    "spei01.nc"
  )
)
```

## UTCI

```{r}

```


# Process Data

## Zambia

### Assign admin area to respondents

```{r}
df_dhs_adm_zmb <- (
  sf_zmb
  %>% st_join(
    sf_dhs_geo_zmb,
    join = st_contains,
    left = TRUE
  )
  %>% as.data.frame()
  %>% select(
    cluster,
    adm1_name,
    adm2_name
  )
)
```

### SPEI

#### Create dataset

```{r}
start_date <- "1960-01-01"
rast_spei_times <- terra::time(rast_spei)
ids <- which(rast_spei_times >= start_date)
```


```{r}
thresholds <- round(quantile(seq_along(ids), probs = c(0, 0.25, 0.5, 0.75, 1)))
sfs_spei <- list()
for (i in seq_along(ids)) {
  if (i %in% thresholds) {
    thresh_idx <- which(thresholds == i)
    print(paste0(names(thresh_idx), " complete"))
  }
  sfs_spei[[i]] <- (
    rast_spei
    %>% terra::subset(ids[[i]])
    %>% terra::crop(terra::ext(coords[["ZMB"]]))
    %>% terra::as.polygons(aggregate = FALSE)
    %>% st_as_sf()
    %>% mutate(
      spei_tile_code = row_number(),
      date = as.Date(rast_spei_times[[ids[[i]]]])
    )
    %>% rename(
      spei = 1
    )
    %>% select(
      spei_tile_code,
      date,
      spei
    )
  )
}
spei_grid_map_zmb <- sfs_spei[[i]] %>% select(spei_tile_code)
df_spei_grid_zmb <- (
  bind_rows(sfs_spei)
  %>% as_tibble()
  %>% select(-geometry)
)
```

#### Assign tile to clusters

```{r}
df_dhs_spei_grid_zmb <- (
  sf_dhs_geo_zmb
  %>% st_join(
    spei_grid_map_zmb,
    join = st_intersects,
    left = TRUE
  )
  %>% as_tibble()
  %>% select(
    cluster,
    spei_tile_code
  )
)
```

#### Calculate rolling stats

```{r}
df_spei_cum_zmb <- (
  df_spei_grid_zmb
  %>% filter(
    spei_tile_code %in% unique(df_dhs_spei_grid_zmb$spei_tile_code)
  )
  %>% mutate(
    date = lubridate::ceiling_date(date, unit = "month") - days(1),
    decade = round(year(date), digits = -1)
  )
  %>% group_by(spei_tile_code, decade)
  %>% mutate(
    hist_mean = mean(spei, na.rm = TRUE),
    hist_sd = sd(spei, na.rm = TRUE),
    p05 = quantile(spei, 0.05, na.rm = TRUE),
    p10 = quantile(spei, 0.1, na.rm = TRUE),
    p15 = quantile(spei, 0.15, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    sd_clim = (spei - hist_mean) / hist_sd,
    below_1sd = as.numeric(sd_clim < -1),
    below_p05 = as.numeric(spei < p05),
    below_p10 = as.numeric(spei < p10),
    below_p15 = as.numeric(spei < p15),
    below_m1 = as.numeric(spei < -1)
  )
  %>% select(-c(decade, hist_mean, hist_sd, sd_clim, p05, p10, p15))
  %>% arrange(spei_tile_code, date)
  %>% group_by(spei_tile_code)
  %>% mutate(
    across(
      -c(date, spei),
      ~ (
        slider::slide_index_dbl(
          .x = (.),
          .i = date,
          .f = ~ sum(.x, na.rm = TRUE),
          .before = ~ .x %m-% months(12),
          .complete = TRUE
        )
      ),
      .names = "{.col}_12m"
    ),
    across(
      -c(date, spei),
      ~ (
        slider::slide_index_dbl(
          .x = (.),
          .i = date,
          .f = ~ sum(.x, na.rm = TRUE),
          .before = ~ .x %m-% months(24),
          .complete = TRUE
        )
      ),
      .names = "{.col}_24m"
    )
  )
  %>% ungroup()
  %>% select(spei_tile_code, date, contains("12m"),contains("24m"))
  %>% filter(!is.na(below_m1_12m))
  %>% pivot_longer(
    cols = -c(spei_tile_code, date),
    names_to = "variable",
    values_to = "value"
  )
  %>% group_by(spei_tile_code, variable)
  %>% mutate(
    mean_value = mean(value, na.rm = TRUE),
    sd_value = sd(value, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    dev_value = value - mean_value,
    # dev_value = if_else(
    #   sd_value == 0,
    #   0,
    #   (value - mean_value) / sd_value
    # ),
    variable = paste0("spei_", variable)
  )
  %>% filter(!is.na(value))
  %>% pivot_wider(
    id_cols = c(spei_tile_code, date),
    names_from = "variable",
    values_from = c(value, dev_value)
  )
)
```

#### Calculate yearly stats

```{r}
df_spei_yearly_zmb <- (
  df_spei_grid_zmb
  %>% filter(
    spei_tile_code %in% unique(df_dhs_spei_grid_zmb$spei_tile_code)
  )
  %>% mutate(
    year = year(date),
    decade = round(year, digits = -1)
  )
  %>% group_by(spei_tile_code, decade)
  %>% mutate(
    hist_mean = mean(spei, na.rm = TRUE),
    hist_sd = sd(spei, na.rm = TRUE),
    p05 = quantile(spei, 0.05, na.rm = TRUE),
    p10 = quantile(spei, 0.1, na.rm = TRUE),
    p15 = quantile(spei, 0.15, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    sd_clim = (spei - hist_mean) / hist_sd,
    below_1sd = as.numeric(sd_clim < -1),
    below_p05 = as.numeric(spei < p05),
    below_p10 = as.numeric(spei < p10),
    below_p15 = as.numeric(spei < p15),
    below_m1 = as.numeric(spei < -1)
  )
  %>% select(-c(decade, hist_mean, hist_sd, sd_clim, p05, p10, p15))
  %>% group_by(spei_tile_code, year)
  %>% summarise(
    mean_spei_year = mean(spei, na.rm = TRUE),
    median_spei_year = median(spei, na.rm = TRUE),
    spei_below_1sd_year = sum(below_1sd, na.rm = TRUE),
    spei_below_m1_year = sum(below_m1, na.rm = TRUE),
    spei_below_p05_year = sum(below_p05, na.rm = TRUE),
    spei_below_p10_year = sum(below_p10, na.rm = TRUE),
    spei_below_p15_year = sum(below_p15, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ungroup()
  %>% select(spei_tile_code, year, contains("year"))
  %>% filter(!is.na(spei_below_1sd_year))
  %>% pivot_longer(
    cols = -c(spei_tile_code, year),
    names_to = "variable",
    values_to = "value"
  )
  %>% group_by(spei_tile_code, variable)
  %>% mutate(
    mean_value = mean(value, na.rm = TRUE),
    sd_value = sd(value, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    dev_value = value - mean_value,
  )
  %>% filter(!is.na(value))
  %>% pivot_wider(
    id_cols = c(spei_tile_code, year),
    names_from = "variable",
    values_from = c(value, dev_value)
  )
)
```

### UTCI

#### Create dataset

```{r}
rast_files <- list.files(file.path(raw_data_dir, "climate/utci/daily/ZMB"), pattern = "^ECMWF_utci_daily_stats", full.names = TRUE, recursive = TRUE)
thresholds <- round(quantile(seq_along(rast_files), probs = c(0, 0.25, 0.5, 0.75, 1)))
sfs_utci <- list()
for (i in seq_along(rast_files)) {
  if (i %in% thresholds) {
    thresh_idx <- which(thresholds == i)
    print(paste0(names(thresh_idx), " complete"))
  }
  rast_utci <- terra::rast(rast_files[[i]])["utci_daily_max"]
  date <- terra::time(rast_utci)[[1]]
  sfs_utci[[i]] <- (
    rast_utci
    %>% terra::as.polygons(aggregate = FALSE)
    %>% st_as_sf()
    %>% mutate(
      utci_tile_code = row_number(),
      date = as.Date(date, format = "%Y-%m-%d"),
      utci_daily_max = utci_daily_max - 273.15
    )
    %>% select(
      utci_tile_code,
      date,
      utci_daily_max
    )
  )
}
utci_grid_map_zmb <- sfs_utci[[i]] %>% select(utci_tile_code)
df_utci_grid_zmb <- (
  bind_rows(sfs_utci)
  %>% as_tibble()
  %>% select(-geometry)
)
```

#### Assign tile to clusters

```{r}
df_dhs_utci_grid_zmb <- (
  sf_dhs_geo_zmb
  %>% st_join(
    utci_grid_map_zmb,
    join = st_intersects,
    left = TRUE
  )
  %>% as_tibble()
  %>% select(
    cluster,
    utci_tile_code
  )
)
```

#### Calculate rolling stats

```{r}
df_utci_cum_zmb <- (
  df_utci_grid_zmb
  %>% filter(
    utci_tile_code %in% unique(df_dhs_utci_grid_zmb$utci_tile_code)
  )
  %>% mutate(
    decade = round(year(date), digits = -1)
  )
  %>% group_by(utci_tile_code, decade)
  %>% mutate(
    hist_mean = mean(utci_daily_max, na.rm = TRUE),
    hist_sd = sd(utci_daily_max, na.rm = TRUE),
    p85 = quantile(utci_daily_max, 0.85, na.rm = TRUE),
    p90 = quantile(utci_daily_max, 0.9, na.rm = TRUE),
    p95 = quantile(utci_daily_max, 0.95, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    sd_clim = (utci_daily_max - hist_mean) / hist_sd,
    above_1sd = as.numeric(sd_clim > 1),
    above_p85 = as.numeric(utci_daily_max > p85),
    above_p90 = as.numeric(utci_daily_max > p90),
    above_p95 = as.numeric(utci_daily_max > p95),
    above_32 = as.numeric(utci_daily_max > 32)
  )
  %>% select(-c(decade, hist_mean, hist_sd, sd_clim, p85, p90, p95))
  %>% arrange(utci_tile_code, date)
  %>% group_by(utci_tile_code)
  %>% mutate(
    across(
      -c(date, utci_daily_max),
      ~ (
        slider::slide_index_dbl(
          .x = (.),
          .i = date,
          .f = ~ sum(.x, na.rm = TRUE),
          .before = ~ .x %m-% months(12),
          .complete = TRUE
        )
      ),
      .names = "{.col}_12m"
    ),
    across(
      -c(date, utci_daily_max),
      ~ (
        slider::slide_index_dbl(
          .x = (.),
          .i = date,
          .f = ~ sum(.x, na.rm = TRUE),
          .before = ~ .x %m-% months(24),
          .complete = TRUE
        )
      ),
      .names = "{.col}_24m"
    )
  )
  %>% ungroup()
  %>% select(utci_tile_code, date, contains("12m"), contains("24m"))
  %>% filter(!is.na(above_1sd_12m))
  %>% pivot_longer(
    cols = -c(utci_tile_code, date),
    names_to = "variable",
    values_to = "value"
  )
  %>% group_by(utci_tile_code, variable)
  %>% mutate(
    mean_value = mean(value, na.rm = TRUE),
    sd_value = sd(value, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    dev_value = value - mean_value,
    # dev_value = if_else(
    #   sd_value == 0,
    #   0,
    #   (value - mean_value) / sd_value
    # ),
    variable = paste0("utci_", variable)
  )
  %>% filter(!is.na(value))
  %>% pivot_wider(
    id_cols = c(utci_tile_code, date),
    names_from = "variable",
    values_from = c(value, dev_value)
  )
)
```

#### Calculate yearly stats

```{r}
df_utci_yearly_zmb <- (
  df_utci_grid_zmb
  %>% filter(
    utci_tile_code %in% unique(df_dhs_utci_grid_zmb$utci_tile_code)
  )
  %>% mutate(
    year = year(date),
    decade = round(year, digits = -1)
  )
  %>% group_by(utci_tile_code, decade)
  %>% mutate(
    hist_mean = mean(utci_daily_max, na.rm = TRUE),
    hist_sd = sd(utci_daily_max, na.rm = TRUE),
    p85 = quantile(utci_daily_max, 0.85, na.rm = TRUE),
    p90 = quantile(utci_daily_max, 0.9, na.rm = TRUE),
    p95 = quantile(utci_daily_max, 0.95, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    sd_clim = (utci_daily_max - hist_mean) / hist_sd,
    above_1sd = as.numeric(sd_clim > 1),
    above_p85 = as.numeric(utci_daily_max > p85),
    above_p90 = as.numeric(utci_daily_max > p90),
    above_p95 = as.numeric(utci_daily_max > p95),
    above_32 = as.numeric(utci_daily_max > 32)
  )
  %>% select(-c(decade, hist_mean, hist_sd, sd_clim, p85, p90, p95))
  %>% group_by(utci_tile_code, year)
  %>% summarise(
    mean_utci_year = mean(utci_daily_max, na.rm = TRUE),
    median_utci_year = median(utci_daily_max, na.rm = TRUE),
    utci_above_1sd_year = sum(above_1sd, na.rm = TRUE),
    utci_above_32_year = sum(above_32, na.rm = TRUE),
    utci_above_p85_year = sum(above_p85, na.rm = TRUE),
    utci_above_p90_year = sum(above_p90, na.rm = TRUE),
    utci_above_p95_year = sum(above_p95, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ungroup()
  %>% select(utci_tile_code, year, contains("year"))
  %>% filter(!is.na(utci_above_1sd_year))
  %>% pivot_longer(
    cols = -c(utci_tile_code, year),
    names_to = "variable",
    values_to = "value"
  )
  %>% group_by(utci_tile_code, variable)
  %>% mutate(
    mean_value = mean(value, na.rm = TRUE),
    sd_value = sd(value, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    dev_value = value - mean_value,
  )
  %>% filter(!is.na(value))
  %>% pivot_wider(
    id_cols = c(utci_tile_code, year),
    names_from = "variable",
    values_from = c(value, dev_value)
  )
)
```

### Assign tiles to respondents

```{r}
df_dhs_surv_tile_zmb <- (
  df_dhs_surv_zmb
  %>% merge(
    df_dhs_spei_grid_zmb,
    by = "cluster",
    all.x = TRUE
  )
  %>% merge(
    df_dhs_utci_grid_zmb,
    by = "cluster",
    all.x = TRUE
  )
)
```

### Assign climate and admin to respondent - Single time

```{r}
df_dhs_surv_adm_clim_zmb <- (
  df_dhs_surv_tile_zmb
  %>% merge(
    df_dhs_adm_zmb,
    by = "cluster",
    all.x = TRUE
  )
  %>% merge(
    df_spei_cum_zmb,
    by = c("spei_tile_code", "date"),
    all.x = TRUE
  )
  %>% merge(
    df_utci_cum_zmb,
    by = c("utci_tile_code", "date"),
    all.x = TRUE
  )
  %>% select(-c(spei_tile_code, utci_tile_code, years_since_sexual_debut, year_of_sexual_debut))
)
```

```{r}
hist(df_dhs_surv_adm_clim_zmb$value_utci_above_p95_12m)
```

```{r}
colMeans(is.na(df_dhs_surv_adm_clim_zmb))
```

### Save

```{r}
write_csv(df_dhs_surv_adm_clim_zmb %>% select(-contains("utci")), file = file.path(proc_data_dir, "dhs_climate_zmb_by_tile.csv"))
```

```{r}
# df_dhs_surv_adm_clim_zmb_old <- (
#   read_csv(
#     file.path(proc_data_dir, "dhs_climate_zmb.csv"),
#     show_col_types = FALSE
#   )
#   %>% select(
#     adm1_name,
#     adm2_name,
#     date,
#     cluster,
#     household,
#     contains("12m"),
#     contains("24m")
#   )
# )
```

```{r}
# df_dhs_surv_adm_clim_zmb_new <- (
#   df_dhs_surv_adm_clim_zmb
#   %>% select(
#     colnames(df_dhs_surv_adm_clim_zmb_old)
#   )
# )
```

```{r}
# df_comp <- (
#   df_dhs_surv_adm_clim_zmb_new
#   %>% pivot_longer(
#     cols = -c("adm1_name", "adm2_name", "date", "cluster", "household"),
#     names_to = "variable",
#     values_to = "value_new"
#   )
#   %>% bind_cols(
#     df_dhs_surv_adm_clim_zmb_old
#     %>% select(contains("12m") | contains("24m"))
#     %>% pivot_longer(
#       cols = everything(),
#       names_to = "variable_old",
#       values_to = "value_old"
#     )
#   )
#   %>% mutate(
#     diff_value = value_new - value_old
#   )
  # %>% select(-c(value_new, value_old, variable_old))
  # %>% pivot_wider(
  #   id_cols = c(adm1_name, adm2_name, date, cluster, household),
  #   names_from = variable,
  #   values_from = diff_value
  # )
)
```


### Assign climate to respondent - Sexual Debut

```{r}
df_sex_debut_adm_zmb <- (
  df_dhs_surv_tile_zmb
  %>% merge(
    df_dhs_adm_zmb,
    by = "cluster",
    all.x = TRUE
  )
  %>% filter(
    num_of_residence_years > years_since_sexual_debut + 1
  )
  %>% select(
    spei_tile_code,
    utci_tile_code,
    adm1_name,
    adm2_name,
    weight,
    rural,
    female,
    wealth_quintile,
    year = year_of_sexual_debut,
    age_sexual_debut
  )
  %>% mutate(
    weight = weight / sum(weight, na.rm = TRUE),
    early_sexual_debut = age_sexual_debut < 15
  )
  %>% merge(
    df_spei_yearly_zmb,
    by = c("spei_tile_code", "year"),
    all.x = TRUE
  )
  %>% merge(
    df_utci_yearly_zmb,
    by = c("utci_tile_code", "year"),
    all.x = TRUE
  )
)
```

# Viz

```{r}
df_sex_debut_viz <- (
  df_sex_debut_adm_zmb
  %>% group_by(adm1_name, year)
  %>% summarise(
    perc_early_sex_debut = sum(early_sexual_debut * weight, na.rm = TRUE) / sum(weight, na.rm = TRUE),
    mean_spei = mean(value_mean_spei_year, na.rm = TRUE),
    spei_below_p10 = mean(value_spei_below_p10_year, na.rm = TRUE),
    dev_spei_below_p10 = mean(dev_value_spei_below_p10_year, na.rm = TRUE),
    dev_mean_utci = mean(dev_value_mean_spei_year, na.rm = TRUE),
    utci_above_p90 = mean(value_utci_above_p90_year, na.rm = TRUE),
    dev_utci_above_p90 = mean(dev_value_utci_above_p90_year, na.rm = TRUE),
    .groups = "drop"
  )
  %>% pivot_longer(
    cols = -c(adm1_name, year),
    names_to = "variable",
    values_to = "value"
  )
)
```

## Sexual Debut

### Counts of Sexual Debuts by year

```{r}
#| fig-width: 5
#| fig-height: 3

(
  df_sex_debut_adm_zmb
  %>% ggplot()
  + geom_histogram(
    mapping = aes(
      x = year,
      fill = early_sexual_debut
    ),
    position = "stack",
    binwidth = 1
  )
  + scale_fill_manual(
    values = c(`FALSE` = "#088BBEFF", `TRUE` = "#EF7C12FF")
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 17),
    plot.margin = margin(l = 10, r = 10, b = 10, t = 10),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.frame = element_blank(),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
    legend.position = "inside",
    legend.position.inside = c(0.25, 0.78)
  )
  + labs(
    x = "Year",
    y = "# Sexual Debuts",
    fill = "Early Sexual Debut"
  )
)

# ggsave(
#   path = fig_dir,
#   filename = "sexual_debut_by_year_zmb.png"
# )
```

### Rate of early sexual debuts by year

```{r}
#| fig-width: 5.075
#| fig-height: 2.5
(
  df_sex_debut_viz
  # %>% filter(year > 1970)
  %>% group_by(year, variable)
  %>% summarise(
    mean_value = mean(value, na.rm = TRUE),
    .groups = "drop"
  )
  %>% filter(
    variable == "perc_early_sex_debut"
  )
  %>% ggplot(
    mapping = aes(x = year, y = mean_value)
  )
  + scale_y_continuous(
    labels = scales::percent
  )
  + geom_line(color = "#1BB6AFFF", linewidth = 1.2)
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 17),
    plot.margin = margin(l = 10, r = 10, b = 10, t = 10),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.frame = element_blank(),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
    legend.position = "inside",
    legend.position.inside = c(0.25, 0.78)
  )
  + labs(
    x = "Year",
    y = "Early Sexual Debuts"
  )
)
ggsave(
  path = fig_dir,
  filename = "rate_of_early_sexual_debut_by_year_zmb.png"
)
```

## By admin 2

### Rate of sexual debut by year

```{r}
p1 <- (
  df_sex_debut_adm_zmb
  %>% filter(
    year > 1990,
    year < 2024
  )
  %>% group_by(adm1_name, year)
  %>% summarise(
    perc_early_sex_debut = mean(early_sexual_debut, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ggplot(
    mapping = aes(x = year, y = perc_early_sex_debut, color = adm1_name)
  )
  + geom_line(linewidth = 1.2)
  + scale_y_continuous(
    labels = scales::percent
  )
  + scale_color_paletteer_d(
    palette = "RColorBrewer::RdYlBu"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 17),
    plot.margin = margin(l = 10, r = 10, b = 10, t = 10),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.frame = element_blank(),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.78)
  )
  + labs(
    x = "Year",
    y = "Early Sexual Debuts",
    color = "District"
  )
)
# ggsave(
#   path = fig_dir,
#   filename = "rate_of_early_sexual_debut_by_year_by_admin1_zmb.png"
# )
```

### SPEI

#### Mean

```{r}
p2 <- (
  df_sex_debut_adm_zmb
  %>% filter(
    year > 1990,
    year < 2024
  )
  %>% group_by(adm1_name, year)
  %>% summarise(
    mean_spei = mean(value_mean_spei_year, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ggplot(
    mapping = aes(x = year, y = mean_spei, color = adm1_name)
  )
  + geom_line(linewidth = 1.2)
  + scale_color_paletteer_d(
    palette = "RColorBrewer::RdYlBu"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 17),
    plot.margin = margin(l = 10, r = 10, b = 10, t = 10),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.frame = element_blank(),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.78)
  )
  + labs(
    x = "Year",
    y = "Mean SPEI",
    color = "District"
  )
)
# ggsave(
#   path = fig_dir,
#   filename = "mean_spei_by_year_by_admin1_zmb.png"
# )
```

#### Months of drought-like conditions

```{r}
p3 <- (
  df_sex_debut_adm_zmb
  %>% filter(
    year > 1990,
    year < 2024
  )
  %>% group_by(adm1_name, year)
  %>% summarise(
    months_low_spei = mean(value_spei_below_p10_year, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ggplot(
    mapping = aes(x = year, y = months_low_spei, color = adm1_name)
  )
  + geom_line(linewidth = 1.2)
  + scale_color_paletteer_d(
    palette = "RColorBrewer::RdYlBu"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 17),
    plot.margin = margin(l = 10, r = 10, b = 10, t = 10),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.frame = element_blank(),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.78)
  )
  + labs(
    x = "Year",
    y = "# Drought Months",
    color = "District"
  )
)
# ggsave(
#   path = fig_dir,
#   filename = "num_spei_below_p10_by_year_by_admin1_zmb.png"
# )
```

### UTCI

#### Deviation from Mean

```{r}
p4 <- (
  df_sex_debut_adm_zmb
  %>% filter(
    year > 1990,
    year < 2024
  )
  %>% group_by(adm1_name, year)
  %>% summarise(
    dev_mean_utci = mean(dev_value_mean_utci_year, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ggplot(
    mapping = aes(x = year, y = dev_mean_utci, color = adm1_name)
  )
  + geom_line(linewidth = 1.2)
  + scale_color_paletteer_d(
    palette = "RColorBrewer::RdYlBu"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 17),
    plot.margin = margin(l = 10, r = 10, b = 10, t = 10),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.frame = element_blank(),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.78)
  )
  + labs(
    x = "Year",
    y = "Dev. Mean UTCI",
    color = "District"
  )
)
# ggsave(
#   path = fig_dir,
#   filename = "mean_utci_by_year_by_admin1_zmb.png"
# )
```

#### Dev. Days of high-heat conditions

```{r}
p5 <- (
  df_sex_debut_adm_zmb
  %>% filter(
    year > 1990,
    year < 2024
  )
  %>% group_by(adm1_name, year)
  %>% summarise(
    dev_days_high_utci = mean(dev_value_utci_above_p90_year, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ggplot(
    mapping = aes(x = year, y = dev_days_high_utci, color = adm1_name)
  )
  + geom_line(linewidth = 1.2)
  + scale_color_paletteer_d(
    palette = "RColorBrewer::RdYlBu"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 17),
    plot.margin = margin(l = 10, r = 10, b = 10, t = 10),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.frame = element_blank(),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.78)
  )
  + labs(
    x = "Year",
    y = "Dev. # Hot Days",
    color = "District"
  )
)
# ggsave(
#   path = fig_dir,
#   filename = "num_utci_above_32_by_year_by_admin1_zmb.png"
# )
```


```{r}
#| fig-width: 8
#| fig-height: 10
(
  p1 / p2 / p3 / p4 / p5
  + plot_layout(
    guides = "collect",
    axes = "collect"
  )
  & theme(
    plot.margin = margin(b = 0, t = 10, l = 10, r = 10)
  )
)
# ggsave(
#   path = fig_dir,
#   filename = "early_sex_debut_spei_by_year_by_admin1_zmb.png"
# )
#   
```

## Whole country

### Rate of sexual debut by year

```{r}
p1 <- (
  df_sex_debut_adm_zmb
  %>% filter(
    year > 1990,
    year < 2024
  )
  %>% group_by(year)
  %>% summarise(
    perc_early_sex_debut = mean(early_sexual_debut, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ggplot(
    mapping = aes(x = year, y = perc_early_sex_debut)
  )
  + geom_line(linewidth = 1.2)
  + scale_y_continuous(
    labels = scales::percent
  )
  + scale_color_paletteer_d(
    palette = "RColorBrewer::RdYlBu"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 17),
    plot.margin = margin(l = 10, r = 10, b = 10, t = 10),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.frame = element_blank(),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
  )
  + labs(
    x = "Year",
    y = "Early Sexual Debuts",
  )
)
# ggsave(
#   path = fig_dir,
#   filename = "rate_of_early_sexual_debut_by_year_zmb.png"
# )
```

### SPEI

#### Mean

```{r}
p2 <- (
  df_sex_debut_adm_zmb
  %>% filter(
    year > 1990,
    year < 2024
  )
  %>% group_by(year)
  %>% summarise(
    mean_spei = mean(value_mean_spei_year, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ggplot(
    mapping = aes(x = year, y = mean_spei)
  )
  + geom_line(linewidth = 1.2)
  + scale_color_paletteer_d(
    palette = "RColorBrewer::RdYlBu"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 17),
    plot.margin = margin(l = 10, r = 10, b = 10, t = 10),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.frame = element_blank(),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
  )
  + labs(
    x = "Year",
    y = "Mean SPEI",
  )
)
# ggsave(
#   path = fig_dir,
#   filename = "mean_spei_by_year_zmb.png"
# )
```

#### Months of drought-like conditions

```{r}
p3 <- (
  df_sex_debut_adm_zmb
  %>% filter(
    year > 1990,
    year < 2024
  )
  %>% group_by(year)
  %>% summarise(
    months_low_spei = mean(value_spei_below_p10_year, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ggplot(
    mapping = aes(x = year, y = months_low_spei)
  )
  + geom_line(linewidth = 1.2)
  + scale_color_paletteer_d(
    palette = "RColorBrewer::RdYlBu"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 17),
    plot.margin = margin(l = 10, r = 10, b = 10, t = 10),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.frame = element_blank(),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
  )
  + labs(
    x = "Year",
    y = "# Drought Months",
  )
)
# ggsave(
#   path = fig_dir,
#   filename = "num_spei_below_p10_by_year_zmb.png"
# )
```

### UTCI

#### Deviation from Mean

```{r}
p4 <- (
  df_sex_debut_adm_zmb
  %>% filter(
    year > 1990,
    year < 2024
  )
  %>% group_by(year)
  %>% summarise(
    dev_mean_utci = mean(dev_value_mean_utci_year, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ggplot(
    mapping = aes(x = year, y = dev_mean_utci)
  )
  + geom_line(linewidth = 1.2)
  + scale_color_paletteer_d(
    palette = "RColorBrewer::RdYlBu"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 17),
    plot.margin = margin(l = 10, r = 10, b = 10, t = 10),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.frame = element_blank(),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
  )
  + labs(
    x = "Year",
    y = "Dev. Mean UTCI",
  )
)
# ggsave(
#   path = fig_dir,
#   filename = "mean_utci_by_year_zmb.png"
# )
```

#### Dev. Days of high-heat conditions

```{r}
p5 <- (
  df_sex_debut_adm_zmb
  %>% filter(
    year > 1990,
    year < 2024
  )
  %>% group_by(year)
  %>% summarise(
    dev_days_high_utci = mean(dev_value_utci_above_p90_year, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ggplot(
    mapping = aes(x = year, y = dev_days_high_utci)
  )
  + geom_line(linewidth = 1.2)
  + scale_color_paletteer_d(
    palette = "RColorBrewer::RdYlBu"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 17),
    plot.margin = margin(l = 10, r = 10, b = 10, t = 10),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.frame = element_blank(),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
  )
  + labs(
    x = "Year",
    y = "Dev. # Hot Days",
  )
)
# ggsave(
#   path = fig_dir,
#   filename = "num_utci_above_32_by_year_zmb.png"
# )
```


```{r}
#| fig-width: 8
#| fig-height: 10
(
  p1 / p2 / p3 / p4 / p5
  + plot_layout(
    guides = "collect",
    axes = "collect"
  )
  & theme(
    plot.margin = margin(b = 0, t = 10, l = 10, r = 10)
  )
)
# ggsave(
#   path = fig_dir,
#   filename = "early_sex_debut_spei_by_year_zmb.png"
# )
#   
```
