---
title: "DHS Zambia Data Processing"
format: html
editor: visual
---

# Environment Setup

```{r setup}
library(tidyverse)
library(sf)
library(readxl)
library(patchwork)
library(paletteer)
library(ggtext)
library(scales)

# Set random seed
nb_name <- "Agile 2025 - Analysis of the Uganda call centre data"
seed <- sum(sapply(str_split(nb_name, pattern = "", simplify = TRUE), FUN = utf8ToInt))

# Folder structure
root <- rprojroot::has_file("agile2025.Rproj")$make_fix_file()
raw_data_dir <- root("data/raw")
proc_data_dir <- root("data/processed/DHS")
dir.create(proc_data_dir, showWarnings = FALSE)
fig_dir <- root("figures")

# Viz settings
col_palette <- c("#dc7726", "#eebb93", "#ecc046", "#f6e0a3", "#0292b7", "#81c9db", "#97DECE", "#35A29F")
```

# Functions

```{r}
# source(file.path(src_dir, "data_processing.R"))
build_spei <- function(rast_spei, adm_shapes, start_date) {
  print(paste0("Starting SPEI data processing"))
  rast_times <- terra::time(rast_spei)
  ids <- which(rast_times >= start_date)
  rast_spei <- rast_spei %>% terra::subset(ids)
  dates <- as.character(terra::time(rast_spei))
  df_spei <- (
    rast_spei
    %>% exactextractr::exact_extract(
      adm_shapes,
      fun = "weighted_mean",
      weights = "area",
      max_cells_in_memory = 176067000,
      append_cols = c("country", "adm1_name", "adm2_name"),
      progress = FALSE
    )
  )
  colnames(df_spei) <- c("country", "adm1_name", "adm2_name", dates)
  df_spei <- (
    df_spei
    %>% pivot_longer(
      cols = -c("country", "adm1_name", "adm2_name"),
      names_to = "date",
      values_to = "spei"
    )
    %>% mutate(
      date = as.Date(date, format = "%Y-%m-%d")
    )
  )
}
```

# Import Data

## Geometries

```{r}
sf_zmb <- (
  st_read(
    file.path(
      raw_data_dir,
      "geographies",
      "zmb_admin_boundaries",
      "zmb_admin2.shp"
    )
  )
  %>% select(
    country = adm0_name,
    adm1_name,
    adm2_name
  )
)
```

## Geolocations

```{r}
sf_dhs_geo_zmb <- (
  st_read(
    file.path(raw_data_dir, "DHS", "ZMGE81FL", "ZMGE81FL.shp")
  )
  %>% rename_with(str_to_lower)
  %>% select(
    cluster = dhsclust
  )
)
```

## Survey Data

```{r}
wealth_map <- c(
  "poorest" = 1,
  "poorer" = 2,
  "middle" = 3,
  "richer" = 4,
  "richest" = 5
)

df_dhs_surv_zmb <- (
  read_csv(
    file.path(
      raw_data_dir,
      "DHS",
      "data_from_Diana_2.csv"
    ),
    show_col_type = FALSE
  )
  %>% mutate(
    date = as.Date(
      paste0(
        year_of_interview,
        sprintf("%02d", month_of_interview),
        sprintf("%02d", day_of_interview)
      ),
      format = "%Y%m%d"
    ),
    date = lubridate::ceiling_date(date, unit = "month") - days(1),
    female = if_else(
      sex == "female",
      1,
      0
    ),
    rural = if_else(
      residence_type == "rural",
      1,
      0
    ),
    wealth_quintile = wealth_map[wealth_quintile],
    num_of_residence_years = as.numeric(case_when(
      num_of_residence_years == "Always" ~ "100",
      num_of_residence_years == "Visitor" ~ "0",
      .default = num_of_residence_years
    )),
    years_since_sexual_debut = age_respondent - age_sexual_debut,
    year_of_sexual_debut = year(date %m-% years(years_since_sexual_debut))
  )
  %>% select(
    cluster = clustre_num,
    date,
    household = household_num,
    weight = weight_household,
    rural,
    female,
    age = age_respondent,
    education,
    marital_status,
    wealth_quintile,
    num_of_residence_years,
    age_sexual_debut,
    years_since_sexual_debut,
    year_of_sexual_debut,
    num_sex_partners,
    condom_last_sex,
    ever_hiv_tested
  )
)
```

```{r}
(
  ggplot(data =   df_dhs_surv_zmb)
  + geom_histogram(
    mapping = aes(
      x = years_since_sexual_debut
    )
  )
)
```

## SPEI

Data obtained using the Climate Data Store API

```{r}
rast_spei <- terra::rast(
  file.path(
    raw_data_dir, 
    "climate/spei",
    "spei01.nc"
  )
)
```

# Process Data

## SPEI

### Assign SPEI to admin areas

```{r}
df_spei <- build_spei(
  rast_spei,
  sf_zmb,
  "1960-01-01"
)
```

### Calculate rolling stats

```{r}
# df_spei_12m <- (
#   df_spei
#   %>% mutate(
#     date = lubridate::ceiling_date(date, unit = "month") - days(1),
#     # decade = round(year(date), digits = -1)
#   )
#   %>% filter(
#     date >= "2000-01-01"
#   )
#   %>% group_by(country, adm1_name, adm2_name)
#   %>% mutate(
#     hist_mean = mean(spei, na.rm = TRUE),
#     hist_sd = sd(spei, na.rm = TRUE)
#   )
#   %>% ungroup()
#   %>% mutate(
#     sd_clim = (spei - hist_mean) / hist_sd,
#     below_1sd = as.numeric(sd_clim < -1),
#     below_m1 = as.numeric(spei < -1)
#   )
#   %>% select(-c(hist_mean, hist_sd, sd_clim))
#   %>% arrange(country, adm1_name, adm2_name, date)
#   %>% group_by(country, adm1_name, adm2_name)
#   %>% mutate(
#     across(
#       -c(date, spei),
#       ~ (
#         slider::slide_index_dbl(
#           .x = (.),
#           .i = date,
#           .f = ~ sum(.x, na.rm = TRUE),
#           .before = ~ .x %m-% months(12),
#           .complete = TRUE
#         )
#       ),
#       .names = "{.col}_12m"
#     ),
#   )
#   %>% ungroup()
#   %>% select(country, adm1_name, adm2_name, date, contains("12m"))
#   %>% filter(!is.na(below_m1_12m))
#   %>% pivot_longer(
#     cols = -c(country, adm1_name, adm2_name, date),
#     names_to = "variable",
#     values_to = "value"
#   )
#   %>% group_by(country, adm1_name, adm2_name, variable)
#   %>% mutate(
#     mean_value = mean(value, na.rm = TRUE),
#     sd_value = sd(value, na.rm = TRUE)
#   )
#   %>% ungroup()
#   %>% mutate(
#     dev_value = value - mean_value,
#     # dev_value = if_else(
#     #   sd_value == 0,
#     #   0,
#     #   (value - mean_value) / sd_value
#     # ),
#     variable = paste0("spei_", variable)
#   )
#   %>% filter(!is.na(value))
#   %>% pivot_wider(
#     id_cols = c(country, adm1_name, adm2_name, date),
#     names_from = "variable",
#     values_from = c(value, dev_value)
#   )
# )
df_spei_cum <- (
  df_spei
  %>% mutate(
    date = lubridate::ceiling_date(date, unit = "month") - days(1),
    decade = round(year(date), digits = -1)
  )
  # %>% filter(
  #   date >= "2000-01-01"
  # )
  %>% group_by(country, adm1_name, adm2_name, decade)
  %>% mutate(
    hist_mean = mean(spei, na.rm = TRUE),
    hist_sd = sd(spei, na.rm = TRUE),
    p05 = quantile(spei, 0.05, na.rm = TRUE),
    p10 = quantile(spei, 0.1, na.rm = TRUE),
    p15 = quantile(spei, 0.15, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    sd_clim = (spei - hist_mean) / hist_sd,
    below_1sd = as.numeric(sd_clim < -1),
    below_p05 = as.numeric(spei < p05),
    below_p10 = as.numeric(spei < p10),
    below_p15 = as.numeric(spei < p15),
    below_m1 = as.numeric(spei < -1)
  )
  %>% select(-c(decade, hist_mean, hist_sd, sd_clim, p05, p10, p15))
  %>% arrange(country, adm1_name, adm2_name, date)
  %>% group_by(country, adm1_name, adm2_name)
  %>% mutate(
    across(
      -c(date, spei),
      ~ (
        slider::slide_index_dbl(
          .x = (.),
          .i = date,
          .f = ~ sum(.x, na.rm = TRUE),
          .before = ~ .x %m-% months(12),
          .complete = TRUE
        )
      ),
      .names = "{.col}_12m"
    ),
    across(
      -c(date, spei),
      ~ (
        slider::slide_index_dbl(
          .x = (.),
          .i = date,
          .f = ~ sum(.x, na.rm = TRUE),
          .before = ~ .x %m-% months(24),
          .complete = TRUE
        )
      ),
      .names = "{.col}_24m"
    )
  )
  %>% ungroup()
  %>% select(country, adm1_name, adm2_name, date, contains("12m"),contains("24m"))
  %>% filter(!is.na(below_m1_12m))
  %>% pivot_longer(
    cols = -c(country, adm1_name, adm2_name, date),
    names_to = "variable",
    values_to = "value"
  )
  %>% group_by(country, adm1_name, adm2_name, variable)
  %>% mutate(
    mean_value = mean(value, na.rm = TRUE),
    sd_value = sd(value, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    dev_value = value - mean_value,
    # dev_value = if_else(
    #   sd_value == 0,
    #   0,
    #   (value - mean_value) / sd_value
    # ),
    variable = paste0("spei_", variable)
  )
  %>% filter(!is.na(value))
  %>% pivot_wider(
    id_cols = c(country, adm1_name, adm2_name, date),
    names_from = "variable",
    values_from = c(value, dev_value)
  )
)
```

## Geolocate respondents

```{r}
df_dhs_adm_zmb <- (
  sf_zmb
  %>% st_join(
    sf_dhs_geo_zmb,
    join = st_contains,
    left = TRUE
  )
  %>% as.data.frame()
  %>% select(
    cluster,
    adm1_name,
    adm2_name
  )
)
```

### Calculate yearly stats

```{r}
df_spei_yearly <- (
  df_spei
  %>% mutate(
    year = year(date),
    decade = round(year, digits = -1)
  )
  %>% group_by(country, adm1_name, adm2_name, decade)
  %>% mutate(
    hist_mean = mean(spei, na.rm = TRUE),
    hist_sd = sd(spei, na.rm = TRUE),
    p05 = quantile(spei, 0.05, na.rm = TRUE),
    p10 = quantile(spei, 0.1, na.rm = TRUE),
    p15 = quantile(spei, 0.15, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    sd_clim = (spei - hist_mean) / hist_sd,
    below_1sd = as.numeric(sd_clim < -1),
    below_p05 = as.numeric(spei < p05),
    below_p10 = as.numeric(spei < p10),
    below_p15 = as.numeric(spei < p15),
    below_m1 = as.numeric(spei < -1)
  )
  %>% select(-c(decade, hist_mean, hist_sd, sd_clim, p05, p10, p15))
  %>% group_by(country, adm1_name, adm2_name, year)
  %>% summarise(
    mean_spei = mean(spei, na.rm = TRUE),
    median_spei = median(spei, na.rm = TRUE),
    spei_below_1sd_year = sum(below_1sd, na.rm = TRUE),
    spei_below_m1_year = sum(below_m1, na.rm = TRUE),
    spei_below_p05_year = sum(below_p05, na.rm = TRUE),
    spei_below_p10_year = sum(below_p10, na.rm = TRUE),
    spei_below_p15_year = sum(below_p15, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ungroup()
)
```

## Assign Admin areas to respondents

```{r}
df_dhs_surv_adm_zmb <- (
  df_dhs_surv_zmb
  %>% merge(
    df_dhs_adm_zmb,
    by = "cluster",
    all.x = TRUE
  )
)
```

## Assign SPEI to respondent - Single time

```{r}
df_dhs_surv_adm_clim_zmb <- (
  df_dhs_surv_adm_zmb
  %>% merge(
    df_spei_cum,
    by = c("adm1_name", "adm2_name", "date"),
    all.x = TRUE
  )
  %>% select(-c(country, years_since_sexual_debut, year_of_sexual_debut))
)
```

```{r}
hist(df_dhs_surv_adm_clim_zmb$value_spei_below_p15_12m)
```

```{r}
colMeans(is.na(df_dhs_surv_adm_clim_zmb))
```

### Save

```{r}
write_csv(df_dhs_surv_adm_clim_zmb, file = file.path(proc_data_dir, "dhs_climate_zmb.csv"))
```

## Assign SPEI to respondent - Sexual Debut

```{r}
df_sex_debut <- (
  df_dhs_surv_adm_zmb
  %>% filter(
    num_of_residence_years > years_since_sexual_debut + 1
  )
  %>% select(
    adm1_name,
    adm2_name,
    weight,
    rural,
    female,
    wealth_quintile,
    year = year_of_sexual_debut,
    age_sexual_debut
  )
  %>% mutate(
    weight = weight / sum(weight, na.rm = TRUE),
    early_sexual_debut = age_sexual_debut < 15
  )
  %>% merge(
    df_spei_yearly,
    by = c("adm1_name", "adm2_name", "year"),
    all.x = TRUE
  )
)
```

# Viz

```{r}
df_sex_debut_viz <- (
  df_sex_debut
  %>% group_by(adm1_name, year)
  %>% summarise(
    perc_early_sex_debut = sum(early_sexual_debut * weight, na.rm = TRUE) / sum(weight, na.rm = TRUE),
    mean_spei = mean(mean_spei, na.rm = TRUE),
    mean_spei_below_p10 = mean(spei_below_p10_year, na.rm = TRUE),
    .groups = "drop"
  )
  %>% pivot_longer(
    cols = -c(adm1_name, year),
    names_to = "variable",
    values_to = "value"
  )
)
```

## Counts of Sexual Debuts by year

```{r}
#| fig-width: 5
#| fig-height: 3

(
  df_sex_debut
  %>% ggplot()
  + geom_histogram(
    mapping = aes(
      x = year,
      fill = early_sexual_debut
    ),
    position = "stack",
    binwidth = 1
  )
  + scale_fill_manual(
    values = c(`FALSE` = "#088BBEFF", `TRUE` = "#EF7C12FF")
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 17),
    plot.margin = margin(l = 10, r = 10, b = 10, t = 10),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.frame = element_blank(),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
    legend.position = "inside",
    legend.position.inside = c(0.25, 0.78)
  )
  + labs(
    x = "Year",
    y = "# Sexual Debuts",
    fill = "Early Sexual Debut"
  )
)

ggsave(
  path = fig_dir,
  filename = "sexual_debut_by_year_zmb.png"
)
```

## Rate of early sexual debuts by year

```{r}
#| fig-width: 5.075
#| fig-height: 2.5
(
  df_sex_debut_viz
  # %>% filter(year > 1970)
  %>% group_by(year, variable)
  %>% summarise(
    mean_value = mean(value, na.rm = TRUE),
    .groups = "drop"
  )
  %>% filter(
    variable == "perc_early_sex_debut"
  )
  %>% ggplot(
    mapping = aes(x = year, y = mean_value)
  )
  + scale_y_continuous(
    labels = scales::percent
  )
  + geom_line(color = "#1BB6AFFF", linewidth = 1.2)
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 17),
    plot.margin = margin(l = 10, r = 10, b = 10, t = 10),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.frame = element_blank(),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
    legend.position = "inside",
    legend.position.inside = c(0.25, 0.78)
  )
  + labs(
    x = "Year",
    y = "Early Sexual Debuts"
  )
)
ggsave(
  path = fig_dir,
  filename = "rate_of_early_sexual_debut_by_year_zmb.png"
)
```

```{r}
#| fig-width: 9
#| fig-height: 7
(
  df_sex_debut_viz
  %>% filter(
    year > 1990,
    year < 2024
  )
  %>% ggplot(
    mapping = aes(x = year, y = value, color = adm1_name)
  )
  + geom_line(linewidth = 1.2)
  + facet_wrap(
    ~variable,
    ncol = 1,
    scales = "free_y"
  )
  + scale_color_paletteer_d(
    palette = "RColorBrewer::RdYlBu"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 17),
    plot.margin = margin(l = 10, r = 10, b = 10, t = 10),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.frame = element_blank(),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.78)
  )
  + labs(
    x = "Year",
    y = "Early Sexual Debuts"
  )
)
ggsave(
  path = fig_dir,
  filename = "rate_of_early_sexual_debut_by_year_by_admin1_zmb.png"
)
```

```{r}
#| fig-width: 9
#| fig-height: 3
p1 <- (
  df_sex_debut
  %>% filter(
    year > 1990,
    year < 2024
  )
  %>% group_by(year, adm1_name)
  %>% summarise(
    perc_early_sex_debut = mean(early_sexual_debut, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ggplot(
    mapping = aes(x = year, y = perc_early_sex_debut, color = adm1_name)
  )
  + geom_line(linewidth = 1.2)
  + scale_y_continuous(
    labels = scales::percent
  )
  + scale_color_paletteer_d(
    palette = "RColorBrewer::RdYlBu"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 17),
    plot.margin = margin(l = 10, r = 10, b = 10, t = 10),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.frame = element_blank(),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.78)
  )
  + labs(
    x = "Year",
    y = "Early Sexual Debuts",
    color = "District"
  )
)
# ggsave(
#   path = fig_dir,
#   filename = "rate_of_early_sexual_debut_by_year_by_admin1_zmb.png"
# )
```

## SPEI

```{r}
#| fig-width: 8.9432
#| fig-height: 3
p2 <- (
  df_spei_yearly
  %>% filter(
    year > 1990,
    year < 2024
  )
  %>% group_by(year, adm1_name)
  %>% summarise(
    mean_spei = mean(mean_spei, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ggplot(
    mapping = aes(x = year, y = mean_spei, color = adm1_name)
  )
  + geom_line(linewidth = 1.2)
  + scale_color_paletteer_d(
    palette = "RColorBrewer::RdYlBu"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 17),
    plot.margin = margin(l = 10, r = 10, b = 10, t = 10),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.frame = element_blank(),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.78)
  )
  + labs(
    x = "Year",
    y = "Mean SPEI",
    color = "District"
  )
)
# ggsave(
#   path = fig_dir,
#   filename = "mean_spei_by_year_by_admin1_zmb.png"
# )
```

```{r}
#| fig-width: 8.72
#| fig-height: 3
p3 <- (
  df_spei_yearly
  %>% filter(
    year > 1990,
    year < 2024
  )
  %>% group_by(year, adm1_name)
  %>% summarise(
    mean_spei = mean(spei_below_p10_year, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ggplot(
    mapping = aes(x = year, y = mean_spei, color = adm1_name)
  )
  + geom_line(linewidth = 1.2)
  + scale_color_paletteer_d(
    palette = "RColorBrewer::RdYlBu"
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    plot.background = element_blank(),
    plot.title = element_text(size = 17),
    plot.margin = margin(l = 10, r = 10, b = 10, t = 10),
    axis.title = element_text(size = 15),
    axis.text = element_text(size = 14),
    legend.background = element_blank(),
    legend.frame = element_blank(),
    legend.title = element_text(size = 15),
    legend.text = element_text(size = 13),
    # legend.position = "inside",
    # legend.position.inside = c(0.25, 0.78)
  )
  + labs(
    x = "Year",
    y = "# Drought Months",
    color = "District"
  )
)
# ggsave(
#   path = fig_dir,
#   filename = "num_spei_below_p10_by_year_by_admin1_zmb.png"
# )
```

```{r}
#| fig-width: 9
#| fig-height: 7
(
  p1 / p2 / p3
  + plot_layout(
    guides = "collect",
    axes = "collect"
  )
  & theme(
    plot.margin = margin(b = 0, t = 10, l = 10, r = 10)
  )
)
ggsave(
  path = fig_dir,
  filename = "early_sex_debut_spei_by_year_by_admin1_zmb.png"
)
  
```

## Map
