---
title: "Process Food Prices"
format: html
editor: visual
---

# Environment Setup

```{r setup}
# Packages
library(tidyverse)
library(sf)
library(patchwork)
library(paletteer)
library(spdep)

# Folder Structure
root <- rprojroot::has_file("agile2025.Rproj")$make_fix_file()
raw_data_dir <- root("data/raw")
proc_data_dir <- root("data/processed/food_prices")
dir.create(proc_data_dir, showWarnings = FALSE)
fig_dir <- root("figures")
dir.create(fig_dir, showWarnings = FALSE)

# Settings
nb_name <-  "Process Food Prices"
seed <- sum(sapply(str_split(nb_name, pattern = "", simplify = TRUE), FUN = utf8ToInt))
set.seed(seed)

# Country Mapping
country_map <- c(
  "MOZ" = "Mozambique",
  "ZWE" = "Zimbabwe",
  "LSO" = "Lesotho",
  "NAM" = "Namibia",
  "ZMB" = "Zambia"
)
```

# Functions

```{r}
neighbors_at_hop <- function(nb_list, start_idx, h) {
  if (h <= 0) return (integer(0))
  if (h == 1) return (nb_list[[start_idx]])
  visited <- integer(0)
  frontier <- unique(nb_list[[start_idx]])
  if (length(frontier) == 0) return(integer(0))
  step <- 1
  # We reconstruct the frontier for starting index `start_idx` after step h
  while (step < h) {
    step <- step + 1
    next_nodes <- if (length(frontier) > 0) unique(unlist(nb_list[frontier], use.names = FALSE)) else integer(0)
    # remove already seen and current frontier to advance
    next_frontier <- setdiff(next_nodes, c(visited, frontier))
    if (length(next_frontier) == 0) return(integer(0))
    visited <- c(visited, frontier)
    frontier <- next_frontier
  }
  setdiff(frontier, start_idx)
}

impute_price_with_nbs <- function(g, y) {
  ct_val <- g[["country"]][1]
  nb <- nb_by_country[[ct_val]]
  vals <- g[["food_price_kg"]]
  nas <- which(is.na(vals))
  if (length(nas) > 0) {
      for (i in nas) {
        for (h in 1:max_hops) {
          candidates <- neighbors_at_hop(nb, i, h)
          if (length(candidates) == 0L) next
          cand_vals <- vals[candidates]
          cand_vals <- cand_vals[!is.na(cand_vals)]
          if (length(cand_vals) > 0L) {
            vals[i] <- mean(cand_vals)
            break
          }
        }
      }
    g[["food_price_kg"]] <- vals
  }
  g %>% select(-c(country, date))
}
```

# Load Data

## Geographies

### Mozambique

```{r}
sf_moz <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Mozambique 2022 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Zimbabwe

```{r}
sf_zwe <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Zimbabwe 2015 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Lesotho

```{r}
sf_lso <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Lesotho 2023 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Namibia

```{r}
sf_nam <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Namibia 2013 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Zambia

```{r}
sf_zmb <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Zambia 2018 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Create combined shapes

```{r}
sf_all <- (
  bind_rows(
    sf_moz,
    sf_zwe,
    sf_lso,
    sf_nam,
    sf_zmb
  )
  %>% st_make_valid()
)
```

## Crop cycles

```{r}
df_crops <- (
  read_csv(
    file.path(
      raw_data_dir, 
      "crop",
      "harvestStat",
      "hvstat_africa_data_v1.0.csv"
    ),
    show_col_types = FALSE
  )
  %>% mutate(
    crop_production_system = if_else(
      country %in% c("Uganda", "Zambia"),
      "All (PS)",
      crop_production_system
    )
  )
  %>% filter(
    country %in% country_map,
    qc_flag == 0,
    product %in% c("Maize", "Sorghum", "Wheat", "Cassava", "Millet", "Rice", "Sweet Potatoes", "Potato", "Banana"),
    !is.na(production),
    harvest_year >= 2000,
    crop_production_system == "All (PS)"
  )
  %>% group_by(country, admin_1, product)
  %>% summarise(
    production = sum(production, na.rm = TRUE),
    .groups = "drop"
  )
  %>% group_by(country, admin_1)
  %>% slice_max(
    order_by = production,
    n = 1,
    with_ties = FALSE,
  )
  %>% ungroup()
  %>% rename(
    adm_name = admin_1
  )
)
```

```{r}
df_crops
```

Maize is usually the most commom product. For simplicity we use this to track price changes

## Food Prices

Data from: https://data.humdata.org/dataset/global-wfp-food-prices

Not all districts are represented

```{r load data}
food_prices_files <- list.files(file.path(raw_data_dir, "food_prices", "wfp_food_prices", fsep = "/"))
list_maize_prices <- list()
for (i in seq(length(food_prices_files))) {
  list_maize_prices[[i]] <- (
    read_csv(
      file.path(raw_data_dir, "food_prices", "wfp_food_prices", food_prices_files[i], fsep = "/"),
      show_col_types = FALSE
    )
    %>% filter(
      pricetype == "Retail",
      !is.na(latitude),
      category %in% c("cereals and tubers", "pulses and nuts", "vegetables and fruits"),
      (
        (commodity == "Maize (white)" & countryiso3 %in% c("MOZ", "ZMB"))
        | (commodity == "Maize" & countryiso3 == "ZWE")
        | (commodity == "Maize meal" & countryiso3 %in% c("LSO", "NAM"))
      )
      # commodity %in% c("Maize meal", "Maize flour")
    )
    %>% mutate(
      admin2 = str_replace(admin2, " Municipality", "")
    )
  )
}
```

```{r}
sf_food_prices <- (
  bind_rows(list_maize_prices)
  %>% filter(
    countryiso3 %in% names(country_map)
  )
  %>% mutate(
    date = as.Date(date, format = "%Y-%m-%d"),
    latitude = as.numeric(latitude),
    longitude = as.numeric(longitude),
    food_price_kg = case_when(
      unit == "400 G" ~ as.numeric(usdprice) * 2.5,
      unit == "KG" ~ as.numeric(usdprice),
      unit == "5 KG" ~ as.numeric(usdprice) / 5,
      unit == "12.5 KG" ~ as.numeric(usdprice) / 12.5,
      unit == "Tin (20L)" ~ as.numeric(usdprice) / 15,
      unit == "25 KG" ~ as.numeric(usdprice) / 25
    )
  )
  %>% st_as_sf(
    coords = c("longitude", "latitude"),
    crs = 4326
  )
  %>% select(
    countryiso3,
    date,
    food_price_kg
  )
)
```

# Process Data

## Combine datasets

```{r}
df_food_prices_adm <- (
  sf_food_prices
  %>% st_join(
    sf_all,
    join = st_nearest_feature,
    left = TRUE
  )
  %>% arrange(
    country,
    adm_code,
    adm_name,
    date
  )
  %>% as_tibble()
  %>% group_by(
    country,
    adm_code,
    adm_name,
    date = lubridate::ceiling_date(date, "month") - lubridate::days(1)
  )
  %>% summarise(
    food_price_kg = mean(food_price_kg, na.rm = TRUE),
    .groups = "drop"
  )
)
```

## Create list of all dates for each country

```{r}
list_adm <- unique(sf_all$adm_name)
date_range <- range(df_food_prices_adm$date)
start_months <- floor_date(date_range[1], unit = "month") + months(0: (12*20))
full_list_dates <- list(ceiling_date(start_months[start_months <= date_range[2]], unit = "month") - days(1))
```


## Add missing dates to dataset

```{r}
sf_food_prices_adm_all_dates <- (
  sf_all
  %>% as_tibble()
  %>% select(-geometry)
  %>% group_by(country, adm_code, adm_name)
  %>% mutate(
    date = full_list_dates
  )
  %>% unnest(date)
  %>% merge(
    df_food_prices_adm,
    by = c("country", "adm_code", "adm_name", "date"),
    all.x = TRUE
  )
  %>% merge(
    sf_all,
    by = c("country", "adm_code", "adm_name"),
    all = TRUE
  )
  %>% arrange(
    country,
    adm_code,
    adm_name,
    date
  )
  %>% st_as_sf()
)
```

```{r}
(
  sf_food_prices_adm_all_dates
  %>% filter(
    date == "2019-04-30"
  )
  %>% ggplot()
  + geom_sf(
    mapping = aes(fill = food_price_kg)
  )
)
```

```{r}
(
  sf_food_prices_adm_all_dates
  %>% filter(is.na(date))
)
```


## Fill in missing values

```{r}
max_hops <- 3
iterative <- TRUE

sf_food_prices_adm_all_dates$.rowid_fill <- seq_len(nrow(sf_food_prices_adm_all_dates))
countries <- unname(country_map)
nb_by_country <- setNames(vector("list", length(countries)), countries)

for (ct in countries) {
  sf_df_country <- sf_food_prices_adm_all_dates %>% filter(country == ct)
  dates_ct <- sf_df_country$date %>% unique()
  geom <- sf_df_country %>% filter(date == dates_ct[[1]])
  nb_by_country[[ct]] <- poly2nb(geom, queen = TRUE)
}

sf_food_prices_adm_interp <- (
  sf_food_prices_adm_all_dates
  %>% group_by(country, date)
  %>% group_modify(impute_price_with_nbs, .keep = TRUE)
  %>% ungroup()
)

df_food_prices_adm_interp <- (
  sf_food_prices_adm_interp[order(sf_food_prices_adm_interp$.rowid_fill), ]
  %>% select(-.rowid_fill)
  %>% as_tibble()
  %>% select(-geometry)
  %>% group_by(country, adm_code, adm_name)
  %>% arrange(date, .by_group = TRUE)
  %>% mutate(
    food_price_kg = zoo::na.approx(food_price_kg, x = date, na.rm = FALSE)
  )
  %>% ungroup()
  %>% filter(
    !is.na(food_price_kg) # removes recent period of times for which we don't have data for some countries
  )
)
```


## Detrend data

### Breakpoints

```{r}
df_food_prices_summ <- (
  df_food_prices_adm_interp
  %>% group_by(country, date)
  %>% summarise(
    food_price_kg = mean(food_price_kg, na.rm = TRUE),
    .groups = "drop"
  )
  %>% ungroup()
  %>% arrange(date)
  %>% group_by(country)
  %>% mutate(
    t = row_number()
  )
  %>% ungroup()
  %>% arrange(country, date)
)

breakpoints_ids <- list()
for (i in seq_along(countries)) {
  df_c <- (
    df_food_prices_summ
    %>% filter(
      country == countries[[i]]
    )
  )
  breakpoints <- strucchange::breakpoints(
    food_price_kg ~ t,
    data = df_c,
    h = 0.15
  )
  breakpoints_ids[[i]] <- breakpoints$breakpoints
}
```

### Remove trends by region

```{r}
list_food_prices_adm_detrend <- list()
for (i in seq_along(countries)) {
  list_food_prices_adm_detrend[[i]] <- (
    df_food_prices_adm_interp
    %>% filter(country == countries[[i]])
    %>% arrange(adm_name, date)
    %>% group_by(adm_name)
    %>% mutate(
      t = row_number()
    )
    %>% ungroup()
    %>% group_by(adm_name)
    %>% group_modify(~{
      fit0 <- lm(food_price_kg ~ t, data = (.x))
      fit_seg <- segmented::segmented(
        fit0,
        seg.Z = ~t,
        psi = breakpoints_ids[[i]],
        control = segmented::seg.control(it.max = 0, display = FALSE)
      )
      (
        .x
        %>% mutate(
          trend_mu = fitted(fit_seg),
          food_price_kg = food_price_kg - trend_mu
        )
      )
    })
    %>% ungroup()
  )
}

df_food_prices_adm_detrend <- bind_rows(list_food_prices_adm_detrend)
```

```{r}
(
  df_food_prices_adm_detrend
  %>% ggplot(
    aes(x = date, y = food_price_kg, color = adm_name)
  )
  + geom_line()
  + facet_wrap(
    ~country
  )
  + guides(
    color = "none"
  )
)
```

## Calculate anomalies

```{r}
df_food_prices_adm_12m <- (
  df_food_prices_adm_detrend
  %>% group_by(country, date)
  %>% mutate(
    mean_country_price = mean(food_price_kg, na.rm = TRUE),
    sd_country_price = sd(food_price_kg, na.rm = TRUE)
  )
  %>% ungroup()
  %>% group_by(country, adm_code, adm_name)
  %>% mutate(
    hist_mean_price = mean(food_price_kg, na.rm = TRUE),
    hist_sd_price = sd(food_price_kg, na.rm = TRUE),
    hist_mean_sd_country_price = mean(sd_country_price, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    price_hist_anomaly = as.numeric(food_price_kg > hist_mean_price + hist_sd_price),
    price_reg_anomaly = as.numeric(food_price_kg > mean_country_price + hist_mean_sd_country_price)
  )
  %>% arrange(country, adm_code, adm_name, date)
  %>% group_by(country, adm_code, adm_name)
  %>% mutate(
    max_food_price_kg_12m = slider::slide_index_dbl( 
      .x = food_price_kg,
      .i = date,
      .f = ~ if (sum(!is.na(.x)) == 0) NA else max(.x, na.rm = TRUE),
      .before = ~ .x %m-% months(12),
      .complete = TRUE
    ),
    food_price_hist_anomaly_12m = slider::slide_index_dbl(
      .x = price_hist_anomaly,
      .i = date,
      .f = ~ sum(.x, na.rm = TRUE),
      .before = ~ .x %m-% months(12),
      .complete = TRUE
    ),
    food_price_reg_anomaly_12m = slider::slide_index_dbl(
      .x = price_reg_anomaly,
      .i = date,
      .f = ~ sum(.x, na.rm = TRUE),
      .before = ~ .x %m-% months(12),
      .complete = TRUE
    )
  )
  %>% ungroup()
  %>% mutate(
    food_price_stress_12m = as.integer(max_food_price_kg_12m > hist_mean_price + hist_sd_price)
  )
  %>% select(country, adm_code, adm_name, date, contains("12m"))
  %>% filter(!is.na(food_price_stress_12m))
  %>% pivot_longer(
    cols = -c(country, adm_code, adm_name, date),
    names_to = "variable",
    values_to = "value"
  ) 
  %>% group_by(country, adm_code, adm_name, variable)
  %>% mutate(
    mean_value = mean(value, na.rm = TRUE),
    sd_value = sd(value, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    dev_value = if_else(
      sd_value == 0,
      0,
      (value - mean_value) / sd_value
    )
  )
  %>% filter(!is.na(value))
  %>% pivot_wider(
    id_cols = c(country, adm_code, adm_name, date),
    names_from = "variable",
    values_from = c(value, dev_value)
  )
)
```

# Visualizations

```{r}
(
  df_food_prices_adm_12m
  %>% filter(
    country == "Namibia"
  )
  %>% ggplot(
    aes(x = date, y = value_max_food_price_kg_12m, color = adm_name)
  )
  + geom_line()
)
```

```{r}
(
  df_food_prices_adm_12m
  %>% filter(
    country == "Namibia"
  )
  %>% ggplot(
    aes(x = date, y = dev_value_max_food_price_kg_12m, color = adm_name)
  )
  + geom_line()
)
```

```{r}
(
  df_food_prices_adm_12m
  %>% filter(
    country == "Namibia"
  )
  %>% ggplot(
    aes(x = date, y = value_food_price_reg_anomaly_12m, color = adm_name)
  )
  + geom_line()
)
```

# Save

```{r}
df_food_prices_adm_interp %>% write_csv(file.path(proc_data_dir, "food_prices_adm.csv"))
df_food_prices_adm_detrend %>% write_csv(file.path(proc_data_dir, "food_prices_adm_detrended.csv"))
df_food_prices_adm_12m %>% write_csv(file.path(proc_data_dir, "food_prices_adm_12m.csv"))
```
