---
title: "Test Weighted UTCI"
format: html
editor: visual
---

# Environment Setup

```{r setup}
# Packages
library(tidyverse)
library(sf)
library(patchwork)
library(paletteer)

# Folder Structure
root <- rprojroot::has_file("agile2025.Rproj")$make_fix_file()
raw_data_dir <- root("data/raw")
proc_data_dir <- root("data/processed")
fig_dir <- root("figures")
dir.create(fig_dir, showWarnings = FALSE)
mod_dir <- root("models")
src_dir <- root("src")

# Settings
nb_name <-  "Process daily climate data"
seed <- sum(sapply(str_split(nb_name, pattern = "", simplify = TRUE), FUN = utf8ToInt))
set.seed(seed)
```

# Functions

```{r}
source(file.path(src_dir, "data_processing.R"))
```

# Data Import

## UTCI

There are 10 UTCI thermal stress categories that correspond to specific human physiological responses to the thermal environment. The categories relate to UTCI values as follows: above +46: extreme heat stress; +38 to +46: very strong heat stress; +32 to +38: strong heat stress; +26 to +32: moderate heat stress; +9 to +26: no thermal stress; +9 to 0: slight cold stress; 0 to -13: moderate cold stress; -13 to -27: strong cold stress; -27 to -40: very strong cold stress; below -40: extreme cold stress.

### Raw UTCI

```{r}
df_utci <- (
  read_csv(
    file.path(proc_data_dir, "UTCI", "utci_max_daily_all_countries.csv"),
    show_col_types = FALSE
  )
)
```

### Population Weighted UTCI daily

```{r}
df_utci_pop_weighted <- (
  read_csv(
    file.path(proc_data_dir, "UTCI", "weighted_utci_max_daily_all_countries.csv"),
    show_col_types = FALSE
  )
  %>% select(-pop)
)
```

### Girl population weighted UTCI

```{r}
df_utci_girl_weighted <- (
  read_csv(
    file.path(
      proc_data_dir,
      "UTCI",
      "age_weighted_utci_max_daily_all_countries.csv"
    ),
    show_col_types = FALSE
  )
  %>% select(-popf_10_19)
)
```

## Temperature (temp_2m_max)

### Raw Temperature

```{r}
df_temp <- (
  read_csv(
    file.path(proc_data_dir, "temp_2m", "temp_max_daily_all_countries.csv"),
    show_col_types = FALSE
  )
)
```

### Population Weighted Temperature

```{r}
df_temp_pop_weighted <- (
  read_csv(
    file.path(proc_data_dir, "temp_2m", "weighted_temp_max_daily_all_countries.csv"),
    show_col_types = FALSE
  )
  %>% select(-pop)
)
```

### Girl population weighted Temperature

```{r}
df_temp_girl_weighted <- (
  read_csv(
    file.path(
      proc_data_dir,
      "temp_2m",
      "age_weighted_temp_max_daily_all_countries.csv"
    ),
    show_col_types = FALSE
  )
  %>% select(-pop)
)
```

## SPEI

### Raw SPEI

```{r}
df_spei <- (
  read_csv(
    file.path(
      proc_data_dir,
      "SPEI",
      "spei_all_countries.csv"
    ),
    show_col_types = FALSE
  )
)
```

### Population weighted SPEI

```{r}
df_spei_pop_weighted <- (
  read_csv(
    file.path(
      proc_data_dir,
      "SPEI",
      "weighted_spei_all_countries.csv"
    ),
    show_col_types = FALSE
  )
  %>% select(-pop)
)
```

### Girl population weighted UTCI

```{r}
df_spei_girl_weighted <- (
  read_csv(
    file.path(
      proc_data_dir,
      "SPEI",
      "age_weighted_spei_all_countries.csv"
    ),
    show_col_types = FALSE
  )
  %>% select(-popf_10_19)
)
```

# Process Data

## UTCI

### Raw UTCI

```{r}
df_utci_12m <- calculate_utci_12m_stats(df_utci)
```

### Population Weighted UTCI

```{r}
df_utci_pop_weighted_12m <- calculate_utci_12m_stats(df_utci_pop_weighted, weights = "pop_weighted")
```

### Girl Population Weighted UTCI

```{r}
df_utci_girl_weighted_12m <- calculate_utci_12m_stats(df_utci_girl_weighted, weights = "girl_weighted")
```

## Temperature

### Raw Temperature

```{r}
df_temp_12m <- calculate_temp_12m_stats(df_temp)
```

```{r}
colSums(is.na(df_temp_12m))
```

### Population Weighted Temperature

```{r}
df_temp_pop_weighted_12m <- calculate_temp_12m_stats(df_temp_pop_weighted, weights = "pop_weighted")
```

```{r}
colSums(is.na(df_weighted_temp_12m))
```

### Girl Population Weighted Temperature

```{r}
df_temp_girl_weighted_12m <- calculate_temp_12m_stats(df_temp_pop_weighted, weights = "girl_weighted")
```

## SPEI

### Raw SPEI

```{r}
df_spei_12m <- calculate_spei_12m_stats(df_spei)
```

### Population Weighted UTCI

```{r}
df_spei_pop_weighted_12m <- calculate_spei_12m_stats(df_spei_pop_weighted, weights = "pop_weighted")
```

### Girl Population Weighted UTCI

```{r}
df_spei_girl_weighted_12m <- calculate_spei_12m_stats(df_spei_girl_weighted, weights = "girl_weighted")
```

# Combine climate variables

```{r}
df_clim_12m <- ( 
  df_utci_12m
  %>% merge(
    df_utci_pop_weighted_12m,
    by = c("country", "adm_code", "adm_name", "date"),
    all.x = TRUE
  )
  %>% merge(
    df_utci_girl_weighted_12m,
    by = c("country", "adm_code", "adm_name", "date"),
    all.x = TRUE
  )
  %>% merge(
    df_temp_12m,
    by = c("country", "adm_code", "adm_name", "date"),
    all.x = TRUE
  )
  %>% merge(
    df_temp_pop_weighted_12m,
    by = c("country", "adm_code", "adm_name", "date"),
    all.x = TRUE
  )
  %>% merge(
    df_spei_12m,
    by = c("country", "adm_code", "adm_name", "date"),
    all.x = TRUE
  )
  %>% merge(
    df_spei_pop_weighted_12m,
    by = c("country", "adm_code", "adm_name", "date"),
    all.x = TRUE
  )
  %>% merge(
    df_spei_girl_weighted_12m,
    by = c("country", "adm_code", "adm_name", "date"),
    all.x = TRUE
  )
  # %>% mutate(
  #   adm_name = case_when(
  #     adm_name == "Harare Chitungwiza" ~ "Harare",
  #     adm_name == "//Karas" ~ "!Karas",
  #     .default = adm_name
  #   ),
  # )
)
```

# Visualizations

```{r}
(
  ggplot(
    data = df_clim_12m %>% filter(date > "2015-01-01"), 
    aes(
      x = dev_value_temp_2m_max_above_1sd_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    position = "stack",
    stat = "density"
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 13),
    )
  + scale_fill_paletteer_d(
    "MetBrewer::Derain",
    na.value = "transparent",
  )
  + labs(
    x = "# days temperature 1 sd above mean (dev)",
    y = "Density",
  )
)
```

## UTCI

### UTCI above 32

```{r}
(
  ggplot(
    data = df_clim_12m %>% filter(date > "2015-01-01"), 
    aes(
      x = value_utci_max_above_32_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    aes(y = after_stat(density)),
    position = "stack",
    bins = 30
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 13),
    )
  + scale_fill_paletteer_d(
    "MetBrewer::Derain",
    na.value = "transparent",
  )
  + labs(
    x = "# days UTCI above 32",
    y = "Density",
  )
)
```

### Weighted UTCI above 32

```{r}
(
  ggplot(
    data = df_clim_12m %>% filter(date > "2015-01-01"), 
    aes(
      x = value_weighted_utci_max_above_32_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    aes(y = after_stat(density)),
    position = "stack",
    bins = 40
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 13),
    )
  + scale_fill_paletteer_d(
    "MetBrewer::Derain",
    na.value = "transparent",
  )
  + labs(
    x = "# days weighted UTCI above 32",
    y = "Density",
  )
)
```

### Dev UTCI above 32

```{r}
(
  ggplot(
    data = df_clim_12m %>% filter(date > "2015-01-01"), 
    aes(
      x = dev_value_utci_max_above_32_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    aes(y = after_stat(density)),
    position = "stack",
    bins = 40
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 13),
    )
  + scale_fill_paletteer_d(
    "MetBrewer::Derain",
    na.value = "transparent",
  )
  + labs(
    x = "# days UTCI above 32 (dev)",
    y = "Density",
  )
)
```

### Dev weighted UTCI above 32

```{r}
(
  ggplot(
    data = df_clim_12m %>% filter(date > "2015-01-01"), 
    aes(
      x = dev_value_weighted_utci_max_above_32_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    aes(y = after_stat(density)),
    position = "stack",
    bins = 40
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 13),
    )
  + scale_fill_paletteer_d(
    "MetBrewer::Derain",
    na.value = "transparent",
  )
  + labs(
    x = "# days weighted UTCI above 32 (dev)",
    y = "Density",
  )
)
```

### UTCI above 1sd

```{r}
(
  ggplot(
    data = df_clim_12m %>% filter(date > "2015-01-01"), 
    aes(
      x = value_utci_max_above_1sd_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    aes(y = after_stat(density)),
    position = "stack",
    bins = 30
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 13),
    )
  + scale_fill_paletteer_d(
    "MetBrewer::Derain",
    na.value = "transparent",
  )
  + labs(
    x = "# days UTCI above 1 sd",
    y = "Density",
  )
)
```

### Weighted UTCI above 1 sd

```{r}
(
  ggplot(
    data = df_clim_12m %>% filter(date > "2015-01-01"), 
    aes(
      x = value_weighted_utci_max_above_1sd_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    aes(y = after_stat(density)),
    position = "stack",
    bins = 30
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 13),
    )
  + scale_fill_paletteer_d(
    "MetBrewer::Derain",
    na.value = "transparent",
  )
  + labs(
    x = "# days weighted UTCI above 1 sd",
    y = "Density",
  )
)
```

### Dev UTCI above 1 sd

```{r}
(
  ggplot(
    data = df_clim_12m %>% filter(date > "2015-01-01"), 
    aes(
      x = dev_value_utci_max_above_1sd_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    aes(y = after_stat(density)),
    position = "stack",
    bins = 40
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 13),
    )
  + scale_fill_paletteer_d(
    "MetBrewer::Derain",
    na.value = "transparent",
  )
  + labs(
    x = "# days UTCI above 1sd (dev)",
    y = "Density",
  )
)
```

### Dev weighted UTCI above 1 sd

```{r}
(
  ggplot(
    data = df_clim_12m %>% filter(date > "2015-01-01"), 
    aes(
      x = dev_value_weighted_utci_max_above_1sd_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    aes(y = after_stat(density)),
    position = "stack",
    bins = 40
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 13),
    )
  + scale_fill_paletteer_d(
    "MetBrewer::Derain",
    na.value = "transparent",
  )
  + labs(
    x = "# days weighted UTCI above 1 sd (dev)",
    y = "Density",
  )
)
```

### UTCI above 32 time series

```{r}
#| fig-width: 9
#| fig-height: 5

(
  df_clim_12m
  %>% filter(
    date > "2010-01-01",
    date < "2025-01-01",
    country == "Mozambique",
    !(adm_name %in% c("Tete", "Zambezia"))
  )
  %>% select(
    date,
    adm_name,
    value_utci_max_above_32_12m,
  )
  %>% ggplot(
    mapping = aes(x = date, y = value_utci_max_above_32_12m)
  )
  + geom_line(
    color = "#0292b7",
    linewidth = 1.3
  )
  + facet_wrap(
    ~ adm_name,
    ncol = 3
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    panel.spacing.x = unit(25, "pt"),
    strip.background = element_blank(),
    strip.text = element_text(size = 14),
    plot.background = element_blank(),
    plot.title = element_text(size = 16),
    plot.margin = margin(t = 0, r = 20, b = 0, l = 10, unit = "pt"),
    axis.ticks = element_line(linewidth = 1),
    axis.ticks.length = unit(4, "pt"),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 15),
  )
  + labs(
    x = "Date",
    y = "# days UTCI max above  (month)"
  )
)
ggsave(
  path = fig_dir,
  filename = "utci_max_above_32_month_moz.png",
  width = 9,
  height = 5
)
```

### Weighted UTCI above 32 time series

```{r}
#| fig-width: 9
#| fig-height: 5

(
  df_clim_12m
  %>% filter(
    date > "2010-01-01",
    date < "2025-01-01",
    country == "Mozambique",
    !(adm_name %in% c("Tete", "Zambezia"))
  )
  %>% select(
    date,
    adm_name,
    value_weighted_utci_max_above_32_12m,
  )
  %>% ggplot(
    mapping = aes(x = date, y = value_weighted_utci_max_above_32_12m)
  )
  + geom_line(
    color = "#0292b7",
    linewidth = 1.3
  )
  + facet_wrap(
    ~ adm_name,
    ncol = 3
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    panel.spacing.x = unit(25, "pt"),
    strip.background = element_blank(),
    strip.text = element_text(size = 14),
    plot.background = element_blank(),
    plot.title = element_text(size = 16),
    plot.margin = margin(t = 0, r = 20, b = 0, l = 10, unit = "pt"),
    axis.ticks = element_line(linewidth = 1),
    axis.ticks.length = unit(4, "pt"),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 15),
  )
  + labs(
    x = "Date",
    y = "# days weighted UTCI max above 32"
  )
)
ggsave(
  path = fig_dir,
  filename = "weighted_utci_max_above_32_moz.png",
  width = 9,
  height = 5
)
```

### Dev UTCI above 32 time series

```{r}
#| fig-width: 9
#| fig-height: 5

(
  df_clim_12m
  %>% filter(
    date > "2010-01-01",
    date < "2025-01-01",
    country == "Mozambique",
    !(adm_name %in% c("Tete", "Zambezia"))
  )
  %>% select(
    date,
    adm_name,
    dev_value_utci_max_above_32_12m,
  )
  %>% ggplot(
    mapping = aes(x = date, y = dev_value_utci_max_above_32_12m)
  )
  + geom_line(
    color = "#dc7726",
    linewidth = 1.3
  )
  + facet_wrap(
    ~ adm_name,
    ncol = 3
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    panel.spacing.x = unit(25, "pt"),
    strip.background = element_blank(),
    strip.text = element_text(size = 14),
    plot.background = element_blank(),
    plot.title = element_text(size = 16),
    plot.margin = margin(t = 0, r = 20, b = 0, l = 10, unit = "pt"),
    axis.ticks = element_line(linewidth = 1),
    axis.ticks.length = unit(4, "pt"),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 15),
  )
  + labs(
    x = "Date",
    y = "# days UTCI max above 32 (dev)"
  )
)
ggsave(
  path = fig_dir,
  filename = "utci_max_above_32_moz.png",
  width = 9,
  height = 5
)
```

### Dev weighted UTCI above 32 time series

```{r}
#| fig-width: 9
#| fig-height: 5

(
  df_clim_12m
  %>% filter(
    date > "2010-01-01",
    date < "2025-01-01",
    country == "Mozambique",
    !(adm_name %in% c("Tete", "Zambezia"))
  )
  %>% select(
    date,
    adm_name,
    dev_value_weighted_utci_max_above_32_12m,
  )
  %>% ggplot(
    mapping = aes(x = date, y = dev_value_weighted_utci_max_above_32_12m)
  )
  + geom_line(
    color = "#dc7726",
    linewidth = 1.3
  )
  + facet_wrap(
    ~ adm_name,
    ncol = 3
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    panel.spacing.x = unit(25, "pt"),
    strip.background = element_blank(),
    strip.text = element_text(size = 14),
    plot.background = element_blank(),
    plot.title = element_text(size = 16),
    plot.margin = margin(t = 0, r = 20, b = 0, l = 10, unit = "pt"),
    axis.ticks = element_line(linewidth = 1),
    axis.ticks.length = unit(4, "pt"),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 15),
  )
  + labs(
    x = "Date",
    y = "# days weighted UTCI max above 32 (dev)"
  )
)
ggsave(
  path = fig_dir,
  filename = "weighted_utci_max_above_32_moz.png",
  width = 9,
  height = 5
)
```

## Temperature

### Temperature above 32

```{r}
(
  ggplot(
    data = df_clim_12m %>% filter(date > "2015-01-01"), 
    aes(
      x = value_temp_2m_max_above_32_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    aes(y = after_stat(density)),
    position = "stack",
    bins = 30
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 13),
    )
  + scale_fill_paletteer_d(
    "MetBrewer::Derain",
    na.value = "transparent",
  )
  + labs(
    x = "# days temperature above 32",
    y = "Density",
  )
)
```

### Weighted temperature above 32

```{r}
(
  ggplot(
    data = df_clim_12m %>% filter(date > "2015-01-01"), 
    aes(
      x = value_weighted_temp_2m_max_above_32_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    aes(y = after_stat(density)),
    position = "stack",
    bins = 40
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 13),
    )
  + scale_fill_paletteer_d(
    "MetBrewer::Derain",
    na.value = "transparent",
  )
  + labs(
    x = "# days weighted temperature above 32",
    y = "Density",
  )
)
```

### Dev temperature above 32

```{r}
(
  ggplot(
    data = df_clim_12m %>% filter(date > "2015-01-01"), 
    aes(
      x = dev_value_temp_2m_max_above_32_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    aes(y = after_stat(density)),
    position = "stack",
    bins = 40
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 13),
    )
  + scale_fill_paletteer_d(
    "MetBrewer::Derain",
    na.value = "transparent",
  )
  + labs(
    x = "# days temperature above 32 (dev)",
    y = "Density",
  )
)
```

### Dev weighted temperature above 32

```{r}
(
  ggplot(
    data = df_clim_12m %>% filter(date > "2015-01-01"), 
    aes(
      x = dev_value_weighted_temp_2m_max_above_32_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    aes(y = after_stat(density)),
    position = "stack",
    bins = 40
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 13),
    )
  + scale_fill_paletteer_d(
    "MetBrewer::Derain",
    na.value = "transparent",
  )
  + labs(
    x = "# days weighted temperature above 32 (dev)",
    y = "Density",
  )
)
```

### Temperature above 1sd

```{r}
(
  ggplot(
    data = df_clim_12m %>% filter(date > "2015-01-01"), 
    aes(
      x = value_temp_2m_max_above_1sd_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    aes(y = after_stat(density)),
    position = "stack",
    bins = 30
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 13),
    )
  + scale_fill_paletteer_d(
    "MetBrewer::Derain",
    na.value = "transparent",
  )
  + labs(
    x = "# days temperature above 1 sd",
    y = "Density",
  )
)
```

### Weighted temperature above 1 sd

```{r}
(
  ggplot(
    data = df_clim_12m %>% filter(date > "2015-01-01"), 
    aes(
      x = value_weighted_temp_2m_max_above_1sd_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    aes(y = after_stat(density)),
    position = "stack",
    bins = 30
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 13),
    )
  + scale_fill_paletteer_d(
    "MetBrewer::Derain",
    na.value = "transparent",
  )
  + labs(
    x = "# days weighted temperature above 1 sd",
    y = "Density",
  )
)
```

### Dev temperature above 1 sd

```{r}
(
  ggplot(
    data = df_clim_12m %>% filter(date > "2015-01-01"), 
    aes(
      x = dev_value_temp_2m_max_above_1sd_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    aes(y = after_stat(density)),
    position = "stack",
    bins = 40
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 13),
    )
  + scale_fill_paletteer_d(
    "MetBrewer::Derain",
    na.value = "transparent",
  )
  + labs(
    x = "# days temperature above 1sd (dev)",
    y = "Density",
  )
)
```

### Dev weighted temperature above 1 sd

```{r}
(
  ggplot(
    data = df_clim_12m %>% filter(date > "2015-01-01"), 
    aes(
      x = dev_value_weighted_temp_2m_max_above_1sd_12m, 
      fill = country
    ),
    color = "black"
  )
  + geom_histogram(
    aes(y = after_stat(density)),
    position = "stack",
    bins = 40
  )
  + theme(
      panel.background = element_blank(),
      panel.grid = element_blank(),
      panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
      plot.background = element_blank(),
      plot.title = element_text(size = 16, hjust = 0.5),
      plot.subtitle = element_text(size = 14, hjust = 0.5),
      axis.title = element_text(size = 16),
      axis.text.y = element_text(size = 15),
      axis.text.x = element_text(size = 15),
      legend.background = element_blank(),
      legend.title = element_blank(),
      legend.text = element_text(size = 13),
    )
  + scale_fill_paletteer_d(
    "MetBrewer::Derain",
    na.value = "transparent",
  )
  + labs(
    x = "# days weighted temperature above 1 sd (dev)",
    y = "Density",
  )
)
```

### Temperature above 1 sd time series

```{r}
#| fig-width: 9
#| fig-height: 5

(
  df_clim_12m
  %>% filter(
    date > "2010-01-01",
    date < "2025-01-01",
    country == "Mozambique",
    !(adm_name %in% c("Tete", "Zambezia"))
  )
  %>% select(
    date,
    adm_name,
    value_temp_2m_max_above_1sd_12m,
  )
  %>% ggplot(
    mapping = aes(x = date, y = value_temp_2m_max_above_1sd_12m)
  )
  + geom_line(
    color = "#0292b7",
    linewidth = 1.3
  )
  + facet_wrap(
    ~ adm_name,
    ncol = 3
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    panel.spacing.x = unit(25, "pt"),
    strip.background = element_blank(),
    strip.text = element_text(size = 14),
    plot.background = element_blank(),
    plot.title = element_text(size = 16),
    plot.margin = margin(t = 0, r = 20, b = 0, l = 10, unit = "pt"),
    axis.ticks = element_line(linewidth = 1),
    axis.ticks.length = unit(4, "pt"),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 15),
  )
  + labs(
    x = "Date",
    y = "# days temperature max above 1 sd (month)"
  )
)
ggsave(
  path = fig_dir,
  filename = "temp_2m_max_above_1sd.png",
  width = 9,
  height = 5
)
```

### Weighted temperature above 1 sd time series

```{r}
#| fig-width: 9
#| fig-height: 5

(
  df_clim_12m
  %>% filter(
    date > "2010-01-01",
    date < "2025-01-01",
    country == "Mozambique",
    !(adm_name %in% c("Tete", "Zambezia"))
  )
  %>% select(
    date,
    adm_name,
    value_weighted_temp_2m_max_above_1sd_12m,
  )
  %>% ggplot(
    mapping = aes(x = date, y = value_weighted_temp_2m_max_above_1sd_12m)
  )
  + geom_line(
    color = "#0292b7",
    linewidth = 1.3
  )
  + facet_wrap(
    ~ adm_name,
    ncol = 3
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    panel.spacing.x = unit(25, "pt"),
    strip.background = element_blank(),
    strip.text = element_text(size = 14),
    plot.background = element_blank(),
    plot.title = element_text(size = 16),
    plot.margin = margin(t = 0, r = 20, b = 0, l = 10, unit = "pt"),
    axis.ticks = element_line(linewidth = 1),
    axis.ticks.length = unit(4, "pt"),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 15),
  )
  + labs(
    x = "Date",
    y = "# days weighted_temperature max above 1 sd"
  )
)
ggsave(
  path = fig_dir,
  filename = "weighted_temp_2m_max_above_1sd.png",
  width = 9,
  height = 5
)
```

### Dev temperature above 1 sd time series

```{r}
#| fig-width: 9
#| fig-height: 5

(
  df_clim_12m
  %>% filter(
    date > "2010-01-01",
    date < "2025-01-01",
    country == "Mozambique",
    !(adm_name %in% c("Tete", "Zambezia"))
  )
  %>% select(
    date,
    adm_name,
    dev_value_temp_2m_max_above_1sd_12m,
  )
  %>% ggplot(
    mapping = aes(x = date, y = dev_value_temp_2m_max_above_1sd_12m)
  )
  + geom_line(
    color = "#dc7726",
    linewidth = 1.3
  )
  + facet_wrap(
    ~ adm_name,
    ncol = 3
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    panel.spacing.x = unit(25, "pt"),
    strip.background = element_blank(),
    strip.text = element_text(size = 14),
    plot.background = element_blank(),
    plot.title = element_text(size = 16),
    plot.margin = margin(t = 0, r = 20, b = 0, l = 10, unit = "pt"),
    axis.ticks = element_line(linewidth = 1),
    axis.ticks.length = unit(4, "pt"),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 15),
  )
  + labs(
    x = "Date",
    y = "# days temperature max above 1 sd (dev)"
  )
)
ggsave(
  path = fig_dir,
  filename = "dev_temp_2m_max_above_1sd.png",
  width = 9,
  height = 5
)
```

### Dev weighted temperature above 1 sd time series

```{r}
#| fig-width: 9
#| fig-height: 5

(
  df_clim_12m
  %>% filter(
    date > "2010-01-01",
    date < "2025-01-01",
    country == "Mozambique",
    !(adm_name %in% c("Tete", "Zambezia"))
  )
  %>% select(
    date,
    adm_name,
    dev_value_weighted_temp_2m_max_above_1sd_12m,
  )
  %>% ggplot(
    mapping = aes(x = date, y = dev_value_weighted_temp_2m_max_above_1sd_12m)
  )
  + geom_line(
    color = "#dc7726",
    linewidth = 1.3
  )
  + facet_wrap(
    ~ adm_name,
    ncol = 3
  )
  + theme(
    panel.background = element_blank(),
    panel.border = element_rect(fill = NA, color = "black", linewidth = 1),
    panel.grid = element_blank(),
    panel.spacing.x = unit(25, "pt"),
    strip.background = element_blank(),
    strip.text = element_text(size = 14),
    plot.background = element_blank(),
    plot.title = element_text(size = 16),
    plot.margin = margin(t = 0, r = 20, b = 0, l = 10, unit = "pt"),
    axis.ticks = element_line(linewidth = 1),
    axis.ticks.length = unit(4, "pt"),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 15),
  )
  + labs(
    x = "Date",
    y = "# days weighted temperature max above 1 sd (dev)"
  )
)
ggsave(
  path = fig_dir,
  filename = "dev_weighted_temp_2m_max_above_1sd.png",
  width = 9,
  height = 5
)
```

# Save data

```{r}
df_clim_12m %>% write_csv(file.path(proc_data_dir, "processed_utci_temp_spei_daily.csv"))
```
