---
title: "Process Food Prices"
format: html
editor: visual
---

# Environment Setup

```{r setup}
# Packages
library(tidyverse)
library(sf)
library(patchwork)
library(paletteer)
library(spdep)

# Folder Structure
root <- rprojroot::has_file("agile2025.Rproj")$make_fix_file()
raw_data_dir <- root("data/raw")
proc_data_dir <- root("data/processed/food_prices")
dir.create(proc_data_dir, showWarnings = FALSE)
fig_dir <- root("figures")
dir.create(fig_dir, showWarnings = FALSE)

# Settings
nb_name <-  "Process Food Prices"
seed <- sum(sapply(str_split(nb_name, pattern = "", simplify = TRUE), FUN = utf8ToInt))
set.seed(seed)

# Country Mapping
country_map <- c(
  "MOZ" = "Mozambique",
  "ZWE" = "Zimbabwe",
  "LSO" = "Lesotho",
  "NAM" = "Namibia",
  "ZMB" = "Zambia"
)
```

# Load Data

## Geographies

### Mozambique

```{r}
sf_moz <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Mozambique 2022 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Zimbabwe

```{r}
sf_zwe <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Zimbabwe 2015 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Lesotho

```{r}
sf_lso <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Lesotho 2023 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Namibia

```{r}
sf_nam <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Namibia 2013 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Zambia

```{r}
sf_zmb <- (
  st_read(
    file.path(raw_data_dir, "geographies", "Zambia 2018 DHS - sdr_subnational_boundaries", "shps", "sdr_subnational_boundaries.shp")
  )
  %>% rename_with(tolower)
  %>% select(
    country = cntrynamee,
    adm_name = dhsregen,
    adm_code = regcode
  )
  %>% arrange(adm_code)
)
```

### Create combined shapes

```{r}
sf_all <- (
  bind_rows(
    sf_moz,
    sf_zwe,
    sf_lso,
    sf_nam,
    sf_zmb
  )
  %>% st_make_valid()
)
```

## Crop cycles

```{r}
df_crops <- (
  read_csv(
    file.path(
      raw_data_dir, 
      "crop",
      "harvestStat",
      "hvstat_africa_data_v1.0.csv"
    ),
    show_col_types = FALSE
  )
  %>% mutate(
    crop_production_system = if_else(
      country %in% c("Uganda", "Zambia"),
      "All (PS)",
      crop_production_system
    )
  )
  %>% filter(
    country %in% country_map,
    qc_flag == 0,
    product %in% c("Maize", "Sorghum", "Wheat", "Cassava", "Millet", "Rice", "Sweet Potatoes", "Potato", "Banana"),
    !is.na(production),
    harvest_year >= 2000,
    crop_production_system == "All (PS)"
  )
  %>% group_by(country, admin_1, product)
  %>% summarise(
    production = sum(production, na.rm = TRUE),
    .groups = "drop"
  )
  %>% group_by(country, admin_1)
  %>% slice_max(
    order_by = production,
    n = 1,
    with_ties = FALSE,
  )
  %>% ungroup()
  %>% rename(
    adm_name = admin_1
  )
)
```

```{r}
df_crops
```

Maize is usually the most commom product. For simplicity we use this to track price changes

## Food Prices

Data from: 
https://microdata.worldbank.org/catalog/4500/get-microdata
https://microdata.worldbank.org/catalog/4483/study-description#study_desc1684335103457 (Data Preview)

```{r}
df_mkt_loc <- (
  read_csv(
    file.path(
      raw_data_dir,
      "food_prices",
      "MOZ_RTFP_mkt_2007_2026-03-24.csv"
    ),
    show_col_types = FALSE
  )
  %>% distinct(
    country,
    adm_name = adm1_name,
    market = mkt_name,
    lat,
    lon
  )
)
```


```{r}
sf_price_index <- (
  read_csv(
    file.path(
      raw_data_dir,
      "food_prices",
      "world_bank_moz_2007_2025.csv"
    ),
    show_col_types = FALSE
  )
  %>% rename_with(str_to_lower)
  %>% filter(
    product == "food_price_index",
    region != "Market Average"
  )
  %>% select(
    country,
    date,
    adm_name = region,
    market,
    max_food_price_index = high
  )
  %>% merge(
    df_mkt_loc,
    by = c("country", "adm_name", "market"),
    all.x = TRUE
  )
  %>% select(
    -c(country, adm_name)
  )
  %>% st_as_sf(
    coords = c("lon", "lat"),
    crs = 4326
  )
)
```

# Process Data

## Combine datasets

```{r}
df_price_index_adm <- (
  sf_price_index
  %>% st_join(
    sf_all,
    join = st_nearest_feature,
    left = TRUE
  )
  %>% arrange(
    country,
    adm_code,
    adm_name,
    date
  )
  %>% as_tibble()
  %>% group_by(
    country,
    adm_code,
    adm_name,
    date = lubridate::ceiling_date(date, "month") - lubridate::days(1)
  )
  %>% summarise(
    max_food_price_index = mean(max_food_price_index, na.rm = TRUE),
    .groups = "drop"
  )
)
```

## Detrend data

### Breakpoints

```{r}
df_price_index_summ <- (
  df_price_index_adm
  %>% group_by(date)
  %>% summarise(
    max_food_price_index = mean(max_food_price_index, na.rm = TRUE)
  )
  %>% ungroup()
  %>% arrange(date)
  %>% mutate(
    t = row_number()
  )
)
  
breakpoints <- strucchange::breakpoints(
  max_food_price_index ~ t,
  data = df_price_index_summ,
  h = 0.15
)
breakpoints_ids <- breakpoints$breakpoints
breakpoints_dates <- df_price_index_summ$date[breakpoints_ids]
```

### Remove trends by region

```{r}
df_price_index_adm_detrend <- (
  df_price_index_adm
  %>% arrange(adm_name, date)
  %>% group_by(adm_name)
  %>% mutate(
    t = row_number()
  )
  %>% ungroup()
  %>% group_by(adm_name)
  %>% group_modify(~{
    fit0 <- lm(max_food_price_index ~ t, data = (.x))
    fit_seg <- segmented::segmented(
      fit0,
      seg.Z = ~t,
      psi = breakpoints_ids,
      control = segmented::seg.control(it.max = 0, display = FALSE)
    )
    (
      .x
      %>% mutate(
        trend_mu = fitted(fit_seg),
        max_food_price_index = max_food_price_index - trend_mu
      )
    )
  })
  %>% ungroup()
)
```

```{r}
(
  df_price_index_adm_detrend
  %>% ggplot(
    aes(x = date, y = max_food_price_index, color = adm_name)
  )
  + geom_line()
)
```

## Calculate anomalies

```{r}
df_price_index_adm_12m <- (
  df_price_index_adm_detrend
  %>% group_by(country, date)
  %>% mutate(
    mean_country_price_index = mean(max_food_price_index, na.rm = TRUE),
    sd_country_price_index = sd(max_food_price_index, na.rm = TRUE)
  )
  %>% ungroup()
  %>% group_by(country, adm_code, adm_name)
  %>% mutate(
    hist_mean_price_index = mean(max_food_price_index, na.rm = TRUE),
    hist_sd_price_index = sd(max_food_price_index, na.rm = TRUE),
    hist_mean_sd_country_price_index = mean(sd_country_price_index, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    price_index_hist_anomaly = as.numeric(max_food_price_index > hist_mean_price_index + hist_sd_price_index),
    price_index_reg_anomaly = as.numeric(max_food_price_index > mean_country_price_index + hist_mean_sd_country_price_index)
  )
  %>% arrange(country, adm_code, adm_name, date)
  %>% group_by(country, adm_code, adm_name)
  %>% mutate(
    max_price_index_12m = slider::slide_index_dbl( 
      .x = max_food_price_index,
      .i = date,
      .f = ~ if (sum(!is.na(.x)) == 0) NA else max(.x, na.rm = TRUE),
      .before = ~ .x %m-% months(12),
      .complete = TRUE
    ),
    price_index_hist_anomaly_12m = slider::slide_index_dbl(
      .x = price_index_hist_anomaly,
      .i = date,
      .f = ~ sum(.x, na.rm = TRUE),
      .before = ~ .x %m-% months(12),
      .complete = TRUE
    ),
    price_index_reg_anomaly_12m = slider::slide_index_dbl(
      .x = price_index_reg_anomaly,
      .i = date,
      .f = ~ sum(.x, na.rm = TRUE),
      .before = ~ .x %m-% months(12),
      .complete = TRUE
    )
  )
  %>% ungroup()
  %>% mutate(
    price_index_stress_12m = as.integer(max_price_index_12m > hist_mean_price_index + hist_sd_price_index)
  )
  %>% select(country, adm_code, adm_name, date, contains("12m"))
  %>% filter(!is.na(price_index_stress_12m))
  %>% pivot_longer(
    cols = -c(country, adm_code, adm_name, date),
    names_to = "variable",
    values_to = "value"
  ) 
  %>% group_by(country, adm_code, adm_name, variable)
  %>% mutate(
    mean_value = mean(value, na.rm = TRUE),
    sd_value = sd(value, na.rm = TRUE)
  )
  %>% ungroup()
  %>% mutate(
    dev_value = if_else(
      sd_value == 0,
      0,
      (value - mean_value) / sd_value
    )
  )
  %>% filter(!is.na(value))
  %>% pivot_wider(
    id_cols = c(country, adm_code, adm_name, date),
    names_from = "variable",
    values_from = c(value, dev_value)
  )
)
```

# Visualizations

```{r}
(
  df_price_index_adm_12m
  %>% ggplot(
    aes(x = date, y = value_price_index_stress_12m, color = adm_name)
  )
  + geom_line()
)
```
```{r}
(
  df_price_index_adm_12m
  %>% ggplot(
    aes(x = date, y = dev_value_price_index_hist_anomaly_12m, color = adm_name)
  )
  + geom_line()
)
```


```{r}
(
  df_price_index_adm_12m
  %>% ggplot(
    aes(x = date, y = dev_value_price_index_reg_anomaly_12m, color = adm_name)
  )
  + geom_line()
)
```

# Save

```{r}
df_price_index_adm %>% write_csv(file.path(proc_data_dir, "world_bank_food_prices_adm.csv"))
df_price_index_adm_detrend %>% write_csv(file.path(proc_data_dir, "world_bank_food_prices_adm_detrended.csv"))
df_price_index_adm_12m %>% write_csv(file.path(proc_data_dir, "world_bank_food_prices_adm_12m.csv"))
```
